### 2014-03-11 paper/main.tex
added paper;

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{fullpage}

\begin{document}
\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
\date{}
\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
  your security assumptions when dealing with untrusted input-}
\maketitle

 \abstract{We present \textit{Nail},
  an interface generator that allows programmers to safely parse and generate protocols defined by a
  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
  internal representation, obviating the need to write semantic actions. Nail also provides
  solutions parsing common patterns such as length and offset fields within binary formats that are
  hard to process with existing parser generators.}
\section{Background}

% For a long time, it has been possible to write a concise and precise specification of what a data
% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
% that format. However, instead of using these tools, programmers have continued to laboriously
% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
% although the underlying algorithms remain the same.

\subsection{Problems handling untrusted input}
Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
code are most likely to be exploitable, because an attacker has total control over the inputs to
that code.

Common wisdom thus says to take special care when preparing code that deals with untrusted input.
However, with currently prevalent software design methods, it is hard to isolate input-handling code
from the rest of the program, as user input is typically passed through the program and processed in
bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.

A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
and the security problems they cause are rampant. Input processing libraries such as libpng
\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
historically contained multiple security
flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.

Besides memory corruption vulnerabilities - against which various mitigations such as static
analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
different implementations of the same format can also differ in their understanding of some edge
cases. The security of many systems relies on different components interpreting the same data
similarly.

Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
  differently, so the code signature sees different bytes at a virtual address than the executable
  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
  files. Signatures are verified with a Java program, but the program is extracted with a C program.
The java program interprets all fields as signed, whereas the C program treats them as unsigned,
allowing one to replace files in a signed archive.}
\cite{saurik-masterkey} code signing and even the
X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
runtime environment or static analysis guarantee absolute memory safety.

Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
provide a full Turing-complete execution environment. Examples include X86 page tables
\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
The more powerful an input
protocol is, the more control an attacker gains over a systems state - and power to manipulate a
systems state is also a degree of undeserved trust. In the offensive research community, this has
been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
operates on an input, analogous to a virtual machine operating on bytecode.

Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.

\subsection{Parser generators}
%%TODO
Luckily, we can avoid all of these problems at once if we have a single, executable specification of
what exactly a programs input is supposed to look like. In fact, definitive grammars have long
been written for all sorts of protocols and are the standard way of describing text-based protocols.
Parser generators can automatically create parsers for such languages, and with common classes of
grammars such as regular expressions or deterministic context free languages, that parser can be
expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}

Specifications of binary protocols, such as RFCs, also contain grammars, although they are
frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.

Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
the security-critical and very well-engineered MIT Kerberos distribution uses parser
generators, but only for handling configuration files. A notable exception is the Mongrel
web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.

 Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
 While previous parser generators could also be used to write grammars for binary
 protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
 inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
 of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
 library. Hammer implements grammars as language-integrated parser combinators, an approach
 popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
 knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
 top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).

Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
action with every rule in the grammar.

Semantic actions are one source of vulnerability within automatically generated parsers, as they are
typically implemented in C or another general purpose programming language and perform dangerous
operations such as memory allocation and copying. If there is a bug in the semantic actions, the
attacker is likely able to craft an input that invokes the semantic actions with controllable
inputs, as the inputs have not been fully verified yet.

\subsection{Motivation}
\paragraph{Motivation}
While it is possible to express short transformations on the input entirely as semantic actions
\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
  walks the parse tree.}, more complicated programs usually construct an internal representation,
which contains all relevant information from the input in a format native to the programming
language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
Haskell programmer records and a LISP programmer property lists. 
 The structure of this internal representation usually resembles the structure of the grammar.

Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
use a parser generator. Once, to write a grammar, another time to write the semantic actions
constructing his favourite intermediate representation. In most languages the programmer also has to
write explicit type definitions, which describe the intermediate object model again. Most
programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
passing user input through their control flow, hopefully remembering to verify each part before
it is accessed.

A similar problem occurs when the programmer wants to generate output. 
Even though not all programs may use the same format for output as they do on input, different
programs might use the same format for input and output and re-using the same grammar is a good way
to save engineering effort and reduce parser differentials.
Some parser generators, e.g.
Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
allow the same grammar to be re-used for generating output from the intermediate representation.
However, those generators require a new set of semantic actions to be written, even though the
relationship between grammar and intermediate model was already expressed in the actions for the
parser.

\section{Design}

\paragraph{Overview}
To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
an internal model, while also describing the grammar of the external protocol. Nail takes the
combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
bytes from an instance of the model.

A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
so parsing the generators output will yield the generators input, but generating the parsers output
does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
  sequence and the actual contents of the transmission.}. However, if some features of Nail are
avoided, there is a proper isomorphism between model and protocol.

In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
by the combinator of conventional Parser Expression Grammars they replace.
\paragraph{Fundamental parsers}
The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
Examples: \texttt{uint4}, \texttt{int32}

Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
\texttt{int32| [1,255,512]}

\paragraph{Constant parser}
Nail also has constant parsers, which do not return a value. The simplest constant parser is an
integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
parser returns the given value and fail otherwise.

\paragraph{Sequence: Structure}
Nails fundamental concept is the structure combinator. It contains a list of named parsers and
unnamed constant parsers. 
When parsing, it parses each parser in sequence and returns a structure containing the result of
each parser.

\begin{tabular}{ll}
\textbf{Nail} & \textbf{Data Model}\\
\begin{minipage}{2in}
\begin{verbatim}
header = {
 id uint16
 qr uint1
 opcode uint4
 aa uint1 
 tc uint1
 rd uint1
 ra uint1
 uint3 = 0
 rcode uint4
}
\end{verbatim} 
\end{minipage}
 & 
\begin{minipage}{2in}
\begin{verbatim}
struct header{
    uint16_t id;
    uint8_t qr;
    uint8_t opcode;
    uint8_t aa;
    uint8_t tc;
    uint8_t rd;
    uint8_t ra;
    uint8_t rcode;
};
\end{verbatim} 
\end{minipage} 
\\
\end{tabular}
\paragraph{Repetition: Many,SepBy}
The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
parsing two values,but not before parsing the first value or after parsing the last.\\
Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
comma-separated lists of numbers).

The many combinator also works on constant parsers, although as described above it can make the
generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
  constant parser in its output.} Example: \texttt{many uint8= "test"}

\paragraph{Semantic choice: choose}
We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
for each choice. The result is saved in a tagged union.

\begin{verbatim}
choose {
   A = uint8 | 1.. 8
   B = uint16 | ..256
}
\end{verbatim}
\paragraph{Wrap}

The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
and then again parses one or more constant parsers.\\ Example: 
\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+

\subsection{Offset fields}
Another problem for parser generators is that binary protocols often contain length and offset
fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
the automaton that will just be skipped - and very cumbersome to express with current parser
generators. Therefore, if languages with offset fields are parsed with parser generators, the only
currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
parser on the fly from semantic actions.

Nail will properly support both offset and length fields and much of the following discussion
applies to both , although the current prototype only implements lengths, which we will focus on.

\paragraph{Dependency fields}
We call length or offset fields \textit{dependency fields}, because during parsing, another parser
depends on them, and while generating output, their value depends on some other structure in the
data model.
Dependency fields appear in a structure combinator as would any other field, but their name begins
with an '@' sign and they have to return an integer value. Dependency fields are not directly
represented in the internal model.

\paragraph{Length fields: n_of}
The length combinator takes a dependency field and a parser, evaluating the parser as often as the
value of the dependency field, returning an array of the values. When generating output, it emits
the array and writes its length to the dependency field.

\subsection{Example Grammars}
To further familiarise the reader with Nail, we provide two example grammars. 
\paragraph{UTF-16}
This grammar recognises valid UTF-16 strings and exposes an array of code points.

\begin{tabular}{ll}
\textbf{Nail} & \textbf{Data Model}\\
\begin{minipage}{3in}
\begin{verbatim}
utfstring = many choose {
         SUPP= {
               lead uint16 | 55296..56319
                          // 0xD800..0xDBFF
               trail uint16 | 56320..57343
                          // 0xDC00..0xDFFF
               }
         BASIC = uint16 | !55296..57343
       }
\end{verbatim}
\end{minipage}
 & 
\begin{minipage}{2in}
\begin{verbatim}
struct utfstring {
 struct {
   enum  {SUPP,BASIC} N_type;
   union {
    struct {
     uint16_t lead;
     uint16_t trail;
    } SUPP;
    uint16_t BASIC;
   };
 }*elem;
 size_t count;
};
\end{verbatim} 
\end{minipage} 
\\
\end{tabular}

\paragraph{DNS packet}
This grammar recognises DNS packets without label compression, as per RFC1035 
\begin{verbatim}
labels = <many {@length  uint8 | 1..255 
                label n_of @length uint8 }
             uint8 = 0>
question= {   labels labels
              qtype uint16 | 1..16 
              qclass uint16 | [1,255]
               }
answer  = {
             labels labels
             rtype uint16 | 1..16
             class uint16 | [1]
             ttl uint32
             @rlength uint16 
             rdata n_of @rlength uint8 
             }
dnspacket = {id uint16
              qr uint1
              opcode uint4
              aa uint1 
              tc uint1
              rd uint1
              ra uint1
              uint3 = 0
              rcode uint4
              @questioncount uint16
              @answercount uint16
              authoritycount uint16   // Ignored
              additionalcount uint16  //Ignored       
              questions n_of @questioncount question         
              responses n_of @answercount answer
            }
\end{verbatim}

\section{Implementation}
The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
features of our parser implementation.

\paragraph{Parsing}
A generated Nail parser makes two passes through the input , the first to validate and recognise the
input, the second to bind this data to the internal model. Currently the parser is a straightforward
top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
\paragraph{Memory allocation} 
Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
result and has to be freed by the programmer.

\paragraph{Intermediate representation}
Most parser generators, such as Bison do not have to dynamically allocate temporary
data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
computation as possible without having validated the input, and we do not want to mix temporary
objects with the results of our parse, we need some intermediate structure to cache parse results.
This structure is also used as a value in the Packrat hash table.

Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
tree is at least an order of magnitude larger than the input, because it stores a large tree node
structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
all necessary information without ever seeing the raw input stream. However, because we also
automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
it as much as we trust the parser and thus expose it to the raw input stream.

Under this premise, the actions need very limited information from the recogniser to correctly
handle the input stream. In particular, the parsers control flow only branches at the choice and
repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
fundamental parsers to a fixed sequence of internal model fields. 
Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
\paragraph{Dependency Fields}
During parsing, dependency fields occur before the context in which they are used. The parser stores
their value and retrieves it afterwards when encountering the combinator that uses them. When
generating output, the dependency field is first filled with a filler value, then later when the
first combinator that determines this fields value is encountered, the field is overwritten. Any
further combinators using this dependency will then validate that the dependency field is correct. 
\paragraph{Bootstrapping}
To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
to recognise Nail grammars. A superset of the grammar language described in this paper is
implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
actual parser generator. Bootstrapping is supported via an LALR implementation.
\section{Evaluation}
In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
framework itself that responds to any valid DNS query with a CNAME record to the domain
``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
structure definitions.

Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
functional toy clone of the host command line tool. However, because our grammar does not yet
support DNS label compression, the latter tool will occasionally reject valid real world DNS
responses. Both clients have functional anti-spoofing measures.

It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
particular for DNS compression and additional hint records than real world DNS servers. 
However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
intend to do so.

\section{Future Work}

The Nail parser generator is work in progress and feedback of all sorts is heavily
encouraged. Source code is available upon request and will be released publicly pending clean-up and
maturity. Short-term future work will include improving the scoping of dependency fields and adding
support for offset fields.

A major problem with the current design of Nail is that the design of the grammar dictates the
interna structure of the software. This makes changing grammars or adding Nail to existing software
awkward. One possible solution to this problem would be to implement a concept similar to Relational
Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
than two sets of semantic actions, while allowing changes in syntax, alternative representations
and adaption to legacy systems.

Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
throughout a network stack - from a user-space TCP stack to a PNG de-compressor.

\subsection{Previous work}
Generating parsers and generators from an executable specification is the core concept of Interface
Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
generators work very well for existing grammars, they do not allow full control over the format of
the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
implements a C like structure model enhanced with data-dependent length fields and constraints.
\begin{comment}

Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
 heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
 an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
 feng shui\cite{fengshui} to help bypass those mitigations.

Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
parsed at the same time (or, given a sufficiently complicated input, two structure within the same
grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
 If there is a buffer overflow past the
end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
placing a known value next to that input.

\end{comment}

\bibliography{main,locasto,login,phrack,short,langsec}
\bibliographystyle{plain}

\end{document}

### 2014-03-11 paper/paper.tex
added paper;

--- 2014-03-11 paper/main.tex
@@
-\documentclass[10pt]{article}
-\usepackage{graphicx}
-\usepackage{url}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{verbatim}
-\usepackage{fullpage}
+\documentclass{article}
 
 \begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
-\date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
+\authors{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek}
+
+\title{Nail: A practical interface generator for binary formats} \subtitle{-How to avoid hitting
   your security assumptions when dealing with untrusted input-}
-\maketitle
 
@@
 
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
-
 \subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
+Code handling un-trusted user inputs is most prone to contain security vulnerabilities. While memory
 corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
@@
 historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
+flaws\foot{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
 
@@
 \cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
+X.509 protocol underlying SSL \cite{pki-layer}. Such vulnerabilities occur even if the language,
 runtime environment or static analysis guarantee absolute memory safety.
@@
 provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
+\cite{bangert-woot}, ELF symbols and relocations\cite{shapiro-woot}.  
 The more powerful an input
@@
 systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
+been generalised into treating a program as a \textit{weird machine}\cite{Bratus-langsec} that
 operates on an input, analogous to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
 
@@
 expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
+limiting the weird machine problem and enabling very efficient implementation.
 
@@
 Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
+the security-critical and very well-engineered Kerberos distribution \cite{kerberos} uses parser
 generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
+webserver \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
 Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
+times to achieve better scalability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
 
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
+ Parser generators for binary protocols were first introduced by the Hammer\cite{hammer} parser.
  While previous parser generators could also be used to write grammars for binary
@@
  library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
+ popularised by Parsec for Haskell\cite{leijen2010parsec}. The parser combinator style (to our
+ knowledge, first  described in \cite{burge-recursive}) is a natural way of concisely expressing
+ top-down grammars \footnote{For more background on the history of expressing grammars, see Bryan
+   Ford's masters thesis\cite{packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
   Hammer then constructs a tree of function pointers which
@@
 \paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
-
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
-
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
-
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
-\begin{verbatim}
-header = {
- id uint16
- qr uint1
- opcode uint4
- aa uint1 
- tc uint1
- rd uint1
- ra uint1
- uint3 = 0
- rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct header{
-    uint16_t id;
-    uint8_t qr;
-    uint8_t opcode;
-    uint8_t aa;
-    uint8_t tc;
-    uint8_t rd;
-    uint8_t ra;
-    uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
-
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
-
-\begin{verbatim}
-choose {
-   A = uint8 | 1.. 8
-   B = uint16 | ..256
-}
-\end{verbatim}
-\paragraph{Wrap}
-
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
-
-\subsection{Offset fields}
+To alleviate these problems, Nail provides a richer set of combinators
+\paragraph{Offset fields}
 Another problem for parser generators is that binary protocols often contain length and offset
@@
 currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
+parser on the fly. However, 
+\paragraph{Hardened parser}
 
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
-
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
-
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
-
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars. 
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035 
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}
-
-\section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
-features of our parser implementation.
-
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
-\paragraph{Memory allocation} 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
-
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
-
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
-
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
-\paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
-\section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
-
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
-
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
-
-\section{Future Work}
-
-The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
-
-A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
-awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
-the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
-than two sets of semantic actions, while allowing changes in syntax, alternative representations
-and adaption to legacy systems.
-
-Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
-formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
-as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
-throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
+\paragraph{Hardening the parser}
+In addition to reducing the risk of memory corruption vulnerabilities within the generated parser,
+Nail also provides mitigations against heap corruption vulnerabilities in other parts of the
+program.
 
@@
 
-\end{comment}
-
-\bibliography{main,locasto,login,phrack,short,langsec}
-\bibliographystyle{plain}
-
 \end{document}

### 2014-03-31 paper/background.tex
split up main.tex into per-section files

--- 2014-03-11 paper/main.tex
@@
-\documentclass[10pt]{article}
-\usepackage{graphicx}
-\usepackage{url}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{verbatim}
-\usepackage{fullpage}
-
-\begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
-\date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
-\maketitle
-
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
 \section{Background}
@@
 parser.
-
-\section{Design}
-
-\paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
-
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
-
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
-
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
-\begin{verbatim}
-header = {
- id uint16
- qr uint1
- opcode uint4
- aa uint1 
- tc uint1
- rd uint1
- ra uint1
- uint3 = 0
- rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct header{
-    uint16_t id;
-    uint8_t qr;
-    uint8_t opcode;
-    uint8_t aa;
-    uint8_t tc;
-    uint8_t rd;
-    uint8_t ra;
-    uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
-
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
-
-\begin{verbatim}
-choose {
-   A = uint8 | 1.. 8
-   B = uint16 | ..256
-}
-\end{verbatim}
-\paragraph{Wrap}
-
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
-
-\subsection{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
-
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
-
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
-
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
-
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars. 
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035 
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}
-
-\section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
-features of our parser implementation.
-
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
-\paragraph{Memory allocation} 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
-
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
-
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
-
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
-\paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
-\section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
-
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
-
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
-
-\section{Future Work}
-
-The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
-
-A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
-awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
-the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
-than two sets of semantic actions, while allowing changes in syntax, alternative representations
-and adaption to legacy systems.
-
-Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
-formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
-as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
-throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}
-
-\bibliography{main,locasto,login,phrack,short,langsec}
-\bibliographystyle{plain}
-
-\end{document}

### 2014-03-31 paper/design.tex
split up main.tex into per-section files

--- 2014-03-11 paper/main.tex
@@
-\documentclass[10pt]{article}
-\usepackage{graphicx}
-\usepackage{url}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{verbatim}
-\usepackage{fullpage}
-
-\begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
-\date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
-\maketitle
-
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
-\section{Background}
-
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
-
-\subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
-
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
-
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
-
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
-operates on an input, analogous to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
-
-\subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
-
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
-
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
-
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
- can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
-
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
-
-\subsection{Motivation}
-\paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
-
-A similar problem occurs when the programmer wants to generate output. 
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
-
 \section{Design}
@@
 \end{verbatim}
-
-\section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
-features of our parser implementation.
-
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
-\paragraph{Memory allocation} 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
-
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
-
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
-
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
-\paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
-\section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
-
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
-
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
-
-\section{Future Work}
-
-The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
-
-A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
-awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
-the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
-than two sets of semantic actions, while allowing changes in syntax, alternative representations
-and adaption to legacy systems.
-
-Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
-formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
-as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
-throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}
-
-\bibliography{main,locasto,login,phrack,short,langsec}
-\bibliographystyle{plain}
-
-\end{document}

### 2014-03-31 paper/eval.tex
split up main.tex into per-section files

--- 2014-03-11 paper/main.tex
@@
-\documentclass[10pt]{article}
-\usepackage{graphicx}
-\usepackage{url}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{verbatim}
-\usepackage{fullpage}
-
-\begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
-\date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
-\maketitle
-
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
-\section{Background}
-
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
-
-\subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
-
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
-
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
-
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
-operates on an input, analogous to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
-
-\subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
-
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
-
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
-
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
- can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
-
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
-
-\subsection{Motivation}
-\paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
-
-A similar problem occurs when the programmer wants to generate output. 
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
-
-\section{Design}
-
-\paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
-
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
-
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
-
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
-\begin{verbatim}
-header = {
- id uint16
- qr uint1
- opcode uint4
- aa uint1 
- tc uint1
- rd uint1
- ra uint1
- uint3 = 0
- rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct header{
-    uint16_t id;
-    uint8_t qr;
-    uint8_t opcode;
-    uint8_t aa;
-    uint8_t tc;
-    uint8_t rd;
-    uint8_t ra;
-    uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
-
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
-
-\begin{verbatim}
-choose {
-   A = uint8 | 1.. 8
-   B = uint16 | ..256
-}
-\end{verbatim}
-\paragraph{Wrap}
-
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
-
-\subsection{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
-
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
-
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
-
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
-
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars. 
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035 
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}
-
-\section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
-features of our parser implementation.
-
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
-\paragraph{Memory allocation} 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
-
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
-
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
-
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
-\paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
 \section{Evaluation}
@@
 intend to do so.
-
-\section{Future Work}
-
-The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
-
-A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
-awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
-the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
-than two sets of semantic actions, while allowing changes in syntax, alternative representations
-and adaption to legacy systems.
-
-Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
-formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
-as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
-throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}
-
-\bibliography{main,locasto,login,phrack,short,langsec}
-\bibliographystyle{plain}
-
-\end{document}

### 2014-03-31 paper/future.tex
split up main.tex into per-section files

--- 2014-03-11 paper/main.tex
@@
-\documentclass[10pt]{article}
-\usepackage{graphicx}
-\usepackage{url}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{verbatim}
-\usepackage{fullpage}
-
-\begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
-\date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
-\maketitle
-
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
-\section{Background}
-
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
-
-\subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
-
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
-
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
-
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
-operates on an input, analogous to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
-
-\subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
-
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
-
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
-
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
- can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
-
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
-
-\subsection{Motivation}
-\paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
-
-A similar problem occurs when the programmer wants to generate output. 
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
-
-\section{Design}
-
-\paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
-
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
-
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
-
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
-\begin{verbatim}
-header = {
- id uint16
- qr uint1
- opcode uint4
- aa uint1 
- tc uint1
- rd uint1
- ra uint1
- uint3 = 0
- rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct header{
-    uint16_t id;
-    uint8_t qr;
-    uint8_t opcode;
-    uint8_t aa;
-    uint8_t tc;
-    uint8_t rd;
-    uint8_t ra;
-    uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
-
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
-
-\begin{verbatim}
-choose {
-   A = uint8 | 1.. 8
-   B = uint16 | ..256
-}
-\end{verbatim}
-\paragraph{Wrap}
-
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
-
-\subsection{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
-
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
-
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
-
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
-
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars. 
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035 
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}
-
-\section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
-features of our parser implementation.
-
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
-\paragraph{Memory allocation} 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
-
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
-
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
-
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
-\paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
-\section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
-
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
-
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
-
 \section{Future Work}
@@
 \end{comment}
-
-\bibliography{main,locasto,login,phrack,short,langsec}
-\bibliographystyle{plain}
-
-\end{document}

### 2014-03-31 paper/impl.tex
split up main.tex into per-section files

--- 2014-03-11 paper/main.tex
@@
-\documentclass[10pt]{article}
-\usepackage{graphicx}
-\usepackage{url}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{underscore}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{verbatim}
-\usepackage{fullpage}
-
-\begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
-\date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
-\maketitle
-
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
-\section{Background}
-
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
-
-\subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
-
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
-
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
-
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
-operates on an input, analogous to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
-
-\subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
-
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
-
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
-
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
- can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
-
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
-
-\subsection{Motivation}
-\paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
-
-A similar problem occurs when the programmer wants to generate output. 
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
-
-\section{Design}
-
-\paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
-
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
-
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
-
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
-\begin{verbatim}
-header = {
- id uint16
- qr uint1
- opcode uint4
- aa uint1 
- tc uint1
- rd uint1
- ra uint1
- uint3 = 0
- rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct header{
-    uint16_t id;
-    uint8_t qr;
-    uint8_t opcode;
-    uint8_t aa;
-    uint8_t tc;
-    uint8_t rd;
-    uint8_t ra;
-    uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
-
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
-
-\begin{verbatim}
-choose {
-   A = uint8 | 1.. 8
-   B = uint16 | ..256
-}
-\end{verbatim}
-\paragraph{Wrap}
-
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
-
-\subsection{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
-
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
-
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
-
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
-
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars. 
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035 
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}
-
 \section{Implementation}
@@
 actual parser generator. Bootstrapping is supported via an LALR implementation.
-\section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
-
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
-
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
-
-\section{Future Work}
-
-The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
-
-A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
-awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
-the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
-than two sets of semantic actions, while allowing changes in syntax, alternative representations
-and adaption to legacy systems.
-
-Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
-formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
-as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
-throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}
-
-\bibliography{main,locasto,login,phrack,short,langsec}
-\bibliographystyle{plain}
-
-\end{document}

### 2014-03-31 paper/main.tex
split up main.tex into per-section files

@@
 \begin{document}
-\author{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek} 
+\author{Julian Bangert and Nickolai Zeldovich}
 \date{}
-\title{Nail: A practical interface generator for binary formats \\ \small-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
+\title{Nail: A practical interface generator for binary formats}
+
 \maketitle
 
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
-\section{Background}
-
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
-
-\subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
-
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
-
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
-
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
-operates on an input, analogous to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
-
-\subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
-
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
-
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
-
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
- can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
-
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
-
-\subsection{Motivation}
-\paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
-
-A similar problem occurs when the programmer wants to generate output. 
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
-
-\section{Design}
-
-\paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
-
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
-
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
-
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
-\begin{verbatim}
-header = {
- id uint16
- qr uint1
- opcode uint4
- aa uint1 
- tc uint1
- rd uint1
- ra uint1
- uint3 = 0
- rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct header{
-    uint16_t id;
-    uint8_t qr;
-    uint8_t opcode;
-    uint8_t aa;
-    uint8_t tc;
-    uint8_t rd;
-    uint8_t ra;
-    uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
-
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
-
-\begin{verbatim}
-choose {
-   A = uint8 | 1.. 8
-   B = uint16 | ..256
-}
-\end{verbatim}
-\paragraph{Wrap}
-
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
-
-\subsection{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
-
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
-
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
-
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
-
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars. 
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
-
-\begin{tabular}{ll}
-\textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-\end{minipage}
- & 
-\begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-\end{minipage} 
-\\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035 
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}
-
-\section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
-features of our parser implementation.
-
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
-\paragraph{Memory allocation} 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
-
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
-
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
-
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
-\paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
-\section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
-
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
-
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
-
-\section{Future Work}
-
-The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
-
-A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
-awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
-the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
-than two sets of semantic actions, while allowing changes in syntax, alternative representations
-and adaption to legacy systems.
-
-Finally, we would like to demonstrate the capabilities of Nail by implementing various binary
-formats 'notorious' for their insecurity in Nail. Nail was designed with the idioms of formats such
-as PDF and PNG in mind. Ultimately, we want to provide examples of successful Nail parsers all
-throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}
+\input{abstract}
+\input{background}
+\input{design}
+\input{impl}
+\input{eval}
+\input{future}
 

### 2014-04-01 paper/concl.tex
stub conclusion

\section{Conclusion}

This paper presented the initial design and implementation of
\textit{Nail}, an XXX...

### 2014-04-01 paper/future.tex
split off relwk

@@
-\section{Future Work}
+\section{Future work}
 
@@
 throughout a network stack - from a user-space TCP stack to a PNG de-compressor.
-
-\subsection{Previous work}
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}

### 2014-04-01 paper/relwk.tex
split off relwk

--- 2014-03-11 paper/paper.tex
@@
-\documentclass{article}
+\section{Related work}
 
-\begin{document}
-\authors{Julian Bangert, Nickolai Zeldovich, Frans Kaashoek}
+Generating parsers and generators from an executable specification is the core concept of Interface
+Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
+generators work very well for existing grammars, they do not allow full control over the format of
+the output, so cannot be used to implement legacy protocols.
+Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
+however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
+implements a C like structure model enhanced with data-dependent length fields and constraints.
 
-\title{Nail: A practical interface generator for binary formats} \subtitle{-How to avoid hitting
-  your security assumptions when dealing with untrusted input-}
-
- \abstract{We present \textit{Nail},
-  an interface generator that allows programmers to safely parse and generate protocols defined by a
-  Parser-Expression based grammar. Nail uses a richer set of parser combinators that induce an
-  internal representation, obviating the need to write semantic actions. Nail also provides
-  solutions parsing common patterns such as length and offset fields within binary formats that are
-  hard to process with existing parser generators.}
-\section{Background}
-
-\subsection{Problems handling untrusted input}
-Code handling un-trusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
-
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
-flaws\foot{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
-
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{pki-layer}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
-
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert-woot}, ELF symbols and relocations\cite{shapiro-woot}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{Bratus-langsec} that
-operates on an input, analogous to a virtual machine operating on bytecode.
-
-\subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation.
-
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
-
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered Kerberos distribution \cite{kerberos} uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-webserver \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scalability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
-
- Parser generators for binary protocols were first introduced by the Hammer\cite{hammer} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{leijen2010parsec}. The parser combinator style (to our
- knowledge, first  described in \cite{burge-recursive}) is a natural way of concisely expressing
- top-down grammars \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
- can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
-
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
-
-\subsection{Motivation}
-\paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
-
-A similar problem occurs when the programmer wants to generate output. 
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
-
-\section{Design}
-
-\paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators
-\paragraph{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly. However, 
-\paragraph{Hardened parser}
-
-\paragraph{Hardening the parser}
-In addition to reducing the risk of memory corruption vulnerabilities within the generated parser,
-Nail also provides mitigations against heap corruption vulnerabilities in other parts of the
-program.
+\begin{comment}
 
@@
 
-\end{document}
+\end{comment}

### 2014-04-01 generator/main.cc
rebase un-pushed changes

@@ -21,6 +21,8 @@ int main(int argc, char**argv)
     size_t outputsize;
     char *out;
     struct grammar *result;
+    std::string filename; 
+    std::string header_filename;
     inputsize = fread(input, 1, sizeof(input), infile(argc,argv));
     //fprintf(stderr, "inputsize=%zu\ninput=", inputsize);
     //fwrite(input, 1, inputsize, stderr);  
@@ -30,31 +32,22 @@ int main(int argc, char**argv)
     //    result = parse_grammar(input,inputsize);
     //    if(result)
     //      print_grammar(result, stderr,0);    exit(1);
+     std::string headerfilename = std::string(argv[1]) + ".h";
+     std::string implfilename = std::string(argv[1]) + ".c";
+     
      if(result) {
-       std::ofstream *outfile = NULL;
-       int i =0;
-       for(i=2;i<argc;i++){
-         if(argv[i][0] != '-'){
-           if(outfile)
-             delete outfile;
-           outfile = new std::ofstream(argv[i]);
-           if(!outfile || !outfile->is_open())
-             error("Cannot open output file %s\n",argv[i]);
-           continue;
-         }
-         else if(!strcmp(argv[i],"-header"))
-           {
-             error("Not implemented");
-             //emit_header(outfile,result);
-           }
-         else if(!strcmp(argv[i],"-parser_hammer")){
-            emit_parser(outfile,result);
-         }
-         else if(!strcmp(argv[i],"-generator"))
-            emit_generator(outfile,result);
-         else
-           error("Unknown command line option %s",argv[i]);
-       }
+       std::ofstream header(headerfilename);
+       std::ofstream impl(implfilename);
+       if(!header.is_open())
+         error("Cannot open output file %s\n",headerfilename.c_str());
+       if(!impl.is_open())
+         error("Cannot open output file %s\n",implfilename.c_str());
+       emit_header(&header,result);
+       impl << "#include \""<< headerfilename << "\""<<std::endl;
+       emit_parser(&impl,result);
+       emit_generator(&impl,result);
+       impl << std::endl;
+       header << std::endl;       
        return 0;
      } else {
        return 1;

### 2014-04-01 generator/mk_gen.cc

@@ -149,6 +149,10 @@ class GenGenerator{
         out << mk_str(p.LENGTH.length) << "=" << count << ";";
       }
       break;
+    case OFFSET:
+      {
+        
+      }
     case ARRAY:
       {
         ValExpr count("count", &val);
@@ -172,10 +176,19 @@ class GenGenerator{
           generator(p.ARRAY.SEPBYONE.separator);
           out << "}";
           generator(p.ARRAY.SEPBYONE.inner->PR,elem);break;
+
         }
         out << "}";
       }
       break;
+    case FIXEDARRAY:{
+      ValExpr iter(boost::str(boost::format("i%d") % num_iters++)); 
+      ArrayElemExpr elem(&val,&iter);
+      out << "for(int "<< iter<<"=0;"<<iter << "<" << intconstant_value(p.FIXEDARRAY.length) << ";" << iter << "++){";
+      generator(p.FIXEDARRAY.inner->PR, elem);
+      out << "}\n";
+    }
+      break; 
     case OPTIONAL:
       {
         out << "if(NULL!="<<val << "){";

### 2014-04-01 generator/mk_parser.cc

@@ -22,25 +22,33 @@ class CDataModel{
   }
 
   // Gets the type for a top-level parser (to be used in the typedef)
-  std::string typedef_type(const parser &p, std::string name){
+  std::string typedef_type(const parser &p, std::string name, std::string *post){
     const parserinner &inner(p.N_type == PR ? p.PR : p.PAREN);
     switch(inner.N_type){
     case INT:
       return int_type(inner.INT.parser);
     case WRAP:
-      return typedef_type(*inner.WRAP.parser,name);
+      return typedef_type(*inner.WRAP.parser,name,post);
     case OPTIONAL:
-      return (boost::format("%s*") % typedef_type(*inner.OPTIONAL,name)).str();
+      return (boost::format("%s*") % typedef_type(*inner.OPTIONAL,name,NULL)).str();
     case NAME:
       return mk_str(inner.NAME);
     case REF:
       return (boost::format("%s *") % mk_str(inner.NAME)).str();
+    case FIXEDARRAY:
+      if(post){
+        *post = (boost::format("[%s]%s") % intconstant_value(inner.FIXEDARRAY.length) % *post).str(); 
+        return typedef_type(*inner.FIXEDARRAY.inner,name,post);
+      } 
+      else
+        return (boost::format("%s *") % typedef_type(*inner.FIXEDARRAY.inner,name,NULL)).str();
     case UNION:
       /*   FOREACH(parser,inner.UNION){
            if(!type_equal(inner.UNION.elem[0],parser))
            throw std::runtime_error("Grammar invalid at UNION");
            } */
-      return typedef_type(*inner.UNION.elem[0],name);
+      return typedef_type(*inner.UNION.elem[0],name,post);
+      //  return (boost::format("%s[%s]") %typedef_type(*inner.FIXEDARRAY.inner,name)% intconstant_value(inner.FIXEDARRAY.length)).str();
     case STRUCT:
     case ARRAY:
     case CHOICE:
@@ -105,12 +113,19 @@ class CDataModel{
         break;
       }
       break;
+    case FIXEDARRAY:
+      emit_type(*p.FIXEDARRAY.inner);
+      out << "["<< intconstant_value(p.FIXEDARRAY.length) << "]";
+      break;
     case OPTIONAL:
       emit_type(*p.OPTIONAL);
       out << "*";
       break;
     case LENGTH:
[…]

### 2014-04-01 generator/nailtool.h

@@ -24,5 +24,6 @@ typedef struct expr{
 #define FOREACH(val,coll) for(__typeof__((coll).elem[0]) *val=(coll).elem;val<(coll).elem + (coll).count;val++)
 extern void emit_generator(std::ostream *out,grammar *grammar);
 extern void emit_parser(std::ostream *out,grammar *grammar);
+extern void emit_header(std::ostream *out,grammar *grammar);
 std::string intconstant_value(const intconstant &val);
 #endif

### 2014-04-01 generator/parser_template.c

@@ -135,15 +135,11 @@ static int n_tr_const(n_trace *trace,pos newoff){
         trace->trace[trace->iter++] = newoff;
         return 0;
 }
-
+#define n_tr_offset n_tr_const
 typedef struct NailArenaPool{
         void *iter;void *end;
         struct NailArenaPool *next;
 } NailArenaPool;
-typedef struct NailArena_{
-        NailArenaPool *current;                
-        size_t blocksize;
-} NailArena ;
 
 void *n_malloc(NailArena *arena, size_t size)
 {

### 2014-04-01 examples/network/tunclient.c
rebased network stack

@@ -1,0 +1,230 @@
+#include <sys/types.h>  
+#include <sys/socket.h>
+#include <sys/stat.h>
+#include <sys/ioctl.h>
+#include <fcntl.h>
+#include <string.h>
+#include <stdlib.h>
+#include <stdio.h>
+#include <unistd.h>
+#include <linux/if.h>
+#include <linux/if_tun.h>
+#include <time.h>
+
+#include <uthash.h>   
+
+#include "net.nail.h"
+#include <hammer/hammer.h>
+
+extern HAllocator system_allocator;
+//http://backreference.org/2010/03/26/tuntap-interface-tutorial/
+int tun_alloc(char *dev, int flags) {
+
+  struct ifreq ifr;
+  int fd, err;
+  char *clonedev = "/dev/net/tun";
+
+  /* Arguments taken by the function:
+   *
+   * char *dev: the name of an interface (or '\0'). MUST have enough
+   *   space to hold the interface name if '\0' is passed
+   * int flags: interface flags (eg, IFF_TUN etc.)
+   */
+
+   /* open the clone device */
+   if( (fd = open(clonedev, O_RDWR)) < 0 ) {
+     return fd;
+   }
+
+   /* preparation of the struct ifr, of type "struct ifreq" */
+   memset(&ifr, 0, sizeof(ifr));
+
+   ifr.ifr_flags = flags;   /* IFF_TUN or IFF_TAP, plus maybe IFF_NO_PI */
+
+   if (*dev) {
+     /* if a device name was specified, put it in the structure; otherwise,
+      * the kernel will try to allocate the "next" device of the
+      * specified type */
+     strncpy(ifr.ifr_name, dev, IFNAMSIZ);
+   }
+
+   /* try to create the device */
+   if( (err = ioctl(fd, TUNSETIFF, (void *) &ifr)) < 0 ) {
+     close(fd);
+     return err;
+   }
+
+  /* if the operation was successful, write back the name of the
+   * interface to the variable "dev", so the caller can know
+   * it. Note that the caller MUST reserve space in *dev (see calling
+   * code below) */
+  strcpy(dev, ifr.ifr_name);
+
+  /* this is the special file descriptor that the caller will use to talk
+   * with the virtual interface */
+  return fd;
+}
+typedef struct arpentry{
+        uint32_t ip;
+        macaddr mac; // All 0 when unused 
+        time_t time;
+        UT_hash_handle hh;
[…]

### 2014-04-01 paper/design.tex
Rebased some changes

@@
 
-In the following, we present the combinators in Nail grouped by the combinators, if possible grouped
-by the combinator of conventional Parser Expression Grammars they replace.
+In the following, we present the combinators in Nail grouped by the combinators of conventional Parser Expression Grammars they replace.
 \paragraph{Fundamental parsers}
@@
 Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
-element of a set of allowed values. \\Examples: \texttt{uint8 | 1..3},\texttt{uint16| ..512},
-\texttt{int32| [1,255,512]}
+element of a set of allowed values. \\
+Examples: \texttt{uint8 | 1..3} (accepts an 8 bit unsigned
+integer between 1 and 3 inclusive) ,\\
+ \texttt{uint16| ..512} (an unsigned 16-bit integer less or
+equal than 512),\\
+\texttt{int32| [1,255,512]} (a signed 32 bit integer with value 1, 255 or 512).
 

### 2014-04-01 paper/intro.tex
split off relwk
Rebased some changes

--- 2014-03-31 paper/background.tex
@@
-\section{Background}
+\section{Introduction}
 
@@
 X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee absolute memory safety.
+runtime environment or static analysis guarantee memory safety.
 
@@
  library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec} and. The parser combinator style (to our
+ popularised by Parsec for Haskell\cite{LeijenMeijer:parsec}. The parser combinator style (to our
  knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing

### 2014-04-01 paper/relwk.tex
Rebased some changes

@@
 Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
+however PacketTypes works only as a parser, not as an output generator and does not support the expressive power of PEGs, but rather
 implements a C like structure model enhanced with data-dependent length fields and constraints.

### 2014-04-01 paper/intro.tex
merged changes

@@
 provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}.  
-The more powerful an input
-protocol is, the more control an attacker gains over a systems state - and power to manipulate a
-systems state is also a degree of undeserved trust. In the offensive research community, this has
+\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}. Even though these
+protocols were never intended to be that expressive - and consequentially hard to parse, implementation details of their
+recognisers - which de facto became a part of the specification and were relied upon by legitimate
+inputs - expose unwanted computational power and have unintended side-effects.
+In the offensive research community, this has
 been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
@@
 Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a hand-written PDF parser could eliminate $98\%$ of known malicious PDFs\cite{Bogk-PDF}.
+one case, a  PDF parser implemented in Coq could eliminate over $95\%$ of known malicious PDFs\cite{Bogk-PDF}.
+However, manually writing parser code does not scale to the number of file formats  and protocols in
+existence and might result in parser code tied to one specific application.
 

### 2014-04-01 paper/future.tex
consistently use a non-breaking space (~) before \cite

@@
 awkward. One possible solution to this problem would be to implement a concept similar to Relational
-Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
+Lenses~\cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
 the rest of the program through an isomorphism. Such an isomorphism would still be much more concise

### 2014-04-01 paper/intro.tex
consistently use a non-breaking space (~) before \cite

@@
 from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.
+bits and pieces as a 'shotgun parser'~\cite{shotgun-parser}.
 
@@
 
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
-\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
+Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
   differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
+  that runs.}~\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
   files. Signatures are verified with a Java program, but the program is extracted with a C program.
 The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}
-\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
+allowing one to replace files in a signed archive.}~\cite{saurik-masterkey} code signing and even the
+X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
 runtime environment or static analysis guarantee memory safety.
@@
 Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables
-\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}. Even though these
+provide a full Turing-complete execution environment. Examples include X86 page tables~\cite{bangert2013page}, ELF symbols and relocations~\cite{shapiro2013weird}. Even though these
 protocols were never intended to be that expressive - and consequentially hard to parse, implementation details of their
@@
 In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
+been generalised into treating a program as a \textit{weird machine}~\cite{bratus2011exploit} that
 operates on an input, analogous to a virtual machine operating on bytecode.
@@
 Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a  PDF parser implemented in Coq could eliminate over $95\%$ of known malicious PDFs\cite{Bogk-PDF}.
+one case, a  PDF parser implemented in Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
 However, manually writing parser code does not scale to the number of file formats  and protocols in
@@
 expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}
+limiting the weird machine problem and enabling very efficient implementation~\cite{Knuth1965607}.
 
@@
 web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.
+Ragel~\cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
+times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations~\cite{patterson-citation}.
 
@@
  library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell\cite{LeijenMeijer:parsec}. The parser combinator style (to our
- knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
+ popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our
+ knowledge, first  described in~\cite{burge1975recursive}) is a natural way of concisely expressing
+ top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For more background on the history of expressing grammars, see Bryan
+   Ford's masters thesis~\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
   Hammer then constructs a tree of function pointers which

### 2014-04-01 paper/relwk.tex
consistently use a non-breaking space (~) before \cite

--- 2014-04-01 paper/relwk.tex
@@
 Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
+Generators, e.g. in CORBA or more recently~\cite{varda2008}. However, while interface
 generators work very well for existing grammars, they do not allow full control over the format of
 the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
-however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
+Very related work has been done at Bell Labs with the PacketTypes system~\cite{mccann2000packet},
+however PacketTypes works only as a parser, not as an output generator and does not support the expressive power of PEGs, but rather
 implements a C like structure model enhanced with data-dependent length fields and constraints.
@@
 Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
- heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
- feng shui\cite{fengshui} to help bypass those mitigations.
+memory locations. Exploit countermeasures  like Address-Space Layout Randomisation~\cite{pax-aslr},
+ heap cookies~\cite{heapcookies} or data execution prevention try to make this harder for
+ an attacker. Attackers have developed techniques such as heap spraying~\cite{heapspray} and heap
+ feng shui~\cite{fengshui} to help bypass those mitigations.
 
-Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
+Nail parsers use two arena allocators~\cite{arena}, one for the temporary syntax tree and one for the
 result. If Nail were to use the normal system allocator, the attacker could cause two packets to be

### 2014-04-01 paper/design.tex
consistent footnote spacing

@@
 The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective. \footnote{By default, the generator will emit exactly one iteration of the
+generator not bijective.\footnote{By default, the generator will emit exactly one iteration of the
   constant parser in its output.} Example: \texttt{many uint8= "test"}

### 2014-04-01 paper/intro.tex
consistent footnote spacing

@@
 A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng
-\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
+and the security problems they cause are rampant. Input processing libraries such as libpng\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
 remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
@@
 generators, but only for handling configuration files. A notable exception is the Mongrel
-web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
+web server\footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
 Ragel~\cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
@@
 \paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions
-\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
+While it is possible to express short transformations on the input entirely as semantic actions\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
   walks the parse tree.}, more complicated programs usually construct an internal representation,
@@
 Some parser generators, e.g.
-Boost.Spirit \footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}},
+Boost.Spirit,\footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}}
 allow the same grammar to be re-used for generating output from the intermediate representation.

### 2014-04-01 paper/intro.tex
line-wrap intro and relwk

@@
 
-% For a long time, it has been possible to write a concise and precise specification of what a data
-% format looks like, feed it to a tool and let it generate a parser and potentially a generator for
-% that format. However, instead of using these tools, programmers have continued to laboriously
-% hand-write bug-prone input/output code. One of the reasons they have done so is that existing tools
-% were geared towards parsing text-based programming languages, not for dealing with binary protocols,
-% although the underlying algorithms remain the same.
+\subsection{Problems handling untrusted input}
 
-\subsection{Problems handling untrusted input}
-Code handling untrusted user inputs is most prone to contain security vulnerabilities. While memory
-corruption errors or fatal logic flaws can occur anywhere in a program, bugs in the input handling
-code are most likely to be exploitable, because an attacker has total control over the inputs to
-that code.
+Code handling untrusted user inputs is most prone to contain security
+vulnerabilities. While memory corruption errors or fatal logic flaws
+can occur anywhere in a program, bugs in the input handling code are
+most likely to be exploitable, because an attacker has total control
+over the inputs to that code.
 
-Common wisdom thus says to take special care when preparing code that deals with untrusted input.
-However, with currently prevalent software design methods, it is hard to isolate input-handling code
-from the rest of the program, as user input is typically passed through the program and processed in
-bits and pieces as a 'shotgun parser'~\cite{shotgun-parser}.
+Common wisdom thus says to take special care when preparing code that
+deals with untrusted input.  However, with currently prevalent software
+design methods, it is hard to isolate input-handling code from the rest
+of the program, as user input is typically passed through the program and
+processed in bits and pieces as a 'shotgun parser'~\cite{shotgun-parser}.
 
-A look at the CVE database or Hacker journal such as \textit{Phrack} shows that such design patterns
-and the security problems they cause are rampant. Input processing libraries such as libpng\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html} shows 24
-remote exploits in the between 2007 and 2013.} or Adobe's PDF and Flash viewers are notoriously
-plagued with input bugs. Even simpler formats such as the \textit{zlib} compression library have
-historically contained multiple security
+A look at the CVE database or Hacker journal such as
+\textit{Phrack} shows that such design patterns and the security
+problems they cause are rampant. Input processing libraries such as
+libpng\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}
+shows 24 remote exploits in the between 2007 and 2013.} or
+Adobe's PDF and Flash viewers are notoriously plagued with
+input bugs. Even simpler formats such as the \textit{zlib}
+compression library have historically contained multiple security
 flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
 
-Besides memory corruption vulnerabilities - against which various mitigations such as static
-analysis, dynamic instrumentation and various memory protection mechanisms have been developed -
-different implementations of the same format can also differ in their understanding of some edge
-cases. The security of many systems relies on different components interpreting the same data
-similarly.
+Besides memory corruption vulnerabilities - against which various
+mitigations such as static analysis, dynamic instrumentation and
+various memory protection mechanisms have been developed - different
+implementations of the same format can also differ in their understanding
+of some edge cases. The security of many systems relies on different
+components interpreting the same data similarly.
 
-Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
-  differently, so the code signature sees different bytes at a virtual address than the executable
-  that runs.}~\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
-  files. Signatures are verified with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C program treats them as unsigned,
-allowing one to replace files in a signed archive.}~\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
-runtime environment or static analysis guarantee memory safety.
+Such parser differential vulnerabilities were found in various
+cryptographic systems, including  iOS\footnote{The XNU kernel and the
+user code-signing verifier interpret executable metadata differently,
+so the code signature sees different bytes at a virtual address than
+the executable that runs.}~\cite{evaders6} and Android\footnote{Android
+applications are distributed as .zip files. Signatures are verified
+with a Java program, but the program is extracted with a C program.
+The java program interprets all fields as signed, whereas the C
+program treats them as unsigned, allowing one to replace files in a
+signed archive.}~\cite{saurik-masterkey} code signing and even the
+X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Such
+vulnerabilities occur even if the language, runtime environment or static
+analysis guarantee memory safety.
 
-Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
-provide a full Turing-complete execution environment. Examples include X86 page tables~\cite{bangert2013page}, ELF symbols and relocations~\cite{shapiro2013weird}. Even though these
-protocols were never intended to be that expressive - and consequentially hard to parse, implementation details of their
-recognisers - which de facto became a part of the specification and were relied upon by legitimate
-inputs - expose unwanted computational power and have unintended side-effects.
-In the offensive research community, this has
-been generalised into treating a program as a \textit{weird machine}~\cite{bratus2011exploit} that
-operates on an input, analogous to a virtual machine operating on bytecode.
+Finally, many protocol implementations are so complicated that even
+without bugs, they inadvertently provide a full Turing-complete execution
+environment. Examples include X86 page tables~\cite{bangert2013page},
+ELF symbols and relocations~\cite{shapiro2013weird}. Even though
+these protocols were never intended to be that expressive -
+and consequentially hard to parse, implementation details of their
+recognisers - which de facto became a part of the specification and were
+relied upon by legitimate inputs - expose unwanted computational power
+and have unintended side-effects.  In the offensive research community,
+this has been generalised into treating a program as a \textit{weird
+machine}~\cite{bratus2011exploit} that operates on an input, analogous
+to a virtual machine operating on bytecode.
 
-Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
-one case, a  PDF parser implemented in Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
-However, manually writing parser code does not scale to the number of file formats  and protocols in
-existence and might result in parser code tied to one specific application.
+Proper input recognition has been shown to be an excellent way of
+eliminating malicious inputs. In one case, a  PDF parser implemented in
+Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
+However, manually writing parser code does not scale to the number of
+file formats  and protocols in existence and might result in parser code
+tied to one specific application.
 
 \subsection{Parser generators}
-%%TODO
-Luckily, we can avoid all of these problems at once if we have a single, executable specification of
-what exactly a programs input is supposed to look like. In fact, definitive grammars have long
-been written for all sorts of protocols and are the standard way of describing text-based protocols.
-Parser generators can automatically create parsers for such languages, and with common classes of
-grammars such as regular expressions or deterministic context free languages, that parser can be
-expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
-limiting the weird machine problem and enabling very efficient implementation~\cite{Knuth1965607}.
 
-Specifications of binary protocols, such as RFCs, also contain grammars, although they are
-frequently translated into prose and diagrams as opposed to the traditional Backus-Naur-Form.
+Luckily, we can avoid all of these problems at once if we have a single,
+executable specification of what exactly a programs input is supposed
+to look like. In fact, definitive grammars have long been written for
+all sorts of protocols and are the standard way of describing text-based
+protocols.  Parser generators can automatically create parsers for such
+languages, and with common classes of grammars such as regular expressions
+or deterministic context free languages, that parser can be expressed
+in a computationally less powerful form, e.g. a finite automaton or a
+push down automaton, limiting the weird machine problem and enabling
+very efficient implementation~\cite{Knuth1965607}.
 
-Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
-the security-critical and very well-engineered MIT Kerberos distribution uses parser
-generators, but only for handling configuration files. A notable exception is the Mongrel
-web server\footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
-Ragel~\cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
-times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations~\cite{patterson-citation}.
+Specifications of binary protocols, such as RFCs, also contain grammars,
+although they are frequently translated into prose and diagrams as
+opposed to the traditional Backus-Naur-Form.
 
- Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
- While previous parser generators could also be used to write grammars for binary
- protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
-   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
- inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
- of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
- library. Hammer implements grammars as language-integrated parser combinators, an approach
- popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our
- knowledge, first  described in~\cite{burge1975recursive}) is a natural way of concisely expressing
- top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For more background on the history of expressing grammars, see Bryan
-   Ford's masters thesis~\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
-  Hammer then constructs a tree of function pointers which
+Unfortunately, grammars are seldom used directly to recognise
+machine-created input. For example, the security-critical and very
+well-engineered MIT Kerberos distribution uses parser generators, but
+only for handling configuration files. A notable exception is the Mongrel
+web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
+for HTTP written in the Ragel~\cite{ragel-paper} regular expression
+language. Mongrel was re-written from scratch multiple times to achieve
+better scaleability and design, yet the grammar could be  re-used across
+all iterations~\cite{patterson-citation}.
+
+Parser generators for binary protocols were first introduced by the
+Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
+While previous parser generators could also be used to write grammars
+for binary protocols\footnote{Theoretically speaking, the alphabet
+over which a grammar is an  abstract set, so most algorithms work
+just as well on an alphabet of $\{0,1\}$}, doing so is impractically
+inconvenient. Hammer allows the programmer to specify a grammar in
+terms of bits and bytes instead of characters. Common concerns, such as
+endianness and bit-packing are transparently hidden by the library. Hammer
+implements grammars as language-integrated parser combinators, an approach
+popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
+parser combinator style (to our knowledge, first  described
+in~\cite{burge1975recursive}) is a natural way of concisely expressing
+top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
+more background on the history of expressing grammars, see Bryan Ford's
+masters thesis~\cite{ford2002packrat}, which also describes the default
+parsing algorithm used by Hammer} by composing them from one or multiple
+sub-parsers.  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).
 
-Working with generic tree structures is tiresome, so the programmer can bind a \textit{semantic action} to a
-sub-parser. Whenever the sub-parser is successfully invoked, the parser executes the semantic action
-on the output of the sub-parser and places the result as a leaf node in the syntax tree. Other
-parser generators frequently skip the AST entirely, requiring the programmer to specify a semantic
-action with every rule in the grammar.
+Working with generic tree structures is tiresome, so the programmer can
+bind a \textit{semantic action} to a sub-parser. Whenever the sub-parser
+is successfully invoked, the parser executes the semantic action on the
+output of the sub-parser and places the result as a leaf node in the
+syntax tree. Other parser generators frequently skip the AST entirely,
+requiring the programmer to specify a semantic action with every rule
+in the grammar.
 
-Semantic actions are one source of vulnerability within automatically generated parsers, as they are
-typically implemented in C or another general purpose programming language and perform dangerous
-operations such as memory allocation and copying. If there is a bug in the semantic actions, the
-attacker is likely able to craft an input that invokes the semantic actions with controllable
-inputs, as the inputs have not been fully verified yet.
+Semantic actions are one source of vulnerability within automatically
+generated parsers, as they are typically implemented in C or another
+general purpose programming language and perform dangerous operations
+such as memory allocation and copying. If there is a bug in the semantic
+actions, the attacker is likely able to craft an input that invokes the
+semantic actions with controllable inputs, as the inputs have not been
+fully verified yet.
 
 \subsection{Motivation}
+
 \paragraph{Motivation}
-While it is possible to express short transformations on the input entirely as semantic actions\footnote{This is in fact the design rationale, to perform computation on the fly as the parser
-  walks the parse tree.}, more complicated programs usually construct an internal representation,
-which contains all relevant information from the input in a format native to the programming
-language used. For example, a C programmer ideally wants to deal with structs and NULL-terminated
-arrays, whereas a C++ programmer might expect STL containers, a Java programmer interfaces, a
-Haskell programmer records and a LISP programmer property lists. 
- The structure of this internal representation usually resembles the structure of the grammar.
 
-Therefore, a typical programmer has to describe the structure of his input twice or even thrice to
-use a parser generator. Once, to write a grammar, another time to write the semantic actions
-constructing his favourite intermediate representation. In most languages the programmer also has to
-write explicit type definitions, which describe the intermediate object model again. Most
-programmers will baulk at this multiplication of efforts and either hand-write a top-down parser
-which still eliminates writing the semantic actions at the slight cost of verbosity or resort to
-passing user input through their control flow, hopefully remembering to verify each part before
-it is accessed.
+While it is possible to express short transformations on the input
+entirely as semantic actions\footnote{This is in fact the design
+rationale, to perform computation on the fly as the parser walks the
+parse tree.}, more complicated programs usually construct an internal
+representation, which contains all relevant information from the input
+in a format native to the programming language used. For example, a C
+programmer ideally wants to deal with structs and NULL-terminated arrays,
+whereas a C++ programmer might expect STL containers, a Java programmer
+interfaces, a Haskell programmer records and a LISP programmer property
+lists.  The structure of this internal representation usually resembles
+the structure of the grammar.
+
+Therefore, a typical programmer has to describe the structure of his input
+twice or even thrice to use a parser generator. Once, to write a grammar,
+another time to write the semantic actions constructing his favourite
+intermediate representation. In most languages the programmer also
+has to write explicit type definitions, which describe the intermediate
+object model again. Most programmers will baulk at this multiplication of
+efforts and either hand-write a top-down parser which still eliminates
+writing the semantic actions at the slight cost of verbosity or resort
+to passing user input through their control flow, hopefully remembering
+to verify each part before it is accessed.
 
 A similar problem occurs when the programmer wants to generate output.
-Even though not all programs may use the same format for output as they do on input, different
-programs might use the same format for input and output and re-using the same grammar is a good way
-to save engineering effort and reduce parser differentials.
-Some parser generators, e.g.
+Even though not all programs may use the same format for output as they
+do on input, different programs might use the same format for input and
+output and re-using the same grammar is a good way to save engineering
+effort and reduce parser differentials.  Some parser generators, e.g.
 Boost.Spirit,\footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}}
-allow the same grammar to be re-used for generating output from the intermediate representation.
-However, those generators require a new set of semantic actions to be written, even though the
-relationship between grammar and intermediate model was already expressed in the actions for the
-parser.
+allow the same grammar to be re-used for generating output from the
+intermediate representation.  However, those generators require a new set
+of semantic actions to be written, even though the relationship between
+grammar and intermediate model was already expressed in the actions for
+the parser.

### 2014-04-01 paper/relwk.tex
line-wrap intro and relwk

@@
 
-Generating parsers and generators from an executable specification is the core concept of Interface
-Generators, e.g. in CORBA or more recently~\cite{varda2008}. However, while interface
-generators work very well for existing grammars, they do not allow full control over the format of
-the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes system~\cite{mccann2000packet},
-however PacketTypes works only as a parser, not as an output generator and does not support the expressive power of PEGs, but rather
-implements a C like structure model enhanced with data-dependent length fields and constraints.
-
-\begin{comment}
-
-Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
-memory locations. Exploit countermeasures  like Address-Space Layout Randomisation~\cite{pax-aslr},
- heap cookies~\cite{heapcookies} or data execution prevention try to make this harder for
- an attacker. Attackers have developed techniques such as heap spraying~\cite{heapspray} and heap
- feng shui~\cite{fengshui} to help bypass those mitigations.
-
-Nail parsers use two arena allocators~\cite{arena}, one for the temporary syntax tree and one for the
-result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
-parsed at the same time (or, given a sufficiently complicated input, two structure within the same
-grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
- If there is a buffer overflow past the
-end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
-placing a known value next to that input.
-
-\end{comment}
+Generating parsers and generators from an executable specification
+is the core concept of Interface Generators, e.g. in CORBA or more
+recently~\cite{varda2008}. However, while interface generators work
+very well for existing grammars, they do not allow full control over the
+format of the output, so cannot be used to implement legacy protocols.
+Very related work has been done at Bell Labs with the PacketTypes
+system~\cite{mccann2000packet}, however PacketTypes works only as a
+parser, not as an output generator and does not support the expressive
+power of PEGs, but rather implements a C like structure model enhanced
+with data-dependent length fields and constraints.

### 2014-04-01 paper/intro.tex
checkpoint

@@
 
-\subsection{Problems handling untrusted input}
-
-Code handling untrusted user inputs is most prone to contain security
-vulnerabilities. While memory corruption errors or fatal logic flaws
-can occur anywhere in a program, bugs in the input handling code are
-most likely to be exploitable, because an attacker has total control
-over the inputs to that code.
-
-Common wisdom thus says to take special care when preparing code that
-deals with untrusted input.  However, with currently prevalent software
-design methods, it is hard to isolate input-handling code from the rest
-of the program, as user input is typically passed through the program and
-processed in bits and pieces as a 'shotgun parser'~\cite{shotgun-parser}.
-
-A look at the CVE database or Hacker journal such as
-\textit{Phrack} shows that such design patterns and the security
-problems they cause are rampant. Input processing libraries such as
-libpng\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}
-shows 24 remote exploits in the between 2007 and 2013.} or
-Adobe's PDF and Flash viewers are notoriously plagued with
-input bugs. Even simpler formats such as the \textit{zlib}
-compression library have historically contained multiple security
-flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.
-
-Besides memory corruption vulnerabilities - against which various
-mitigations such as static analysis, dynamic instrumentation and
-various memory protection mechanisms have been developed - different
-implementations of the same format can also differ in their understanding
-of some edge cases. The security of many systems relies on different
-components interpreting the same data similarly.
-
-Such parser differential vulnerabilities were found in various
-cryptographic systems, including  iOS\footnote{The XNU kernel and the
-user code-signing verifier interpret executable metadata differently,
-so the code signature sees different bytes at a virtual address than
-the executable that runs.}~\cite{evaders6} and Android\footnote{Android
-applications are distributed as .zip files. Signatures are verified
-with a Java program, but the program is extracted with a C program.
-The java program interprets all fields as signed, whereas the C
-program treats them as unsigned, allowing one to replace files in a
-signed archive.}~\cite{saurik-masterkey} code signing and even the
-X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Such
-vulnerabilities occur even if the language, runtime environment or static
-analysis guarantee memory safety.
-
-Finally, many protocol implementations are so complicated that even
-without bugs, they inadvertently provide a full Turing-complete execution
-environment. Examples include X86 page tables~\cite{bangert2013page},
-ELF symbols and relocations~\cite{shapiro2013weird}. Even though
-these protocols were never intended to be that expressive -
-and consequentially hard to parse, implementation details of their
-recognisers - which de facto became a part of the specification and were
-relied upon by legitimate inputs - expose unwanted computational power
-and have unintended side-effects.  In the offensive research community,
-this has been generalised into treating a program as a \textit{weird
-machine}~\cite{bratus2011exploit} that operates on an input, analogous
-to a virtual machine operating on bytecode.
-
-Proper input recognition has been shown to be an excellent way of
-eliminating malicious inputs. In one case, a  PDF parser implemented in
-Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
-However, manually writing parser code does not scale to the number of
-file formats  and protocols in existence and might result in parser code
-tied to one specific application.
-
-\subsection{Parser generators}
+Code that handles untrusted inputs, such as processing network
+data or parsing a file, is error-prone and is often exploited by
+attackers.  This is in part because attackers have precise control
+over the inputs to that code, and can craft inputs that trigger
+subtle corner cases in input processing.  For example, the libpng
+library has had 24 remotely exploitable vulnerabilities from 2007 to
+2013,\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}}
+and Adobe's PDF and Flash viewers have been notoriously
+plagued by input processing vulnerabilities.  Even relatively
+simple formats, such as those used by the zlib compression
+library, have had input processing vulnerabilities in the
+past.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}
 

### 2014-04-01 paper/relwk.tex
checkpoint

@@
 \section{Related work}
+
+\paragraph{Parsing frameworks.}
 
@@
 with data-dependent length fields and constraints.
+
+\paragraph{Language safety.}
+
+Input processing vulnerabilities fall into two broad classes.  The first
+class is memory safety bugs, such as buffer overflows, which allow an
+adversary to corrupt the application's memory using specially crafted
+inputs.  These mistakes arise in lower-level languages that do not
+provide memory safety guarantees (such as C), and can be partially
+mitigated by a wide range of techniques, such as static analysis,
+dynamic instrumentation, and address space layout randomization, that
+make it more difficult for an adversary to exploit these bugs.
+
+The second class is logic errors, where application code misinterprets
+input data.  This can lead to serious security consequences when two
+systems disagree on the meaning of a network packet or a signed message,
+as in iOS\footnote{The XNU kernel and the user code-signing verifier
+interpret executable metadata differently, so the code signature
+sees different bytes at a virtual address than the executable that
+runs.}~\cite{evaders6} and Android\footnote{Android applications
+are distributed as .zip files. Signatures are verified with a Java
+program, but the program is extracted with a C program.  The java
+program interprets all fields as signed, whereas the C program
+treats them as unsigned, allowing one to replace files in a signed
+archive.}~\cite{saurik-masterkey} code signing and even the X.509 protocol
+underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
+highly application-specific, and are difficult to mitigate using existing
+techniques, and these mistakes can occur even in high-level languages
+that guarantee memory safety.
+
+--- XXX integrate the following paragraphs somewhere into the logical flow ---
+
+With currently prevalent software design methods, it is hard to isolate
+input-handling code from the rest of the program, as user input is
+typically passed through the program and processed in bits and pieces as a
+``shotgun parser''~\cite{shotgun-parser}.
+
+Finally, many protocol implementations are so complicated that even
+without bugs, they inadvertently provide a full Turing-complete execution
+environment. Examples include X86 page tables~\cite{bangert2013page},
+ELF symbols and relocations~\cite{shapiro2013weird}. Even though
+these protocols were never intended to be that expressive -
+and consequentially hard to parse, implementation details of their
+recognisers - which de facto became a part of the specification and were
+relied upon by legitimate inputs - expose unwanted computational power
+and have unintended side-effects.  In the offensive research community,
+this has been generalised into treating a program as a \textit{weird
+machine}~\cite{bratus2011exploit} that operates on an input, analogous
+to a virtual machine operating on bytecode.
+
+Proper input recognition has been shown to be an excellent way of
+eliminating malicious inputs. In one case, a  PDF parser implemented in
+Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
+However, manually writing parser code does not scale to the number of
+file formats  and protocols in existence and might result in parser code
+tied to one specific application.

### 2014-04-01 paper/intro.tex
another par

@@
 
-Luckily, we can avoid all of these problems at once if we have a single,
-executable specification of what exactly a programs input is supposed
-to look like. In fact, definitive grammars have long been written for
-all sorts of protocols and are the standard way of describing text-based
-protocols.  Parser generators can automatically create parsers for such
-languages, and with common classes of grammars such as regular expressions
-or deterministic context free languages, that parser can be expressed
-in a computationally less powerful form, e.g. a finite automaton or a
-push down automaton, limiting the weird machine problem and enabling
-very efficient implementation~\cite{Knuth1965607}.
-
-Specifications of binary protocols, such as RFCs, also contain grammars,
-although they are frequently translated into prose and diagrams as
-opposed to the traditional Backus-Naur-Form.
+A promising approach to avoid such vulnerabilities is to specify
+a precise grammar for the input data format, and to build a tool for
+parsing inputs given a particular grammar, such as the traditional {\tt
+lex} and {\tt yacc} generators.  Applications that use such a tool to
+process input data will not need to write error-prone input processing
+code, and as long as the tool is bug-free, the application will be safe
+from input processing vulnerabilities.
 

### 2014-04-01 paper/relwk.tex
another par

@@
 
-Finally, many protocol implementations are so complicated that even
+Many protocol implementations are so complicated that even
 without bugs, they inadvertently provide a full Turing-complete execution
@@
 
+Parser generators can automatically create parsers for such languages,
+and with common classes of grammars such as regular expressions or
+deterministic context free languages, that parser can be expressed in a
+computationally less powerful form, e.g. a finite automaton or a push
+down automaton, limiting the weird machine problem and enabling very
+efficient implementation~\cite{Knuth1965607}.
+
 Proper input recognition has been shown to be an excellent way of
@@
 tied to one specific application.
+
+In fact, definitive grammars have long been written for all sorts of
+protocols and are the standard way of describing text-based protocols.
+Specifications of binary protocols, such as RFCs, also contain grammars,
+although they are frequently translated into prose and diagrams as
+opposed to the traditional Backus-Naur-Form.

### 2014-04-01 paper/intro.tex
move more relwk from intro to relwk

@@
 
-Unfortunately, grammars are seldom used directly to recognise
-machine-created input. For example, the security-critical and very
-well-engineered MIT Kerberos distribution uses parser generators, but
-only for handling configuration files. A notable exception is the Mongrel
-web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
-for HTTP written in the Ragel~\cite{ragel-paper} regular expression
-language. Mongrel was re-written from scratch multiple times to achieve
-better scaleability and design, yet the grammar could be  re-used across
-all iterations~\cite{patterson-citation}.
-
-Parser generators for binary protocols were first introduced by the
-Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
-While previous parser generators could also be used to write grammars
-for binary protocols\footnote{Theoretically speaking, the alphabet
-over which a grammar is an  abstract set, so most algorithms work
-just as well on an alphabet of $\{0,1\}$}, doing so is impractically
-inconvenient. Hammer allows the programmer to specify a grammar in
-terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing are transparently hidden by the library. Hammer
-implements grammars as language-integrated parser combinators, an approach
-popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
-parser combinator style (to our knowledge, first  described
-in~\cite{burge1975recursive}) is a natural way of concisely expressing
-top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
-more background on the history of expressing grammars, see Bryan Ford's
-masters thesis~\cite{ford2002packrat}, which also describes the default
-parsing algorithm used by Hammer} by composing them from one or multiple
-sub-parsers.  Hammer then constructs a tree of function pointers which
-can be invoked to parse a given input into an abstract syntax tree (AST).
-
 Working with generic tree structures is tiresome, so the programmer can
@@
 fully verified yet.
-
-\subsection{Motivation}
-
-\paragraph{Motivation}
 

### 2014-04-01 paper/relwk.tex
move more relwk from intro to relwk

@@
 with data-dependent length fields and constraints.
+
+Parser generators for binary protocols were first introduced by the
+Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
+While previous parser generators could also be used to write grammars
+for binary protocols\footnote{Theoretically speaking, the alphabet
+over which a grammar is an  abstract set, so most algorithms work
+just as well on an alphabet of $\{0,1\}$}, doing so is impractically
+inconvenient. Hammer allows the programmer to specify a grammar in
+terms of bits and bytes instead of characters. Common concerns, such as
+endianness and bit-packing are transparently hidden by the library. Hammer
+implements grammars as language-integrated parser combinators, an approach
+popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
+parser combinator style (to our knowledge, first  described
+in~\cite{burge1975recursive}) is a natural way of concisely expressing
+top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
+more background on the history of expressing grammars, see Bryan Ford's
+masters thesis~\cite{ford2002packrat}, which also describes the default
+parsing algorithm used by Hammer} by composing them from one or multiple
+sub-parsers.  Hammer then constructs a tree of function pointers which
+can be invoked to parse a given input into an abstract syntax tree (AST).
 
@@
 that guarantee memory safety.
+
+\paragraph{Application use of parsers.}
+
+Unfortunately, grammars are seldom used directly to recognise
+machine-created input. For example, the security-critical and very
+well-engineered MIT Kerberos distribution uses parser generators, but
+only for handling configuration files. A notable exception is the Mongrel
+web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
+for HTTP written in the Ragel~\cite{ragel-paper} regular expression
+language. Mongrel was re-written from scratch multiple times to achieve
+better scaleability and design, yet the grammar could be  re-used across
+all iterations~\cite{patterson-citation}.
 

### 2014-04-01 paper/intro.tex
one more par

@@
 
-A promising approach to avoid such vulnerabilities is to specify
-a precise grammar for the input data format, and to build a tool for
-parsing inputs given a particular grammar, such as the traditional {\tt
-lex} and {\tt yacc} generators.  Applications that use such a tool to
-process input data will not need to write error-prone input processing
-code, and as long as the tool is bug-free, the application will be safe
-from input processing vulnerabilities.
+A promising approach to avoid such vulnerabilities is to specify a
+precise grammar for the input data format, and to use a parser generator,
+such as {\tt lex} and {\tt yacc}, to synthesize the input processing
+code.  Developers that use a parser generator will not need to write
+error-prone input processing code on their own, and as long as the
+parser generator is bug-free, the application will be safe from input
+processing vulnerabilities.  Unfortunately, applying this approach in
+practice, using state-of-the-art parser generators, requires addressing
+several challenges.
 
-Working with generic tree structures is tiresome, so the programmer can
-bind a \textit{semantic action} to a sub-parser. Whenever the sub-parser
-is successfully invoked, the parser executes the semantic action on the
-output of the sub-parser and places the result as a leaf node in the
-syntax tree. Other parser generators frequently skip the AST entirely,
-requiring the programmer to specify a semantic action with every rule
-in the grammar.
-
-Semantic actions are one source of vulnerability within automatically
-generated parsers, as they are typically implemented in C or another
-general purpose programming language and perform dangerous operations
-such as memory allocation and copying. If there is a bug in the semantic
-actions, the attacker is likely able to craft an input that invokes the
-semantic actions with controllable inputs, as the inputs have not been
-fully verified yet.
+First, parser generators typically parse inputs into an abstract syntax
+tree (AST) that corresponds to the grammar.  In order to produce a data
+structure that the rest of the application code can easily process,
+application developers must write explicit {\em semantic actions} that
+update the application's internal representation of the data based on
+each AST node.  Writing these semantic actions by hand is error-prone,
+much like other input processing code, and mistakes can result in memory
+corruption bugs or misinterpreted inputs.
 

### 2014-04-01 paper/intro.tex
slow progress

@@
 much like other input processing code, and mistakes can result in memory
-corruption bugs or misinterpreted inputs.
-
-While it is possible to express short transformations on the input
-entirely as semantic actions\footnote{This is in fact the design
-rationale, to perform computation on the fly as the parser walks the
-parse tree.}, more complicated programs usually construct an internal
-representation, which contains all relevant information from the input
-in a format native to the programming language used. For example, a C
-programmer ideally wants to deal with structs and NULL-terminated arrays,
-whereas a C++ programmer might expect STL containers, a Java programmer
-interfaces, a Haskell programmer records and a LISP programmer property
-lists.  The structure of this internal representation usually resembles
-the structure of the grammar.
-
-Therefore, a typical programmer has to describe the structure of his input
-twice or even thrice to use a parser generator. Once, to write a grammar,
-another time to write the semantic actions constructing his favourite
-intermediate representation. In most languages the programmer also
-has to write explicit type definitions, which describe the intermediate
-object model again. Most programmers will baulk at this multiplication of
-efforts and either hand-write a top-down parser which still eliminates
-writing the semantic actions at the slight cost of verbosity or resort
-to passing user input through their control flow, hopefully remembering
-to verify each part before it is accessed.
+corruption bugs or misinterpreted inputs.  Writing these semantic actions
+also requires the programmer to describe the structure of the input three
+times---once to describe the grammar, once to describe the internal data
+structure, and once again in the semantic actions that translate the
+grammar into the data structure---leading to another potential source
+of bugs and inconsistencies.
 

### 2014-04-01 paper/relwk.tex
slow progress

@@
 opposed to the traditional Backus-Naur-Form.
+
+While it is possible to express short transformations on the input
+entirely as semantic actions\footnote{This is in fact the design
+rationale, to perform computation on the fly as the parser walks the
+parse tree.}, more complicated programs usually construct an internal
+representation, which contains all relevant information from the input
+in a format native to the programming language used. For example, a C
+programmer ideally wants to deal with structs and NULL-terminated arrays,
+whereas a C++ programmer might expect STL containers, a Java programmer
+interfaces, a Haskell programmer records and a LISP programmer property
+lists.  The structure of this internal representation usually resembles
+the structure of the grammar.

### 2014-04-01 paper/intro.tex
add a challenge (maybe) about cconstraints in the input format

@@
 
-A promising approach to avoid such vulnerabilities is to specify a
-precise grammar for the input data format, and to use a parser generator,
-such as {\tt lex} and {\tt yacc}, to synthesize the input processing
-code.  Developers that use a parser generator will not need to write
-error-prone input processing code on their own, and as long as the
-parser generator is bug-free, the application will be safe from input
-processing vulnerabilities.  Unfortunately, applying this approach in
-practice, using state-of-the-art parser generators, requires addressing
-several challenges.
+A promising approach to avoid such vulnerabilities is to specify
+a precise grammar for the input data format, and to use a parser
+generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
+processing code.  Developers that use a parser generator do not need
+to write error-prone input processing code on their own, and as long
+as the parser generator is bug-free, the application will be safe from
+input processing vulnerabilities.  Unfortunately, applying this approach
+in practice, using state-of-the-art parser generators, still requires
+too much manual programmer effort, making it error-prone, as we describe
+next.
 
@@
 
-A similar problem occurs when the programmer wants to generate output.
-Even though not all programs may use the same format for output as they
-do on input, different programs might use the same format for input and
-output and re-using the same grammar is a good way to save engineering
-effort and reduce parser differentials.  Some parser generators, e.g.
+Second, data formats can have constraints XXX todo.
+
+Third, applications often need to produce output in the same
+format as their input---for example, applications might both
+read and write files, or both receive and send network packets.
+Most parser generators just focus on parsing an input, rather
+than producing an output, thus requiring the programmer to manually
+construct outputs, which is error-prone.  Some parser generators, such as
 Boost.Spirit,\footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}}
-allow the same grammar to be re-used for generating output from the
-intermediate representation.  However, those generators require a new set
-of semantic actions to be written, even though the relationship between
-grammar and intermediate model was already expressed in the actions for
-the parser.
+allow reusing the grammar for generating output from the internal
+representation.  However, those generators require yet another set of
+semantic actions to be written, transforming the internal representation
+into an AST\@.
+
+This paper presents the initial design and implementation of Nail, a
+parser generator that greatly reduces the programmer effort required to
+use a grammar-based parser.  Nail addresses the above three challenges
+with several key ideas.  First, Nail XXX todo.

### 2014-04-01 paper/impl.tex
consistently use a non-breaking space (~) before \cite
rough intro pass done, with some sizable holes left

@@
 \section{Implementation}
-The current prototype of the Nail parser generator supports the C programming language and top down parsers. Options for C++ STL data models and emitting Packrat parser
-\cite{packrat-parsing:icfp02} are under heavy development. In this section, we will discuss some particular
+\label{s:impl}
+
+The current prototype of the Nail parser generator supports the C programming language and top down parsers.
+Options for C++ STL data models and emitting Packrat parser~\cite{packrat-parsing:icfp02} are under heavy development.
+In this section, we will discuss some particular
 features of our parser implementation.

### 2014-04-01 paper/intro.tex
rough intro pass done, with some sizable holes left

@@
 Second, data formats can have constraints XXX todo.
+Lengths.
+Counts.
+Checksums.
 
@@
 use a grammar-based parser.  Nail addresses the above three challenges
-with several key ideas.  First, Nail XXX todo.
+with several key ideas.  First, Nail XXX todo.  Bijections.  Reduced
+expressiveness.  Constraints.  External interaction (checksum, ports,
+etc).
+
+We have implemented an initial prototype of Nail in C\@.  Our experience
+so far suggests that it is a promising approach: we were able to construct
+a succinct grammar for DNS packets, and write a small DNS server that
+uses Nail for all packet input and output, and operates purely on
+Nail-generated data structures, with no need for any semantic actions.
+
+The rest of this paper is organized as follows.  \S\ref{s:relwk}
+puts Nail in the context of related work.  \S\ref{s:design} describes
+Nail's design.  \S\ref{s:impl} discusses our initial implementation
+of Nail.  \S\ref{s:eval} provides some initial evaluation results.
+\S\ref{s:future} suggests several directions for future work, and
+\S\ref{s:concl} concludes.

### 2014-04-01 paper/intro.tex
turn Boost.Spirit into a citation

@@
 construct outputs, which is error-prone.  Some parser generators, such as
-Boost.Spirit,\footnote{\url{http://www.boost.org/doc/libs/1_55_0/libs/spirit/doc/html/index.html}}
+Boost.Spirit~\cite{boost-spirit},
 allow reusing the grammar for generating output from the internal

### 2014-04-01 paper/intro.tex
another paragraph

@@
 a precise grammar for the input data format, and to use a parser
-generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
+generator, such as {\tt lex} and {\tt yacc}, to synthesise the input
 processing code.  Developers that use a parser generator do not need
@@
 
-Second, data formats can have constraints XXX todo.
-Lengths.
-Counts.
-Checksums.
-
-Third, applications often need to produce output in the same
+Second,  applications often need to produce output in the same
 format as their input---for example, applications might both
@@
 into an AST\@.
+
+Third, data formats can have structural dependencies, such as offset, length and checksum fields
+that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
+manually write control code, such as re-positioning the parser's input stream, looping over the
+invocation of a sub-parser or computing a
+checksum over raw input bytes.  Besides leaving much room for errors with offset arithmetic, such
+code is usually not reusable when generating output.
 

### 2014-04-01 paper/intro.tex
intro and title

@@
 
-Third, data formats can have structural dependencies, such as offset, length and checksum fields
+Third, data formats, especially binary formats such as PNG or PDF, can have \emph{structural dependencies}, such as offset, length and checksum fields
 that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
@@
 
-This paper presents the initial design and implementation of Nail, a
-parser generator that greatly reduces the programmer effort required to
-use a grammar-based parser.  Nail addresses the above three challenges
-with several key ideas.  First, Nail XXX todo.  Bijections.  Reduced
-expressiveness.  Constraints.  External interaction (checksum, ports,
-etc).
+This paper presents the initial design and implementation of Nail, a parser
+generator that greatly reduces the programmer effort required to use a
+grammar-based parser. Nail addresses the above three challenges with several key
+ideas. First, Nail reduces the expressiveness of its grammar language by
+removing semantic actions. Existing parser generators allow arbitrary
+computation to transform between the AST and the parser output.%  Instead, Nail
+% derives the structure of its output automatically from the grammar, forcing the
+% programmer to clearly separate syntactic validation and semantic processing.
 
-We have implemented an initial prototype of Nail in C\@.  Our experience
+Second, this allows the parser to construct a \emph{syntactic bijection} between input
+and output. That is, parsing is the inverse of generating output. However,
+Nail allows \emph{constants} in the external format to have multiple
+representations if they should not affect the semantics of the data. For
+example, in a text protocol, the amount of white-space separating tokens should
+not affect the meaning of the data and is consequently not exposed to the
+application by Nail in the internal representation. However, as long as
+constants are only used for their intended purpose of representing syntax-only
+features, the generated output will have the same semantics as the parsed input.
+
+Third, Nail provides support for structural dependencies, transparently handling
+offset and length fields. XXX todo
+
+%  XXX todo.  Bijections.  Reduced
+% expressiveness.  Constraints.  External interaction (checksum, ports,
+% etc).
+
+We have implemented an initial prototype of Nail for C\@.  Our experience
 so far suggests that it is a promising approach: we were able to construct

### 2014-04-01 paper/main.tex
some typesetting tweaks
affiliation
acks
fixing up some bib stuff
kill another unused .bib file
stub conclusion
split off relwk
intro and title

@@
-\documentclass[10pt]{article}
+\documentclass[10pt,twocolumn]{article}
+\usepackage[10pt,nocopyright]{sigmin}
+\usepackage[letterpaper, breaklinks, pdfborder={0 0 0}]{hyperref}
+\usepackage[square,comma,numbers,sort&compress]{natbib}
+\usepackage{times}
+\usepackage{microtype}
+\usepackage{booktabs}
+\usepackage{mathptmx}
 \usepackage{graphicx}
@@
 
+\renewcommand{\ttdefault}{pxtt}   %% better \tt font
+
+%% Ensure ligatures (e.g., ``fine official flag'') can be copy/pasted from PDF.
+\input{glyphtounicode}
+\pdfgentounicode=1
+
 \begin{document}
-\author{Julian Bangert and Nickolai Zeldovich}
+\author{Julian Bangert and Nickolai Zeldovich \\ \em MIT CSAIL}
 \date{}
-\title{Nail: A practical interface generator for binary formats}
+\title{Nail: A practical interface generator for data formats}
 
@@
 \input{abstract}
-\input{background}
+\input{intro}
+\input{relwk}
 \input{design}
@@
 \input{future}
+\input{concl}
+\input{ack}
 
-\bibliography{main,locasto,login,phrack,short,langsec}
+\bibliography{n-str,main,locasto,phrack,langsec,n,n-conf}
 \bibliographystyle{plain}

### 2014-04-01 paper/intro.tex
small intro tweaks

@@
 generator that greatly reduces the programmer effort required to use a
-grammar-based parser. Nail addresses the above three challenges with several key
-ideas. First, Nail reduces the expressiveness of its grammar language by
-removing semantic actions. Existing parser generators allow arbitrary
-computation to transform between the AST and the parser output.%  Instead, Nail
+grammar-based parser. Nail addresses the above three challenges with
+several key ideas, as follows.
+
+First, Nail reduces the expressiveness of its grammar language by
+removing semantic actions. Although this limits the kinds of grammars that
+Nail can support, compared to parser generators that allow arbitrary
+computation in semantic actions, many widely used formats can be still
+captured by Nail grammars, and the lack of arbitrary semantic actions
+enables Nail to automate parsing input and producing output.
+
+%  Instead, Nail
 % derives the structure of its output automatically from the grammar, forcing the
@@
 
-Second, this allows the parser to construct a \emph{syntactic bijection} between input
-and output. That is, parsing is the inverse of generating output. However,
-Nail allows \emph{constants} in the external format to have multiple
-representations if they should not affect the semantics of the data. For
-example, in a text protocol, the amount of white-space separating tokens should
-not affect the meaning of the data and is consequently not exposed to the
-application by Nail in the internal representation. However, as long as
-constants are only used for their intended purpose of representing syntax-only
-features, the generated output will have the same semantics as the parsed input.
+Second, this allows the parser to construct a \emph{syntactic bijection}
+between input and output---that is, parsing is the inverse of generating
+output.  The bijection does not guarantee that all aspects of the data
+format are preserved; for example, in a text protocol, the amount of
+white-space separating tokens should not affect the meaning of the
+data, and is consequently not exposed to the application by Nail in the
+internal representation.  More generally, Nail allows \emph{constants}
+in the external format to have multiple representations if they do not
+affect the semantics of the data.  As long as constants are only used
+for their intended purpose of representing syntax-only features, the
+generated output will have the same semantics as the parsed input.
 

### 2014-04-01 paper/relwk.tex
rough intro pass done, with some sizable holes left
Make hammer a citation
related work cleaned up

--- 2014-04-01 paper/relwk.tex
@@
 \section{Related work}
-
-\paragraph{Parsing frameworks.}
-
-Generating parsers and generators from an executable specification
-is the core concept of Interface Generators, e.g. in CORBA or more
-recently~\cite{varda2008}. However, while interface generators work
-very well for existing grammars, they do not allow full control over the
-format of the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes
-system~\cite{mccann2000packet}, however PacketTypes works only as a
-parser, not as an output generator and does not support the expressive
-power of PEGs, but rather implements a C like structure model enhanced
-with data-dependent length fields and constraints.
-
-Parser generators for binary protocols were first introduced by the
-Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
-While previous parser generators could also be used to write grammars
-for binary protocols\footnote{Theoretically speaking, the alphabet
-over which a grammar is an  abstract set, so most algorithms work
-just as well on an alphabet of $\{0,1\}$}, doing so is impractically
-inconvenient. Hammer allows the programmer to specify a grammar in
-terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing are transparently hidden by the library. Hammer
-implements grammars as language-integrated parser combinators, an approach
-popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
-parser combinator style (to our knowledge, first  described
-in~\cite{burge1975recursive}) is a natural way of concisely expressing
-top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
-more background on the history of expressing grammars, see Bryan Ford's
-masters thesis~\cite{ford2002packrat}, which also describes the default
-parsing algorithm used by Hammer} by composing them from one or multiple
-sub-parsers.  Hammer then constructs a tree of function pointers which
-can be invoked to parse a given input into an abstract syntax tree (AST).
+\label{s:relwk}
 
@@
 are distributed as .zip files. Signatures are verified with a Java
-program, but the program is extracted with a C program.  The java
+program, but the program is extracted with a C program.  The Java
 program interprets all fields as signed, whereas the C program
 treats them as unsigned, allowing one to replace files in a signed
-archive.}~\cite{saurik-masterkey} code signing and even the X.509 protocol
+archive, thereby undermining Android's security model.}~\cite{saurik-masterkey} code signing and even the X.509 protocol
 underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
@@
 
+A subclass of logic errors are so-called \textit{weird machines}, where implementation side effects
+or under-specified parser behaviour leads to a protocol inadvertently becoming a Turing complete
+execution environment. Frequently, this execution environment can either then directly manipulate
+data in unwanted ways or be used to make exploiting another bug feasible\footnote{For example, by
+  compiling a return-oriented-programming exploit from code fragments discovered on the fly}.
+ Examples include X86 page tables~\cite{bangert2013page},
+ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive research community,
+this has been generalised into treating a program as a \textit{weird
+machine}~\cite{bratus2011exploit} that operates on an input, analogous
+to a virtual machine operating on bytecode.
+
+% With currently prevalent software design methods, it is hard to isolate
+% input-handling code from the rest of the program, as user input is
+% typically passed through the program and processed in bits and pieces as a
+% ``shotgun parser''~\cite{shotgun-parser}. Thus, partially invalid input might be partly 
+% processed by the program until the invalid sections are encountered.
+
+\paragraph{Parsing frameworks.}
+Proper input recognition has  been shown to be an excellent way of
+eliminating malicious inputs. In one case, a  PDF parser implemented in
+Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
+However, manually writing parser code does not scale to the number of
+file formats  and protocols in existence and might result in parser code
+tied to one specific application.
+
+Generating parsers and generators from an executable specification
+is the core concept of Interface Generators, e.g. in CORBA or more
+recently~\cite{varda2008}. However, while interface generators work
+very well for existing grammars, they do not allow full control over the
+format of the output, so cannot be used to implement legacy protocols.
+Very related work has been done at Bell Labs with the PacketTypes
+system~\cite{mccann2000packet}, however PacketTypes works only as a
+parser, not as an output generator and does not support the expressive
+power of PEGs, but rather implements a C like structure model enhanced
+with data-dependent length fields and constraints.
+
+Parser generators for binary protocols were first introduced by the
+Hammer~\cite{hammer-parser} parser.
+While previous parser generators could also be used to write grammars
+for binary protocols\footnote{Theoretically speaking, the alphabet
+over which a grammar is an  abstract set, so most algorithms work
+just as well on an alphabet of $\{0,1\}$}, doing so is impractically
+inconvenient. Hammer allows the programmer to specify a grammar in
+terms of bits and bytes instead of characters. Common concerns, such as
+endianness and bit-packing are transparently hidden by the library. Hammer
+implements grammars as language-integrated parser combinators, an approach
+popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
+parser combinator style (to our knowledge, first  described
+in~\cite{burge1975recursive}) is a natural way of concisely expressing
+top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
+more background on the history of expressing grammars, see Bryan Ford's
+masters thesis~\cite{ford2002packrat}, which also describes the default
+parsing algorithm used by Hammer} by composing them from one or multiple
+sub-parsers.  Hammer then constructs a tree of function pointers which
+can be invoked to parse a given input into an abstract syntax tree (AST).
+
 \paragraph{Application use of parsers.}
+Generated parsers have long been used to parse human input, such as programming languages and
+configuration files. Frequently, such languages are often specified with a formal grammar in an
+executable form.
 
-Unfortunately, grammars are seldom used directly to recognise
+Unfortunately,  parser frameworks are seldom used  to recognise
 machine-created input. For example, the security-critical and very
@@
 
-With currently prevalent software design methods, it is hard to isolate
-input-handling code from the rest of the program, as user input is
-typically passed through the program and processed in bits and pieces as a
-``shotgun parser''~\cite{shotgun-parser}.
-
-Many protocol implementations are so complicated that even
-without bugs, they inadvertently provide a full Turing-complete execution
-environment. Examples include X86 page tables~\cite{bangert2013page},
-ELF symbols and relocations~\cite{shapiro2013weird}. Even though
-these protocols were never intended to be that expressive -
-and consequentially hard to parse, implementation details of their
-recognisers - which de facto became a part of the specification and were
-relied upon by legitimate inputs - expose unwanted computational power
-and have unintended side-effects.  In the offensive research community,
-this has been generalised into treating a program as a \textit{weird
-machine}~\cite{bratus2011exploit} that operates on an input, analogous
-to a virtual machine operating on bytecode.
-
-Parser generators can automatically create parsers for such languages,
-and with common classes of grammars such as regular expressions or
-deterministic context free languages, that parser can be expressed in a
-computationally less powerful form, e.g. a finite automaton or a push
-down automaton, limiting the weird machine problem and enabling very
-efficient implementation~\cite{Knuth1965607}.
-
-Proper input recognition has been shown to be an excellent way of
-eliminating malicious inputs. In one case, a  PDF parser implemented in
-Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
-However, manually writing parser code does not scale to the number of
-file formats  and protocols in existence and might result in parser code
-tied to one specific application.
-
-In fact, definitive grammars have long been written for all sorts of
-protocols and are the standard way of describing text-based protocols.
-Specifications of binary protocols, such as RFCs, also contain grammars,
-although they are frequently translated into prose and diagrams as
-opposed to the traditional Backus-Naur-Form.
+%%This is supposed to go into design
+% While it is possible to express short transformations on the input
+% entirely as semantic actions\footnote{This is in fact the design
+% rationale, to perform computation on the fly as the parser walks the
+% parse tree.}, more complicated programs usually construct an internal
+% representation, which contains all relevant information from the input
+% in a format native to the programming language used. For example, a C
+% programmer ideally wants to deal with structs and NULL-terminated arrays,
+% whereas a C++ programmer might expect STL containers, a Java programmer
+% interfaces, a Haskell programmer records and a LISP programmer property
+% lists.  The structure of this internal representation usually resembles
+% the structure of the grammar.

### 2014-04-01 paper/design.tex
rough intro pass done, with some sizable holes left
Intro

@@
 \section{Design}
+\label{s:design}
 
 \paragraph{Overview}
-To alleviate these problems, Nail provides a richer set of combinators that induce the structure of
-an internal model, while also describing the grammar of the external protocol. Nail takes the
-combinators and produces type declarations for an internal model, the \textit{parser}, a function to parse a sequence of
+
+To alleviate these problems, Nail provides a richer set of combinators that not only describes the
+grammar of the external protocol, but also induce an internal model.
+
+A central design decision of Nail is that there is a semantic bijection between the model exposed to
+the programmer and the bytes input and output up to syntactic equivalence. More precisely, the
+parser is the generators inverse, so parsing the generators output will yield the generators input,
+but generating the parsers output does not necessarily yield the parsers input. To understand why
+this makes sense, consider a grammar for a text language that tolerates white space or a binary
+protocol that tolerates arbitrarily long padding\footnote{Say, the physical layer of most
+  communication protocols is a possibly infinite sequence of symbols that are syntactically nil
+  followed by a pre-determined synchronisation sequence and the actual contents of the
+  transmission.}. In that case, the program semantics should be independent of the number of padding
+elements in the input, and Nail therefore does not expose that information to the programmer. We
+call such discarded fields \emph{constants}. Currently, Nail always makes a default choice when
+there are multiple options to express a constant, however Nail could be extended to allow a
+grammar-specific plug-in to make these choices, say for faster alignment, streaming applications
+when data is not ready or visual appearance in human-readable protocols. However, if some features
+of Nail are avoided, there is a proper isomorphism between model and protocol.
+
+Nail evaluates the combinators and produces type declarations for the internal model, the \textit{parser}, a function to parse a sequence of
 bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
@@
 
-A central design decision of Nail is that there is a  bijection between the model exposed to the programmer and the bytes
-input and output up to syntactic equivalence. More precisely, the parser is the generators inverse,
-so parsing the generators output will yield the generators input, but generating the parsers output
-does not necessarily yield the parsers input. To understand why this makes sense, consider a grammar
-for a text language that tolerates white space or a binary protocol that tolerates arbitrarily long
-padding\footnote{Say, the physical layer of most communication protocols is a possibly infinite
-  sequence of symbols that are syntactically nil followed by a pre-determined synchronisation
-  sequence and the actual contents of the transmission.}. However, if some features of Nail are
-avoided, there is a proper isomorphism between model and protocol.
-
-In the following, we present the combinators in Nail grouped by the combinators of conventional Parser Expression Grammars they replace.
+In the following, we present the combinators in Nail.
 \paragraph{Fundamental parsers}
@@
 \paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of named parsers and
-unnamed constant parsers. 
-When parsing, it parses each parser in sequence and returns a structure containing the result of
-each parser.
+Nails fundamental concept is the structure combinator. It contains a list of
+named parsers and unnamed constant parsers. When parsing, it parses each parser
+in sequence and returns a structure containing the result of each parser.
 

### 2014-04-01 paper/intro.tex
Intro

--- 2014-04-01 paper/intro.tex
@@
 grammar-based parser. Nail addresses the above three challenges with several key
-ideas. First, Nail reduces the expressiveness of its grammar language by
-removing semantic actions. Existing parser generators allow arbitrary
-computation to transform between the AST and the parser output.%  Instead, Nail
-% derives the structure of its output automatically from the grammar, forcing the
-% programmer to clearly separate syntactic validation and semantic processing.
+ideas.
 
-Second, this allows the parser to construct a \emph{syntactic bijection} between input
-and output. That is, parsing is the inverse of generating output. However,
-Nail allows \emph{constants} in the external format to have multiple
-representations if they should not affect the semantics of the data. For
-example, in a text protocol, the amount of white-space separating tokens should
-not affect the meaning of the data and is consequently not exposed to the
-application by Nail in the internal representation. However, as long as
+First, Nail reduces the expressiveness of its grammar language by removing
+semantic actions. Existing parser generators allow arbitrary computation to
+transform between the AST and the parser output. Instead, Nail derives the
+structure of its output automatically from the grammar, forcing the programmer
+to clearly separate syntactic validation and semantic processing.
+
+Second, this well-defined internal representation allows Nail to inverse the
+parser and generate output. However, Nail allows \emph{constants} in the
+external format to have multiple representations if they should not affect the
+semantics of the data. For example, in a text protocol, the amount of
+white-space separating tokens should not affect the meaning of the data and
+consequently Nail does not expose it to the application. As long as
 constants are only used for their intended purpose of representing syntax-only
@@
 Third, Nail provides support for structural dependencies, transparently handling
-offset and length fields. XXX todo
+offset and length fields. By hiding the offset and length values from the
+programmer, Nail ensures they remain consistent even if the data is changed by
+the application.
 

### 2014-04-02 paper/design.tex
Design

@@
 
-To alleviate these problems, Nail provides a richer set of combinators that not only describes the
-grammar of the external protocol, but also induce an internal model.
+To alleviate these problems, Nail provides a richer set of combinators that not
+only describes the grammar of the external protocol, but also induce an internal
+model.
 
-A central design decision of Nail is that there is a semantic bijection between the model exposed to
-the programmer and the bytes input and output up to syntactic equivalence. More precisely, the
-parser is the generators inverse, so parsing the generators output will yield the generators input,
-but generating the parsers output does not necessarily yield the parsers input. To understand why
-this makes sense, consider a grammar for a text language that tolerates white space or a binary
-protocol that tolerates arbitrarily long padding\footnote{Say, the physical layer of most
-  communication protocols is a possibly infinite sequence of symbols that are syntactically nil
-  followed by a pre-determined synchronisation sequence and the actual contents of the
-  transmission.}. In that case, the program semantics should be independent of the number of padding
-elements in the input, and Nail therefore does not expose that information to the programmer. We
-call such discarded fields \emph{constants}. Currently, Nail always makes a default choice when
-there are multiple options to express a constant, however Nail could be extended to allow a
-grammar-specific plug-in to make these choices, say for faster alignment, streaming applications
-when data is not ready or visual appearance in human-readable protocols. However, if some features
-of Nail are avoided, there is a proper isomorphism between model and protocol.
+A central design decision of Nail is that there is a semantic bijection between
+the model exposed to the programmer and the bytes input and output up to
+syntactic equivalence for unambiguous grammars. More precisely, the parser is
+the generators inverse, so parsing the generators output will yield the
+generators input, but generating the parsers output does not necessarily yield
+the parsers input. To understand why this makes sense, consider a grammar for a
+text language that tolerates white space or a binary protocol that tolerates
+arbitrarily long padding\footnote{Say, the physical layer of most communication
+  protocols is a possibly infinite sequence of symbols that are syntactically
+  nil followed by a pre-determined synchronisation sequence and the actual
+  contents of the transmission.}. In that case, the program semantics should be
+independent of the number of padding elements in the input, and Nail therefore
+does not expose that information to the programmer. We call such discarded
+fields \emph{constants}. Currently, Nail always makes a default choice when
+there are multiple options to express a constant, however Nail could be extended
+to allow a grammar-specific plug-in to make these choices, say for faster
+alignment, streaming applications when data is not ready or visual appearance in
+human-readable protocols. Similarly, Nail does not preserve the layout of
+objects referred to by their offsets. If the grammar contains no constants and
+offset fields, there is a proper isomorphism between model and protocol.
 
-Nail evaluates the combinators and produces type declarations for the internal model, the \textit{parser}, a function to parse a sequence of
-bytes into an instance of the model and the \textit{generator}, a function to create a sequence of
-bytes from an instance of the model.
+Nail evaluates the combinators and produces type declarations for the internal
+model, the \textit{parser}, a function to parse a sequence of bytes into an
+instance of the model and the \textit{generator}, a function to create a
+sequence of bytes from an instance of the model. In the following, we present
+Nail's combinators.
 
-In the following, we present the combinators in Nail.
 \paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and unsigned integers with
-arbitrary lengths up to 64 bits. Note that is possible to define parsers for sub-byte lengths, e.g.
-to parse the 4-bit data offset within the TCP header. Integer parsers return their value in the
+The elementary parsers of Nail are the same as those of Hammer, signed and
+unsigned integers with arbitrary lengths up to 64 bits. Note that is possible to
+define parsers for sub-byte lengths, e.g. to parse the 4-bit data offset within
+the TCP header. Integer parsers return their value in the
 smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
@@
 
-Integer parsers can be constrained to fall either within an (inclusive) range of values or be an
+Integer parsers can be constrained to fall either within an (inclusive) range of
+values or be an
 element of a set of allowed values. \\
-Examples: \texttt{uint8 | 1..3} (accepts an 8 bit unsigned
-integer between 1 and 3 inclusive) ,\\
- \texttt{uint16| ..512} (an unsigned 16-bit integer less or
-equal than 512),\\
-\texttt{int32| [1,255,512]} (a signed 32 bit integer with value 1, 255 or 512).
-
-\paragraph{Constant parser}
-Nail also has constant parsers, which do not return a value. The simplest constant parser is an
-integer with fixed value, e.g. \texttt{uint8=0}. Constant parsers succeed if the inner integer
-parser returns the given value and fail otherwise.
+Examples:\\
+\texttt{uint8 | 1..3} (8 bit unsigned integer $1\leq x \leq 3$) ,\\
+\texttt{uint16| ..512} (unsigned 16-bit integer $\leq 512$),\\
+\texttt{int32 | [1,255,512]} (signed 32 bit integer with value 1, 255 or 512).
+Invalid values or prematurely reaching the end of input raise an error when
+parsing input or generating output and abort the procedure.
 
@@
 Nails fundamental concept is the structure combinator. It contains a list of
-named parsers and unnamed constant parsers. When parsing, it parses each parser
+named parsers and unnamed constant parsers  . The parser invokes each field 
 in sequence and returns a structure containing the result of each parser.
@@
   \textbf{Nail} & \textbf{Data Model}\\
-\begin{minipage}{2in}
+  \begin{minipage}{1.2in}
 \begin{verbatim}
@@
   & 
-\begin{minipage}{2in}
+  \begin{minipage}{1.2in}
 \begin{verbatim}
@@
 \paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails, returning an array of
-the inner parsers results. SepBy additionally takes a constant parser, which it applies in between
+The many combinator takes a parser and applies it repeatedly until it fails,
+returning an array of the inner parsers results. SepBy additionally takes a
+constant parser, which it applies in between
 parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples \texttt{many uint8}, \texttt{sepBy uint8=',' (many uint8| '0'..'9')} (which recognises
-comma-separated lists of numbers).
-
-The many combinator also works on constant parsers, although as described above it can make the
-generator not bijective.\footnote{By default, the generator will emit exactly one iteration of the
-  constant parser in its output.} Example: \texttt{many uint8= "test"}
+Examples: \texttt{many uint8}\\ \texttt{sepBy uint8=',' (many uint8| '0'..'9')}
+(which recognises comma-separated lists of numbers).
 
 \paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator, which picks the first choice that succeeds, with a tag
-for each choice. The result is saved in a tagged union.
+We extend PEG's ordered choice combinator with a tag for each choice. The parser
+attempts to parse each option in the order they are specified in the field and
+stores the result in a tagged union. If an option fails, the parser backtracks
+to the beginning of the choice combinator's input. Care must be taken that the
+choices do not overlap, because Nail always picks to the first successfully
+parsed choice. If two options overlap, generated output for the latter option is
+not necessarily understood identically by the parser. However, actual data
+formats are usually not ambiguous in this sense.
 
@@
 \end{verbatim}
-\paragraph{Wrap}
 
-The wrap combinator parses one or more constant parser, then an inner parser whose value it returns
-and then again parses one or more constant parsers.\\ Example: 
-\verb+<uint8='"'; many8 uint8|'a'..'z'; uint8='"'  >+
+\subsection{Constant parsers}
+Nail also features \emph{constant parsers}, which are not affecting the internal
+representation. Constant parsers can appear instead of structure fields and as
+separators in the SepBy combinator.
 
-\subsection{Offset fields}
-Another problem for parser generators is that binary protocols often contain length and offset
-fields. While conventional parsing algorithms can deal with bounded offset fields - a finite
-automaton can count a bounded integer, and we can feed the (finite) input multiple times to the
-finite automaton. However, this implementation is both time-inefficient - why feed many bytes into
-the automaton that will just be skipped - and very cumbersome to express with current parser
-generators. Therefore, if languages with offset fields are parsed with parser generators, the only
-currently feasible way is to add 'ad-hoc' hacks such as changing the input pointer of the generated
-parser on the fly from semantic actions.
+The simplest constant parser is an integer or array of integers with fixed
+value, e.g. \texttt{uint8=0}, \texttt{many uint8=[1,2]}. A convenience notation
+for strings is also supported: \texttt{many uint8 = "foo"}.
 
-Nail will properly support both offset and length fields and much of the following discussion
-applies to both , although the current prototype only implements lengths, which we will focus on.
+\paragraph{Wrap combinator.} When implementing real protocols with Nail, we often found that many structures
+consist of many constant parsers and only one named field. This pattern is
+common in binary protocols which use fixed headers to denote the type of data
+structure to be parsed.  In order to keep the internal representation cleaner,
+we introduced the wrap combinator, which takes a sequence of parsers containing
+exactly one non-constant parser. The parser and generator act as though the wrap
+combinator was a sequence of parsers, but the data model does not wrap the
+single value in another structure, making the application code more concise.\\
+Example: \\ \verb+<uint8='"';many int8|'a'..'z'; uint8='"'>+  parses a quoted
+lowercase word into an array - excluding the quotation marks.
 
-\paragraph{Dependency fields}
-We call length or offset fields \textit{dependency fields}, because during parsing, another parser
-depends on them, and while generating output, their value depends on some other structure in the
-data model.
-Dependency fields appear in a structure combinator as would any other field, but their name begins
-with an '@' sign and they have to return an integer value. Dependency fields are not directly
-represented in the internal model.
+\paragraph{Constant combinators.}
+In some protocols, there might be many ways to represent the same constant field
+and there is no semantic difference between different syntaxes.  Nail therefore
+supports choice (\texttt{|| uint8 = 0 || uint16 = 0xFFFF}) and repetition
+(\texttt{many (uint8= ' ')}). This is particularly useful for handling padding or whitespace.
 
+As discussed above, choosing to use these combinators on constant parsers
+removes the bijection between byte-strings and our data model, as there are
+multiple byte-strings that correspond to the same internal data structure and
+the generator has to choose one of these representations. As such, constant
+combinators are the generator dual of ambiguous choice combinators in the
+parser, because they lead to ambiguities in the generator.
+
+\subsection{Dependent fields}
+Another problem for parser generators is that binary protocols often contain
+length and offset fields. While conventional parsing algorithms can deal with
+bounded offset fields - a finite automaton can count a bounded integer, and we
+can feed the (finite) input multiple times to the finite automaton. However,
+this implementation is both time-inefficient - why feed many bytes into the
+automaton that will just be skipped - and very cumbersome to express with
+current parser generators. Therefore, if languages with offset fields are parsed
+with parser generators, the only currently feasible way is to add 'ad-hoc' hacks
+such as changing the input pointer of the generated parser on the fly from
+semantic actions.
+
+Nail will properly support both offset and length fields and much of the
+following discussion applies to both , although the current prototype only
+implements lengths, which we will focus on.
+
+We call length or offset fields \textit{dependent fields}, because during
+parsing, another parser depends on them, and while generating output, their
+value depends on some other structure in the data model. Dependency fields
+appear in a structure combinator as would any other integer field, but their
+name begins with an '@' sign. Dependency fields are not exposed in the data
+model, but transparently computed.
+
+A dependency field has to appear in the grammar before it can be used. 
 \paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the parser as often as the
-value of the dependency field, returning an array of the values. When generating output, it emits
-the array and writes its length to the dependency field.
+The length combinator takes a dependency field and a parser, evaluating the
+parser as often as the value of the dependency field, returning an array of the
+values. When generating output, it emits the array and writes its length to the
+dependency field.
+\paragraph{Offsets: offset} 
+The offset combinator takes a dependency field and a parser. It moves the parser
+to the position specified in the offset field and invokes the inner parser, and
+then moves the input back to its original position.
+While generating output, all structures referred to by offset are generated
+after the main structure and the dependent offset fields patched up.
 
+\paragraph{Checksums: raw_depend}
+In many data formats, some values depend on external representation, such as
+checksums and cryptographic signatures. While it would be possible to extend
+our constraint language to be powerful enough to support such constructs, we
+would essentially be building a separate, Turing complete language that has all
+the same pitfalls existing programs have. Therefore, we intend to  allow the programmer to
+very carefully escape Nail's programming model and write a function that takes a
+pointer to the dependent value and a range of bytes. 
+This feature is not implemented in the current prototype. \\
+Example:\\
+ \verb+{ data many uint8; @checksum uint32+\\
+ \verb+raw_depend @checksum data crc32 }+\\
+ \verb+/*bool crc32(uint32_t *out, uint8_t *in);*/+
+Because this feature compromises Nail's security guarantees, it should only be
+used in limited circumstances and with carefully prepared checksum functions. 
 \subsection{Example Grammars}
@@
 \paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code points.
+This grammar recognises valid UTF-16 strings and exposes an array of code
+points.
 

### 2014-04-02 paper/design.tex
github link. misc

@@
 used in limited circumstances and with carefully prepared checksum functions.
-\subsection{Example Grammars}
-To further familiarise the reader with Nail, we provide two example grammars.
-\paragraph{UTF-16}
-This grammar recognises valid UTF-16 strings and exposes an array of code
-points.
-
-\begin{tabular}{ll}
-  \textbf{Nail} & \textbf{Data Model}\\
-  \begin{minipage}{3in}
-\begin{verbatim}
-utfstring = many choose {
-         SUPP= {
-               lead uint16 | 55296..56319
-                          // 0xD800..0xDBFF
-               trail uint16 | 56320..57343
-                          // 0xDC00..0xDFFF
-               }
-         BASIC = uint16 | !55296..57343
-       }
-\end{verbatim}
-  \end{minipage}
-  & 
-  \begin{minipage}{2in}
-\begin{verbatim}
-struct utfstring {
- struct {
-   enum  {SUPP,BASIC} N_type;
-   union {
-    struct {
-     uint16_t lead;
-     uint16_t trail;
-    } SUPP;
-    uint16_t BASIC;
-   };
- }*elem;
- size_t count;
-};
-\end{verbatim} 
-  \end{minipage}
-  \\
-\end{tabular}
-
-\paragraph{DNS packet}
-This grammar recognises DNS packets without label compression, as per RFC1035
-\begin{verbatim}
-labels = <many {@length  uint8 | 1..255 
-                label n_of @length uint8 }
-             uint8 = 0>
-question= {   labels labels
-              qtype uint16 | 1..16 
-              qclass uint16 | [1,255]
-               }
-answer  = {
-             labels labels
-             rtype uint16 | 1..16
-             class uint16 | [1]
-             ttl uint32
-             @rlength uint16 
-             rdata n_of @rlength uint8 
-             }
-dnspacket = {id uint16
-              qr uint1
-              opcode uint4
-              aa uint1 
-              tc uint1
-              rd uint1
-              ra uint1
-              uint3 = 0
-              rcode uint4
-              @questioncount uint16
-              @answercount uint16
-              authoritycount uint16   // Ignored
-              additionalcount uint16  //Ignored       
-              questions n_of @questioncount question         
-              responses n_of @answercount answer
-            }
-\end{verbatim}

### 2014-04-02 paper/eval.tex
rough intro pass done, with some sizable holes left
github link. misc

@@
 \section{Evaluation}
+
+\label{s:eval}
+In the following section we will present two short Nail grammars for UTF-16
+encoded strings and for a subset of DNS packets sent to and from an
+authoritative name server. Then we present a functioning toy DNS server that we
+have implemented to demonstrate the feasibility of using Nail's internal
+representation.
+
+As Nail is work in progress, many parts of the implementation, syntax and design
+are not complete yet and we do not yet have meaningful performance or security metrics.
+
+\subsection{Example Grammars}
+\paragraph{UTF-16.}
+This grammar recognises valid UTF-16 strings and exposes an array of code
+points.
+\hspace{-2em}\begin{tabular}{ll}
+  \textbf{Nail} & \textbf{Data Model}\\
+  \begin{minipage}{1.3in}
+\begin{verbatim}
+utf = many choose {
+ SUPP= {
+  lead uint16  |
+     0xD800..0xDBFF
+  trail uint16 | 
+     0xDC00..0xDFFF
+ }
+ BASIC = uint16 |
+    !0xD800..0xDFFF
+}
+\end{verbatim}
+  \end{minipage}
+  & 
+  \begin{minipage}{2in}
+\begin{verbatim}
+struct utfstring {
+ struct {
+   enum  {SUPP,BASIC} 
+         N_type;
+   union {
+    struct {
+     uint16_t lead;
+     uint16_t trail;
+    } SUPP;
+    uint16_t BASIC;
+   };
+ }*elem;
+ size_t count;
+};
+\end{verbatim} 
+  \end{minipage}
+  \\
+\end{tabular}
+
+\paragraph{DNS packet}
+This grammar recognises DNS packets without label compression, as per RFC1035.
+\begin{verbatim}
+labels = <many {@length  uint8 | 1..255 
+                label n_of @length uint8 }
+             uint8 = 0>
+question= {   labels labels
+              qtype uint16 | 1..16 
+              qclass uint16 | [1,255]
+          }
+answer  = {   labels labels
+              rtype uint16 | 1..16
+              class uint16 | [1]
+              ttl uint32
+              @rlength uint16 
+              rdata n_of @rlength uint8 
+          }
+dnspacket = { id uint16
+              qr uint1
+              opcode uint4
+              aa uint1 
+              tc uint1
+              rd uint1
+              ra uint1
+              uint3 = 0
+              rcode uint4
+              @qc uint16
+              @ac uint16
+              uint16 = 0   // authority 
+              uint16 = 0 // additional
+// We don't support authority or 
+// additional sections in the prototype
+              questions n_of @qc question         
+              responses n_of @ac answer
+            }
+\end{verbatim}
+
+Furthermore, our GitHub\footnote{\url{https://github.com/jbangert/nail/tree/master/examples}}
+repository contains other Nail grammars, such as a TAP network stack that processes Ethernet, ARP,
+ICMP, IP and UDP and the grammar for Nail itself.
+
+\subsection{Applications}
+
 In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
@@
 ``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
+structure definitions, with only 52 lines of code defining the grammar with Hammer's combinators.
 
@@
 It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
+particular for DNS compression and additional hint records real-world DNS servers send. 
 However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims

### 2014-04-02 paper/impl.tex
github link. misc
Implementation

@@
 
-The current prototype of the Nail parser generator supports the C programming language and top down parsers.
-Options for C++ STL data models and emitting Packrat parser~\cite{packrat-parsing:icfp02} are under heavy development.
-In this section, we will discuss some particular
-features of our parser implementation.
+The current prototype of the Nail parser generator supports the C programming
+language and top down parsers. Options for C++ STL data models and emitting
+Packrat parser~\cite{packrat-parsing:icfp02} are under heavy development. In
+this section, we will discuss some particular features of our parser
+implementation.
 
+The source code of our implementation, together with the examples described in
+\S\ref{s:eval} are available on GitHub\footnote{\url{https://github.com/jbangert/nail}}.
 \paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to validate and recognise the
-input, the second to bind this data to the internal model. Currently the parser is a straightforward
-top down parser, although facilities have been made to add Packrat parsing to achieve linear parsing times.
+A generated Nail parser makes two passes through the input , the first to
+validate and recognise the input, the second to bind this data to the internal
+model. Currently the parser is a straightforward top down parser, although
+facilities have been made to add Packrat parsing to achieve linear parsing
+times.
 \paragraph{Memory allocation}
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed size blocks from the system malloc. However, Nail extends upon Hammers
-approach and uses two arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only for allocating the
-result and has to be freed by the programmer.
+Many security vulnerabilities can occur when heap allocations are done
+improperly. Therefore, just like Hammer, Nail avoids using the heap as much as
+possible, using a custom arena allocator and allocating only fixed size blocks
+from the system malloc. However, Nail extends upon Hammers approach and uses two
+arenas for each parsed input. One arena is used for intermediate results and is
+freed(and zeroed) after parsing completes, whereas the other arena is used only
+for allocating the result and has to be freed by the programmer.
 
 \paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate temporary
-data, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible without having validated the input, and we do not want to mix temporary
-objects with the results of our parse, we need some intermediate structure to cache parse results.
-This structure is also used as a value in the Packrat hash table.
+Most parser generators, such as Bison do not have to dynamically allocate
+temporary data, as they evaluate a semantic action on every rule. However, as
+our goal is to perform as little computation as possible without having
+validated the input, and we do not want to mix temporary objects with the
+results of our parse, we use an append-only trace to store intermediate parser
+results.
 
-Hammer solves this problem by storing a full abstract syntax tree. However, this abstract syntax
-tree is at least an order of magnitude larger than the input, because it stores a large tree node
-structure for each input byte and for each rule reduced. This allows Hammer semantic actions to get
-all necessary information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic actions, we can trust
-it as much as we trust the parser and thus expose it to the raw input stream.
+Hammer solves this problem by storing a full abstract syntax tree. However, this
+abstract syntax tree is at least an order of magnitude larger than the input,
+because it stores a large tree node structure for each input byte and for each
+rule reduced. This allows Hammer semantic actions to get all necessary
+information without ever seeing the raw input stream. However, because we also
+automatically generate our second pass, which corresponds to Hammers semantic
+actions, we can trust it as much as we trust the parser and thus expose it to
+the raw input stream.
 
-Under this premise, the actions need very limited information from the recogniser to correctly
-handle the input stream. In particular, the parsers control flow only branches at the choice and
-repeat combinators. For all other combinators, the action just has to assign a fixed sequence of
-fundamental parsers to a fixed sequence of internal model fields. 
-Thus we can represent the parsers success as a sequence of selected choices and counted repeats and
-use this to reconstruct the syntax of the input. %TODO: This sentence is HORRIBLE. Re-work 
+Under this premise, the actions need very limited information from the
+recogniser to correctly handle the input stream. In particular, the parsers
+control flow only branches at the choice, repetition and constant
+combinators.
+Thus, for each of those combinators, we store the minimum amount of information
+required to reconstruct the syntactic structure of the input. The trace is an
+array of integers.
+ Whenever the parser encounters a choice, it appends two integers to the trace. After it
+successfully parses a choice, the parser writes the number of that choice and
+the length of the trace when it began parsing that choice. When backtracking in
+the input, the parser does not backtrack in the trace. This means that offsets
+within the trace can be used for a Packrat hash-table to memoise backtrack-heavy
+parsers.
+
+When encountering a repetition combinator, the parser records the number of
+times the inner parser succeeded and when encountering a constant parser of
+variable size, it records how much input was consumed by the constant parser.
+
+In a second pass, the parser then allocates the internal representation from an
+arena allocator and binds the fields to values from the input, while following
+the trace to determine how many array fields to parse and which choices to pick.
 \paragraph{Dependency Fields}
-During parsing, dependency fields occur before the context in which they are used. The parser stores
-their value and retrieves it afterwards when encountering the combinator that uses them. When
-generating output, the dependency field is first filled with a filler value, then later when the
-first combinator that determines this fields value is encountered, the field is overwritten. Any
-further combinators using this dependency will then validate that the dependency field is correct. 
+During parsing, dependency fields occur before the context in which they are
+used. The parser stores their value and retrieves it afterwards when
+encountering the combinator that uses them. When generating output, the
+dependency field is first filled with a filler value, then later when the first
+combinator that determines this fields value is encountered, the field is
+overwritten. Any further combinators using this dependency will then validate
+that the dependency field is correct.
 \paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser generator uses a Nail parser
-to recognise Nail grammars. A superset of the grammar language described in this paper is
-implemented in 100 lines of Nail, which feed into about a thousand lines of C++ that implement the
-actual parser generator. Bootstrapping is supported via an LALR implementation.
+To demonstrate the feasibility of the nail parser generator, our parser
+generator uses a Nail parser to recognise Nail grammars. A superset of the
+grammar language described in this paper is implemented in 100 lines of Nail,
+which feed into about a thousand lines of C++ that implement the actual parser
+generator. Bootstrapping is supported via an implementation of the Nail language
+in the Lemon parser generator, a variant of Yacc.

### 2014-04-02 paper/concl.tex
Conclusion, eval, future work

@@
 \section{Conclusion}
+\label{s:concl}
+This paper presented the initial design and implementation of \textit{Nail}, a usable interface
+generator for data formats. Nail allows programmers to avoid memory corruption and ambiguity
+vulnerabilities while reducing effort in parsing and generating real world protocols and file
+formats. We have demonstrated Nail's expressive power by implementing DNS and Ethernet.
 
-This paper presented the initial design and implementation of
-\textit{Nail}, an XXX...
+Our code is open source on GitHub\footnote{\url{https://github.com/jbangert/nail/}}.

### 2014-04-02 paper/eval.tex
Conclusion, eval, future work

@@
 
-Furthermore, our GitHub\footnote{\url{https://github.com/jbangert/nail/tree/master/examples}}
+Furthermore, our GitHub
 repository contains other Nail grammars, such as a TAP network stack that processes Ethernet, ARP,

### 2014-04-02 paper/future.tex
rough intro pass done, with some sizable holes left
Conclusion, eval, future work

@@
 \section{Future work}
+\label{s:future}
 
 The Nail parser generator is work in progress and feedback of all sorts is heavily
-encouraged. Source code is available upon request and will be released publicly pending clean-up and
-maturity. Short-term future work will include improving the scoping of dependency fields and adding
-support for offset fields.
+encouraged.
 
 A major problem with the current design of Nail is that the design of the grammar dictates the
-interna structure of the software. This makes changing grammars or adding Nail to existing software
+internal structure of the software. This makes changing grammars or adding Nail to existing software
 awkward. One possible solution to this problem would be to implement a concept similar to Relational

### 2014-04-02 paper/intro.tex
minor nits in the intro

@@
 
-Third, data formats, especially binary formats such as PNG or PDF, can have \emph{structural dependencies}, such as offset, length and checksum fields
+Third, data formats, especially binary formats such as PNG or PDF, can have \emph{structural dependencies}, such as offset, length, and checksum fields
 that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
 manually write control code, such as re-positioning the parser's input stream, looping over the
-invocation of a sub-parser or computing a
+invocation of a sub-parser, or computing a
 checksum over raw input bytes.  Besides leaving much room for errors with offset arithmetic, such
@@
 
-Second, this well-defined internal representation allows Nail to inverse the
+Second, this well-defined internal representation allows Nail to invert the
 parser and generate output. However, Nail allows \emph{constants} in the

### 2014-04-02 paper/relwk.tex
forgot to remove todo notice
some relwk nits; one XXX regarding hammer vs nail

@@
 make it more difficult for an adversary to exploit these bugs.
+Nail helps developers of lower-level languages avoid these bugs
+in the first place.
 
@@
 that guarantee memory safety.
+By allowing developers to specify their data format just once, Nail
+avoids logic errors and inconsistencies in parsing and output generation.
 
-A subclass of logic errors are so-called \textit{weird machines}, where implementation side effects
-or under-specified parser behaviour leads to a protocol inadvertently becoming a Turing complete
-execution environment. Frequently, this execution environment can either then directly manipulate
-data in unwanted ways or be used to make exploiting another bug feasible\footnote{For example, by
-  compiling a return-oriented-programming exploit from code fragments discovered on the fly}.
- Examples include X86 page tables~\cite{bangert2013page},
-ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive research community,
-this has been generalised into treating a program as a \textit{weird
-machine}~\cite{bratus2011exploit} that operates on an input, analogous
-to a virtual machine operating on bytecode.
+A subclass of logic errors are so-called \textit{weird machines}, where
+implementation side effects or under-specified parser behaviour leads to a
+protocol or data format inadvertently becoming a Turing-complete execution
+environment, even though the original grammar did not require it.
+Frequently, this execution environment can either
+then directly manipulate data in unwanted ways or be used to make
+exploiting another bug feasible.\footnote{For example, by compiling a
+return-oriented-programming exploit from code fragments discovered on
+the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
+ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
+research community, this has been generalised into treating a program
+as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
+an input, analogous to a virtual machine operating on bytecode.
+Nail avoids these problems by having the parser precisely match the
+specified grammar, eliminating under-specified behavior.
 
@@
 parser, not as an output generator and does not support the expressive
-power of PEGs, but rather implements a C like structure model enhanced
-with data-dependent length fields and constraints.
+power of parsing expression grammars (PEGs), but rather implements a
+C-like structure model enhanced with data-dependent length fields and
+constraints.
 
@@
 While previous parser generators could also be used to write grammars
-for binary protocols\footnote{Theoretically speaking, the alphabet
+for binary protocols,\footnote{Theoretically speaking, the alphabet
 over which a grammar is an  abstract set, so most algorithms work
-just as well on an alphabet of $\{0,1\}$}, doing so is impractically
+just as well on an alphabet of $\{0,1\}$.} doing so is practically
 inconvenient. Hammer allows the programmer to specify a grammar in
@@
 masters thesis~\cite{ford2002packrat}, which also describes the default
-parsing algorithm used by Hammer} by composing them from one or multiple
+parsing algorithm used by Hammer.} by composing them from one or multiple
 sub-parsers.  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).
+\XXX[how does Nail relate to Hammer, in terms of what's new, what it
+solves, etc?]
 
@@
 Unfortunately,  parser frameworks are seldom used  to recognise
-machine-created input. For example, the security-critical and very
+machine-created input. For example, the security-critical and
 well-engineered MIT Kerberos distribution uses parser generators, but
@@
 all iterations~\cite{patterson-citation}.
-
---- XXX integrate the following paragraphs somewhere into the logical flow ---
 

### 2014-04-02 paper/design.tex
pass over design up to 3.2

@@
 
-\paragraph{Overview}
+\subsection{Overview}
 
-To alleviate these problems, Nail provides a richer set of combinators that not
-only describes the grammar of the external protocol, but also induce an internal
+To integrate the data format and the internal representation, Nail
+provides a rich set of combinators that not
+only describe the grammar of the external protocol, but also induce an internal
 model.
 
-A central design decision of Nail is that there is a semantic bijection between
-the model exposed to the programmer and the bytes input and output up to
+A central design decision of Nail is that there is a \emph{semantic bijection} between
+the model exposed to the programmer and the byte-level input and output, up to
 syntactic equivalence for unambiguous grammars. More precisely, the parser is
-the generators inverse, so parsing the generators output will yield the
-generators input, but generating the parsers output does not necessarily yield
+the generator's inverse, so parsing the generator's output will yield the
+generator's input, but generating the parsers output does not necessarily yield
 the parsers input. To understand why this makes sense, consider a grammar for a
-text language that tolerates white space or a binary protocol that tolerates
-arbitrarily long padding\footnote{Say, the physical layer of most communication
+text language that tolerates white space, or a binary protocol that tolerates
+arbitrarily long padding.\footnote{Say, the physical layer of most communication
   protocols is a possibly infinite sequence of symbols that are syntactically
   nil followed by a pre-determined synchronisation sequence and the actual
-  contents of the transmission.}. In that case, the program semantics should be
+  contents of the transmission.} In that case, the program semantics should be
 independent of the number of padding elements in the input, and Nail therefore
@@
 to allow a grammar-specific plug-in to make these choices, say for faster
-alignment, streaming applications when data is not ready or visual appearance in
+alignment, streaming applications when data is not ready, or visual appearance in
 human-readable protocols. Similarly, Nail does not preserve the layout of
@@
 
-Nail evaluates the combinators and produces type declarations for the internal
-model, the \textit{parser}, a function to parse a sequence of bytes into an
-instance of the model and the \textit{generator}, a function to create a
-sequence of bytes from an instance of the model. In the following, we present
-Nail's combinators.
+Nail evaluates the combinators and produces:
 
-\paragraph{Fundamental parsers}
-The elementary parsers of Nail are the same as those of Hammer, signed and
-unsigned integers with arbitrary lengths up to 64 bits. Note that is possible to
-define parsers for sub-byte lengths, e.g. to parse the 4-bit data offset within
-the TCP header. Integer parsers return their value in the
-smallest appropriately sized machine integer type, e.g. a 24 bit integer is stored in a 32-bit wide variable.\\
-Examples: \texttt{uint4}, \texttt{int32}
+\begin{itemize}
+\item type declarations for the internal model,
+\item the \textit{parser}, a function to parse a sequence of bytes into an
+instance of the model, and
+\item the \textit{generator}, a function to create a
+sequence of bytes from an instance of the model.
+\end{itemize}
 
-Integer parsers can be constrained to fall either within an (inclusive) range of
-values or be an
-element of a set of allowed values. \\
-Examples:\\
-\texttt{uint8 | 1..3} (8 bit unsigned integer $1\leq x \leq 3$) ,\\
-\texttt{uint16| ..512} (unsigned 16-bit integer $\leq 512$),\\
-\texttt{int32 | [1,255,512]} (signed 32 bit integer with value 1, 255 or 512).
-Invalid values or prematurely reaching the end of input raise an error when
-parsing input or generating output and abort the procedure.
+\noindent
+In the rest of this section, we present Nail's combinators.
 
-\paragraph{Sequence: Structure}
-Nails fundamental concept is the structure combinator. It contains a list of
-named parsers and unnamed constant parsers  . The parser invokes each field 
+\begin{figure}[tb]
+\begin{tabular}{ll}
+\toprule
+\bf Syntax & \bf Semantics \\
+\midrule
+\texttt{int32} & 32-bit signed integer \\
+\multirow{2}{*}{\texttt{uint4}}
+  & 4-bit unsigned integer, \\
+  & returned as an 8-bit value \\
+\multirow{2}{*}{\texttt{uint8 | 1..3}}
+  & 8-bit unsigned integer, \\
+  & $1\leq x \leq 3$ \\
+\multirow{2}{*}{\texttt{uint16 | ..512}}
+  & unsigned 16-bit integer, \\
+  & $x \leq 512$ \\
+\multirow{2}{*}{\texttt{int32 | [1,255,512]}}
+  & signed 32 bit integer, \\
+  & $x \in \{ 1, 255, 512 \}$ \\
+\bottomrule
+\end{tabular}
+\caption{Example Nail grammars for integer values.}
+\label{fig:range}
+\end{figure}
+
+\paragraph{Fundamental parsers.}
+The elementary parsers of Nail are the same as those of Hammer, signed
+and unsigned integers with arbitrary lengths up to 64 bits.  Note that is
+possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
+data offset within the TCP header; in Nail's syntax.  Integer parsers
+return their value in the smallest appropriately-sized machine integer
+type; e.g., a 24-bit integer is stored in a 32-bit wide variable.
+
+Integer parsers can be constrained to fall either within an (inclusive)
+range of values or be an element of a set of allowed values.
+Figure~\ref{fig:range} shows several examples.
+Invalid values or prematurely reaching the end of input will raise an
+error when parsing input or generating output, and abort the procedure.
+
+\paragraph{Sequence.}
+
+Nail's fundamental concept is the structure combinator. It contains a list of
+named parsers and unnamed constant parsers. The parser invokes each field 
 in sequence and returns a structure containing the result of each parser.
+For example, Figure~\ref{fig:dns-struct} shows the structure combinator
+from a part of the grammar for DNS packets, along with the data model
+corresponding to it.
 
-\begin{tabular}{ll}
-  \textbf{Nail} & \textbf{Data Model}\\
-  \begin{minipage}{1.2in}
+\begin{figure}[tb]
+
+\begin{minipage}{0.45\columnwidth}
 \begin{verbatim}
@@
 \end{minipage}
-  & 
-  \begin{minipage}{1.2in}
+~
+\begin{minipage}{0.45\columnwidth}
 \begin{verbatim}
-struct header{
+struct header {
   uint16_t id;
@@
 \end{minipage}
-  \\
-\end{tabular}
-\paragraph{Repetition: Many,SepBy}
-The many combinator takes a parser and applies it repeatedly until it fails,
-returning an array of the inner parsers results. SepBy additionally takes a
-constant parser, which it applies in between
-parsing two values,but not before parsing the first value or after parsing the last.\\
-Examples: \texttt{many uint8}\\ \texttt{sepBy uint8=',' (many uint8| '0'..'9')}
-(which recognises comma-separated lists of numbers).
 
-\paragraph{Semantic choice: choose}
-We extend PEG's ordered choice combinator with a tag for each choice. The parser
+\caption{Nail grammar (left) and data model (right) for a part of the
+grammar for DNS packets.  The \texttt{uint3 = 0} grammar represents
+3 bits of padding (filled with zeroes).}
+\label{fig:dns-struct}
+\end{figure}
+
+\paragraph{Repetition.}
+
+The \emph{many} combinator takes a parser and applies it repeatedly
+until it fails, returning an array of the inner parsers results. The
+\emph{sepBy} combinator
+additionally takes a constant parser, which it applies in between parsing
+two values, but not before parsing the first value or after parsing the
+last.
+For example, \texttt{many uint8} represents an array of 8-bit unsigned
+integers, and \texttt{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
+comma-separated lists of decimal numbers.
+
+\paragraph{Semantic choice.}
+
+We extend a parsing expression grammar's ordered choice combinator with
+a tag for each choice. The parser
 attempts to parse each option in the order they are specified in the field and
@@
 to the beginning of the choice combinator's input. Care must be taken that the
-choices do not overlap, because Nail always picks to the first successfully
+choices do not overlap, because Nail always picks the first successfully
 parsed choice. If two options overlap, generated output for the latter option is
@@
 formats are usually not ambiguous in this sense.
+Figure~\ref{fig:choice} demonstrates a simple choice combinator.
 
+\begin{figure}[tb]
 \begin{verbatim}
 choose {
-   A = uint8 | 1.. 8
+  A = uint8 | 1..8
   B = uint16 | ..256
@@
 \end{verbatim}
+\caption{A simple choice combinator that parses either an 8-bit unsigned
+integer with a value between 1 and 8 (option A), or a 16-bit unsigned
+integer with a value of at most 256 (option B).}
+\label{fig:choice}
+\end{figure}
 
 \subsection{Constant parsers}
+
 Nail also features \emph{constant parsers}, which are not affecting the internal

### 2014-04-02 paper/design.tex
done with pass over section 3 (design), another XXX note

@@
 
-Nail also features \emph{constant parsers}, which are not affecting the internal
+Nail also features \emph{constant parsers}, which do not affect the internal
 representation. Constant parsers can appear instead of structure fields and as
-separators in the SepBy combinator.
+separators in the sepBy combinator.
 
 The simplest constant parser is an integer or array of integers with fixed
-value, e.g. \texttt{uint8=0}, \texttt{many uint8=[1,2]}. A convenience notation
+value; for example, \texttt{uint8=0}, or \texttt{many uint8=[1,2]}. A convenience notation
 for strings is also supported: \texttt{many uint8 = "foo"}.
 
-\paragraph{Wrap combinator.} When implementing real protocols with Nail, we often found that many structures
-consist of many constant parsers and only one named field. This pattern is
+\paragraph{Wrap combinator.} When implementing real protocols with Nail, we often found that
+structures that consist of many constant parsers and only one named field. This pattern is
 common in binary protocols which use fixed headers to denote the type of data
@@
 combinator was a sequence of parsers, but the data model does not wrap the
-single value in another structure, making the application code more concise.\\
-Example: \\ \verb+<uint8='"';many int8|'a'..'z'; uint8='"'>+  parses a quoted
-lowercase word into an array - excluding the quotation marks.
+single value in another structure, making the application-visible representation
+(and thus application code) more concise.
+
+For example, \texttt{<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
+lowercase word into an array, excluding the quotation marks.
 
@@
 In some protocols, there might be many ways to represent the same constant field
-and there is no semantic difference between different syntaxes.  Nail therefore
-supports choice (\texttt{|| uint8 = 0 || uint16 = 0xFFFF}) and repetition
-(\texttt{many (uint8= ' ')}). This is particularly useful for handling padding or whitespace.
+and there is no semantic difference between the different syntactic representations.
+To support this pattern, Nail allows developers to use choice and repetition
+combinators together with constant fields, such as
+\texttt{many (uint8=' ')} (representing any number of space characters), or
+\texttt{|| uint8 = 0 || uint16 = 0xFFFF} (\XXX[what is this doing?  what is this
+syntax?  it looks notably different from Figure~\ref{fig:choice}]).
+This is particularly useful for handling padding or whitespace.
 
@@
 Another problem for parser generators is that binary protocols often contain
-length and offset fields. While conventional parsing algorithms can deal with
-bounded offset fields - a finite automaton can count a bounded integer, and we
+length and offset fields. Conventional parsing algorithms can, in principle,
+deal with
+bounded offset fields: a finite automaton can count a bounded integer, and we
 can feed the (finite) input multiple times to the finite automaton. However,
-this implementation is both time-inefficient - why feed many bytes into the
-automaton that will just be skipped - and very cumbersome to express with
-current parser generators. Therefore, if languages with offset fields are parsed
-with parser generators, the only currently feasible way is to add 'ad-hoc' hacks
-such as changing the input pointer of the generated parser on the fly from
-semantic actions.
+this implementation is both time-inefficient (it feeds many bytes into the
+automaton that will just be skipped) and very cumbersome to express with
+current parser generators. Therefore, if languages with offset fields need
+to be parsed
+with parser generators, the only currently practical way is to add ad-hoc hacks
+such as changing the input pointer of the generated parser on the fly, as part
+of the code in the semantic action for the offset field.
 
 Nail will properly support both offset and length fields and much of the
-following discussion applies to both , although the current prototype only
+following discussion applies to both, although the current prototype only
 implements lengths, which we will focus on.
@@
 value depends on some other structure in the data model. Dependency fields
-appear in a structure combinator as would any other integer field, but their
-name begins with an '@' sign. Dependency fields are not exposed in the data
-model, but transparently computed.
+appear in a structure combinator as would any other integer field, but
+their name begins with an \texttt{@} sign.  A dependency field has to
+appear in the grammar before it can be used.
 
-A dependency field has to appear in the grammar before it can be used. 
-\paragraph{Length fields: n_of}
-The length combinator takes a dependency field and a parser, evaluating the
-parser as often as the value of the dependency field, returning an array of the
-values. When generating output, it emits the array and writes its length to the
-dependency field.
-\paragraph{Offsets: offset} 
-The offset combinator takes a dependency field and a parser. It moves the parser
-to the position specified in the offset field and invokes the inner parser, and
-then moves the input back to its original position.
-While generating output, all structures referred to by offset are generated
-after the main structure and the dependent offset fields patched up.
+Dependency fields are not exposed in the data model, but instead are
+transparently computed.  This frees the developer from checking that
+these fields are correct (for input) or having to keep their values in
+sync with the rest of the data structure (for output).
 
-\paragraph{Checksums: raw_depend}
-In many data formats, some values depend on external representation, such as
-checksums and cryptographic signatures. While it would be possible to extend
-our constraint language to be powerful enough to support such constructs, we
-would essentially be building a separate, Turing complete language that has all
-the same pitfalls existing programs have. Therefore, we intend to  allow the programmer to
-very carefully escape Nail's programming model and write a function that takes a
-pointer to the dependent value and a range of bytes. 
-This feature is not implemented in the current prototype. \\
-Example:\\
- \verb+{ data many uint8; @checksum uint32+\\
- \verb+raw_depend @checksum data crc32 }+\\
- \verb+/*bool crc32(uint32_t *out, uint8_t *in);*/+
-Because this feature compromises Nail's security guarantees, it should only be
-used in limited circumstances and with carefully prepared checksum functions.
+\paragraph{Length fields.}
+
+The length combinator, \texttt{n_of}, takes a dependency field $n$
+and a parser, evaluates the parser exactly $n$ times (i.e., setting
+the number of iterations to be the $n$ field's value), and returns
+an array of the parser's values. When generating output, it emits
+the array and writes its length to the dependency field.
+
+\paragraph{Offsets.}
+
+The \texttt{offset} combinator takes a dependency field and a parser. It
+moves the parser to the position specified in the offset field and
+invokes the inner parser, and then moves the input back to its original
+position.  While generating output, all structures referred to by offset
+are generated after the main structure and the dependent offset fields
+are patched up.
+
+\paragraph{Checksums.}
+
+In many data formats, some values depend on external representation,
+such as checksums and cryptographic signatures.  While it would be
+possible to extend our constraint language to be powerful enough to
+support such constructs, we would essentially be building a separate,
+Turing-complete language that has all the same pitfalls existing programs
+have.  Therefore, we intend to allow the programmer to carefully escape
+Nail's programming model and write a function that takes a pointer to
+the dependent value and a range of bytes, using the \texttt{raw_depend}
+combinator.
+
+For example, we imagine the following grammar could be used to represent
+a sequence of bytes \texttt{data} followed by its CRC32 checksum:
+
+\begin{verbatim}
+data many uint8; @checksum uint32
+raw_depend @checksum data crc32
+\end{verbatim}
+
+\noindent
+where \texttt{crc32} is a function supplied by the application, with
+the following signature:
+
+\begin{verbatim}
+bool crc32(uint32_t *out, uint8_t *in);
+\end{verbatim}
+
+Because this feature compromises Nail's security guarantees, it should
+only be used in limited circumstances and with carefully prepared checksum
+functions.  This feature is not implemented in the current prototype.

### 2014-04-02 paper/concl.tex
minor nits in the intro
some relwk nits; one XXX regarding hammer vs nail
conclusion

--- 2014-04-01 paper/concl.tex
@@
 \section{Conclusion}
+\label{s:concl}
 
 This paper presented the initial design and implementation of
-\textit{Nail}, an XXX...
+\textit{Nail}, a parser generator that aims to minimize the amount
+of programmer effort.  Nail achieves this by reducing the expressive
+power of the grammar, maintaining a \emph{semantic bijection} between
+data formats and internal representations, and allowing programmers to
+specify \emph{structural dependencies} on the data format.  This avoids
+the need for programmers to write explicit semantic actions for input
+and output processing.  We hope that by reducing programmer effort
+required to use a parser generator, Nail will enable more application
+developers to use generated parsers in practice.  Our prototype source
+code is available at \url{https://github.com/jbangert/nail}.

### 2014-04-02 paper/impl.tex
Conclusion, eval, future work
conclusion

@@
 The current prototype of the Nail parser generator supports the C programming
-language and top down parsers. Options for C++ STL data models and emitting
-Packrat parser~\cite{packrat-parsing:icfp02} are under heavy development. In
+language and top-down parsers. Options for C++ STL data models and emitting
+Packrat parser~\cite{packrat-parsing:icfp02} are under development. In
 this section, we will discuss some particular features of our parser
@@
 The source code of our implementation, together with the examples described in
-\S\ref{s:eval} are available on GitHub\footnote{\url{https://github.com/jbangert/nail}}.
+\S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
+
 \paragraph{Parsing}

### 2014-04-02 paper/impl.tex
some small impl nits

@@
 
-\paragraph{Parsing}
-A generated Nail parser makes two passes through the input , the first to
-validate and recognise the input, the second to bind this data to the internal
-model. Currently the parser is a straightforward top down parser, although
+\paragraph{Parsing.}
+
+A generated Nail parser makes two passes through the input: the first to
+validate and recognise the input, and the second to bind this data to the internal
+model. Currently the parser is a straightforward top-down parser, although
 facilities have been made to add Packrat parsing to achieve linear parsing
 times.
-\paragraph{Memory allocation}
+
+\paragraph{Memory allocation.}
+
 Many security vulnerabilities can occur when heap allocations are done
 improperly. Therefore, just like Hammer, Nail avoids using the heap as much as
-possible, using a custom arena allocator and allocating only fixed size blocks
-from the system malloc. However, Nail extends upon Hammers approach and uses two
+possible, using a custom arena allocator and allocating only fixed-size blocks
+from the system allocator (malloc). However, Nail extends upon Hammer's approach and uses two
 arenas for each parsed input. One arena is used for intermediate results and is
-freed(and zeroed) after parsing completes, whereas the other arena is used only
-for allocating the result and has to be freed by the programmer.
+freed (and zeroed) after parsing completes, whereas the other arena is used only
+for allocating the result, and has to be freed by the programmer.
 
-\paragraph{Intermediate representation}
-Most parser generators, such as Bison do not have to dynamically allocate
+\paragraph{Intermediate representation.}
+
+Most parser generators, such as Bison, do not have to dynamically allocate
 temporary data, as they evaluate a semantic action on every rule. However, as
-our goal is to perform as little computation as possible without having
-validated the input, and we do not want to mix temporary objects with the
+our goal is to perform as little computation as possible before the input
+has been validated, and we do not want to mix temporary objects with the
 results of our parse, we use an append-only trace to store intermediate parser
@@
 because it stores a large tree node structure for each input byte and for each
-rule reduced. This allows Hammer semantic actions to get all necessary
+rule reduced. This allows Hammer semantic actions to get all of the necessary
 information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammers semantic
-actions, we can trust it as much as we trust the parser and thus expose it to
-the raw input stream.
+automatically generate our second pass, which corresponds to Hammer's semantic
+actions, we can trust it as much as we trust the parser, and thus can expose it
+to the raw input stream.
 
-Under this premise, the actions need very limited information from the
-recogniser to correctly handle the input stream. In particular, the parsers
-control flow only branches at the choice, repetition and constant
-combinators.
+Under this premise, the actions need limited information from the
+recogniser to correctly handle the input stream. In particular, the parser's
+control flow branches only at the choice, repetition, and constant combinators.
 Thus, for each of those combinators, we store the minimum amount of information
@@
 the input, the parser does not backtrack in the trace. This means that offsets
-within the trace can be used for a Packrat hash-table to memoise backtrack-heavy
+within the trace can be used for a Packrat hash-table to memoize backtrack-heavy
 parsers.
@@
 When encountering a repetition combinator, the parser records the number of
-times the inner parser succeeded and when encountering a constant parser of
+times the inner parser succeeded, and when encountering a constant parser of
 variable size, it records how much input was consumed by the constant parser.
@@
 the trace to determine how many array fields to parse and which choices to pick.
-\paragraph{Dependency Fields}
+
+\paragraph{Dependency fields.}
+
 During parsing, dependency fields occur before the context in which they are
-used. The parser stores their value and retrieves it afterwards when
+used. The parser stores their values and retrieves them afterwards when
 encountering the combinator that uses them. When generating output, the
@@
 that the dependency field is correct.
-\paragraph{Bootstrapping}
-To demonstrate the feasibility of the nail parser generator, our parser
+
+\paragraph{Bootstrapping.}
+
+To demonstrate the feasibility of the Nail parser generator, our parser
 generator uses a Nail parser to recognise Nail grammars. A superset of the
 grammar language described in this paper is implemented in 100 lines of Nail,
-which feed into about a thousand lines of C++ that implement the actual parser
+which feed into about 1,000 lines of C++ that implement the actual parser
 generator. Bootstrapping is supported via an implementation of the Nail language

### 2014-04-02 paper/eval.tex
future work pass

--- 2014-03-31 paper/eval.tex
@@
 \section{Evaluation}
-In order to test the Nail framework, we cloned the test DNS server from the Hammer distribution to
-Nail. Hammer ships with a toy DNS server written in 683 lines of code, excluding the hammer
-framework itself that responds to any valid DNS query with a CNAME record to the domain
-``spargelze.it''.  Most of this code is taken up with custom validators, semantic actions and data
-structure definitions.
+\label{s:eval}
 
-Our DNS server measures 148 lines of C and 48 lines of Nail grammar, and supports a custom zone file format
-with A,NS,MX and CNAME records. The same grammar is used together with 98 lines of C to implement a
-functional toy clone of the host command line tool. However, because our grammar does not yet
-support DNS label compression, the latter tool will occasionally reject valid real world DNS
-responses. Both clients have functional anti-spoofing measures.
+\begin{figure*}
+\begin{minipage}{0.45\paperwidth}
+\begin{verbatim}
+utfstring = many choose {
+  SUPP = {
+    lead  uint16 |  0xD800..0xDBFF
+    trail uint16 |  0xDC00..0xDFFF
+  }
+  BASIC = uint16 | !0xD800..0xDFFF
+}
+\end{verbatim}
+\end{minipage}
+~
+\begin{minipage}{0.45\paperwidth}
+\begin{verbatim}
+struct utfstring {
+  struct {
+    enum { SUPP, BASIC } N_type;
+    union {
+      struct {
+        uint16_t lead;
+        uint16_t trail;
+      } SUPP;
+      uint16_t BASIC;
+    };
+  } *elem;
+  size_t count;
+};
+\end{verbatim} 
+\end{minipage}
+\caption{Nail grammar (left) and data model (right) for UTF-16 strings.}
+\label{fig:utf16}
+\end{figure*}
 
-It is hard to compare our efforts to a real world DNS server, as we have less functionality, in
-particular for DNS compression and additional hint records than real world DNS servers. 
-However, the closest in functionality and intent is Dan Bernstein's djbdns\footnote{\url{http://cr.yp.to/djbdns.html}}, which aims
-to be a minimalist, highly secure DNS server. The latest release of DJBDNS, however, including
-various support tools is about ten thousand lines of C as measured by SLOCCOUNT. We are confident
-that it is possible to a feature-par version with Nail that is an order of magnitude smaller and
-intend to do so.
+\begin{figure}
+\begin{verbatim}
+labels = <many { @length  uint8 | 1..255 
+                 label n_of @length uint8 }
+          uint8 = 0>
+
+question = {
+  labels labels
+  qtype uint16 | 1..16 
+  qclass uint16 | [1,255]
+}
+
+answer = {
+  labels labels
+  rtype uint16 | 1..16
+  class uint16 | [1]
+  ttl uint32
+  @rlength uint16 
+  rdata n_of @rlength uint8 
+}
+
+dnspacket = {
+  id uint16
+  qr uint1
+  opcode uint4
+  aa uint1 
+  tc uint1
+  rd uint1
+  ra uint1
+  uint3 = 0
+  rcode uint4
+  @qc uint16
+  @ac uint16
+  uint16 = 0    // authority 
+  uint16 = 0    // additional
+  // We don't support authority or 
+  // additional sections in the prototype
+  questions n_of @qc question
+  responses n_of @ac answer
+}
+\end{verbatim}
+\caption{Nail grammar for DNS packets, used by our prototype DNS server.}
+\label{fig:dns-full}
+\end{figure}
+
+In our preliminary evaluation of Nail, we try to answer two questions:
+
+\begin{itemize}
+\item Can Nail grammars support real-world data formats?
+\item How much programmer effort is required to build an
+      application that uses Nail for data input and output?
+\end{itemize}
+
+\paragraph{Data formats.}
+
+To answer the first question, we implemented two Nail grammars:
+one for UTF-16 encoded strings, exposing an array of code points
+(shown in Figure~\ref{fig:utf16}), and another for a subset of DNS
+packets sent to and from an authoritative name server, without label
+compression, as per RFC1035 (shown in Figure~\ref{fig:dns-full}).
+Furthermore, our GitHub repository contains other Nail
+grammars, such as a TAP network stack that processes Ethernet,
+ARP, ICMP, IP, and UDP packets, and the grammar for Nail
+itself.\footnote{\url{https://github.com/jbangert/nail/tree/master/examples}}
+The results suggest that Nail is a good fit for these data formats.
+
+\paragraph{Programmer effort.}
+
+To answer the second question, we implemented a functioning toy DNS
+server.  In particular, we cloned the test DNS server from the Hammer
+distribution to Nail.  Hammer ships with a toy DNS server written in 683
+lines of code, excluding the Hammer framework itself, that responds to
+any valid DNS query with a CNAME record to the domain ``spargelze.it''.
+Most of this code is taken up with custom validators, semantic actions,
+and data structure definitions, with only 52 lines of code defining the
+grammar with Hammer's combinators.
+
+Our DNS server consists of 148 lines of C code and 48 lines of Nail
+grammar, and supports a custom zone file format with A, NS, MX, and
+CNAME records. The same grammar is used, together with 98 lines of C,
+to implement a functional toy clone of the \texttt{host} command-line
+tool. However, because our grammar does not yet support DNS label
+compression, the latter tool will occasionally reject valid real-world
+DNS responses. Both clients have functional anti-spoofing measures.
+
+It is hard to compare the programming effort required to implement
+our toy DNS server to that of a real world DNS server, since
+we have less functionality, in particular for DNS compression
+and additional hint records that real-world DNS servers send.
+However, the closest in functionality and intent is Dan Bernstein's
+djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
+a minimalist, highly secure DNS server. The latest release of djbdns,
+including various support tools, is about 10,000 lines of C as measured by
+\texttt{sloccount}.  We expect that it is possible to build a feature-par
+version with Nail that is an order of magnitude smaller and intend to
+do so.
+
+\paragraph{Other issues.}
+
+As Nail is work in progress, many parts of the implementation, syntax and
+design are not complete yet and we do not yet have meaningful performance
+or security metrics.

### 2014-04-02 paper/impl.tex
make spell

@@
 A generated Nail parser makes two passes through the input: the first to
-validate and recognise the input, and the second to bind this data to the internal
+validate and recognize the input, and the second to bind this data to the internal
 model. Currently the parser is a straightforward top-down parser, although
@@
 Under this premise, the actions need limited information from the
-recogniser to correctly handle the input stream. In particular, the parser's
+recognizer to correctly handle the input stream. In particular, the parser's
 control flow branches only at the choice, repetition, and constant combinators.
@@
 To demonstrate the feasibility of the Nail parser generator, our parser
-generator uses a Nail parser to recognise Nail grammars. A superset of the
+generator uses a Nail parser to recognize Nail grammars. A superset of the
 grammar language described in this paper is implemented in 100 lines of Nail,

### 2014-04-02 paper/relwk.tex
make spell

@@
 A subclass of logic errors are so-called \textit{weird machines}, where
-implementation side effects or under-specified parser behaviour leads to a
+implementation side effects or under-specified parser behavior leads to a
 protocol or data format inadvertently becoming a Turing-complete execution
@@
 ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
-research community, this has been generalised into treating a program
+research community, this has been generalized into treating a program
 as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
@@
 implements grammars as language-integrated parser combinators, an approach
-popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
+popularized by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
 parser combinator style (to our knowledge, first  described
@@
 
-Unfortunately,  parser frameworks are seldom used  to recognise
+Unfortunately,  parser frameworks are seldom used  to recognize
 machine-created input. For example, the security-critical and
@@
 language. Mongrel was re-written from scratch multiple times to achieve
-better scaleability and design, yet the grammar could be  re-used across
+better scalability and design, yet the grammar could be  re-used across
 all iterations~\cite{patterson-citation}.

### 2014-04-02 paper/bin/abbrv.pl

@@ -1,0 +1,11 @@
+#!/usr/bin/perl
+# Detects potential abbreviations at the end of a sentence.
+
+while (<>) {
+    s/%.*//;
+    next unless s/([A-Z]+\.)/\e[7m$1\e[m/g;
+
+    s/^([^\e]*\n)+//mg;
+    s/^/$ARGV: /mg;
+    print;
+}

### 2014-04-02 paper/eval.tex
figure 4 spacing
spurious space
improve figure placement

@@
 
-\begin{figure*}
-\begin{minipage}{0.45\paperwidth}
-\begin{verbatim}
-utfstring = many choose {
-  SUPP = {
-    lead  uint16 |  0xD800..0xDBFF
-    trail uint16 |  0xDC00..0xDFFF
-  }
-  BASIC = uint16 | !0xD800..0xDFFF
-}
-\end{verbatim}
-\end{minipage}
-~
-\begin{minipage}{0.45\paperwidth}
-\begin{verbatim}
-struct utfstring {
-  struct {
-    enum { SUPP, BASIC } N_type;
-    union {
-      struct {
-        uint16_t lead;
-        uint16_t trail;
-      } SUPP;
-      uint16_t BASIC;
-    };
-  } *elem;
-  size_t count;
-};
-\end{verbatim} 
-\end{minipage}
-\caption{Nail grammar (left) and data model (right) for UTF-16 strings.}
-\label{fig:utf16}
-\end{figure*}
-
-\begin{figure}
-\begin{verbatim}
-labels = <many { @length  uint8 | 1..255 
-                 label n_of @length uint8 }
-          uint8 = 0>
-
-question = {
-  labels labels
-  qtype uint16 | 1..16 
-  qclass uint16 | [1,255]
-}
-
-answer = {
-  labels labels
-  rtype uint16 | 1..16
-  class uint16 | [1]
-  ttl uint32
-  @rlength uint16 
-  rdata n_of @rlength uint8 
-}
-
-dnspacket = {
-  id uint16
-  qr uint1
-  opcode uint4
-  aa uint1 
-  tc uint1
-  rd uint1
-  ra uint1
-  uint3 = 0
-  rcode uint4
-  @qc uint16
-  @ac uint16
-  uint16 = 0    // authority 
-  uint16 = 0    // additional
-  // We don't support authority or 
-  // additional sections in the prototype
-  questions n_of @qc question
-  responses n_of @ac answer
-}
-\end{verbatim}
-\caption{Nail grammar for DNS packets, used by our prototype DNS server.}
-\label{fig:dns-full}
-\end{figure}
+\input{fig-dnsfull}
 

### 2014-04-03 paper/design.tex

@@
   protocols is a possibly infinite sequence of symbols that are syntactically
-  nil followed by a pre-determined synchronisation sequence and the actual
+  nil followed by a pre-determined synchronization sequence and the actual
   contents of the transmission.} In that case, the program semantics should be
@@
 data offset within the TCP header; in Nail's syntax.  Integer parsers
-return their value in the smallest appropriately-sized machine integer
+return their value in the smallest appropriately sized machine integer
 type; e.g., a 24-bit integer is stored in a 32-bit wide variable.

### 2014-04-03 paper/impl.tex
improve figure placement

@@
 \S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
+
+\input{fig-utf16}
 
@@
 arenas for each parsed input. One arena is used for intermediate results and is
-freed (and zeroed) after parsing completes, whereas the other arena is used only
+released (and zeroed) after parsing completes, whereas the other arena is used only
 for allocating the result, and has to be freed by the programmer.

### 2014-04-03 paper/intro.tex

@@
 a precise grammar for the input data format, and to use a parser
-generator, such as {\tt lex} and {\tt yacc}, to synthesise the input
+generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
 processing code.  Developers that use a parser generator do not need

### 2014-04-03 paper/relwk.tex
Fixed XXX

@@
 can be invoked to parse a given input into an abstract syntax tree (AST).
-\XXX[how does Nail relate to Hammer, in terms of what's new, what it
-solves, etc?]
+
+Nail improves upon Hammer in three ways. First, Nail  generates output besides recognising input.
+Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
+Nail's structural dependencies allow it to work with protocols Hammer cannot recognise, such as
+protocols with offset fields or length fields (Hammer has extremely limited support for length
+fields. It can parse arrays immediately preceeded by their length).
 

### 2014-04-03 paper/relwk.tex
minor

@@
 protocols with offset fields or length fields (Hammer has extremely limited support for length
-fields. It can parse arrays immediately preceeded by their length).
+fields: it can parse arrays immediately preceeded by their length).
 

### 2014-04-09 paper/langsec14/relwk.tex
move paper into langsec14

--- 2014-04-02 paper/relwk.tex
@@
 can be invoked to parse a given input into an abstract syntax tree (AST).
-\XXX[how does Nail relate to Hammer, in terms of what's new, what it
-solves, etc?]
+
+Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
+Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
+Nail's structural dependencies allow it to work with protocols Hammer cannot recognize, such as
+protocols with offset fields or length fields (Hammer has extremely limited support for length
+fields: it can parse arrays immediately preceded by their length).
 

### 2014-04-09 paper/osdi14/concl.tex
copy langsec paper to osdi14

\section{Conclusion}
\label{s:concl}

This paper presented the initial design and implementation of
\textit{Nail}, an interface generator for data formats.  Nail helps
programmers to avoid memory corruption and ambiguity vulnerabilities
while reducing effort in parsing and generating real-world protocols
and file formats.  Nail achieves this by reducing the expressive
power of the grammar, maintaining a \emph{semantic bijection} between
data formats and internal representations, and allowing programmers to
specify \emph{structural dependencies} in the data format.  Preliminary
experience with implementing a DNS server using Nail suggests that this
is a promising approach.

The source code for our prototype of Nail is available at
\url{https://github.com/jbangert/nail}.

### 2014-04-09 paper/osdi14/bin/get-tex-files.sh

@@ -1,0 +1,24 @@
+#!/bin/sh
+
+FLIST="$1"
+
+while :;
+do
+    LAST="$FLIST"
+    for F in $LAST; do
+	NL=`cat $F | grep \\input{ | grep -v ^% | cut -d{ -f2 | cut -d} -f1`
+	for F2 in $NL; do
+            if [ -e "$F2" ]; then
+	        FLIST="$FLIST $F2"
+            else
+	        FLIST="$FLIST $F2.tex"
+            fi
+	done
+    done
+    FLIST=`ls $FLIST 2>/dev/null | sort | uniq`
+
+    if [ "$LAST" = "$FLIST" ]; then
+	echo $LAST
+	exit 0
+    fi
+done

### 2014-04-17 generator/mk_gen.cc
compiles except for templates

@@ -95,7 +95,7 @@ class GenGenerator{
       FOREACH(field,p.structure){
         if(field->N_type != DEPENDENCY)
           continue;
-        out<< "NO dependencies in generator yet!\n";
+        out<< "//NO dependencies in generator yet!\n";
       }
       out << "out->index = end_of_struct.index;\n";
       out << "out->bit_offset = end_of_struct.bit_offset;\n}";
@@ -151,7 +151,7 @@ class GenGenerator{
       break;
     case APPLY: 
       {
-        assert(!"Implemented");
+        //assert(!"Implemented");
       }
       break;
     case ARRAY:
@@ -232,14 +232,15 @@ public:
       if(definition->N_type == CONSTANTDEF){
         std::string name = mk_str(definition->constantdef.name);
         out<<"void gen_"<<name<<"(HBitWriter* out){\n";
-        generator(definition->constantdef.definition);
+        //        generator(definition->constantdef.definition);
         out << "}";
       }
       else if(definition->N_type==PARSER){
         std::string name = mk_str(definition->parser.name);
         out << "void gen_" << (name)<<"(HBitWriter *out,"<< name << " * val){";
         ValExpr outval("val",NULL,1);
-        generator(definition->parser.definition.pr,outval);
+        
+        //        generator(definition->parser.definition.pr,outval);
         out << "}";
       }          
     }

### 2014-04-17 generator/mk_parser.cc

@@ -312,10 +312,11 @@ public:
     case APPLY:
       {
         out << "{/*Apply*/ "
-            <<"NailStream original_stream = str_current;\n #error \" not implemented \"";
-        out << "str_current = n_tr_read_stream(n_trace);";
+            <<"NailStream *original_stream = stream;\n;";
+        out << "stream = (NailStream *)tr;";
+        out << "tr+= sizeof(NailStream *) / sizeof(*tr);";
         action(p.apply.inner->pr, lval);
-        out << "str_current = original_stream;";
+        out << "stream = original_stream;";
         out <<"}\n";
       }
       break;
@@ -422,21 +423,26 @@ public:
         action(def->parser.definition.pr,outexpr);
         out << "*trace = tr;";
         out<< "return 0;}";
-        out << name << "*parse_" << name << "(NailArena *arena, const uint8_t *data, size_t size){\n"
-            << "NailStream stream = {.data = data, .pos= 0, .size = size*8};\n"
-            <<"n_trace trace;\n"
-            <<"pos *tr_ptr;\n pos pos;\n"
-            << name <<"* retval;\n"
-            << "n_trace_init(&trace,4096,4096);\n"
-            << "tr_ptr = trace.trace;"
-            << "if(parser_fail(peg_"<<name<<"(&trace,&stream))) return NULL;"
-            << "retval =  (typeof(retval))n_malloc(arena,sizeof(*retval));\n"
-            <<"if(!retval) return NULL;\n"
-            << "stream.pos = 0;\n"
-            << "if(bind_"<<name<<"(arena,retval,&stream,&tr_ptr,trace.trace) < 0) return NULL;\n"
-            << "n_trace_release(&trace);\n"
-            <<"return retval;"
-            <<"}";
+        if(!def->parser.parameters || def->parser.parameters->count==0){
+          out << name << "*parse_" << name << "(NailArena *arena, const uint8_t *data, size_t size){\n"
+              << "NailStream stream = {.data = data, .pos= 0, .size = size*8};\n"
+              << "NailArena tmp_arena;"
+              << "NailArena_init(&tmp_arena, 4096);"
+              <<"n_trace trace;\n"
+              <<"pos *tr_ptr;\n pos pos;\n"
+              << name <<"* retval;\n"
+              << "n_trace_init(&trace,4096,4096);\n"
+              << "tr_ptr = trace.trace;"
[…]

### 2014-04-17 generator/new_grammar.nail.h

@@ -1,7 +1,7 @@
 #include <stdint.h>
-#include <string.h>
+ #include <string.h>
 #include <assert.h>
-enum N_types {_NAIL_NULL,CONSTANTDEF,PARSER,PR,PAREN,APPLY,NAME,REF,NUNION,OPTIONAL,LENGTH,FIXEDARRAY,ARRAY,CHOICE,WRAP,STRUCTURE,INTEGER,DDEPENDENCY,DSTREAM,PSTREAM,PDEPENDENCY,SEPBY,SEPBYONE,MANY,MANYONE,TRANSFORM,FIELD,DEPENDENCY,CONSTANT,NEGATE,SET,SINGLE,VALUE,RANGE,CUNION,CSTRUCT,CREF,CINT,CREPEAT,CARRAY,VALUES,STRING,SIGN,UNSIGN,DIRECT,ESCAPE,NUMBER,HEX,ASCII};
+enum N_types {_NAIL_NULL,CONSTANTDEF,PARSER,PR,PAREN,NAME,REF,NUNION,OPTIONAL,APPLY,LENGTH,FIXEDARRAY,ARRAY,CHOICE,WRAP,STRUCTURE,INTEGER,DDEPENDENCY,DSTREAM,PSTREAM,PDEPENDENCY,SEPBY,SEPBYONE,MANY,MANYONE,TRANSFORM,FIELD,DEPENDENCY,CONSTANT,NEGATE,SET,SINGLE,VALUE,RANGE,CUNION,CSTRUCT,CREF,CINT,CREPEAT,CARRAY,VALUES,STRING,SIGN,UNSIGN,DIRECT,ESCAPE,NUMBER,HEX,ASCII};
 typedef struct number number;
 typedef struct varidentifier varidentifier;
 typedef struct constidentifier constidentifier;
@@ -31,299 +31,289 @@ typedef struct parserinner parserinner;
 typedef struct parser parser;
 typedef struct definition definition;
 typedef struct grammar grammar;
-struct number {
-    uint8_t*elem;
-    size_t count;
+struct number{
+uint8_t*elem;
+ size_t count;
 };
-struct varidentifier {
-    uint8_t*elem;
-    size_t count;
+struct varidentifier{
+uint8_t*elem;
+ size_t count;
 };
-struct constidentifier {
-    uint8_t*elem;
-    size_t count;
+struct constidentifier{
+uint8_t*elem;
+ size_t count;
 };
-struct streamidentifier {
[…]

### 2014-04-17 generator/parser_template.c

@@ -139,6 +139,14 @@ static pos n_tr_begin_choice(n_trace *trace){
         trace->iter+= 2;
         return trace->iter - 2;
 }
+static int n_tr_stream(n_trace *trace, const NailStream *stream){
+        assert(sizeof(stream) % sizeof(pos) == 0);
+        if(n_trace_grow(trace,sizeof(stream)/sizeof(pos)))
+                return -1;
+        *(const NailStream **)(trace->trace + trace->iter) = stream;
+        trace->iter += sizeof(stream)/sizeof(pos);
+        return 0;
+}
 static pos n_tr_memo_choice(n_trace *trace){
         return trace->iter;
 }

### 2014-04-17 generator/expr.hpp
Compiles without warnings. Self-hosting

@@ -5,6 +5,7 @@
 #include <string>
 struct Expr{
 public:
+  virtual ~Expr(){}
   virtual int is_ptr() const = 0;
   //  virtual void make_ptr() = 0;
   virtual void print_val(std::ostream &str) const = 0;

### 2014-04-17 generator/new_grammar.nail.h

@@ -1,5 +1,5 @@
 #include <stdint.h>
- #include <string.h>
+#include <string.h>
 #include <assert.h>
 enum N_types {_NAIL_NULL,CONSTANTDEF,PARSER,PR,PAREN,NAME,REF,NUNION,OPTIONAL,APPLY,LENGTH,FIXEDARRAY,ARRAY,CHOICE,WRAP,STRUCTURE,INTEGER,DDEPENDENCY,DSTREAM,PSTREAM,PDEPENDENCY,SEPBY,SEPBYONE,MANY,MANYONE,TRANSFORM,FIELD,DEPENDENCY,CONSTANT,NEGATE,SET,SINGLE,VALUE,RANGE,CUNION,CSTRUCT,CREF,CINT,CREPEAT,CARRAY,VALUES,STRING,SIGN,UNSIGN,DIRECT,ESCAPE,NUMBER,HEX,ASCII};
 typedef struct number number;
@@ -31,289 +31,299 @@ typedef struct parserinner parserinner;
 typedef struct parser parser;
 typedef struct definition definition;
 typedef struct grammar grammar;
-struct number{
-uint8_t*elem;
- size_t count;
+struct number {
+    uint8_t*elem;
+    size_t count;
 };
-struct varidentifier{
-uint8_t*elem;
- size_t count;
+struct varidentifier {
+    uint8_t*elem;
+    size_t count;
 };
-struct constidentifier{
-uint8_t*elem;
- size_t count;
+struct constidentifier {
+    uint8_t*elem;
+    size_t count;
 };
-struct streamidentifier{
-uint8_t*elem;
- size_t count;
+struct streamidentifier {
+    uint8_t*elem;
+    size_t count;
 };
-struct dependencyidentifier{
-uint8_t*elem;
- size_t count;
+struct dependencyidentifier {
+    uint8_t*elem;
+    size_t count;
 };
-struct intconstant{
-enum N_types N_type;
-union {
-struct {
-enum N_types N_type;
-union {
-uint8_t escape;
-uint8_t direct;
+struct intconstant {
+    enum N_types N_type;
+    union {
+        struct {
+            enum N_types N_type;
+            union {
+                uint8_t escape;
+                uint8_t direct;
+            };
+        } ascii;
+        struct {
+            uint8_t*elem;
+            size_t count;
+        } hex;
+        number number;
+    };
 };
-} ascii;
-struct {
-uint8_t*elem;
- size_t count;
-} hex;
-number number;
+struct intp {
[…]

### 2014-04-17 generator/parser_template.c

@@ -89,6 +89,13 @@ static void n_trace_release(n_trace *out){
         out->iter = 0;
         out->grow = 0;
 }
+static pos n_trace_getpos(n_trace *tr){
+        return tr->iter;
+}
+static void n_tr_setpos(n_trace *tr,pos offset){
+        assert(offset<tr->capacity);
+        tr->iter = offset;
+}
 static int n_trace_grow(n_trace *out, int space){
         if(out->capacity - space>= out->iter){
                 return 0;

### 2014-04-17 generator/mk_parser.cc
Can parse a simple offset

@@ -314,7 +314,10 @@ public:
         out << "{/*Apply*/ "
             <<"NailStream *original_stream = stream;\n;";
         out << "stream = (NailStream *)tr;";
-        out << "tr+= sizeof(NailStream *) / sizeof(*tr);";
+#ifdef DEBUG_OUT
+        out << "fprintf(stderr,\"%d = stream \\n\",tr-trace_begin, stream);\n";
+#endif
+        out << "tr+= sizeof(NailStream) / sizeof(*tr);";
         action(p.apply.inner->pr, lval);
         out << "stream = original_stream;";
         out <<"}\n";
@@ -739,6 +742,7 @@ class CPrimitiveParser{
               out << "NailStream str_" << mk_str(*stream) <<";\n";
               // str_current cannot appear on the left
             }
+            out << "if(";
             out << mk_str(field->transform.cfunction) << "_parse(tmp_arena"; //TODO: use temp arena
 
             FOREACH(stream, field->transform.left){           
@@ -761,7 +765,7 @@ class CPrimitiveParser{
                 break;
               }
             }
-            out << ");\n";
+            out << ")) {" << fail << "}";
             header << ");\n";
             FOREACH(stream, field->transform.left){
               newscope.add_stream_definition(mk_str(*stream));
@@ -848,6 +852,7 @@ class CPrimitiveParser{
           << "str_current = orig_str;\n";
       out << "}";        
     }
+      break;
     case OPTIONAL:
       peg_optional(*parser.pr.optional,fail,scope);
       break;

### 2014-04-17 generator/parser_template.c

@@ -148,10 +148,11 @@ static pos n_tr_begin_choice(n_trace *trace){
 }
 static int n_tr_stream(n_trace *trace, const NailStream *stream){
         assert(sizeof(stream) % sizeof(pos) == 0);
-        if(n_trace_grow(trace,sizeof(stream)/sizeof(pos)))
+        if(n_trace_grow(trace,sizeof(*stream)/sizeof(pos)))
                 return -1;
-        *(const NailStream **)(trace->trace + trace->iter) = stream;
-        trace->iter += sizeof(stream)/sizeof(pos);
+        *(NailStream *)(trace->trace + trace->iter) = *stream;
+        fprintf(stderr,"%d = stream\n",trace->iter,stream);
+        trace->iter += sizeof(*stream)/sizeof(pos);  
         return 0;
 }
 static pos n_tr_memo_choice(n_trace *trace){

### 2014-04-17 tests/harness.sh

@@ -1,15 +1,21 @@
-#!/bin/sh
+#!/bin/sh 
 cd $(dirname $1)
 TEST=$(basename $1)
 
 echo "Running $TEST"
-NAIL=`../../generator/nail $TEST   2>&1`  || ( echo "Failed generating $case \n"  $NAIL; exit -1) || exit -1 
+NAIL=`../../generator/nail $TEST   2>&1`  || ( echo "Failed generating $case \n"  $NAIL; exit -1) || echo $NAIL
+CFILE="${TEST%.*}"
+if [ -e $CFILE.c ]; then
+    echo "Adding transform functions from $CFILE"
+    cat  $CFILE.c >> $TEST.c
+fi
 cat ../test_harness.c >> $TEST.c
 astyle $TEST.c > /dev/null
+
 rm $TEST.c.orig
 COMPILER=`gcc  -ggdb -DXYCONST=0 -lhammer -std=gnu99 -o $TEST-test $TEST.c  2>&1` || (echo "Failed compiling $case \n $COMPILER"; exit -1 ) || exit -1 
     
-for case in in.*
+for case in $CFILE.in.*
  do 
    OUT=`./$TEST-test < $case 2>&1` || ( echo "Failed  testcase $case \n" $OUT; exit -1) || exit -1 
  done

### 2014-04-17 tests/test_harness.c

@@ -1,6 +1,7 @@
 #line 1 "test_harness.c"
 #include <stdio.h>
 #include <stdint.h>
+#include <hammer/internal.h>
 extern HAllocator system_allocator;
 #define TOKENPASTE(x,y) x ## y
 #define CAT(x,y) TOKENPASTE(x,y)

### 2014-04-17 tests/transform/make_protozip.rb

@@ -1,0 +1,19 @@
+#!/usr/bin/env ruby
+offset = 0
+out = ""
+
+header = [ARGV.size].pack("S>")
+ARGV.each do |file|
+  contents = File.read(file)
+  out+= contents
+  filesiz = contents.bytesize
+  header += [offset, filesiz].pack("L>L>") + file+ [0].pack("C")
+  offset = out.bytesize()
+end
+while header.bytesize()<1024
+  header += [0].pack("C")
+end
+
+print out
+print header
+

### 2014-04-17 tests/transform/trans1.c

@@ -1,0 +1,27 @@
+//Note that all nail offsets are in bits
+int offsetsize_parse(NailArena *tmp,NailStream *str_filestream,NailStream *contentstream,uint32_t* offset,uint32_t* filesize){
+        if(*offset > contentstream->size / 8)
+                return -1;
+        if( *filesize * 8 > contentstream->size  - *offset *8)
+                return -1;
+        str_filestream->data = contentstream->data + *offset;
+        str_filestream->pos  = 0;
+        str_filestream->size = *filesize * 8;
+        return 0;
+}
+
+int splitzip_parse(NailArena *tmp,NailStream *str_header,NailStream *str_content,NailStream *current){
+        const int headersize = 1024*8;
+        if(current->size <= headersize)
+                return -1;
+        if(current->pos != 0)
+                return -1;
+        str_header->data = current->data + (current->size - headersize)/8;
+        str_header->pos = 0;
+        str_header->size = headersize;
+        str_content->data = current->data;
+        str_content->pos = 0;
+        str_content->size = current->size - headersize;
+        current->pos = current->size; //We consumed all of current. 
+        return 0;
+}

### 2014-04-17 paper/osdi14/future.tex

\section{Future work}
\label{s:future}

The Nail parser generator is currently a work-in-progress, and we would
appreciate feedback on our initial design and prototype implementation.
Short-term next steps include improving the scoping of dependency fields
and adding support for offset fields.

One problem with the current design of Nail is that the design of the
grammar dictates the internal structure of the software.  This makes
changing grammars or adding Nail to existing software awkward.
One possible solution to this problem would be to implement a concept
similar to relational lenses~\cite{bohannon2006relational}, which would
allow the data model to be ``seen'' by the rest of the program through
an isomorphism.  Such an isomorphism would still be much more concise
than two sets of semantic actions, while allowing changes in syntax,
alternative representations, and adaption to legacy systems.

Finally, we would like to demonstrate the capabilities of Nail by
implementing various binary formats ``notorious'' for their insecurity
in Nail. Nail was designed with the idioms of formats such as PDF and
PNG in mind. Ultimately, we want to provide examples of successful Nail
parsers throughout a network stack, from a user-space TCP stack to a
PNG de-compressor.

### 2014-04-17 tests/test_harness.c
Modified test harness to show partial success

@@ -24,6 +24,7 @@ int main(){
     XYZZY *object; 
     object = CAT(parse_,XYZZY)(&arena,input,inputsize);
     if(object){
+            printf("\n Successfully parsed! Trying to generate\n");
             const char *buf;
             size_t siz;
             HBitWriter *bw;

### 2014-04-22 examples/zip/extract.c
fixed extract.c

@@ -3,6 +3,8 @@
 #include <sys/mman.h>
 #include <unistd.h>
 #include <fcntl.h>
+#include <string.h>
+#include <errno.h>
 #include <stdio.h>
 #include "zip.nail.h"
 char *mmap_file(char *filename, size_t *size){
@@ -10,8 +12,8 @@ char *mmap_file(char *filename, size_t *size){
         int fil  = open(filename, O_RDONLY);
         if(fil < 0 ) return NULL;
         if(fstat(fil,&stat) < 0) return NULL;
-        char *retval = mmap(NULL,stat.st_size,PROT_READ, 0, fil, 0);
-        if(!retval) return NULL;
+        char *retval = mmap(NULL,stat.st_size,PROT_READ, MAP_SHARED, fil, 0);
+        if(retval == MAP_FAILED) return NULL;
         *size = stat.st_size;
         close(fil);
         return retval;                
@@ -22,11 +24,11 @@ int main(int argc,char **argv){
                 zip_file *zip;
                 size_t size;
                 char *filedata = mmap_file(argv[f],&size);
-                if(!filedata){fprintf(stderr,"Err: Cannot open %s\n",filedata); continue;}
+                if(!filedata){fprintf(stderr,"Err: Cannot open %s %s\n",argv[f], strerror(errno)); continue;}
                 NailArena_init(&arena,1024);
                 zip = parse_zip_file(&arena,filedata,size);
                 if(!zip){
-                        fprintf(stderr,"Err: Invalid ZIP file %s\n",filedata); 
+                        fprintf(stderr,"Err: Invalid ZIP file %s\n",argv[f]); 
                 } else {
                 }
                 NailArena_release(&arena);

### 2014-04-23 examples/zip/zip.c
Fixed finding end-of-file header in zip

@@ -65,7 +65,7 @@ int zip_compression_generate(NailArena *tmp, NailStream *uncompressed, NailStrea
 static int is_valid_directory(uint8_t *begin, uint8_t *end){
         if(end - begin<= 21 ) return 0;
         if(*(uint32_t *)begin != 0x06054b50 ) return 0;
-        if(*(uint16_t *)(begin+20) + 20 != end-begin) return 0;
+        if(*(uint16_t *)(begin+20) + 22 != end-begin) return 0;
         return 1;
 }
 int zip_end_of_directory_parse(NailArena *tmp, NailStream *filestream, NailStream *directory, NailStream *entire_file){

### 2014-04-23 generator/new_grammar.nail.c
endian syntax

@@ -2856,6 +2856,224 @@ choice_17_PARSER_out:
         goto choice_17_succ;
 choice_17_CONSTANTDEF_out:
         stream_reposition(str_current, back);
+        choice = n_tr_memo_choice(trace);
+        {   NailStreamPos back = stream_getpos(str_current);
+            pos choice_begin = n_tr_begin_choice(trace);
+            pos choice;
+            if(parser_fail(choice_begin)) {
+                goto choice_17_ENDIAN_out;
+            }
+            choice = n_tr_memo_choice(trace);
+            if(parser_fail(peg_WHITE(str_current))) {
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(n_tr_const(trace,str_current))) {
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(stream_check(str_current,8))) {
+                goto choice_18_LITTLE_out;
+            }
+            if( read_unsigned_bits(str_current,8)!= '!') {
+                stream_backup(str_current,8);
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(stream_check(str_current,8))) {
+                goto choice_18_LITTLE_out;
+            }
+            if( read_unsigned_bits(str_current,8)!= 'L') {
+                stream_backup(str_current,8);
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(stream_check(str_current,8))) {
+                goto choice_18_LITTLE_out;
+            }
+            if( read_unsigned_bits(str_current,8)!= 'I') {
+                stream_backup(str_current,8);
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(stream_check(str_current,8))) {
+                goto choice_18_LITTLE_out;
+            }
+            if( read_unsigned_bits(str_current,8)!= 'T') {
+                stream_backup(str_current,8);
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(stream_check(str_current,8))) {
+                goto choice_18_LITTLE_out;
+            }
+            if( read_unsigned_bits(str_current,8)!= 'T') {
+                stream_backup(str_current,8);
+                goto choice_18_LITTLE_out;
+            }
+            if(parser_fail(stream_check(str_current,8))) {
+                goto choice_18_LITTLE_out;
+            }
+            if( read_unsigned_bits(str_current,8)!= 'L') {
+                stream_backup(str_current,8);
[…]

### 2014-04-23 generator/new_grammar.nail.h

@@ -1,7 +1,7 @@
 #include <stdint.h>
 #include <string.h>
 #include <assert.h>
-enum N_types {_NAIL_NULL,CONSTANTDEF,PARSER,PR,PAREN,NAME,REF,NUNION,OPTIONAL,APPLY,LENGTH,FIXEDARRAY,ARRAY,CHOICE,WRAP,STRUCTURE,INTEGER,DDEPENDENCY,DSTREAM,PSTREAM,PDEPENDENCY,SEPBY,SEPBYONE,MANY,MANYONE,FIELD,TRANSFORM,DEPENDENCY,CONSTANT,NEGATE,SET,SINGLE,VALUE,RANGE,CUNION,CSTRUCT,CREF,CINT,CREPEAT,CARRAY,VALUES,STRING,SIGN,UNSIGN,DIRECT,ESCAPE,NUMBER,HEX,ASCII};
+enum N_types {_NAIL_NULL,BIG,LITTLE,ENDIAN,CONSTANTDEF,PARSER,PR,PAREN,NAME,REF,NUNION,OPTIONAL,APPLY,LENGTH,FIXEDARRAY,ARRAY,CHOICE,WRAP,STRUCTURE,INTEGER,DDEPENDENCY,DSTREAM,PSTREAM,PDEPENDENCY,SEPBY,SEPBYONE,MANY,MANYONE,FIELD,TRANSFORM,DEPENDENCY,CONSTANT,NEGATE,SET,SINGLE,VALUE,RANGE,CUNION,CSTRUCT,CREF,CINT,CREPEAT,CARRAY,VALUES,STRING,SIGN,UNSIGN,DIRECT,ESCAPE,NUMBER,HEX,ASCII};
 typedef struct number number;
 typedef struct varidentifier varidentifier;
 typedef struct constidentifier constidentifier;
@@ -302,6 +302,17 @@ struct definition {
             constparser definition;
         }
         constantdef;
+        struct {
+            enum N_types N_type;
+            union {
+                struct {
+                }
+                little;
+                struct {
+                }
+                big;
+            };
+        } endian;
     };
 };
 struct grammar {

### 2014-04-23 generator/mk_parser.cc
Parser now supports endianness

@@ -207,13 +207,18 @@ std::string intconstant_value(const intconstant &val){
   }
 }
 
-static std::string int_expr(const char * stream,unsigned int width){
+static std::string int_expr(const char * stream,unsigned int width,Endian endian){
   if(width>=64){
     throw std::runtime_error("More than 64 bits in integer");
   }
   //TODO: Support little endian?
-  return boost::str(boost::format("read_unsigned_bits(%s,%d)")%stream % width);
+  if(endian == Big){
+    return boost::str(boost::format("read_unsigned_bits(%s,%d)")%stream % width);
+  } else { 
+    return boost::str(boost::format("read_unsigned_bits_littleendian(%s,%d)")%stream % width);    
+  }
 }
+
 class CAction{
   std::ostream &out;
   int num_iters;
@@ -226,11 +231,12 @@ class CAction{
         << "tr++;\n";
   }
 public:
-  void action (const parserinner &p, Expr &lval){
+  bool big_endian;
+  void action (const parserinner &p, Expr &lval, Endian end){
     switch(p.N_type){
     case INTEGER:{
       int width = boost::lexical_cast<int>(mk_str(p.integer.parser.unsign));
-      out << lval << "=" << int_expr("stream",width) << ";\n";
+      out << lval << "=" << int_expr("stream",width,end) << ";\n";
     }
       break;
     case STRUCTURE:
@@ -242,7 +248,7 @@ public:
         case FIELD:{
           std::string name = mk_str(field->field.name);
           ValExpr fieldname(name,&lval);
-          action(field->field.parser->pr,fieldname);
+          action(field->field.parser->pr,fieldname,end);
           break;
         }
         case DEPENDENCY:{
@@ -267,7 +273,7 @@ public:
           action_constparser();
         }
       }
-      action(p.wrap.parser->pr,lval);
+      action(p.wrap.parser->pr,lval,end);
       if(p.wrap.constafter){
         FOREACH(c,*p.wrap.constafter){
           action_constparser();
@@ -288,7 +294,7 @@ public:
           std::string tag (mk_str(c->tag));
           boost::algorithm::to_lower(tag);
           ValExpr expr(tag,&lval);
-          action(c->parser->pr, expr );
+          action(c->parser->pr, expr,end );
           out << "break;\n";
         }
[…]

### 2014-04-23 generator/nailtool.h

@@ -113,4 +113,6 @@ std::string parameter_definition(const definition &def, Scope &scope );
 std::string parameter_invocation(const parameterlist *parameters, const Scope &scope);
 
 void constraint(std::ostream &out,std::string val,  intconstraint &c);
+
+typedef enum {Big, Little} Endian;
 #endif

### 2014-04-23 generator/parser_template.c
read_unsigned_bits readded

@@ -9,6 +9,37 @@ typedef struct{
         pos capacity,iter,grow;
 } n_trace; 
 #define parser_fail(i) __builtin_expect(i<0,0)
+static uint64_t read_unsigned_bits_littleendian(NailStream *stream, unsigned count) {
+    uint64_t retval = 0;
+    unsigned int out_idx=0;
+    size_t pos = stream->pos;
+    char bit_offset = 7-stream->bit_offset;
+    const uint8_t *data = stream->data;
+    while(count>0) {
+        if(bit_offset == 0 && (count &7) ==0) {
+            retval|= data[pos] << out_idx;
+            out_idx+=8;
+            pos ++;
+            count-=8;
+        }
+        else {
+            //This can use a lot of performance love
+//TODO: test this
+            retval |= ((data[pos] >> (bit_offset)) & 1) << out_idx;
+            out_idx++;
+            count--;
+            bit_offset--;
+            if(bit_offset <0) {
+                bit_offset += 8;
+                pos++;
+            }
+        }
+    }
+    stream->pos = pos;
+    stream->bit_offset = bit_offset;
+    return retval;
+
+}
 
 static uint64_t read_unsigned_bits(NailStream *stream, unsigned count){ 
         uint64_t retval = 0;

### 2014-04-23 examples/zip/extract.c
Can list zip file contents

@@ -7,6 +7,7 @@
 #include <errno.h>
 #include <stdio.h>
 #include "zip.nail.h"
+#define FOREACH(val,coll) for(__typeof__((coll).elem[0]) *val=(coll).elem;val<(coll).elem + (coll).count;val++)
 char *mmap_file(char *filename, size_t *size){
         struct stat stat;
         int fil  = open(filename, O_RDONLY);
@@ -30,6 +31,9 @@ int main(int argc,char **argv){
                 if(!zip){
                         fprintf(stderr,"Err: Invalid ZIP file %s\n",argv[f]); 
                 } else {
+                        FOREACH(file, zip->contents.files){
+                                printf("%.*s = %d\n", file->filename.count, file->filename.elem, file->contents.contents.count);
+                        }
                 }
                 NailArena_release(&arena);
                 munmap(filedata,size);

### 2014-04-23 examples/zip/zip.c

@@ -13,11 +13,12 @@ depend_parse(uint16)
 depend_parse(uint32)
 
 int size_u32_parse(NailArena *foo, NailStream *fragment, NailStream *current, uint32_t *size){
-        if(current->pos + *size >= current->size)
+        if(current->pos + *size > current->size)
                 return -1;
         fragment->data = current->data + current->pos;
         fragment->bit_offset = current->bit_offset;
         fragment->size = *size;
+        fragment->pos = 0;
         return 0;
 }
 int size_u32_generate(NailArena *tmp, NailStream *fragment, NailStream *current, uint32_t *size){ 
@@ -85,6 +86,7 @@ int zip_end_of_directory_parse(NailArena *tmp, NailStream *filestream, NailStrea
                         directory->bit_offset = 0;
                         directory->size = entire_file->size - off;
                         directory->pos = 0;
+                        entire_file->pos = entire_file->size;
                         return 0;
                 }
         }

### 2014-04-23 generator/parser_template.c

@@ -13,7 +13,7 @@ static uint64_t read_unsigned_bits_littleendian(NailStream *stream, unsigned cou
     uint64_t retval = 0;
     unsigned int out_idx=0;
     size_t pos = stream->pos;
-    char bit_offset = 7-stream->bit_offset;
+    char bit_offset = stream->bit_offset;
     const uint8_t *data = stream->data;
     while(count>0) {
         if(bit_offset == 0 && (count &7) ==0) {
@@ -28,9 +28,9 @@ static uint64_t read_unsigned_bits_littleendian(NailStream *stream, unsigned cou
             retval |= ((data[pos] >> (bit_offset)) & 1) << out_idx;
             out_idx++;
             count--;
-            bit_offset--;
-            if(bit_offset <0) {
-                bit_offset += 8;
+            bit_offset++;
+            if(bit_offset >7) {
+                bit_offset -= 8;
                 pos++;
             }
         }

### 2014-04-23 paper/osdi14/abstract.tex
Suggestion for better abstract

\begin{abstract}

  We present \textit{Nail}, an interface generator that allows programmers to safely parse and
  generate protocols defined by a Parser-Expression based grammar. Nail uses a richer set of parser
  combinators that induce an internal representation, obviating the need to write semantic actions.
  Nail also introduces \textit{dependency fields} and \textit{stream transformations} that capture
  protocol features such as size and offset fields, check-sums and compressed data that are hard to
  express in existing protocol languages.

\end{abstract}

### 2014-04-24 paper/osdi14/design.tex
Some paper changes

--- 2014-04-03 paper/design.tex
@@
 \subsection{Overview}
+
+Nail grammars describe not only a
 
@@
 
-\paragraph{Fundamental parsers.}
+\paragraph{Integers and Constraints}
+Nail's fundamental 
 The elementary parsers of Nail are the same as those of Hammer, signed
@@
 \texttt{many (uint8=' ')} (representing any number of space characters), or
-\texttt{|| uint8 = 0 || uint16 = 0xFFFF} (\XXX[what is this doing?  what is this
-syntax?  it looks notably different from Figure~\ref{fig:choice}]).
+\texttt{|| uint8 = 0x90 || uint16 = 0x1F0F} (parsing two of the many representations of NOP on the
+x86 architecture). Note that constant may have varying lengths.
 This is particularly useful for handling padding or whitespace.

### 2014-04-24 paper/osdi14/eval.tex
Some paper changes

--- 2014-04-02 paper/eval.tex
@@
 \label{s:eval}
-
+\input{fig-ethernet}
 \input{fig-dnsfull}
-
-In our preliminary evaluation of Nail, we try to answer two questions:
+\input{fig-zip}
+In our preliminary evaluation of Nail, we try to answer four questions:
 
@@
       application that uses Nail for data input and output?
+\item Do applications handling input and output with Nail suffer from less security vulnerabilities?
+\item How fast are systems using Nail parsers as opposed to hand-written input handling? 
 \end{itemize}
 
-\paragraph{Data formats.}
+\subsection{Data formats.}
+\label{s:eval-formats}
+To answer the first question, we implemented grammars for four protocols with particularly
+challenging features and show that Nail is a good fit for these data formats. In
+\S\ref{s:eval-effort} we will demonstrate that Nail is easier to use than hand-written parsers and
+more convenient than state-of-the-art parser generators for formats that they can handle.
 
-To answer the first question, we implemented two Nail grammars:
-one for UTF-16 encoded strings, exposing an array of code points
-(shown in Figure~\ref{fig:utf16}), and another for a subset of DNS
-packets sent to and from an authoritative name server, without label
-compression, as per RFC1035 (shown in Figure~\ref{fig:dns-full}).
-Furthermore, our GitHub repository contains other Nail
-grammars, such as a TAP network stack that processes Ethernet,
-ARP, ICMP, IP, and UDP packets, and the grammar for Nail
-itself.\footnote{\url{https://github.com/jbangert/nail/tree/master/examples}}
+In particular, we show extracts from grammars for the following protocols: 
+\begin{itemize}
+\item A subset of Ethernet network packets, containing ARP, ICMP and UDP/IP packets. See Figure~\ref{fig:ethernet-extract}.
+\item DNS packets as described in RFC 1035, featuring label compression. See Figure~\ref{fig:dns-full}.
+\item ZIP files, with compression, check-sums and a complicated procedure to identify the header.
+  See Figure~\ref{fig:zip-extract}.
+\end{itemize}
+
+one for  DNS queries and respones, as per RFC1035 (shown in, one for the ZIP
+archive format and a grammar for Ethernet frames containing ARP, ICMP and UDP packets. We also
+parse Nail grammars with Nail, demonstrating that our system can also handle text-based protocols
+and languages.
+
 The results suggest that Nail is a good fit for these data formats.
 
-\paragraph{Programmer effort.}
+\subsection{Programmer effort.}
+\label{s:eval-effort}
+To answer the second question, we implemented small example applications based on the above
+grammars and compared code size with comparable applications written without parser generators. We
+also compare our DNS implementation to the example DNS application provided with the Hammer parser generator.
 
-To answer the second question, we implemented a functioning toy DNS
+ a functioning toy DNS
 server.  In particular, we cloned the test DNS server from the Hammer
@@
 
+\paragraph{Security}
+
 \paragraph{Other issues.}

### 2014-04-24 tests/harness.sh
Print testcase name

@@ -17,6 +17,7 @@ COMPILER=`gcc  -ggdb -DXYCONST=0 -lhammer -std=gnu99 -o $TEST-test $TEST.c  2>&1
     
 for case in $CFILE.in.*
  do 
+    echo "$case"
    OUT=`./$TEST-test < $case 2>&1` || ( echo "Failed  testcase $case \n" $OUT; exit -1) || exit -1 
  done
 

### 2014-04-24 generator/mk_parser.cc
Allow stream aliasing in transform

@@ -707,7 +707,7 @@ class CPrimitiveParser{
           {
             header << "extern int " << mk_str(field->transform.cfunction) << "_parse(NailArena *tmp";
             FOREACH(stream, field->transform.left){
-              out << ";\nNailStream str_" << mk_str(*stream) <<";\n"; // If declaration occurs right
+              out << ";\nNailStream *str_" << mk_str(*stream) <<";\n"; // If declaration occurs right
                                                                       // after a label, we need an
                                                                       // empty statement
               // str_current cannot appear on the left
@@ -716,7 +716,7 @@ class CPrimitiveParser{
             out << mk_str(field->transform.cfunction) << "_parse(tmp_arena"; //TODO: use temp arena
 
             FOREACH(stream, field->transform.left){           
-              header << ",NailStream *str_" << mk_str(*stream);
+              header << ",NailStream **str_" << mk_str(*stream);
               out << ", &str_"  << mk_str(*stream);
             }      
             FOREACH(param, field->transform.right){
@@ -738,7 +738,7 @@ class CPrimitiveParser{
             out << ")) {" << fail << "}";
             header << ");\n";
             FOREACH(stream, field->transform.left){
-              newscope.add_stream_definition(mk_str(*stream));
+              newscope.add_stream_parameter(mk_str(*stream));
             }
             out << "n_tr_stream(trace,str_current);";
             

### 2014-04-24 generator/parser_template.h

@@ -16,7 +16,10 @@ struct NailStream {
 
 typedef struct NailStream NailStream;
 typedef size_t NailStreamPos;
-extern int NailOutStream_new(NailStream *str,size_t siz);
+static NailStream * NailStream_alloc(NailArena *arena) {
+        return (NailStream *)n_malloc(arena, sizeof(NailStream));
+} 
+extern int NailOutStream_init(NailStream *str,size_t siz);
 extern void NailOutStream_release(NailStream *str);
 const uint8_t * NailOutStream_buffer(NailStream *str,size_t *siz);
 extern int NailOutStream_grow(NailStream *stream, size_t count);

### 2014-04-24 tests/transform/trans1.c
Changed trans1

@@ -1,13 +1,17 @@
 //Note that all nail offsets are in bits
-int offsetsize_parse(NailArena *tmp,NailStream *str_filestream,NailStream *contentstream,uint32_t* offset,uint32_t* filesize){
+int offsetsize_parse(NailArena *tmp,NailStream **str_filestream,NailStream *contentstream,uint32_t* offset,uint32_t* filesize){
         if(*offset > contentstream->size)
                 return -1;
         if(*filesize > contentstream->size  - *offset)
                 return -1;
-        str_filestream->data = contentstream->data + *offset;
-        str_filestream->pos  = 0;
-        str_filestream->size = *filesize;
-        str_filestream->bit_offset = 0;
+        NailStream *filestream = NailStream_alloc(tmp);
+        if(!filestream) 
+                return -1;
+        *str_filestream = filestream;
+        filestream->data = contentstream->data + *offset;
+        filestream->pos  = 0;
+        filestream->size = *filesize;
+        filestream->bit_offset = 0;
         return 0;
 }
 int offsetsize_generate(NailArena *tmp, NailStream *str_filestream, NailStream *contentstream, uint32_t *offset, uint32_t *filesize){
@@ -23,18 +27,22 @@ int offsetsize_generate(NailArena *tmp, NailStream *str_filestream, NailStream *
         return 0;
 }
 static const int headersize = 1024;
-int splitzip_parse(NailArena *tmp,NailStream *str_header,NailStream *str_content,NailStream *current){
+int splitzip_parse(NailArena *tmp,NailStream **str_header,NailStream **str_content,NailStream *current){
         if(current->size <= headersize)
                 return -1;
         if(current->pos != 0)
                 return -1;
-        str_header->data = current->data + current->size - headersize;
-        str_header->pos = 0;
-        str_header->size = headersize;
-        str_header->bit_offset = 0;
-        str_content->data = current->data;
-        str_content->pos = 0;
-        str_content->size = current->size - headersize;
+        if(!(str_header[0] = NailStream_alloc(tmp)))
+           return -1;
+        str_header[0]->data = current->data + current->size - headersize;
+        str_header[0]->pos = 0;
+        str_header[0]->size = headersize;
+        str_header[0]->bit_offset = 0;
+        if(!(str_content[0] = NailStream_alloc(tmp)))
+           return -1;
+        str_content[0]->data = current->data;
+        str_content[0]->pos = 0;
+        str_content[0]->size = current->size - headersize;
[…]

### 2014-04-24 generator/nailtool.h
Proper exceptions

@@ -5,6 +5,7 @@
 #include <cstring>
 #include <cstdio>
 #include <cstdlib>
+#include <stdexcept>
 #include <map>
 #include <boost/format.hpp>
 #include <boost/lexical_cast.hpp>
@@ -37,7 +38,10 @@ std::string intconstant_value(const intconstant &val);
 bool parameter_type_check(parameterlist *param, parameterdefinitionlist *def);
 
 #define mk_str(x) std::string((const char *)(x).elem,(x).count)
-
+class NailException : std::logic_error { 
+public:
+NailException(const std::string &message) : std::logic_error(message){}              
+};
 struct Dependency{
   std::string name;
   std::string typedef_type;
@@ -51,7 +55,7 @@ public:
   Scope(Scope *p_parent = NULL) : parent(p_parent) {}
   void add_stream_parameter(std::string value){
     if(!streams.emplace(value, std::string("str_")+value).second){
-      throw new std::string("Duplicate stream "+value);
+      throw NailException("Duplicate stream "+value);
     }
   }
   void add_dependency_parameter(std::string value, std::string type){
@@ -59,12 +63,12 @@ public:
     dep.name = std::string("dep_")+value;
     dep.typedef_type = type;
     if(!dependencies.emplace(value,dep).second){
-      throw new std::string("Duplicate dependency "+value);
+      throw NailException("Duplicate dependency "+value);
     }
   }
   void add_stream_definition(std::string value){
     if(!streams.emplace(value, std::string("&str_")+value).second){
-      throw new std::string("Duplicate dependency "+value);
+      throw NailException("Duplicate dependency "+value);
     }
   }
   void add_dependency_definition(std::string value,std::string type){
@@ -72,14 +76,14 @@ public:
     dep.name = std::string("&dep_")+value;
     dep.typedef_type = type;
     if(!dependencies.emplace(value,dep).second){
-      throw new std::string("Duplicate dependency "+value);
+      throw NailException("Duplicate dependency "+value);
     }
   }
   std::string dependency_type(std::string name) const{
     auto i = dependencies.find(name);
     if(i==dependencies.end()){
       if(!parent)
-        throw std::string("Undefined reference to ") +  name;
+              throw NailException(std::string("Undefined reference to ") +  name);
       else
         return parent->dependency_type(name);
     }
@@ -89,7 +93,7 @@ public:
     auto i = dependencies.find(name);
     if(i==dependencies.end()){
       if(!parent)
[…]

### 2014-04-24 generator/main.cc
Better error messages in nailtool

@@ -36,8 +36,11 @@ int main(int argc, char**argv)
     //      print_grammar(result, stderr,0);    exit(1);
      std::string headerfilename = std::string(argv[1]) + ".h";
      std::string implfilename = std::string(argv[1]) + ".c";
-     
-     if(result) {
+     if(!result){
+       error("Syntax error in grammar\n");
+       return 1;
+     }
+     try {
        std::ofstream header(headerfilename);
        std::ofstream impl(implfilename);
        if(!header.is_open())
@@ -51,8 +54,8 @@ int main(int argc, char**argv)
        impl << std::endl;
        header << std::endl;       
        return 0;
-     } else {
-       error("Syntax error in grammar\n");
+     } catch ( std::exception const &e){
+       error("An error occured:%s\n",e.what());
        return 1;
      }
 }

### 2014-04-24 generator/mk_parser.cc

@@ -180,7 +180,7 @@ public:
       }
       outs << "};\n";
       outs << out.str()<< std::endl;
-    } catch(std::exception &e){
+    } catch(const std::exception &e){
       std::cerr << "Exception while generating parser" << e.what()<<std::endl;
       exit(-1);
     }

### 2014-04-24 generator/nailtool.h

@@ -38,7 +38,7 @@ std::string intconstant_value(const intconstant &val);
 bool parameter_type_check(parameterlist *param, parameterdefinitionlist *def);
 
 #define mk_str(x) std::string((const char *)(x).elem,(x).count)
-class NailException : std::logic_error { 
+struct NailException : public std::logic_error { 
 public:
 NailException(const std::string &message) : std::logic_error(message){}              
 };

### 2014-04-24 generator/main.cc
emit transforms into header

@@ -49,7 +49,7 @@ int main(int argc, char**argv)
          error("Cannot open output file %s\n",implfilename.c_str());
        emit_header(&header,result);
        impl << "#include \""<< headerfilename << "\""<<std::endl;
-       emit_parser(&impl,result);
+       emit_parser(&impl,&header,result);
        emit_generator(&impl,&header,result);
        impl << std::endl;
        header << std::endl;       

### 2014-04-24 generator/mk_parser.cc

@@ -468,7 +468,7 @@ public:
 
 class CPrimitiveParser{
   std::stringstream out;
-  std::ostream &final;
+  std::ostream &final,&hdr;
   std::stringstream header;
   int nr_choice,nr_option;
   int nr_many;
@@ -840,7 +840,7 @@ class CPrimitiveParser{
     }
   }
 public:
-  CPrimitiveParser(std::ostream *os) : final(*os), nr_choice(1),nr_many(0),nr_const(0),nr_dep(0),num_iters(1), dependency_action(&out,"tmp_arena"), end(Big){}
+  CPrimitiveParser(std::ostream *os, std::ostream *header) : final(*os), hdr(*header),nr_choice(1),nr_many(0),nr_const(0),nr_dep(0),num_iters(1), dependency_action(&out,"tmp_arena"), end(Big){}
   void peg(const definition &def) {
     if(def.N_type == ENDIAN){  
       end = def.endian.N_type == BIG ? Big : Little;
@@ -875,12 +875,13 @@ public:
     }
     out << std::endl;
     final << header.str() << out.str() << std::endl;
+    hdr << header.str() << std::endl;
 
   }
 };
-void emit_parser(std::ostream *out, grammar *grammar){
+void emit_parser(std::ostream *out, std::ostream *header,grammar *grammar){
 
-  CPrimitiveParser p(out);
+  CPrimitiveParser p(out,header);
   CAction a(out,"arena");
   *out << std::string(parser_template_start,parser_template_end - parser_template_start);
   p.emit_parser(*grammar);

### 2014-04-24 generator/nailtool.h

@@ -31,7 +31,7 @@ typedef struct expr{
 
 #define FOREACH(val,coll) for(__typeof__((coll).elem[0]) *val=(coll).elem;val<(coll).elem + (coll).count;val++)
 extern void emit_generator(std::ostream *out,std::ostream *header,grammar *grammar);
-extern void emit_parser(std::ostream *out,grammar *grammar);
+extern void emit_parser(std::ostream *out,std::ostream *header,grammar *grammar);
 extern void emit_header(std::ostream *out,grammar *grammar);
 std::string intconstant_value(const intconstant &val);
 

### 2014-04-24 generator/mk_parser.cc
Remove DEBUG_OUT

@@ -5,7 +5,7 @@
 #include <boost/algorithm/string/case_conv.hpp>
 #include <list>
 #define MAP(f,collection) FOREACH(iter,collection){ f(*iter);}
-#define DEBUG_OUT
+#undef DEBUG_OUT
 //static const std::string parser_template(_binary_parser_template_c_start,_binary_parser_template_c_end - _binary_parser_template_c_start);
 class CDataModel{
   std::stringstream out;

### 2014-04-24 generator/mk_parser.cc
Forward declarations

@@ -470,6 +470,7 @@ class CPrimitiveParser{
   std::stringstream out;
   std::ostream &final,&hdr;
   std::stringstream header;
+  std::stringstream forward_declarations;
   int nr_choice,nr_option;
   int nr_many;
   int nr_const;
@@ -856,7 +857,7 @@ public:
       out << "return 0;\n"
           << "fail:\n return -1;\n";
       out << "}\n";
-      header << "static pos peg_" << mk_str(def.parser.name) <<"(NailArena *tmp_arena,n_trace *trace,NailStream *str_current"<<params<<");\n";
+      forward_declarations << "static pos peg_" << mk_str(def.parser.name) <<"(NailArena *tmp_arena,n_trace *trace,NailStream *str_current"<<params<<");\n";
 
     }
     else if(def.N_type == CONSTANTDEF){
@@ -866,7 +867,7 @@ public:
       out << "return 0;\n"
           << "fail: return -1;\n"
           << "}";
-        header << "static pos peg_" << mk_str(def.constantdef.name) <<"(NailStream *str_current);\n";
+      forward_declarations << "static pos peg_" << mk_str(def.constantdef.name) <<"(NailStream *str_current);\n";
     }
   }
   void emit_parser(const grammar &gram){
@@ -874,7 +875,7 @@ public:
       peg(*def);
     }
     out << std::endl;
-    final << header.str() << out.str() << std::endl;
+    final << header.str() << forward_declarations.str() << out.str() << std::endl;
     hdr << header.str() << std::endl;
 
   }

### 2014-04-24 generator/parser_template.c
Trace debug

@@ -246,7 +246,9 @@ static pos n_tr_memo_many(n_trace *trace){
 static void n_tr_write_many(n_trace *trace, pos where, pos count){
         trace->trace[where] = count;
         trace->trace[where+1] = trace->iter;
+#ifdef NAIL_DEBUG
         fprintf(stderr,"%d = many %d %d\n", where, count,trace->iter);
+#endif
 }
 
 static pos n_tr_begin_choice(n_trace *trace){
@@ -264,7 +266,9 @@ static int n_tr_stream(n_trace *trace, const NailStream *stream){
         if(parser_fail(n_trace_grow(trace,sizeof(*stream)/sizeof(pos))))
                 return -1;
         *(NailStream *)(trace->trace + trace->iter) = *stream;
+#ifdef NAIL_DEBUG
         fprintf(stderr,"%d = stream\n",trace->iter,stream);
+#endif
         trace->iter += sizeof(*stream)/sizeof(pos);  
         return 0;
 }
@@ -274,13 +278,17 @@ static pos n_tr_memo_choice(n_trace *trace){
 static void n_tr_pick_choice(n_trace *trace, pos where, pos which_choice, pos  choice_begin){
         trace->trace[where] = which_choice;
         trace->trace[where + 1] = choice_begin;
+#ifdef NAIL_DEBUG
         fprintf(stderr,"%d = pick %d %d\n",where, which_choice,choice_begin);
+#endif
 }
 static int n_tr_const(n_trace *trace,NailStream *stream){
         if(parser_fail(n_trace_grow(trace,1)))
                 return -1;
         NailStreamPos newoff = stream_getpos(stream);
+#ifdef NAIL_DEBUG
         fprintf(stderr,"%d = const %d \n",trace->iter, newoff);        
+#endif
         trace->trace[trace->iter++] = newoff;
         return 0;
 }

### 2014-04-24 generator/new_grammar.nail.c
Removed nail trace output!

@@ -10,6 +10,37 @@ typedef struct {
     pos capacity,iter,grow;
 } n_trace;
 #define parser_fail(i) __builtin_expect(i<0,0)
+static uint64_t read_unsigned_bits_littleendian(NailStream *stream, unsigned count) {
+    uint64_t retval = 0;
+    unsigned int out_idx=0;
+    size_t pos = stream->pos;
+    char bit_offset = stream->bit_offset;
+    const uint8_t *data = stream->data;
+    while(count>0) {
+        if(bit_offset == 0 && (count &7) ==0) {
+            retval|= data[pos] << out_idx;
+            out_idx+=8;
+            pos ++;
+            count-=8;
+        }
+        else {
+            //This can use a lot of performance love
+//TODO: test this
+            retval |= ((data[pos] >> (bit_offset)) & 1) << out_idx;
+            out_idx++;
+            count--;
+            bit_offset++;
+            if(bit_offset >7) {
+                bit_offset -= 8;
+                pos++;
+            }
+        }
+    }
+    stream->pos = pos;
+    stream->bit_offset = bit_offset;
+    return retval;
+
+}
 
 static uint64_t read_unsigned_bits(NailStream *stream, unsigned count) {
     uint64_t retval = 0;
@@ -216,7 +247,9 @@ static pos n_tr_memo_many(n_trace *trace) {
 static void n_tr_write_many(n_trace *trace, pos where, pos count) {
     trace->trace[where] = count;
     trace->trace[where+1] = trace->iter;
+#ifdef NAIL_DEBUG
     fprintf(stderr,"%d = many %d %d\n", where, count,trace->iter);
+#endif
 }
 
 static pos n_tr_begin_choice(n_trace *trace) {
@@ -234,7 +267,9 @@ static int n_tr_stream(n_trace *trace, const NailStream *stream) {
     if(parser_fail(n_trace_grow(trace,sizeof(*stream)/sizeof(pos))))
         return -1;
     *(NailStream *)(trace->trace + trace->iter) = *stream;
+#ifdef NAIL_DEBUG
     fprintf(stderr,"%d = stream\n",trace->iter,stream);
+#endif
     trace->iter += sizeof(*stream)/sizeof(pos);
     return 0;
 }
@@ -244,13 +279,17 @@ static pos n_tr_memo_choice(n_trace *trace) {
[…]

### 2014-04-24 generator/new_grammar.nail.h

@@ -339,7 +339,10 @@ struct NailStream {
 
 typedef struct NailStream NailStream;
 typedef size_t NailStreamPos;
-extern int NailOutStream_new(NailStream *str,size_t siz);
+static NailStream * NailStream_alloc(NailArena *arena) {
+    return (NailStream *)n_malloc(arena, sizeof(NailStream));
+}
+extern int NailOutStream_init(NailStream *str,size_t siz);
 extern void NailOutStream_release(NailStream *str);
 const uint8_t * NailOutStream_buffer(NailStream *str,size_t *siz);
 extern int NailOutStream_grow(NailStream *stream, size_t count);
@@ -372,6 +375,7 @@ parserinner*parse_parserinner(NailArena *arena, const uint8_t *data, size_t size
 parser*parse_parser(NailArena *arena, const uint8_t *data, size_t size);
 definition*parse_definition(NailArena *arena, const uint8_t *data, size_t size);
 grammar*parse_grammar(NailArena *arena, const uint8_t *data, size_t size);
+
 int gen_number(NailArena *tmp_arena,NailStream *out,number * val);
 int gen_varidentifier(NailArena *tmp_arena,NailStream *out,varidentifier * val);
 int gen_constidentifier(NailArena *tmp_arena,NailStream *out,constidentifier * val);

### 2014-04-24 generator/mk_parser.cc
revert stream aliasing

@@ -708,7 +708,7 @@ class CPrimitiveParser{
           {
             header << "extern int " << mk_str(field->transform.cfunction) << "_parse(NailArena *tmp";
             FOREACH(stream, field->transform.left){
-              out << ";\nNailStream *str_" << mk_str(*stream) <<";\n"; // If declaration occurs right
+              out << ";\nNailStream str_" << mk_str(*stream) <<";\n"; // If declaration occurs right
                                                                       // after a label, we need an
                                                                       // empty statement
               // str_current cannot appear on the left
@@ -717,7 +717,7 @@ class CPrimitiveParser{
             out << mk_str(field->transform.cfunction) << "_parse(tmp_arena"; //TODO: use temp arena
 
             FOREACH(stream, field->transform.left){           
-              header << ",NailStream **str_" << mk_str(*stream);
+              header << ",NailStream *str_" << mk_str(*stream);
               out << ", &str_"  << mk_str(*stream);
             }      
             FOREACH(param, field->transform.right){
@@ -739,7 +739,7 @@ class CPrimitiveParser{
             out << ")) {" << fail << "}";
             header << ");\n";
             FOREACH(stream, field->transform.left){
-              newscope.add_stream_parameter(mk_str(*stream));
+              newscope.add_stream_definition(mk_str(*stream));
             }
             out << "n_tr_stream(trace,str_current);";
             

### 2014-04-24 tests/transform/trans1.c

@@ -1,17 +1,13 @@
 //Note that all nail offsets are in bits
-int offsetsize_parse(NailArena *tmp,NailStream **str_filestream,NailStream *contentstream,uint32_t* offset,uint32_t* filesize){
+int offsetsize_parse(NailArena *tmp,NailStream *str_filestream,NailStream *contentstream,uint32_t* offset,uint32_t* filesize){
         if(*offset > contentstream->size)
                 return -1;
         if(*filesize > contentstream->size  - *offset)
                 return -1;
-        NailStream *filestream = NailStream_alloc(tmp);
-        if(!filestream) 
-                return -1;
-        *str_filestream = filestream;
-        filestream->data = contentstream->data + *offset;
-        filestream->pos  = 0;
-        filestream->size = *filesize;
-        filestream->bit_offset = 0;
+        str_filestream->data = contentstream->data + *offset;
+        str_filestream->pos  = 0;
+        str_filestream->size = *filesize;
+        str_filestream->bit_offset = 0;
         return 0;
 }
 int offsetsize_generate(NailArena *tmp, NailStream *str_filestream, NailStream *contentstream, uint32_t *offset, uint32_t *filesize){
@@ -27,22 +23,18 @@ int offsetsize_generate(NailArena *tmp, NailStream *str_filestream, NailStream *
         return 0;
 }
 static const int headersize = 1024;
-int splitzip_parse(NailArena *tmp,NailStream **str_header,NailStream **str_content,NailStream *current){
+int splitzip_parse(NailArena *tmp,NailStream *str_header,NailStream *str_content,NailStream *current){
         if(current->size <= headersize)
                 return -1;
         if(current->pos != 0)
                 return -1;
-        if(!(str_header[0] = NailStream_alloc(tmp)))
-           return -1;
-        str_header[0]->data = current->data + current->size - headersize;
-        str_header[0]->pos = 0;
-        str_header[0]->size = headersize;
-        str_header[0]->bit_offset = 0;
-        if(!(str_content[0] = NailStream_alloc(tmp)))
-           return -1;
-        str_content[0]->data = current->data;
-        str_content[0]->pos = 0;
-        str_content[0]->size = current->size - headersize;
+        str_header->data = current->data + current->size - headersize;
+        str_header->pos = 0;
+        str_header->size = headersize;
+        str_header->bit_offset = 0;
+        str_content->data = current->data;
+        str_content->pos = 0;
+        str_content->size = current->size - headersize;
[…]

### 2014-04-24 examples/zip/zip.c
zip now does compression

@@ -1,4 +1,5 @@
 #include "zip.nail.h"
+#include <zlib.h>
 #define depend_parse(type) \
         int type ## _depend_parse(NailArena *foo,type ##_t *left, type ##_t *right){ \
                 if(*left != *right) return -1;                      \
@@ -52,15 +53,45 @@ int offset_u32_generate(NailArena *tmp, NailStream *fragment, NailStream *curren
         return 0;
 }
 
-int zip_compression_parse(NailArena *tmp, NailStream *uncompressed, NailStream *compressed, uint16_t *compression_method, uint32_t *crc32){
-        *uncompressed = *compressed; //TODO: do decompression here
-        //TODO: CHeck CRC32
+int zip_compression_parse(NailArena *tmp, NailStream *uncompressed, NailStream *compressed, uint16_t *compression_method, uint32_t *uncompressed_size){
+        switch(*compression_method){
+        case 0:
+                if(*uncompressed_size != compressed->size)
+                        return -1;
+                *uncompressed = *compressed; //TODO: do decompression here
+                break;
+        case 8:{
+                z_stream strm;
+                strm.zalloc = strm.zfree = strm.opaque = Z_NULL;
+                strm.avail_in = compressed->size;
+                strm.next_in = compressed->data;
+                if(inflateInit2(&strm,-MAX_WBITS) != Z_OK)
+                        return -1;
+                uncompressed->data = n_malloc(tmp,*uncompressed_size);
+                uncompressed->pos = 0;
+                uncompressed->bit_offset = 0;
+                uncompressed->size = *uncompressed_size;
+                strm.next_out =  uncompressed->data;
+                if(!strm.next_out) 
+                        return -1;
+                strm.avail_out = *uncompressed_size;
+                int ret = inflate(&strm,Z_FINISH);
+                if(ret != Z_STREAM_END || strm.avail_in != 0){
+                        (void)inflateEnd(&strm);
+                        return -1;
+                }
+                inflateEnd(&strm);
+        }
+                break;
+        default:
+                return -1;
+        }
         return 0;
 }
-int zip_compression_generate(NailArena *tmp, NailStream *uncompressed, NailStream *compressed, uint16_t *compression_method, uint32_t *crc32){
[…]

### 2014-04-25 paper/osdi14/eval.tex
Paper extract

@@
 \subsection{Programmer effort.}
+
+%code size unzip 6.0
+%2821 lines
 \label{s:eval-effort}
 To answer the second question, we implemented small example applications based on the above
-grammars and compared code size with comparable applications written without parser generators. We
-also compare our DNS implementation to the example DNS application provided with the Hammer parser generator.
+grammars and compared code size with comparable applications that process data manually. We will
+also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
+example for the Hammer parser generator. All code size measurements for C were performed by
+SlocCount\cite{sloccount}. \XXX[code size is primitive, but we show orders of magnitude in the same language].
 
- a functioning toy DNS
-server.  In particular, we cloned the test DNS server from the Hammer
-distribution to Nail.  Hammer ships with a toy DNS server written in 683
-lines of code, excluding the Hammer framework itself, that responds to
+%%DNS: 
+\paragraph{DNS.}
+
+The first application we built with Nail, is a toy DNS server inspired by the example shipped with
+Hammers DNS server consists of 683 lines of C, excluding the Hammer framework itself, that responds to
 any valid DNS query with a CNAME record to the domain ``spargelze.it''.
@@
 
-Our DNS server consists of 148 lines of C code and 48 lines of Nail
-grammar, and supports a custom zone file format with A, NS, MX, and
-CNAME records. The same grammar is used, together with 98 lines of C,
+Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
+grammar and 64 lines of C implementing DNS label compression. \XXX[Make this smaller]
+The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
+(16 lines) supporting A, NS, MX and CNAME records.
+
+Our simple DNS server, which parses a zone file, listens to incoming DNS requests, parses them and
+generates appropriate responses with this Nail grammar, is implemented in 149 lines of C.
+
+ The same grammar is used, together with 98 lines of C,
 to implement a functional toy clone of the \texttt{host} command-line
-tool. However, because our grammar does not yet support DNS label
-compression, the latter tool will occasionally reject valid real-world
-DNS responses. Both clients have functional anti-spoofing measures.
+tool.
 
@@
 our toy DNS server to that of a real world DNS server, since
-we have less functionality, in particular for DNS compression
-and additional hint records that real-world DNS servers send.
+we have less , in particular the additional hint records that real-world DNS servers send and a
+complete lack of configuration options.
 However, the closest in functionality and intent is Dan Bernstein's
@@
 including various support tools, is about 10,000 lines of C as measured by
-\texttt{sloccount}.  We expect that it is possible to build a feature-par
-version with Nail that is an order of magnitude smaller and intend to
-do so.
+\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
+features we don't have]
 
-\paragraph{Security}
+\paragraph{ZIP.}
 
-\paragraph{Other issues.}
+Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}. The ZIP
+format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
+security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
+Android security has been completely by-passed multiple times through parsing differentials in
+handling ZIP files\cite{saurik-masterkey}. \XXX[mention talk by Julia Wolf about zip].
 
-As Nail is work in progress, many parts of the implementation, syntax and
-design are not complete yet and we do not yet have meaningful performance
-or security metrics.
+ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
+header}. This header contains a variable length comment, so it has to be located by scanning
+backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
+implementations disagree on how to find this header in confusing situations, such as when the
+comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
+This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
+array of one \emph{directory entry header} per file in the archive.
+Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
+in addition to the offset of a \emph{local file header}. The local file header duplicates most
+information from the directory entry header and the compressed file contents follow it immediately.
+
+Duplicating information made sense when ZIP files were
+stored on floppy disks with slow seek times and high fault rates and memory constraints made it
+impossible to keep the ZIP directory in memory or the archive was split across multiple disks.
+However, care must be taken that the meta-data is consistent. For example, vulnerabilities could occur
+if the length in the central directory is used to allocate memory and the length in the local
+directory is used to extract without checking that they are equal first.
+
+\XXX[condense]
+
+We wrote a grammar for ZIP, which consists of 92 lines of Nail and 78 lines of C implementing two
+stream transforms, one for the DEFLATE compression algorithm with the help of the zlib library and
+one for finding the end-of-directory header. This grammar avoids the two pitfalls mentioned above:
+Nail forces the programmer to specify an explicit procedure for finding the end-of-directory record
+and we use dependent fields to enforce that the duplicated information is consistent.
+
+Our grammar is also very efficient to use. Using our grammar, we build a ZIP file extractor in 50
+lines of C. Because more recent versions of ZIP have added more features, such as large file support
+and encryption, the closest existing tool in functionality is the historic version 5.4 of the Info-Zip unzip
+utility\cite{infozip} that is shipped with most Linux distributions. The entire unzip distribution
+is about 46.000 lines of code, which is mostly optimised implementation of various compression
+algorithms and other configuration and portability code.
+
+However, unzip isolates the equivalent of our Nail tool in the file extract.c, which parses the ZIP
+metadata and calls various decompression routines in other files. This file measures over 1,600
+lines of C.
+
+\paragraph{Network stack.}
+Write up this stack.
+\subsection{Security.} 
+Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.
+
+\subsection{Performance.}
+
+Compare DNS requests per second of BIND, DJBDNS and NailDNS on a small zonefile.
+Compare extraction performance for ZIP.

### 2014-04-25 paper/osdi14/relwk.tex
we should remember to mention the workshop paper in the osdi submission
Paper extract

--- 2014-04-09 paper/langsec14/relwk.tex
@@
 \label{s:relwk}
+
+\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]
+
+\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
+     but requires explicit encode/decode rules.  actually kind-of like
+     pack/unpack.]
 
@@
 treats them as unsigned, allowing one to replace files in a signed
-archive, thereby undermining Android's security model.}~\cite{saurik-masterkey} code signing and even the X.509 protocol
+archive, thereby undermining Android's security model.}~\cite{saurik-masterkey}\XXX{Duplicate, we
+already mention this in the eval} code signing and even the X.509 protocol
 underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are

### 2014-04-25 examples/zip/extract.c

@@ -32,7 +32,15 @@ int main(int argc,char **argv){
                         fprintf(stderr,"Err: Invalid ZIP file %s\n",argv[f]); 
                 } else {
                         FOREACH(file, zip->contents.files){
-                                printf("%.*s = %d\n", file->filename.count, file->filename.elem, file->contents.contents.count);
+                                char name[255];
+                                char dirname[255]="out-";
+                                strncpy(dirname+4,argv[f],sizeof(dirname)-4);
+                                mkdir(dirname,0777);
+                                snprintf(name,sizeof(name), "%s/%.*s",dirname,file->filename.count, file->filename.elem);
+                                printf("%s = %d\n", name, file->contents.contents.count);
+                                FILE *f = fopen(name,"w");
+                                fwrite(file->contents.contents.elem,file->contents.contents.count,1,f);
+                                fclose(f);
                         }
                 }
                 NailArena_release(&arena);

### 2014-04-28 examples/dns/bench/config.h
add queryperf from BIND 9

@@ -1,0 +1,38 @@
+/* config.h.  Generated from config.h.in by configure.  */
+/* config.h.in.  Generated from configure.in by autoheader.  */
+
+/* Define to 1 if you have the `getaddrinfo' function. */
+#define HAVE_GETADDRINFO 1
+
+/* Define to 1 if you have the `gethostbyname2' function. */
+#define HAVE_GETHOSTBYNAME2 1
+
+/* Define to 1 if you have the `getnameinfo' function. */
+#define HAVE_GETNAMEINFO 1
+
+/* Define to 1 if you have the `nsl' library (-lnsl). */
+#define HAVE_LIBNSL 1
+
+/* Define to 1 if you have the `socket' library (-lsocket). */
+/* #undef HAVE_LIBSOCKET */
+
+/* Define to 1 if `struct sockaddr' has element `sa_len'. */
+/* #undef HAVE_SA_LEN */
+
+/* Define to the address where bug reports for this package should be sent. */
+#define PACKAGE_BUGREPORT ""
+
+/* Define to the full name of this package. */
+#define PACKAGE_NAME ""
+
+/* Define to the full name and version of this package. */
+#define PACKAGE_STRING ""
+
+/* Define to the one symbol short name of this package. */
+#define PACKAGE_TARNAME ""
+
+/* Define to the version of this package. */
+#define PACKAGE_VERSION ""
+
+/* Define to `int' if `socklen_t' does not exist. */
+/* #undef socklen_t */

### 2014-04-29 generator/mk_parser.cc
Fixed memory leak in tmp_arena

@@ -361,7 +361,7 @@ public:
         out << count << "=" << "*(tr++);\n"
             << "save = *(tr++);\n";
         out <<data << "= " << "(typeof("<<data<<"))n_malloc("<<arena<<"," << count << "* sizeof(*"<<data<<"));\n"
-            << "if(!"<< data<< "){return 0;}\n";
+            << "if(!"<< data<< "){return -1;}\n";
 
         std::string iter = boost::str(boost::format("i%d") % num_iters++);
         out << "for(pos "<<iter<<"=0;"<<iter<<"<"<<count<<";"<<iter<<"++){";
@@ -449,15 +449,17 @@ public:
               <<"pos *tr_ptr;\n pos pos;\n"
               << name <<"* retval;\n"
               << "n_trace_init(&trace,4096,4096);\n"
-              << "tr_ptr = trace.trace;"
-              << "if(parser_fail(peg_"<<name<<"(&tmp_arena,&trace,&stream))) return NULL;"
-              << "if(stream.pos != stream.size) return NULL; "
+              << "if(parser_fail(peg_"<<name<<"(&tmp_arena,&trace,&stream))) goto fail;"
+              << "if(stream.pos != stream.size) goto fail; "
               << "retval =  (typeof(retval))n_malloc(arena,sizeof(*retval));\n"
-              <<"if(!retval) return NULL;\n"
+              <<"if(!retval) goto fail;\n"
               << "stream.pos = 0;\n"
-              << "if(bind_"<<name<<"(arena,retval,&stream,&tr_ptr,trace.trace) < 0) return NULL;\n"
-              << "n_trace_release(&trace);\n"
+              << "tr_ptr = trace.trace;"
+              << "if(bind_"<<name<<"(arena,retval,&stream,&tr_ptr,trace.trace) < 0) goto fail;\n"
+              << "out: n_trace_release(&trace);\n"
+              << "NailArena_release(&tmp_arena);"
               <<"return retval;"
+              << "fail: retval = NULL; goto out;"
               <<"}";
         }
       }

### 2014-04-29 examples/dns/bench/visualize.py
Performance stuff on DNS

@@ -1,0 +1,20 @@
+#!/usr/bin/python
+import sys
+
+import numpy as np
+from pylab import *
+
+def labelstr(path):
+    return os.path.splitext(path)[0]
+files = sys.argv
+del files[0]
+data = []
+labels = [] 
+for file in files:
+    print(file)
+    data.append( loadtxt(file))
+    labels.append(labelstr(file))
+figure(1)
+boxplot(data)
+xticks(range(0,len(labels)),labels)
+

### 2014-04-29 examples/dns/bench/visualize.py
Abandoned efforts to make plots

@@ -1,6 +1,6 @@
 #!/usr/bin/python
 import sys
-
+import os
 import numpy as np
 from pylab import *
 
@@ -12,9 +12,10 @@ data = []
 labels = [] 
 for file in files:
     print(file)
-    data.append( loadtxt(file))
+    data.append(loadtxt(file))
     labels.append(labelstr(file))
 figure(1)
-boxplot(data)
+boxplot(data, vert=False)
 xticks(range(0,len(labels)),labels)
 
+show()

### 2014-04-29 examples/dns/server.c
Remove DNS binary Add hashtable

@@ -31,6 +31,33 @@ int domain_cmp(domain *a, labels *b){
         }
         return 0;
 }
+int hashsize(zone *zon){
+        return 3*zon->count/2;
+}
+unsigned long domain_hash(domain *a){ // djb2
+        unsigned long hash = 5381;
+        FOREACH(label, *a){
+                FOREACH(ch,*label){
+                        hash = ((hash<<5) + hash) + *ch; 
+                }
+        }
+        return hash;
+}
+struct hashentry{
+        unsigned long hash;
+        definition *def;
+};
+struct hashentry *zone_hashtable(zone *zon){
+        struct hashentry * table= calloc(sizeof(struct hashentry),hashsize(zon)); //COLLISIONS
+        FOREACH(domain,*zon){
+                unsigned long hash = domain_hash(domain);
+                int entry = hash%hashsize(zon);
+                while(table[entry].def!=NULL) entry = (entry + 1) % hashsize(zon);
+                table[entry].hash = hash;
+                table[entry].def= domain;
+        }
+        return table;
+}
 //responses are a bit hackish - should really put those into the grammar too
 void domain_response(NailArena *tmp_arena,answer *rr, domain *dom){
         labels *l =(labels *)dom;
@@ -52,7 +79,7 @@ int natoi(char *s, int n)
     }
     return x;
 }
-int dns_respond(NailStream *stream,NailArena * arena,struct dnspacket *query, zone *zone)
+int dns_respond(NailStream *stream,NailArena * arena,struct dnspacket *query, zone *zone, struct hashentry *hashtable)
 {
         dnspacket response;// = n_malloc(arena,sizeof(*response));
         char *retval;
@@ -69,8 +96,12 @@ int dns_respond(NailStream *stream,NailArena * arena,struct dnspacket *query, zo
         narray_alloc(response.responses,arena,query->questions.count);
         for(int i=0;i<query->questions.count;i++){
                 question *q =  &response.questions.elem[i];
-                definition *j;// We really should do a hashtable. Oh well.
-                for(j=zone->elem + 0;j<zone->elem + zone->count;j++){
+                unsigned long hash = domain_hash((domain *) &q->labels.labels);
+                int entry = hash%hashsize(zone);
+                for(;hashtable[entry].def != NULL;entry = (entry+1)%hashsize(zone) ){
+                        definition *j = hashtable[entry].def;
+                        if(hashtable[entry].hash != hash)
[…]

### 2014-04-29 paper/osdi14/design.tex
Misc paper changes

@@
 
-Nail grammars describe not only a
+Nail grammars describe both the \emph{external format} and an \emph{internal representation} of a
+protocol. Typical grammar languages describe only the former and leave it up to the programmer to
+construct an internal representation, typically in the form of an abstract syntax tree.
 
-To integrate the data format and the internal representation, Nail
-provides a rich set of combinators that not
-only describe the grammar of the external protocol, but also induce an internal
-model.
+Our grammars are based on Parser Expression Grammars, describing
 
-A central design decision of Nail is that there is a \emph{semantic bijection} between
-the model exposed to the programmer and the byte-level input and output, up to
-syntactic equivalence for unambiguous grammars. More precisely, the parser is
-the generator's inverse, so parsing the generator's output will yield the
-generator's input, but generating the parsers output does not necessarily yield
-the parsers input. To understand why this makes sense, consider a grammar for a
-text language that tolerates white space, or a binary protocol that tolerates
-arbitrarily long padding.\footnote{Say, the physical layer of most communication
-  protocols is a possibly infinite sequence of symbols that are syntactically
-  nil followed by a pre-determined synchronization sequence and the actual
-  contents of the transmission.} In that case, the program semantics should be
-independent of the number of padding elements in the input, and Nail therefore
-does not expose that information to the programmer. We call such discarded
-fields \emph{constants}. Currently, Nail always makes a default choice when
-there are multiple options to express a constant, however Nail could be extended
-to allow a grammar-specific plug-in to make these choices, say for faster
-alignment, streaming applications when data is not ready, or visual appearance in
-human-readable protocols. Similarly, Nail does not preserve the layout of
-objects referred to by their offsets. If the grammar contains no constants and
-offset fields, there is a proper isomorphism between model and protocol.
-
-Nail evaluates the combinators and produces:
+Nail produces the following from a single, descriptive grammmar:
 
@@
 \noindent
-In the rest of this section, we present Nail's combinators.
+In the rest of this section, we present the design of the Nail language.
 
+\subsection{Basics}
+Nail grammars contain parsers
 \begin{figure}[tb]

### 2014-04-29 paper/osdi14/eval.tex
Misc paper changes

@@
 \input{fig-zip}
-In our preliminary evaluation of Nail, we try to answer four questions:
+In our evaluation of Nail, we try to answer four questions:
 
 \begin{itemize}
-\item Can Nail grammars support real-world data formats?
+\item Can Nail grammars support real-world data formats and are Nail's concepts critical to handling
+  them?
 \item How much programmer effort is required to build an
@@
 To answer the first question, we implemented grammars for four protocols with particularly
-challenging features and show that Nail is a good fit for these data formats. In
-\S\ref{s:eval-effort} we will demonstrate that Nail is easier to use than hand-written parsers and
-more convenient than state-of-the-art parser generators for formats that they can handle.
+challenging features and show that Nail is a good fit for these data formats and some aspects of
+these formats would be very hard to parse with other parser generators.
 
@@
 \begin{itemize}
-\item A subset of Ethernet network packets, containing ARP, ICMP and UDP/IP packets. See Figure~\ref{fig:ethernet-extract}.
-\item DNS packets as described in RFC 1035, featuring label compression. See Figure~\ref{fig:dns-full}.
-\item ZIP files, with compression, check-sums and a complicated procedure to identify the header.
+\item DNS messages.
+\item Ethernet packets containing ARP, ICMP and UDP.
+\item ZIP files.
   See Figure~\ref{fig:zip-extract}.
@@
 
-one for  DNS queries and respones, as per RFC1035 (shown in, one for the ZIP
-archive format and a grammar for Ethernet frames containing ARP, ICMP and UDP packets. We also
-parse Nail grammars with Nail, demonstrating that our system can also handle text-based protocols
-and languages.
+\paragraph{DNS.}
+Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
+corresponds to the diagrams in Section 4 of that document. Each DNS packet consists of a header, a
+set of question records and a set of answer records. Domain names in both queries and answers are
+encoded as a sequence of Pascal-style strings. Each label is preceded by a length field, and the
+domain name is terminated by a $0$ octet. The client's query packet contains one question
+record to the server for each DNS look-up it wants to perform. In addition to the appropriate
+response records, the server's reply contains a duplicate of the client's question section, which
+allows stateless DNS clients and caches.
 
-The results suggest that Nail is a good fit for these data formats.
+To reduce the size overhead of including each domain name twice in a DNS reply, once in the
+question section, at least once in the response record, DNS includes a label compression scheme. If
+a domain name suffix is repeated, instead of repeating that suffix, the DNS packet may write a
+two-bit marker sequence followed by a 14-bit offset into the packet where that suffix was previously
+encoded.
 
-\subsection{Programmer effort.}
+Processing these encodings in existing tools, such as Bison or Hammer, would at best be very
+awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
+Keeping track of the position all recognised labels would not be enough, as the offset field may
+refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
+server used as a tutorial example for Hammer does not support compression.
 
-%code size unzip 6.0
-%2821 lines
-\label{s:eval-effort}
-To answer the second question, we implemented small example applications based on the above
-grammars and compared code size with comparable applications that process data manually. We will
-also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
-example for the Hammer parser generator. All code size measurements for C were performed by
-SlocCount\cite{sloccount}. \XXX[code size is primitive, but we show orders of magnitude in the same language].
+Nail handles compression with a stream transform. When parsing, this transform decompresses the DNS
+label stream by following the offset pointers. When generating, this transform receives the current
+suffix as an input and can scan the packet so far for previous occurrences.
 
-%%DNS: 
-\paragraph{DNS.}
+\paragraph{Ethernet}
+\XXX[TODO]
+\paragraph{ZIP Files.}
 
-The first application we built with Nail, is a toy DNS server inspired by the example shipped with
-Hammers DNS server consists of 683 lines of C, excluding the Hammer framework itself, that responds to
-any valid DNS query with a CNAME record to the domain ``spargelze.it''.
-Most of this code is taken up with custom validators, semantic actions,
-and data structure definitions, with only 52 lines of code defining the
-grammar with Hammer's combinators.
-
-Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
-grammar and 64 lines of C implementing DNS label compression. \XXX[Make this smaller]
-The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
-(16 lines) supporting A, NS, MX and CNAME records.
-
-Our simple DNS server, which parses a zone file, listens to incoming DNS requests, parses them and
-generates appropriate responses with this Nail grammar, is implemented in 149 lines of C.
-
- The same grammar is used, together with 98 lines of C,
-to implement a functional toy clone of the \texttt{host} command-line
-tool.
-
-It is hard to compare the programming effort required to implement
-our toy DNS server to that of a real world DNS server, since
-we have less , in particular the additional hint records that real-world DNS servers send and a
-complete lack of configuration options.
-However, the closest in functionality and intent is Dan Bernstein's
-djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
-a minimalist, highly secure DNS server. The latest release of djbdns,
-including various support tools, is about 10,000 lines of C as measured by
-\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
-features we don't have]
-
-\paragraph{ZIP.}
-
-Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}. The ZIP
-format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
-security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
-Android security has been completely by-passed multiple times through parsing differentials in
-handling ZIP files\cite{saurik-masterkey}. \XXX[mention talk by Julia Wolf about zip].
-
+Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
 ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
@@
 
-\XXX[condense]
+Furthermore, Nail transparently handles ZIP's data compression and checksums with stream
+transformations. The implementation of one of these transforms are shown in
+Figure~\ref{fig:zip-compress}. For formats such as Office documents or Android applications that are based on ZIP
+containers, our ZIP grammar could be modified by including a more concrete grammar for file
+contents, which currently are just an array of bytes.
+
+With existing parser languages, the grammar can not guarantee correctness of checksums, length
+fields or offsets, which places additional verification burden on the programmer. Furthermore,
+existing parser generators cannot decompress data as they are parsing. This is merely inconvenient
+for writing a grammar for plain ZIP files, but this prevents the grammar from validating the data
+within a ZIP file.
+
+The results suggest that Nail is a good fit for these data formats.
+
+\subsection{Programmer effort.}
+
+%code size unzip 6.0
+%2821 lines
+\label{s:eval-effort}
+To answer the second question, we implemented small example applications based on the above
+grammars and compared code size with comparable applications that process data manually. We will
+also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
+example for the Hammer parser generator. All code size measurements for C were performed by
+SlocCount\cite{sloccount}. \XXX[code size is primitive, but we show orders of magnitude in the same language].
+
+%%DNS: 
+\paragraph{DNS.}
+
+Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
+grammar and 64 lines of C implementing DNS label compression. \XXX[Make this smaller]
+The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
+(16 lines) supporting A, NS, MX and CNAME records.
+
+With this grammar, we built a simple authoritative DNS server, which parses a zone file, listens to incoming DNS
+requests, parses them and generates appropriate responses, is implemented in 183 lines of C. The
+same grammar is used, together with 98 lines of C, to implement a minimal clone of the
+\texttt{host} command-line tool. Most of this code consists of a custom hashtable and system
+interface code, such as listening to sockets and reading the zonefile.
+
+The Hammer parser framework\cite{hammer-dns} ships with a toy DNS server that responds to
+any valid DNS query with a CNAME record to the domain ``spargelze.it''. 
+The server consists of 683 lines of C, mostly custom validators, semantic actions,
+and data structure definitions, with only 52 lines of code defining the
+grammar with Hammer's combinators.
+
+It is hard to compare the programming effort required to implement
+our DNS server to that of a real world DNS server, since we implement less functionality.
+ In particular, we do not send the additional glue records real-world DNS servers send and
+we did not implement any configuration options.
+
+However, the closest in functionality and intent is Dan Bernstein's
+djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
+a minimalist, highly secure DNS server. The latest release of djbdns,
+including various support tools, is about 10,000 lines of C as measured by
+\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
+features we don't have]
+
+\paragraph{ZIP.}
 
@@
 stream transforms, one for the DEFLATE compression algorithm with the help of the zlib library and
-one for finding the end-of-directory header. This grammar avoids the two pitfalls mentioned above:
-Nail forces the programmer to specify an explicit procedure for finding the end-of-directory record
-and we use dependent fields to enforce that the duplicated information is consistent.
+one for finding the end-of-directory header.
 
@@
 \paragraph{Network stack.}
-Write up this stack.
+\XXX[Write this up]
 \subsection{Security.} 
+ \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
+security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
+Android security has been completely by-passed multiple times through parsing differentials in
+handling ZIP files\cite{saurik-masterkey}. mention talk by Julia Wolf about zip]. 
 Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.
@@
 \subsection{Performance.}
+\input{fig-perf-dns}
+To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
+benchmarked our DNS server against BIND \XXX[and our network stack against \texttt{lwip-tap}, a userspace
+router implemented with the LWIP network stack\cite{lwip}].
 
-Compare DNS requests per second of BIND, DJBDNS and NailDNS on a small zonefile.
-Compare extraction performance for ZIP.
+\paragraph{DNS}
+We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind8}, one of
+the oldest and most popular DNS daemons, on a load that resembles the authoritative name server at a
+hosting company. First, we generated domain names consisting of one or
+two labels randomly selected from an English dictionary and one label that is one of three popular
+Top Level Domains: ``com'', ``net'' or ``org''. Second, we randomly selected 90\% of these domains and
+created a zone file that mapped these domain names to the local IP address.
+
+Finally, we used the \texttt{queryperf} tool provided with BIND to query each domain between zero
+and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
+core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
+queries outstanding at once. The queryperf tool was configured to
+repeat the same, randomised sequence of queries for one minute and the average throughput and
+response latency were measured.  We repeated each test seven times  with 50.000 and one million
+domain names, restarting each daemon in between. We also performed one initial dry run to warm the
+file system cache for the zone file.
+\XXX[Add 1 million results.They are largely the same]
+
+Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
+of optimisation by a factor of 3. \XXX[Why the hell?] This demonstrates that even though Nail was
+not developed focusing on performance, applications built on Nail can compete with real-world
+hand-written parsers.

### 2014-04-29 paper/osdi14/intro.tex
Misc paper changes

--- 2014-04-03 paper/intro.tex
@@
 over the inputs to that code, and can craft inputs that trigger
-subtle corner cases in input processing.  For example, the libpng
+subtle corner cases in input processing.  For example, the libpng image decompression
 library has had 24 remotely exploitable vulnerabilities from 2007 to
@@
 
-Third, data formats, especially binary formats such as PNG or PDF, can have \emph{structural dependencies}, such as offset, length, and checksum fields
-that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
-manually write control code, such as re-positioning the parser's input stream, looping over the
-invocation of a sub-parser, or computing a
-checksum over raw input bytes.  Besides leaving much room for errors with offset arithmetic, such
-code is usually not reusable when generating output.
+Third, data formats, for example PNG or PDF, contain features that are hard to represent with
+existing parser generators. Some features, such as length and checksum fields, require the parser to
+keep additional state. More complicated formats incorporate offset fields that require the parser to
+reposition its input stream or consume the same input multiple times. In some cases, such as with
+compressed data, very complicated transformations on the external format are necessary before it can
+be understood. State-of-the-Art parser generators provide at most limited, unelegant facilities for
+such features, so programmers are forced to manually writing control code to handle them, such as
+moving the parsers input stream, looping over the invocation of a sub-parser or computing a checksum
+over raw input bytes. Besides leaving much room for errors with offset arithmetic, such code is
+usually not reusable when generating output.
+
+\XXX[Motivate Stream transforms better]
 
@@
 Second, this well-defined internal representation allows Nail to invert the
-parser and generate output. However, Nail allows \emph{constants} in the
+parser and generate output, while maintaining a \emph{semantic bijection}.
+A bijection in the traditional sense often does not make sense when parsing data formats. 
+Consider a grammar for a
+text language that tolerates white space, or a binary protocol that tolerates
+arbitrarily long padding.\footnote{Say, the physical layer of most communication
+  protocols is a possibly infinite sequence of symbols that are syntactically
+  nil followed by a pre-determined synchronization sequence and the actual
+  contents of the transmission.} Program semantics should be
+independent of the number of padding elements in the input, and Nail therefore
+does not expose that information to the programmer. We call such discarded
+fields \emph{constants}. 
+Similarly, Nail does not preserve the layout of
+objects referred to by their offsets. If the grammar consists of a simple protocol without offset
+fields, constants and the like, there is a normal bijection between the internal model and valid
+parser inputs.
+
+\XXX[Currently, Nail always makes a default choice when
+there are multiple options to express a constant, however Nail could be extended
+to allow a grammar-specific plug-in to make these choices, say for faster
+alignment, streaming applications when data is not ready, or visual appearance in
+human-readable protocols.]
+
+ However, Nail allows \emph{constants} in the
 external format to have multiple representations if they should not affect the
@@
 
-Third, Nail provides support for structural dependencies, transparently handling
-offset and length fields. By hiding the offset and length values from the
-programmer, Nail ensures they remain consistent even if the data is changed by
-the application.
+Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{stream transforms}, to
+elegantly handle problematic features of data formats. Dependent fields are fields in the protocol
+whose value depends in some way on the value or layout of other parts of the protocol, such as
+length, offset and checksum fields. \XXX[Stream transforms allow the programmer to write a function
+that produces one or more virtual streams from one or more virtual streams and the values of
+dependent fields.]
 
-%  XXX todo.  Bijections.  Reduced
-% expressiveness.  Constraints.  External interaction (checksum, ports,
-% etc).
-
-We have implemented an initial prototype of Nail for C\@.  Our experience
+We have implemented a prototype of Nail for C\@.  Our experience
 so far suggests that it is a promising approach: we were able to construct
-a succinct grammar for DNS packets, and write a small DNS server that
-uses Nail for all packet input and output, and operates purely on
-Nail-generated data structures, with no need for any semantic actions.
+a succinct grammar for a set of common file formats and protocols that demonstrate Nail's ability to
+handle tricky constructs and build a series of sample applications with these grammars.
 
@@
 Nail's design.  \S\ref{s:impl} discusses our initial implementation
-of Nail.  \S\ref{s:eval} provides some initial evaluation results.
+of Nail.  \S\ref{s:eval} provides evaluation results.
 \S\ref{s:future} suggests several directions for future work, and

### 2014-04-29 paper/osdi14/design.tex
Wrote dependencies

@@
 construct an internal representation, typically in the form of an abstract syntax tree.
-
-Our grammars are based on Parser Expression Grammars, describing
 
@@
 \subsection{Basics}
-Nail grammars contain parsers
+Nail grammars are a set of rules. Each rule assigns a parser, which can be a reference to another
+rule, a trivial parser describing an integer of a certain width or a combinator that combines
+multiple parsers in some way. We provide the most common combinators familiar from other parser
+combinator libraries, such as Parsec\cite{LeijenMeijer:parsec} and Hammer\cite{hammer-parser}.
+However, we extend the sequence and choice combinators normally found in parser combinator libraries
+with \emph{semantic labels} that name fields in the internal models. The syntax described by a Nail
+grammar does not depend on these semantic labels.
+
+\paragraph{Integers and Constraints}
 \begin{figure}[tb]
@@
   & $1\leq x \leq 3$ \\
-\multirow{2}{*}{\texttt{uint16 | ..512}}
+\multirow{2}{*}{\texttt{uint16 | [1,2]}}
   & unsigned 16-bit integer, \\
-  & $x \leq 512$ \\
-\multirow{2}{*}{\texttt{int32 | [1,255,512]}}
+  & $x \in \{1,2\}$ \\
+\multirow{2}{*}{\texttt{int32 | [1,5..255,512]}}
   & signed 32 bit integer, \\
-  & $x \in \{ 1, 255, 512 \}$ \\
+  & $x \in \{ 1, 512 \} \vee 5\leq x\leq 255$ \\
 \bottomrule
@@
 
-\paragraph{Integers and Constraints}
-Nail's fundamental 
-The elementary parsers of Nail are the same as those of Hammer, signed
-and unsigned integers with arbitrary lengths up to 64 bits.  Note that is
-possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
-data offset within the TCP header; in Nail's syntax.  Integer parsers
-return their value in the smallest appropriately sized machine integer
-type; e.g., a 24-bit integer is stored in a 32-bit wide variable.
+Nail's fundamental parsers recognise signed or unsigned with arbitrary lengths up to 64 bits, an
+approach popularised by the Hammer framework.\cite{hammer-parser}  
+Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
+data offset within the TCP header.  In the internal representation, integer values are represented
+by the nearest usable integer size, for example a 24-bit unsigned integer will be stored in an
+unsigned 32-bit double word.
 
-Integer parsers can be constrained to fall either within an (inclusive)
-range of values or be an element of a set of allowed values.
+The grammar can also express a constraint on an integer field. The current version of Nail expresses
+constraints as a set of permissible values or value ranges. Extending the Nail language and
+implementation to support richer constraints languages would be relatively trivial, however we have
+found that the current syntax covers permissible values within existing protocols correctly and
+concisely.
+
 Figure~\ref{fig:range} shows several examples.
-Invalid values or prematurely reaching the end of input will raise an
-error when parsing input or generating output, and abort the procedure.
 
-\paragraph{Sequence.}
+\paragraph{S.}
 
-Nail's fundamental concept is the structure combinator. It contains a list of
-named parsers and unnamed constant parsers. The parser invokes each field 
-in sequence and returns a structure containing the result of each parser.
-For example, Figure~\ref{fig:dns-struct} shows the structure combinator
-from a part of the grammar for DNS packets, along with the data model
-corresponding to it.
+The \emph{many} combinator takes a parser and applies it repeatedly
+until it fails, returning an array of the inner parsers results. The
+\emph{sepBy} combinator
+additionally takes a constant parser, which it applies in between parsing
+two values, but not before parsing the first value or after parsing the
+last.
+For example, \texttt{many uint8} represents an array of 8-bit unsigned
+integers, and \texttt{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
+comma-separated lists of decimal numbers.
+
+\paragraph{Structures.}
 
@@
 \begin{verbatim}
-header = {
+dns_header = {
   id uint16
@@
 \begin{verbatim}
-struct header {
+struct dns_header {
   uint16_t id;
@@
 
-\paragraph{Repetition.}
+Nail provides a structure combinator with semantic labels instead of the sequence combinator that
+other parser combinator libraries use to capture structures in data formats. 
+The structure combinator consists of a sequence of fields, typically consisting of a label and a
+parser that describes the contents of that field. Other field types will be described below.
+ The syntax of the structure combinator is inspired by the Go language\cite{golang}, with field names preceding their definition.
+In our opinion, this syntax is easier to read than the widespread C-style structure and type
+declarations.
 
-The \emph{many} combinator takes a parser and applies it repeatedly
-until it fails, returning an array of the inner parsers results. The
-\emph{sepBy} combinator
-additionally takes a constant parser, which it applies in between parsing
-two values, but not before parsing the first value or after parsing the
-last.
-For example, \texttt{many uint8} represents an array of 8-bit unsigned
-integers, and \texttt{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
-comma-separated lists of decimal numbers.
+A sample grammar for a DNS header and the corresponding C data type the Nail parser produces, are
+shown in Figure~\ref{fig:dns-struct}.
 
-\paragraph{Semantic choice.}
+\paragraph{Constants.}
+In some cases, not all bytes in a structure actually contain information, such as magic numbers or
+reserved fields. Those fields can be represented in Nail grammars by constant fields in structures.
+Constant fields do not correspond to a field in the internal model, but they are validated during
+parsing and generated during output.  Constants can either have integer values, such as
+\texttt{uint3=0} or string values for text-based protocols, e.g. \texttt{many uint8 = ``Foo''}.
 
-We extend a parsing expression grammar's ordered choice combinator with
-a tag for each choice. The parser
-attempts to parse each option in the order they are specified in the field and
-stores the result in a tagged union. If an option fails, the parser backtracks
-to the beginning of the choice combinator's input. Care must be taken that the
-choices do not overlap, because Nail always picks the first successfully
-parsed choice. If two options overlap, generated output for the latter option is
-not necessarily understood identically by the parser. However, actual data
-formats are usually not ambiguous in this sense.
-Figure~\ref{fig:choice} demonstrates a simple choice combinator.
+In some protocols, there might be many ways to represent the same constant field
+and there is no semantic difference between the different syntactic representations.
+Nail therefore allows repeated constants, such as \texttt{many (uint8=' ')}, which parses any number of space characters, or
+\texttt{|| uint8 = 0x90 || uint16 = 0x1F0F}, which parses two of the many representations for X86
+NOP instructions, which are used as padding between basic blocks in an executable.
+
+As discussed above, choosing to use these combinators on constant parsers
+removes the bijection between byte-strings and our data model, as there are
+multiple byte-strings that correspond to the same internal data structure and
+the generator has to choose one of these representations.
+\paragraph{Wrap combinator.} 
+When implementing real protocols with Nail, we often found that
+structures that consist of many constant parsers and only one named field. This pattern is
+common in binary protocols which use fixed headers to denote the type of data
+structure to be parsed.  In order to keep the internal representation cleaner,
+we introduced the wrap combinator, which takes a sequence of parsers containing
+exactly one non-constant parser. The parser and generator act as though the wrap
+combinator was a sequence of parsers, but the data model does not wrap the
+single value in another structure, making the application-visible representation
+(and thus application code) more concise.
+
+For example, \texttt{word= <uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
+lowercase word into an array.
+
+\paragraph{References.}
+To support recursive structures, we include a reference combinator \texttt{*}, which is
+syntactically equivalent to directly including the parser, but in the internal model adds a level of
+indirection to the value. This allows recursive structures to be parsed, for example, balanced
+parentheses could be parsed as follows: \texttt{paren = < uint8 ="("; optional *paren ;uint8=")">}
+
+\paragraph{Choices.}
+Our choice combinator takes multiple labelled parsers and tries each of them in turn, backtracking
+to the initial position if an option fails. This behaviour for the choice combinator, the defining
+feature of Parser Expression Grammars\cite{ford2002packrat}, resolves any ambiguities explicitly.
+Internally, the choice combinator represents choices as tagged unions.
+
+If two options the same string could be parser by two options, generated output for the latter
+option is not necessarily understood identically by the parser. However, actual data formats are
+usually not ambiguous in this sense. Figure~\ref{fig:choice} demonstrates a simple choice
+combinator.
+
+Nail also includes an \texttt{optional} combinator, which attempts to recognise a value, but succeeds
+without consuming input when it cannot recognise that value. Syntactically, \texttt{optional} is
+equivalent to a choice between the parser and an empty structure, but in the internal model it is
+more concisely represented as a reference that is null when the parser fails.
 
@@
 
-\subsection{Constant parsers}
+\subsection{Dependent fields}
 
-Nail also features \emph{constant parsers}, which do not affect the internal
-representation. Constant parsers can appear instead of structure fields and as
-separators in the sepBy combinator.
+Data formats often contain values that are determined by other values or the layout of information,
+such as checksums, duplicated information  or offset and  length fields.
+We represent such as values with \emph{dependent fields} and handle them transparently during
+parsing and generation without exposing them to the internal model.
 
-The simplest constant parser is an integer or array of integers with fixed
-value; for example, \texttt{uint8=0}, or \texttt{many uint8=[1,2]}. A convenience notation
-for strings is also supported: \texttt{many uint8 = "foo"}.
+Dependent fields are defined within a structure like normal fields, but their name starts with an \@-sign. 
+A dependent field is in scope and can be referred to by the definition of all subsequent fields in the same structure.
 
-\paragraph{Wrap combinator.} When implementing real protocols with Nail, we often found that
-structures that consist of many constant parsers and only one named field. This pattern is
-common in binary protocols which use fixed headers to denote the type of data
-structure to be parsed.  In order to keep the internal representation cleaner,
-we introduced the wrap combinator, which takes a sequence of parsers containing
-exactly one non-constant parser. The parser and generator act as though the wrap
-combinator was a sequence of parsers, but the data model does not wrap the
-single value in another structure, making the application-visible representation
-(and thus application code) more concise.
+During parsing, dependent fields are parsed just like other fields, but instead of storing their
+value in a structure that will be returned to the programmer, the parser stores the value in a
+temporary location. Thereafter, the value can be referenced in the grammar.
 
-For example, \texttt{<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
-lowercase word into an array, excluding the quotation marks.
+When generating output, Nail visits a dependent field twice. When generating the fields of a
+structure, it will reserve space for the dependent field in the output, but not write a value. When
+the dependent field goes out of scope, the generator will backtrack to this space and write the
+dependent fields value.
 
-\paragraph{Constant combinators.}
-In some protocols, there might be many ways to represent the same constant field
-and there is no semantic difference between the different syntactic representations.
-To support this pattern, Nail allows developers to use choice and repetition
-combinators together with constant fields, such as
-\texttt{many (uint8=' ')} (representing any number of space characters), or
-\texttt{|| uint8 = 0x90 || uint16 = 0x1F0F} (parsing two of the many representations of NOP on the
-x86 architecture). Note that constant may have varying lengths.
-This is particularly useful for handling padding or whitespace.
+Nail only provides one built-in combinator, \texttt{n_of} that uses dependency fields, which is used
+for count fields. For example, DNS labels are encoded as Pascal-style strings - a length followed by
+the value of the label. The Nail grammar for DNS packets in Figure~\ref{fig:dns-full} represents
+this pattern as follows:
 
-As discussed above, choosing to use these combinators on constant parsers
-removes the bijection between byte-strings and our data model, as there are
-multiple byte-strings that correspond to the same internal data structure and
-the generator has to choose one of these representations. As such, constant
-combinators are the generator dual of ambiguous choice combinators in the
-parser, because they lead to ambiguities in the generator.
+\begin{figure}[tb]
+\begin{verbatim}
+{ @length uint8 | 1..64
+  label n_of @length uint8 }
+\end{verbatim}
+\caption{Nail grammar for uncompressed DNS labels.}
+\end{figure}
 
-\subsection{Dependent fields}
-Another problem for parser generators is that binary protocols often contain
-length and offset fields. Conventional parsing algorithms can, in principle,
-deal with
-bounded offset fields: a finite automaton can count a bounded integer, and we
-can feed the (finite) input multiple times to the finite automaton. However,
-this implementation is both time-inefficient (it feeds many bytes into the
-automaton that will just be skipped) and very cumbersome to express with
-current parser generators. Therefore, if languages with offset fields need
-to be parsed
-with parser generators, the only currently practical way is to add ad-hoc hacks
-such as changing the input pointer of the generated parser on the fly, as part
-of the code in the semantic action for the offset field.
+Other dependencies, such as offset fields or checksums are not handled directly by combinators, but
+through stream transformation.
 
-Nail will properly support both offset and length fields and much of the
-following discussion applies to both, although the current prototype only
-implements lengths, which we will focus on.
+\subsection{Input streams and Transformations.}
 
-We call length or offset fields \textit{dependent fields}, because during
-parsing, another parser depends on them, and while generating output, their
-value depends on some other structure in the data model. Dependency fields
-appear in a structure combinator as would any other integer field, but
-their name begins with an \texttt{@} sign.  A dependency field has to
-appear in the grammar before it can be used.
+% Data formats often contain redundant information, such as checksums or duplicated values.
+% Conventional parsers expose all this information and validate it manually, which can lead to bugs.
 
-Dependency fields are not exposed in the data model, but instead are
-transparently computed.  This frees the developer from checking that
-these fields are correct (for input) or having to keep their values in
-sync with the rest of the data structure (for output).
+% Another problem for interface generators is that binary protocols often contain length and offset
+% fields. Conventional parsing algorithms can, in principle, deal with bounded offset fields - a finite
+% automaton can count a bounded integer, by introducing a state for each value. However, such grammars
+% are very large, making the state-machine implementation inefficient and the grammar very complicated.
 
-\paragraph{Length fields.}
+% Nail will properly support both offset and length fields and much of the
+% following discussion applies to both, although the current prototype only
+% implements lengths, which we will focus on.
 
-The length combinator, \texttt{n_of}, takes a dependency field $n$
-and a parser, evaluates the parser exactly $n$ times (i.e., setting
-the number of iterations to be the $n$ field's value), and returns
-an array of the parser's values. When generating output, it emits
-the array and writes its length to the dependency field.
+% We call length or offset fields \textit{dependent fields}, because during
+% parsing, another parser depends on them, and while generating output, their
+% value depends on some other structure in the data model. Dependency fields
+% appear in a structure combinator as would any other integer field, but
+% their name begins with an \texttt{@} sign.  A dependency field has to
+% appear in the grammar before it can be used.
 
-\paragraph{Offsets.}
+% Dependency fields are not exposed in the data model, but instead are
+% transparently computed.  This frees the developer from checking that
+% these fields are correct (for input) or having to keep their values in
+% sync with the rest of the data structure (for output).
 
-The \texttt{offset} combinator takes a dependency field and a parser. It
-moves the parser to the position specified in the offset field and
-invokes the inner parser, and then moves the input back to its original
-position.  While generating output, all structures referred to by offset
-are generated after the main structure and the dependent offset fields
-are patched up.
+% \paragraph{Length fields.}
 
-\paragraph{Checksums.}
+% The length combinator, \texttt{n_of}, takes a dependency field $n$
+% and a parser, evaluates the parser exactly $n$ times (i.e., setting
+% the number of iterations to be the $n$ field's value), and returns
+% an array of the parser's values. When generating output, it emits
+% the array and writes its length to the dependency field.
 
-In many data formats, some values depend on external representation,
-such as checksums and cryptographic signatures.  While it would be
-possible to extend our constraint language to be powerful enough to
-support such constructs, we would essentially be building a separate,
-Turing-complete language that has all the same pitfalls existing programs
-have.  Therefore, we intend to allow the programmer to carefully escape
-Nail's programming model and write a function that takes a pointer to
-the dependent value and a range of bytes, using the \texttt{raw_depend}
-combinator.
+% \paragraph{Offsets.}
 
-For example, we imagine the following grammar could be used to represent
-a sequence of bytes \texttt{data} followed by its CRC32 checksum:
+% The \texttt{offset} combinator takes a dependency field and a parser. It
+% moves the parser to the position specified in the offset field and
+% invokes the inner parser, and then moves the input back to its original
+% position.  While generating output, all structures referred to by offset
+% are generated after the main structure and the dependent offset fields
+% are patched up.
 
-\begin{verbatim}
-data many uint8; @checksum uint32
-raw_depend @checksum data crc32
-\end{verbatim}
+% \paragraph{Checksums.}
 
-\noindent
-where \texttt{crc32} is a function supplied by the application, with
-the following signature:
+% In many data formats, some values depend on external representation,
+% such as checksums and cryptographic signatures.  While it would be
+% possible to extend our constraint language to be powerful enough to
+% support such constructs, we would essentially be building a separate,
+% Turing-complete language that has all the same pitfalls existing programs
+% have.  Therefore, we intend to allow the programmer to carefully escape
+% Nail's programming model and write a function that takes a pointer to
+% the dependent value and a range of bytes, using the \texttt{raw_depend}
+% combinator.
 
-\begin{verbatim}
-bool crc32(uint32_t *out, uint8_t *in);
-\end{verbatim}
+% For example, we imagine the following grammar could be used to represent
+% a sequence of bytes \texttt{data} followed by its CRC32 checksum:
 
-Because this feature compromises Nail's security guarantees, it should
-only be used in limited circumstances and with carefully prepared checksum
-functions.  This feature is not implemented in the current prototype.
+% \begin{verbatim}
+% data many uint8; @checksum uint32
+% raw_depend @checksum data crc32
+% \end{verbatim}
+
+% \noindent
+% where \texttt{crc32} is a function supplied by the application, with
+% the following signature:
+
+% \begin{verbatim}
+% bool crc32(uint32_t *out, uint8_t *in);
+% \end{verbatim}
+
+% Because this feature compromises Nail's security guarantees, it should
+% only be used in limited circumstances and with carefully prepared checksum
+% functions.  This feature is not implemented in the current prototype.

### 2014-04-29 paper/osdi14/impl.tex
Wrote dependencies

--- 2014-04-02 paper/impl.tex
@@
 language and top-down parsers. Options for C++ STL data models and emitting
-Packrat parser~\cite{packrat-parsing:icfp02} are under development. In
+Packrat parsers~\cite{packrat-parsing:icfp02} are under development. In
 this section, we will discuss some particular features of our parser
@@
 model. Currently the parser is a straightforward top-down parser, although
-facilities have been made to add Packrat parsing to achieve linear parsing
+preparations have been made to add Packrat parsing to achieve linear parsing
 times.
@@
 
-Many security vulnerabilities can occur when heap allocations are done
-improperly. Therefore, just like Hammer, Nail avoids using the heap as much as
-possible, using a custom arena allocator and allocating only fixed-size blocks
-from the system allocator (malloc). However, Nail extends upon Hammer's approach and uses two
-arenas for each parsed input. One arena is used for intermediate results and is
-freed (and zeroed) after parsing completes, whereas the other arena is used only
-for allocating the result, and has to be freed by the programmer.
+Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
+like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
+allocating only fixed-size blocks from the system allocator (malloc). However, Nail extends upon
+Hammer's approach and uses two arenas for each parsed input. One arena is used for intermediate
+structures, including dependency fields and any temporary streams, and is released (and zeroed)
+after parsing completes, whereas the other arena is used only for allocating the result, and has to
+be freed by the programmer.
 
@@
 
-Most parser generators, such as Bison, do not have to dynamically allocate
-temporary data, as they evaluate a semantic action on every rule. However, as
-our goal is to perform as little computation as possible before the input
-has been validated, and we do not want to mix temporary objects with the
-results of our parse, we use an append-only trace to store intermediate parser
+Most parser generators, such as Bison, do not have to dynamically allocate temporary data on the
+heap, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
+computation as possible before the input has been validated, and we do not want to mix temporary
+objects with the results of our parse, we use an append-only trace to store intermediate parser
 results.
@@
 
-\paragraph{Dependency fields.}
+% \paragraph{.}
 
-During parsing, dependency fields occur before the context in which they are
-used. The parser stores their values and retrieves them afterwards when
-encountering the combinator that uses them. When generating output, the
-dependency field is first filled with a filler value, then later when the first
-combinator that determines this fields value is encountered, the field is
-overwritten. Any further combinators using this dependency will then validate
-that the dependency field is correct.
+% During parsing, dependency fields occur before the context in which they are
+% used. The parser stores their values and retrieves them afterwards when
+% encountering the combinator that uses them. When generating output, the
+% dependency field is first filled with a filler value, then later when the first
+% combinator that determines this fields value is encountered, the field is
+% overwritten. Any further combinators using this dependency will then validate
+% that the dependency field is correct.
 

### 2014-04-29 paper/osdi14/intro.tex
Wrote dependencies

@@
 
-Third, data formats, for example PNG or PDF, contain features that are hard to represent with
+Third, many data formats contain redundancies, such as repeating information in multiple structures.
+Applications usually do not explicitly check for consistency, and if different applications use
+different instances of the same value, an attacker can craft input that causes inconsistencies.
+Furthermore, security vulnerability can occur when an application assumes two repetitions of the
+same data to be consistent.
+
+Finally, data formats, for example PNG or PDF, contain features that are hard to represent with
 existing parser generators. Some features, such as length and checksum fields, require the parser to
@@
 
-\XXX[Motivate Stream transforms better]
-
 This paper presents the initial design and implementation of Nail, a parser
 generator that greatly reduces the programmer effort required to use a
-grammar-based parser. Nail addresses the above three challenges with several key
+grammar-based parser. Nail addresses the above four challenges with several key
 ideas.
@@
 
-\XXX[Currently, Nail always makes a default choice when
-there are multiple options to express a constant, however Nail could be extended
-to allow a grammar-specific plug-in to make these choices, say for faster
-alignment, streaming applications when data is not ready, or visual appearance in
-human-readable protocols.]
+% \XXX[Currently, Nail always makes a default choice when
+% there are multiple options to express a constant, however Nail could be extended
+% to allow a grammar-specific plug-in to make these choices, say for faster
+% alignment, streaming applications when data is not ready, or visual appearance in
+% human-readable protocols.]
 
- However, Nail allows \emph{constants} in the
-external format to have multiple representations if they should not affect the
-semantics of the data. For example, in a text protocol, the amount of
-white-space separating tokens should not affect the meaning of the data and
-consequently Nail does not expose it to the application. As long as
-constants are only used for their intended purpose of representing syntax-only
-features, the generated output will have the same semantics as the parsed input.
+%  However, Nail allows \emph{constants} in the
+% external format to have multiple representations if they should not affect the
+% semantics of the data. For example, in a text protocol, the amount of
+% white-space separating tokens should not affect the meaning of the data and
+% consequently Nail does not expose it to the application. As long as
+% constants are only used for their intended purpose of representing syntax-only
+% features, the generated output will have the same semantics as the parsed input.
 
-Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{stream transforms}, to
+Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{structural transformations}, to
 elegantly handle problematic features of data formats. Dependent fields are fields in the protocol
 whose value depends in some way on the value or layout of other parts of the protocol, such as
-length, offset and checksum fields. \XXX[Stream transforms allow the programmer to write a function
-that produces one or more virtual streams from one or more virtual streams and the values of
-dependent fields.]
+length, offset and checksum fields. Structural transformations allow the programmer to write a pair
+of functions that transforms the \XXX[TODO]
 

### 2014-04-30 paper/osdi14/eval.tex
minor stuff

@@
 \label{s:eval}
+
 \input{fig-ethernet}
@@
 \input{fig-zip}
-In our evaluation of Nail, we try to answer four questions:
+
+Our evaluation of Nail answers four questions:
 
 \begin{itemize}
-\item Can Nail grammars support real-world data formats and are Nail's concepts critical to handling
-  them?
+
+\item Can Nail grammars support real-world data formats, and
+      are Nail's techniques critical to handling these formats?
+
 \item How much programmer effort is required to build an
       application that uses Nail for data input and output?
-\item Do applications handling input and output with Nail suffer from less security vulnerabilities?
-\item How fast are systems using Nail parsers as opposed to hand-written input handling? 
+
+\item Does using Nail for handling input and output improve
+      application security?
+
+\item Does Nail achieve acceptable performance?
+
 \end{itemize}
 
-\subsection{Data formats.}
+\subsection{Data formats}
 \label{s:eval-formats}
@@
 
-\subsection{Programmer effort.}
+\subsection{Programmer effort}
 
@@
 \XXX[Write this up]
-\subsection{Security.} 
+\subsection{Security} 
  \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
@@
 
-\subsection{Performance.}
+\subsection{Performance}
 \input{fig-perf-dns}

### 2014-04-30 paper/osdi14/design.tex
Design

@@
 
-\subsection{Overview}
-
+\subsection{Goals}
+\paragraph{Internal Model.}
 Nail grammars describe both the \emph{external format} and an \emph{internal representation} of a
-protocol. Typical grammar languages describe only the former and leave it up to the programmer to
+protocol, as opposed to existing grammar languages describe only the former and leave it up to the programmer to
 construct an internal representation, typically in the form of an abstract syntax tree.
@@
 Nail produces the following from a single, descriptive grammmar: 
-
 \begin{itemize}
@@
 \noindent
-In the rest of this section, we present the design of the Nail language.
 
+\paragraph{Semantic Bijecction.}
+Second, this well-defined internal representation allows Nail to invert the
+parser and generate output, while maintaining a \emph{semantic bijection}, that is, when generating
+the corresponding output to a parser input, the two byte streams will have the same meaning, i.e. be
+interpreted the same by Nail, but might not be identical.
+A bijection in the traditional sense often does not make sense when parsing data formats. 
+Consider a grammar for a text language that tolerates white space, or a binary protocol that tolerates
+arbitrarily long padding.\footnote{Say, the physical layer of most communication
+  protocols is a possibly infinite sequence of symbols that are syntactically
+  nil followed by a pre-determined synchronization sequence and the actual
+  contents of the transmission.} Program semantics should be
+independent of the number of padding elements in the input, and Nail therefore
+does not expose that information to the programmer. We call such discarded
+fields \emph{constants}. 
+Similarly, Nail does not preserve the layout of
+objects referred to by their offsets. If the grammar consists of a simple protocol without offset
+fields, constants and the like, there is a normal bijection between the internal model and valid
+parser inputs.
+\paragraph{Hide redundant information.}
+Nail's internal model is designed to hide unneeded information from the application. The more
+information Nail automatically deduces, the less the application can produce invalid data.
+Therefore, Nail introduces \emph{dependent fields}, which contain data that can be computed during
+generation and needs to be kept as additional state during parsing.
+
+\paragraph{Allow parser extension.}
+Real-world protocols contain complicated ways of encoding data. Representing these in an
+intentionally limited model such as our parser language is complicated. Therefore, Nail introduces
+\emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
+generator. Nail parsers and generators interact with data through an abstract stream, which allows
+reading and writing of bits as re-positioning. Transformations allow the programmer to write
+functions in a general purpose language that consume streams and define new temporary streams, while
+also reading or writing the value of dependent fields.
+
+Initial versions of Nail included an offset combinator, which consumed a dependent field and applied
+a parser at the offset specified therein. However, it proved impossible to foresee all the ways in which a protocol could encode an offset, for example
+some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
+end of the file or a fixed offset. In nested grammars, offsets are also not necessarily from
+the beginning of a file or packet.
+
+We will introduce transformations, which allow the programmer to write arbitrary functions that can
+handle such structures and streams, which are a generic abstraction for input and output data that
+allow the decoded data to be integrated with the rest of the generated Nail parser.
+
+\XXX[In the rest of this section, we first present the design of the Nail language and then show how
+Nail's features match these goals.]
 \subsection{Basics}
@@
 Dependent fields are defined within a structure like normal fields, but their name starts with an \@-sign. 
-A dependent field is in scope and can be referred to by the definition of all subsequent fields in the same structure.
+A dependent field is in scope and can be referred to by the definition of all subsequent fields in
+the same structure. Dependency fields can be passed to rule invocations as parameters.
 
@@
 
-% Data formats often contain redundant information, such as checksums or duplicated values.
-% Conventional parsers expose all this information and validate it manually, which can lead to bugs.
+In order to handle complicated input functions, Nail allows the programmer to
+break out of Nail's generated code in a controlled fashion, by manipulating dependent fields and
+streams. 
+By default, Nail implements streams with in-memory buffers, but the programmer can override this implementation.
+While parsing or generating a Nail grammar, there is always an implicit \emph{current stream} and
+referred to as \texttt{\$current} in the Nail language.
 
-% Another problem for interface generators is that binary protocols often contain length and offset
-% fields. Conventional parsing algorithms can, in principle, deal with bounded offset fields - a finite
-% automaton can count a bounded integer, by introducing a state for each value. However, such grammars
-% are very large, making the state-machine implementation inefficient and the grammar very complicated.
+The programmer can define additional temporary streams by writing a \emph{transformation}, a pair of
+functions, one for parsing and one for generating output. Temporary streams follow the naming and
+scoping convention of dependent fields, but their names start with a \$-sign.
 
-% Nail will properly support both offset and length fields and much of the
-% following discussion applies to both, although the current prototype only
-% implements lengths, which we will focus on.
+The parsing function of the transform takes any number of streams, either temporary or the current,
+and outputs any number of temporary streams. The parsing function may reposition and read from the
+input streams and read the values of dependent fields, but not change their values. The generating
+function has to be an inverse of the parsing function. It takes the same set of temporary streams
+the parsing function returned and is supposed to change the values of the same set of values the
+generating function produced.
 
-% We call length or offset fields \textit{dependent fields}, because during
-% parsing, another parser depends on them, and while generating output, their
-% value depends on some other structure in the data model. Dependency fields
-% appear in a structure combinator as would any other integer field, but
-% their name begins with an \texttt{@} sign.  A dependency field has to
-% appear in the grammar before it can be used.
+The programmer has to guarantee memory safety and unambiguity of the stream transformations. To
+reduce both programmer effort and the risk of unsafe operations, we provide implementations of
+transformations for many common features, such as checksums and size and offset fields. Furthermore,
+we provide library functions that can be used to safely operate on streams, such as splitting and
+concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
+efficiently duplicated and split.
 
-% Dependency fields are not exposed in the data model, but instead are
-% transparently computed.  This frees the developer from checking that
-% these fields are correct (for input) or having to keep their values in
-% sync with the rest of the data structure (for output).
+Transformations can handle a wide variety of patterns in data formats, including the following: 
+\paragraph{Offsets.}
+We provide a transform for handling offset fields that is invoked as follows:
+\texttt{$fragment transform offset\_u32($file_stream, @offset)}. 
+They are implemented by two  functions:
+\begin{figure}[tb]
+\begin{verbatim}
+int offset_u32_parse(NailStream *out_fragment,NailStream *in_current,const uint32_t *off){
+   out_fragment = suffix of in_current at offset off
+}
+int offset_u32_generate(NailStream *in_fragment, NailStream *out_current,uint32_t *off){
+   off = position of out_current
+   append in_fragment to out_current
+}
+\end{verbatim}
+\caption{Pseudocode for the two C functions implementing an offset transformation.}
+\end{figure}
 
-% \paragraph{Length fields.}
+\paragraph{Sizes.}
+A similar transform handles size fields. Just like the offset transform, it takes two parameters, a
+stream and a dependent field, but instead of returning the suffix of the current stream after an
+offset, it returns a slice of the given size from the current stream starting at its current
+position. When generating, it appends the fragment stream to the current stream and writes the size
+of the fragment to the dependent field.
 
-% The length combinator, \texttt{n_of}, takes a dependency field $n$
-% and a parser, evaluates the parser exactly $n$ times (i.e., setting
-% the number of iterations to be the $n$ field's value), and returns
-% an array of the parser's values. When generating output, it emits
-% the array and writes its length to the dependency field.
+\paragraph{Compressed data.}
+Transforms can handle encoded or compressed data. \XXX[More]
+\paragraph{Checksum}
+Checksums can also be implemented using transform \XXX[How?]
 
-% \paragraph{Offsets.}
-
-% The \texttt{offset} combinator takes a dependency field and a parser. It
-% moves the parser to the position specified in the offset field and
-% invokes the inner parser, and then moves the input back to its original
-% position.  While generating output, all structures referred to by offset
-% are generated after the main structure and the dependent offset fields
-% are patched up.
-
-% \paragraph{Checksums.}
-
-% In many data formats, some values depend on external representation,
-% such as checksums and cryptographic signatures.  While it would be
-% possible to extend our constraint language to be powerful enough to
-% support such constructs, we would essentially be building a separate,
-% Turing-complete language that has all the same pitfalls existing programs
-% have.  Therefore, we intend to allow the programmer to carefully escape
-% Nail's programming model and write a function that takes a pointer to
-% the dependent value and a range of bytes, using the \texttt{raw_depend}
-% combinator.
+\noindent A real-world example with many different transforms is explained in Section~\ref{s:eval-format-zip}.
 

### 2014-04-30 paper/osdi14/intro.tex
Design

@@
 
-Second, this well-defined internal representation allows Nail to invert the
-parser and generate output, while maintaining a \emph{semantic bijection}.
-A bijection in the traditional sense often does not make sense when parsing data formats. 
-Consider a grammar for a
-text language that tolerates white space, or a binary protocol that tolerates
-arbitrarily long padding.\footnote{Say, the physical layer of most communication
-  protocols is a possibly infinite sequence of symbols that are syntactically
-  nil followed by a pre-determined synchronization sequence and the actual
-  contents of the transmission.} Program semantics should be
-independent of the number of padding elements in the input, and Nail therefore
-does not expose that information to the programmer. We call such discarded
-fields \emph{constants}. 
-Similarly, Nail does not preserve the layout of
-objects referred to by their offsets. If the grammar consists of a simple protocol without offset
-fields, constants and the like, there is a normal bijection between the internal model and valid
-parser inputs.
+Second, this well-defined internal representation allows Nail to generate the external format from
+the internal representation, without requiring the programmer to write additional code.
 
@@
 
-Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{structural transformations}, to
-elegantly handle problematic features of data formats. Dependent fields are fields in the protocol
-whose value depends in some way on the value or layout of other parts of the protocol, such as
-length, offset and checksum fields. Structural transformations allow the programmer to write a pair
-of functions that transforms the \XXX[TODO]
+Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{transformations}, to
+elegantly handle problematic structures, such as offset fields or checksums. Dependent fields are
+fields in the protocol whose value depends in some way on the value or layout of other parts of the
+format. Transformations allow the programmer to escape the generated code to modify the raw data and
+interact with dependent fields in a controlled manner.
 

### 2014-04-30 paper/osdi14/eval.tex
Design
ZIP grammar from hell

--- 2014-04-29 paper/osdi14/eval.tex
@@
 \input{fig-ethernet}
-\input{fig-dnsfull}
-\input{fig-zip}
+
 In our evaluation of Nail, we try to answer four questions:
@@
 \item ZIP files.
-  See Figure~\ref{fig:zip-extract}.
+
 \end{itemize}
@@
 \paragraph{DNS.}
+
+\input{fig-dnsfull}
 Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
@@
 \XXX[TODO]
-\paragraph{ZIP Files.}
-
+\subsubsection{ZIP Files.}
+\input{fig-zip}
+\label{s:eval-format-zip}
 Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
@@
 
-Furthermore, Nail transparently handles ZIP's data compression and checksums with stream
-transformations. The implementation of one of these transforms are shown in
-Figure~\ref{fig:zip-compress}. For formats such as Office documents or Android applications that are based on ZIP
-containers, our ZIP grammar could be modified by including a more concrete grammar for file
-contents, which currently are just an array of bytes.
+We have implemented a grammar for ZIP. A strongly abbreviated version is shown in
+Figure~\ref{fig:zip-extract}.
 
-With existing parser languages, the grammar can not guarantee correctness of checksums, length
-fields or offsets, which places additional verification burden on the programmer. Furthermore,
-existing parser generators cannot decompress data as they are parsing. This is merely inconvenient
-for writing a grammar for plain ZIP files, but this prevents the grammar from validating the data
-within a ZIP file.
+Our ZIP grammar is a good example of transformations can capture complicated syntax in real-world
+file format. The \texttt{zip\_file} grammar first splits the entire file stream into two streams
+based on the \texttt{zip\_end\_of\_directory} transform.
+\texttt{zip\_end\_of\_directory_parse(NailStream *files, NailStream *dir, NailStream (current))}
+finds the end of directory header as described above by scanning the file backwards and splits the
+file into two streams, one containing the end-of-directory header and one containing the file
+contents. The header stream is then parsed. The file stream, to which all the offsets in the
+metadata refer, is passed as an argument to the \texttt{end\_of\_directory} parser.
+The directory header is then parsed and the offset and size transforms provided with Nail are used
+to isolate the actual directory from the file stream. Each directory entry in turn points to a local
+file header.
 
-The results suggest that Nail is a good fit for these data formats.
+The \texttt{file} grammar for a zip file entry is particularly interesting. It receives the file
+size, checksum and compression method of the file from the directory header. However, this same
+information is duplicated in the file entry, so the grammar uses a standard \texttt{depend}
+transform to check whether the two values are equal. Unlike most other transforms, \texttt{depend}
+does not consume or produce strings, it only checks that two dependent fields are equal when parsing
+and assigns the value of the second field to the first when generating. Immediately following the
+file entry is the compressed data. Because most compression algorithms operate on unbounded streams
+of data, Nail decompresses data in two steps.
+First, it isolates the compressed data from the rest of the stream by using the size transform. In
+this case, size operates on the current stream, which means it will consume data starting at the
+current position of the parser in the input.
+Second, Nail invokes a custom transform that implements the appropriate compression and decompression
+functions based on the specified compression method. These functions are otherwise oblivious about
+the layout or metadata of the file.
+
+Existing parser languages are not able to handle a file format of this complexity.
+
+\noindent The results suggest that Nail is a good fit for these data formats.
 
@@
 \subsection{Security.} 
+\XXX[Add PROTOS genome CVE's. count]
  \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous

### 2014-04-30 paper/osdi14/intro.tex
ZIP grammar from hell

@@
 Furthermore, security vulnerability can occur when an application assumes two repetitions of the
-same data to be consistent.
+same data to be consistent, such as allocating a buffer based on one size field and copying into
+that buffer based on another.
 

### 2014-04-30 paper/osdi14/design.tex
More design

@@
 
-\begin{figure}[tb]
 \begin{verbatim}
@@
 \end{verbatim}
-\caption{Nail grammar for uncompressed DNS labels.}
-\end{figure}
 
 Other dependencies, such as offset fields or checksums are not handled directly by combinators, but
-through stream transformation.
+through  transformations.
 
@@
 
-The programmer can define additional temporary streams by writing a \emph{transformation}, a pair of
-functions, one for parsing and one for generating output. Temporary streams follow the naming and
+The programmer can define additional streams by writing a \emph{transformation}, a pair of
+functions, one that generates new streams from existing streams during parsing and one that
+combines streams during output generation. Temporary streams follow the naming and
 scoping convention of dependent fields, but their names start with a \$-sign.
 
-The parsing function of the transform takes any number of streams, either temporary or the current,
+The parsing function of the transformation takes any number of streams, either temporary or the current,
 and outputs any number of temporary streams. The parsing function may reposition and read from the
@@
 \paragraph{Offsets.}
-We provide a transform for handling offset fields that is invoked as follows:
-\texttt{$fragment transform offset\_u32($file_stream, @offset)}. 
-They are implemented by two  functions:
-\begin{figure}[tb]
+We provide a transformation for handling offset fields that is invoked as follows:
+\texttt{\$fragment transform offset\_u32(\$file\_stream, \@offset)}. 
+They are implemented by these two pseudocode functions:
 \begin{verbatim}
-int offset_u32_parse(NailStream *out_fragment,NailStream *in_current,const uint32_t *off){
-   out_fragment = suffix of in_current at offset off
+int offset_u32_parse(NailArena *tmp,
+ NailStream *out_str,
+ NailStream *in_current,const uint32_t *off){
+ out_str = suffix of in_current at offset off
 }
-int offset_u32_generate(NailStream *in_fragment, NailStream *out_current,uint32_t *off){
+int offset_u32_generate(NailArena *tmp,
+  NailStream *in_fragment,
+  NailStream *out_current,
+  uint32_t *off){
    off = position of out_current
@@
 \end{verbatim}
-\caption{Pseudocode for the two C functions implementing an offset transformation.}
-\end{figure}
 
 \paragraph{Sizes.}
-A similar transform handles size fields. Just like the offset transform, it takes two parameters, a
+A similar transformation handles size fields. Just like the offset transform, it takes two parameters, a
 stream and a dependent field, but instead of returning the suffix of the current stream after an
@@
 \paragraph{Compressed data.}
-Transforms can handle encoded or compressed data. \XXX[More]
-\paragraph{Checksum}
-Checksums can also be implemented using transform \XXX[How?]
+Encoded, compressed or encrypted data can be handled transparently by writing a custom
+transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
+versa. Usually, this involves calling an existing library.
+
+\paragraph{Checksums.}
+Checksums can be verified and computed in a transformation that takes a stream and a dependent
+field. In some cases, a checksum is calculated over a buffer that contains the checksum itself, with
+the checksum being set to some particular value. Because the functions implementing a transformation
+are passed a pointer to any dependent fields, the checksum function can set the checksums initial
+value before calculating the checksum over the entire buffer, including the checksum.
 

### 2014-04-30 examples/network/tunclient.c
Partial, and borked grammar for network

@@ -94,17 +94,18 @@ void make_ip(ethernet *eth, struct phy_iface *iface){
 void send_frame(ethernet *packet,struct phy_iface *iface){
         const char *buf;
         size_t len;
-        HBitWriter *writer = h_bit_writer_new(&system_allocator);
-        gen_ethernet(writer,packet);
+        NailStream out;
+        NailOutStream_init(&out,4096);
+        gen_ethernet(&out,packet);
         buf =  h_bit_writer_get_buffer(writer,&len);
         write(iface->tun_fd,buf,len);
-        h_bit_writer_free(writer);
+        NailOutStream_release(&out);
 }
 void prepare_ethernet(ethernet *eth, struct phy_iface *iface){
         switch(eth->payload.N_type){
         case ARP:
-                memcpy(eth->dest,eth->payload.ARP.targethost,sizeof(eth->dest));
-                memcpy(eth->src,eth->payload.ARP.senderhost,sizeof(eth->src));
+                memcpy(eth->dest,eth->payload.arp.targethost,sizeof(eth->dest));
+                memcpy(eth->src,eth->payload.arp.senderhost,sizeof(eth->src));
 
                 //TODO: Prepare VLAN trunking. Is a host ever supposed to see VLAN trunking? 
                 break;
@@ -124,12 +125,12 @@ void process_arp(arpfour *arp, struct phy_iface *i, ethernet *eth){
                         //Send response, with our MAC
                         ethernet response;
                         response.payload.N_type = ARP;
-                        response.payload.ARP.operation = 2;
+                        response.payload.arp.operation = 2;
                         response.vlan = NULL;
-                        memcpy(response.payload.ARP.senderhost,i->mac,sizeof(macaddr));
-                        response.payload.ARP.senderip = i->ip.ip;
-                        memcpy(response.payload.ARP.targethost,arp->senderhost,sizeof(macaddr));
-                        response.payload.ARP.targetip = arp->senderip;
+                        memcpy(response.payload.arp.senderhost,i->mac,sizeof(macaddr));
+                        response.payload.arp.senderip = i->ip.ip;
+                        memcpy(response.payload.arp.targethost,arp->senderhost,sizeof(macaddr));
+                        response.payload.arp.targetip = arp->senderip;
                         prepare_ethernet(&response,i);
                         send_frame(&response,i);
                 }
[…]

### 2014-04-30 paper/osdi14/eval.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

@@
 example for the Hammer parser generator. All code size measurements for C were performed by
-SlocCount\cite{sloccount}. \XXX[code size is primitive, but we show orders of magnitude in the same language].
+SlocCount\cite{sloccount}.
 
@@
 Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
-grammar and 64 lines of C implementing DNS label compression. \XXX[Make this smaller]
+grammar and 64 lines of C implementing DNS label compression.
 The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
@@
 \subsection{Security.} 
-\XXX[Add PROTOS genome CVE's. count]
- \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
-security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
-Android security has been completely by-passed multiple times through parsing differentials in
-handling ZIP files\cite{saurik-masterkey}. mention talk by Julia Wolf about zip]. 
-Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.
+We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
+analyse a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
+immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
+server.
 
+The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
+security vulnerabilities in most implementations of ZIP and other archive formats that are directly
+related to input handling.
+
+Between May 2011 and May 2014, the CVE database\cite{cve-database} lists 83 vulnerabilities when searching for the
+string ``ZIP''. Of these, 42 were completely unrelated to the archive format, for example reporting
+vulnerabilities revealing Zip codes of other users. 26 vulnerabilities were related to ZIP archives,
+but not to input processing, such as trusting arbitrary update ZIP files sent to a server or
+vulnerabilities that can be accessed through manipulating files in a ZIP container. This includes
+several vulnerabilities allowing another files to replaced while they are extracted from ZIP
+files because insufficiently secure temporary directories are used. Of the 15 remaining
+vulnerabilities pertaining to ZIP files, 14 vulnerabilities were related to input processing. The
+ other vulnerability, which falls outside of Nails purview, concerns a ZIP packer choosing a weak
+ legacy cryptosystem even when configured to use AES.
+
+ Nail parsers are immune by design to vulnerabilities like these 14. Ten of those vulnerabilities
+ are memory corruption attacks, that could also be avoided by using safer languages or better
+ verification tools. Most of the buffer overflows seem to result from inconsistent offset fields
+ being manipulated without checking. Because Nail's generated code checks offsets before reading and
+ does not expose any untrusted pointers to the application, it protects against memory corruption
+ attacks by design.
+
+The remaining four vulnerabilities do not involve memory corruption, but rather parser ambiguities.
+Two of them relate to implementations mis-parsing ZIP files, particularly in virus scanners. One
+other vulnerability is in a Python ZIP library, which uses both the size field in the file entry and
+in the header, without checking for their consistency. This results in a denial-of-service through
+endless loops or incorrectly extracted ZIP files.
+
+A similar attack was used by the fourth vulnerability, the infamous Android master key bug
+\cite{saurik-masterkey}, that completely by-passed Android security multiple times through parser
+differentials between the ZIP handler that checks signatures for privileged applications and the ZIP
+implementations that ultimately extracts those files. Thus, valid  application bundles could be
+modified to include malicious applications without breaking their signatures. Instead of both
+components containing an ad-hoc archive implementation, a single Nail grammar would be reusable
+across both implementations. 
 \subsection{Performance.}
@@
 To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
-benchmarked our DNS server against BIND \XXX[and our network stack against \texttt{lwip-tap}, a userspace
-router implemented with the LWIP network stack\cite{lwip}].
+benchmarked our DNS server against BIND.
 
-\paragraph{DNS}
 We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind8}, one of
@@
 repeat the same, randomised sequence of queries for one minute and the average throughput and
-response latency were measured.  We repeated each test seven times  with 50.000 and one million
+response latency were measured.  We repeated each test seven times  with 50.000
 domain names, restarting each daemon in between. We also performed one initial dry run to warm the
-file system cache for the zone file.
-\XXX[Add 1 million results.They are largely the same]
+file system cache for the zone file. We repeated the tests with 1 million domain names and found
+that the servers performed almost identically.
 
 Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
-of optimisation by a factor of 3. \XXX[Why the hell?] This demonstrates that even though Nail was
+of optimisation by a factor of 3.  This demonstrates that even though Nail was
 not developed focusing on performance, applications built on Nail can compete with real-world

### 2014-04-30 paper/osdi14/impl.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

@@
 The current prototype of the Nail parser generator supports the C programming
-language and top-down parsers. Options for C++ STL data models and emitting
+language and top-down parsers. The implementation parses Nail grammars with Nail itself, using a 130
+line Nail grammar feeding into a 2000 line C++ program that implements the parser. Bootstrapping is
+performed with a subset of the grammar implemented using conventional grammars. 
+Options for C++ STL data models and emitting
 Packrat parsers~\cite{packrat-parsing:icfp02} are under development. In
@@
 % that the dependency field is correct.
-
-\paragraph{Bootstrapping.}
-
-To demonstrate the feasibility of the Nail parser generator, our parser
-generator uses a Nail parser to recognize Nail grammars. A superset of the
-grammar language described in this paper is implemented in 100 lines of Nail,
-which feed into about 1,000 lines of C++ that implement the actual parser
-generator. Bootstrapping is supported via an implementation of the Nail language
-in the Lemon parser generator, a variant of Yacc.

### 2014-04-30 paper/osdi14/relwk.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

@@
 
-\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]
-
-\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
-     but requires explicit encode/decode rules.  actually kind-of like
-     pack/unpack.]
-
-\paragraph{Language safety.}
+\paragraph{Motivation.}
+\XXX[Move to the intro, into its own section, or kill?]
 
@@
 sees different bytes at a virtual address than the executable that
-runs.}~\cite{evaders6} and Android\footnote{Android applications
-are distributed as .zip files. Signatures are verified with a Java
-program, but the program is extracted with a C program.  The Java
-program interprets all fields as signed, whereas the C program
-treats them as unsigned, allowing one to replace files in a signed
-archive, thereby undermining Android's security model.}~\cite{saurik-masterkey}\XXX{Duplicate, we
-already mention this in the eval} code signing and even the X.509 protocol
+runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing and even the X.509 protocol
 underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
@@
 
-A subclass of logic errors are so-called \textit{weird machines}, where
-implementation side effects or under-specified parser behavior leads to a
-protocol or data format inadvertently becoming a Turing-complete execution
-environment, even though the original grammar did not require it.
-Frequently, this execution environment can either
-then directly manipulate data in unwanted ways or be used to make
-exploiting another bug feasible.\footnote{For example, by compiling a
-return-oriented-programming exploit from code fragments discovered on
-the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
-ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
-research community, this has been generalized into treating a program
-as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
-an input, analogous to a virtual machine operating on bytecode.
-Nail avoids these problems by having the parser precisely match the
-specified grammar, eliminating under-specified behavior.
+% A subclass of logic errors are so-called \textit{weird machines}, where
+% implementation side effects or under-specified parser behavior leads to a
+% protocol or data format inadvertently becoming a Turing-complete execution
+% environment, even though the original grammar did not require it.
+% Frequently, this execution environment can either
+% then directly manipulate data in unwanted ways or be used to make
+% exploiting another bug feasible.\footnote{For example, by compiling a
+% return-oriented-programming exploit from code fragments discovered on
+% the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
+% ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
+% research community, this has been generalized into treating a program
+% as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
+% an input, analogous to a virtual machine operating on bytecode.
+% Nail avoids these problems by having the parser precisely match the
+% specified grammar, eliminating under-specified behavior.
 
@@
 
-\paragraph{Parsing frameworks.}
+\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]
+
+\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
+     but requires explicit encode/decode rules.  actually kind-of like
+     pack/unpack.]
+
+\XXX[Parsifal at langsec '14]
+\paragraph{Related Work}
 Proper input recognition has  been shown to be an excellent way of
@@
 Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
-Nail's structural dependencies allow it to work with protocols Hammer cannot recognize, such as
+Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
 protocols with offset fields or length fields (Hammer has extremely limited support for length

### 2014-04-30 paper/osdi14/eval.tex

@@
 \begin{itemize}
-\item Can Nail grammars support real-world data formats and are Nail's concepts critical to handling
-  them?
+
+\item Can Nail grammars support real-world data formats, and
+      are Nail's techniques critical to handling these formats?
+
 \item How much programmer effort is required to build an
       application that uses Nail for data input and output?
-\item Do applications handling input and output with Nail suffer from less security vulnerabilities?
-\item How fast are systems using Nail parsers as opposed to hand-written input handling? 
+
+\item Does using Nail for handling input and output improve
+      application security?
+
+\item Does Nail achieve acceptable performance?
+
 \end{itemize}
 
-\subsection{Data formats.}
+\subsection{Data formats}
 \label{s:eval-formats}

### 2014-04-30 paper/osdi14/eval.tex
two more subsubsections
mostly misc fmt stuff

@@
 
-\paragraph{DNS.}
+\subsubsection{DNS.}
 
 \input{fig-dnsfull}
+
+\begin{figure}
+\smaller[0.5]
+\input{code/dns-xform}
+\caption{Signatures of stream transform functions for handling DNS label compression.}
+\label{fig:dns-xform}
+\end{figure}
+
 Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
@@
 
-\paragraph{Ethernet}
+\subsubsection{Ethernet}
 \XXX[TODO]
@@
 Our ZIP grammar is a good example of transformations can capture complicated syntax in real-world
-file format. The \texttt{zip\_file} grammar first splits the entire file stream into two streams
-based on the \texttt{zip\_end\_of\_directory} transform.
-\texttt{zip\_end\_of\_directory_parse(NailStream *files, NailStream *dir, NailStream (current))}
+file format. The \cc{zip\_file} grammar first splits the entire file stream into two streams
+based on the \cc{zip\_end\_of\_directory} transform.
+\cc{zip\_end\_of\_directory_parse(NailStream *files, NailStream *dir, NailStream (current))}
 finds the end of directory header as described above by scanning the file backwards and splits the
@@
 contents. The header stream is then parsed. The file stream, to which all the offsets in the
-metadata refer, is passed as an argument to the \texttt{end\_of\_directory} parser.
+metadata refer, is passed as an argument to the \cc{end\_of\_directory} parser.
 The directory header is then parsed and the offset and size transforms provided with Nail are used
@@
 
-The \texttt{file} grammar for a zip file entry is particularly interesting. It receives the file
+The \cc{file} grammar for a zip file entry is particularly interesting. It receives the file
 size, checksum and compression method of the file from the directory header. However, this same
-information is duplicated in the file entry, so the grammar uses a standard \texttt{depend}
-transform to check whether the two values are equal. Unlike most other transforms, \texttt{depend}
+information is duplicated in the file entry, so the grammar uses a standard \cc{depend}
+transform to check whether the two values are equal. Unlike most other transforms, \cc{depend}
 does not consume or produce strings, it only checks that two dependent fields are equal when parsing
@@
 
-Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
-grammar and 64 lines of C implementing DNS label compression.
+Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full}, consists of a 48 line Nail
+grammar and 64 lines of C implementing DNS label compression, whose signatures are shown in
+Figure~\ref{fig:dns-xform}.
 The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
@@
 same grammar is used, together with 98 lines of C, to implement a minimal clone of the
-\texttt{host} command-line tool. Most of this code consists of a custom hashtable and system
+\cc{host} command-line tool. Most of this code consists of a custom hashtable and system
 interface code, such as listening to sockets and reading the zonefile.
@@
 including various support tools, is about 10,000 lines of C as measured by
-\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
+\cc{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
 features we don't have]
@@
 
-Finally, we used the \texttt{queryperf} tool provided with BIND to query each domain between zero
+Finally, we used the \cc{queryperf} tool provided with BIND to query each domain between zero
 and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 

### 2014-04-30 paper/osdi14/code/fmt.tex

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

### 2014-04-30 paper/osdi14/design.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

@@
 \begin{figure}[tb]
-\begin{tabular}{ll}
+\begin{tabular}{@{}ll@{}}
 \toprule
@@
 \midrule
-\texttt{int32} & 32-bit signed integer \\
-\multirow{2}{*}{\texttt{uint4}}
+\cc{int32} & 32-bit signed integer \\
+\multirow{2}{*}{\cc{uint4}}
   & 4-bit unsigned integer, \\
   & returned as an 8-bit value \\
-\multirow{2}{*}{\texttt{uint8 | 1..3}}
+\multirow{2}{*}{\cc{uint8 | 1..3}}
   & 8-bit unsigned integer, \\
   & $1\leq x \leq 3$ \\
-\multirow{2}{*}{\texttt{uint16 | [1,2]}}
+\multirow{2}{*}{\cc{uint16 | [1,2]}}
   & unsigned 16-bit integer, \\
   & $x \in \{1,2\}$ \\
-\multirow{2}{*}{\texttt{int32 | [1,5..255,512]}}
+\multirow{2}{*}{\cc{int32 | [1,5..255,512]}}
   & signed 32 bit integer, \\
@@
 
-\paragraph{S.}
+\paragraph{Repetition.}
 
@@
 last.
-For example, \texttt{many uint8} represents an array of 8-bit unsigned
-integers, and \texttt{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
+For example, \cc{many uint8} represents an array of 8-bit unsigned
+integers, and \cc{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
 comma-separated lists of decimal numbers.
@@
 \begin{figure}[tb]
-
+\smaller[0.5]
 \begin{minipage}{0.45\columnwidth}
@@
 \caption{Nail grammar (left) and data model (right) for a part of the
-grammar for DNS packets.  The \texttt{uint3 = 0} grammar represents
+grammar for DNS packets.  The \cc{uint3 = 0} grammar represents
 3 bits of padding (filled with zeroes).}
@@
 parsing and generated during output.  Constants can either have integer values, such as
-\texttt{uint3=0} or string values for text-based protocols, e.g. \texttt{many uint8 = ``Foo''}.
+\cc{uint3 = 0} or string values for text-based protocols, e.g. \cc{many uint8 = ``Foo''}.
 
@@
 and there is no semantic difference between the different syntactic representations.
-Nail therefore allows repeated constants, such as \texttt{many (uint8=' ')}, which parses any number of space characters, or
-\texttt{|| uint8 = 0x90 || uint16 = 0x1F0F}, which parses two of the many representations for X86
+Nail therefore allows repeated constants, such as \cc{many (uint8=' ')}, which parses any number of space characters, or
+\cc{|| uint8 = 0x90 || uint16 = 0x1F0F}, which parses two of the many representations for X86
 NOP instructions, which are used as padding between basic blocks in an executable.
@@
 
-For example, \texttt{word= <uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
+For example, \cc{word=<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
 lowercase word into an array.
@@
 \paragraph{References.}
-To support recursive structures, we include a reference combinator \texttt{*}, which is
+To support recursive structures, we include a reference combinator \cc{*}, which is
 syntactically equivalent to directly including the parser, but in the internal model adds a level of
 indirection to the value. This allows recursive structures to be parsed, for example, balanced
-parentheses could be parsed as follows: \texttt{paren = < uint8 ="("; optional *paren ;uint8=")">}
+parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren ;uint8=")">}
 
@@
 option is not necessarily understood identically by the parser. However, actual data formats are
-usually not ambiguous in this sense. Figure~\ref{fig:choice} demonstrates a simple choice
+usually not ambiguous in this sense.
+For example, Figure~\ref{fig:choice} demonstrates a simple choice
 combinator.
 
-Nail also includes an \texttt{optional} combinator, which attempts to recognise a value, but succeeds
-without consuming input when it cannot recognise that value. Syntactically, \texttt{optional} is
-equivalent to a choice between the parser and an empty structure, but in the internal model it is
-more concisely represented as a reference that is null when the parser fails.
-
-\begin{figure}[tb]
+\begin{figure}[h!]
+\smaller[0.5]
 \begin{verbatim}
@@
 \end{figure}
+
+Nail also includes an \cc{optional} combinator, which attempts to recognise a value, but succeeds
+without consuming input when it cannot recognise that value. Syntactically, \cc{optional} is
+equivalent to a choice between the parser and an empty structure, but in the internal model it is
+more concisely represented as a reference that is null when the parser fails.
 
@@
 
-Nail only provides one built-in combinator, \texttt{n_of} that uses dependency fields, which is used
-for count fields. For example, DNS labels are encoded as Pascal-style strings - a length followed by
-the value of the label. The Nail grammar for DNS packets in Figure~\ref{fig:dns-full} represents
+Nail only provides one built-in combinator, \cc{n_of} that uses dependency fields, which is used
+for count fields. For example, DNS labels are encoded as Pascal-style strings: a length followed by
+the value of the label. The Nail grammar for DNS labels can represent
 this pattern as follows:
 
+{
+\smaller[0.5]
 \begin{verbatim}
-{ @length uint8 | 1..64
-  label n_of @length uint8 }
+  {
+    @length uint8 | 1..64
+    label n_of @length uint8
+  }
 \end{verbatim}
+}
 
@@
 While parsing or generating a Nail grammar, there is always an implicit \emph{current stream} and
-referred to as \texttt{\$current} in the Nail language.
+referred to as \cc{\$current} in the Nail language.
 
@@
 We provide a transformation for handling offset fields that is invoked as follows:
-\texttt{\$fragment transform offset\_u32(\$file\_stream, \@offset)}. 
-They are implemented by these two pseudocode functions:
-\begin{verbatim}
-int offset_u32_parse(NailArena *tmp,
- NailStream *out_str,
- NailStream *in_current,const uint32_t *off){
- out_str = suffix of in_current at offset off
-}
-int offset_u32_generate(NailArena *tmp,
-  NailStream *in_fragment,
-  NailStream *out_current,
-  uint32_t *off){
-   off = position of out_current
-   append in_fragment to out_current
-}
-\end{verbatim}
+\cc{\$fragment transform offset\_u32(\$file\_stream, \@offset)}. 
+They are implemented by two functions, as shown in Figure~\ref{fig:xform-sig}.
+
+\begin{figure}[h]
+\smaller[0.5]
+\input{code/offset-xform}
+\caption{Pseudocode for two functions that implement the offset transform.}
+\label{fig:xform-sig}
+\end{figure}
 

### 2014-04-30 paper/osdi14/eval.tex
more xxx
added tables

@@
 challenging features and show that Nail is a good fit for these data formats and some aspects of
-these formats would be very hard to parse with other parser generators.
+these formats would be very hard to parse with other parser generators. The following table lists
+our example grammars, along with any features that make those protocols hard or impossible to handle
+with existing grammar languages and the size of the grammar and its associated transform functions
+measured in lines of code.
 
-In particular, we show extracts from grammars for the following protocols: 
-\begin{itemize}
-\item DNS messages.
-\item Ethernet packets containing ARP, ICMP and UDP.
-\item ZIP files.
+\begin{tabular}{l|l|l}
+\textbf{Protocol} & \textbf{LOC} & \textbf{Problematic features} \\ \hline
+DNS messages &  48+64 & Label compression\\
+& & Count fields \\
+ZIP archives & 92 + 78 & Checksums, \\ 
+& & compression, offsets, \\
+& &  variable length trailer\\ 
+Ethernet frames & 16 & \\
+ARP &  10  & \\
+IP &  25 & Total Length field, options \\
+UDP & 7 & Checksum, Length field \\
+ICMP & 5 & Checksum \\
 
-\end{itemize}
+\end{tabular}
 
@@
 
+\noindent We implemented the following applications and compare them to applications written without
+Nail:
+
+\noindent\begin{tabular}{l | l | l }
+\textbf{System} & \textbf{Nail LoC}  & \textbf{LoC w/o Nail} \\
+DNS server & 183 + 48 + 64 & 683 (Hammer parser)\\
+&&  ~10.000 (DJBDNS)\\ 
+DNS \texttt{host} tool & 98 + 48 + 64  & \XXX[TODO]\\
+\texttt{unzip} & 50 + 92 + 78 & 1600 (Info-Zip) \\
+
+\end{tabular}
 %%DNS: 
@@
 However, the closest in functionality and intent is Dan Bernstein's
-djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
-a minimalist, highly secure DNS server. The latest release of djbdns,
-including various support tools, is about 10,000 lines of C as measured by
-\cc{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
-features we don't have]
+djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be a minimalist, highly secure,
+authoritative-only DNS server. The latest release of djbdns, including various support tools, is
+about 10,000 lines of C as measured by \cc{sloccount}. DJ
 
@@
 across both implementations.
+
+\XXX[Add fuzzing results - no crashes]
 \subsection{Performance.}

### 2014-04-30 paper/osdi14/design.tex

@@
 
-\subsection{Goals}
+Nail's goals are to reduce programmer effort required to safely interact
+with data formats.  In particular, this means:
+
+\begin{itemize}
+
+\item Using the grammar to define both the \emph{external format}
+      and the \emph{internal representation}.
+
+\item Both parsing inputs into internal representations, as well as
+      generating outputs from internal representations, without
+      requiring the programmer to write any semantic actions.
+
+\item Eliminating redundancy in internal representations, such as
+      storing both an explicit length field and an implicit length
+      of a container data structure, to ensure that programmers
+      see a consistent view of the data.
+
+\item Allow programmers to define grammars for complex real-world
+      data formats through well-defined extensibility mechanisms.
+
+\end{itemize}
+
+\subsection{Overview}
+
 \paragraph{Internal Model.}

### 2014-04-30 paper/osdi14/intro.tex
intro

@@
 
-Finally, data formats, for example PNG or PDF, contain features that are hard to represent with
-existing parser generators. Some features, such as length and checksum fields, require the parser to
-keep additional state. More complicated formats incorporate offset fields that require the parser to
-reposition its input stream or consume the same input multiple times. In some cases, such as with
-compressed data, very complicated transformations on the external format are necessary before it can
-be understood. State-of-the-Art parser generators provide at most limited, unelegant facilities for
-such features, so programmers are forced to manually writing control code to handle them, such as
-moving the parsers input stream, looping over the invocation of a sub-parser or computing a checksum
-over raw input bytes. Besides leaving much room for errors with offset arithmetic, such code is
-usually not reusable when generating output.
+Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
+existing parser generators. Those parsers cannot directly deal with length or checksum fields
+directly, so either the programmer has to write potentially unsafe code to deal with such features
+or build contrived grammar constructs, such as introducing one grammar rule for each possible value
+of a length fields. Offset fields usually require the programmer to manipulate a parser's internal
+state to re-position its input. More complicated transformations on the external format, such as
+compressed data, cannot be implemented at all.
 
-This paper presents the initial design and implementation of Nail, a parser
-generator that greatly reduces the programmer effort required to use a
-grammar-based parser. Nail addresses the above four challenges with several key
-ideas.
+This paper presents the design and implementation of Nail, a parser generator that greatly reduces
+the programmer effort required to use grammars. Nail addresses the above four challenges with
+several key ideas.
 
-First, Nail reduces the expressiveness of its grammar language by removing
-semantic actions. Existing parser generators allow arbitrary computation to
-transform between the AST and the parser output. Instead, Nail derives the
-structure of its output automatically from the grammar, forcing the programmer
+First, Nail grammars define both a format's external representation and an internal object model.
+This removes the semantic actions and type declarations that programmers have to write with existing
+parser generator. While this somewhat reduces the flexibility of the internal model, it forces the programmer
 to clearly separate syntactic validation and semantic processing.

### 2014-04-30 paper/osdi14/relwk.tex
Related work

@@
 
-\paragraph{Motivation.}
-\XXX[Move to the intro, into its own section, or kill?]
+% \XXX[Move to the intro, into its own section, or kill?]
+
+\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]
+
+\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
+     but requires explicit encode/decode rules.  actually kind-of like
+     pack/unpack.]
+
+\XXX[Parsifal at langsec '14]
+\paragraph{Parsers}
+
+Generating parsers and generators from an executable specification
+is the core concept of Interface Generators, e.g. in CORBA or more
+recently~\cite{varda2008}. However, while interface generators work
+very well for existing grammars, they do not allow full control over the
+format of the output, so cannot be used to implement legacy protocols.
+Very related work has been done at Bell Labs with the PacketTypes
+system~\cite{mccann2000packet}, however PacketTypes works only as a
+parser, not as an output generator and does not support the expressive
+power of parsing expression grammars (PEGs), but rather implements a
+C-like structure model enhanced with data-dependent length fields and
+constraints.
+
+Parser generators for binary protocols were first introduced by the
+Hammer~\cite{hammer-parser} parser.
+While previous parser generators could also be used to write grammars
+for binary protocols,\footnote{Theoretically speaking, the alphabet
+over which a grammar is an  abstract set, so most algorithms work
+just as well on an alphabet of $\{0,1\}$.} doing so is practically
+inconvenient. Hammer allows the programmer to specify a grammar in
+terms of bits and bytes instead of characters. Common concerns, such as
+endianness and bit-packing are transparently hidden by the library.
+
+Hammer implements grammars as language-integrated parser combinators, an approach popularized by
+Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
+described in~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
+grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For more background on the history of
+  expressing grammars, see Bryan Ford's masters thesis~\cite{ford2002packrat}, which also describes
+  the default parsing algorithm used by Hammer.} by composing them from one or multiple sub-parsers.
+Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
+an abstract syntax tree (AST).
+
+Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
+Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
+Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
+protocols with offset fields or length fields (Hammer has extremely limited support for length
+fields: it can parse arrays immediately preceded by their length).
+
+\paragraph{Application use of parsers.}
+Generated parsers have long been used to parse human input, such as programming languages and
+configuration files. Frequently, such languages are often specified with a formal grammar in an
+executable form. Unfortunately,  parser frameworks are seldom used  to recognize
+machine-created input.
+
+A notable exception is the Mongrel
+web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
+for HTTP written in the Ragel~\cite{ragel-paper} regular expression
+language. Mongrel was re-written from scratch multiple times to achieve
+better scalability and design, yet the grammar could be  re-used across
+all iterations~\cite{patterson-citation}.
+
+\paragraph{Other vulnerability mitigations.}
 
@@
 
-The second class is logic errors, where application code misinterprets
-input data.  This can lead to serious security consequences when two
-systems disagree on the meaning of a network packet or a signed message,
-as in iOS\footnote{The XNU kernel and the user code-signing verifier
-interpret executable metadata differently, so the code signature
-sees different bytes at a virtual address than the executable that
-runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing and even the X.509 protocol
-underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
-highly application-specific, and are difficult to mitigate using existing
-techniques, and these mistakes can occur even in high-level languages
-that guarantee memory safety.
-By allowing developers to specify their data format just once, Nail
-avoids logic errors and inconsistencies in parsing and output generation.
+The second class is logic errors, where application code misinterprets input data. Safe languages
+and exploit mitigation technologies do not help against such vulnerabilities. This can lead to
+serious security consequences when two systems disagree on the meaning of a network packet or a
+signed message, as in iOS\footnote{The XNU kernel and the user code-signing verifier interpret
+  executable metadata differently, so the code signature sees different bytes at a virtual address
+  than the executable that runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing
+and even the X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Logic errors are often
+causing memory corruption bugs as well, for example when a protocol stores the size of a buffer in
+two places and the implementation does not check their consistency. These mistakes are
+highly application-specific, and are difficult to mitigate using existing techniques, and these
+mistakes can occur even in high-level languages that guarantee memory safety. By allowing developers
+to specify their data format just once, Nail avoids logic errors and inconsistencies in parsing and
+output generation.
 
@@
 
-\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]
-
-\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
-     but requires explicit encode/decode rules.  actually kind-of like
-     pack/unpack.]
-
-\XXX[Parsifal at langsec '14]
-\paragraph{Related Work}
 Proper input recognition has  been shown to be an excellent way of
-eliminating malicious inputs. In one case, a  PDF parser implemented in
+eliminating ambiguous inputs. In one case, a  PDF parser implemented in
 Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
@@
 tied to one specific application.
-
-Generating parsers and generators from an executable specification
-is the core concept of Interface Generators, e.g. in CORBA or more
-recently~\cite{varda2008}. However, while interface generators work
-very well for existing grammars, they do not allow full control over the
-format of the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes
-system~\cite{mccann2000packet}, however PacketTypes works only as a
-parser, not as an output generator and does not support the expressive
-power of parsing expression grammars (PEGs), but rather implements a
-C-like structure model enhanced with data-dependent length fields and
-constraints.
-
-Parser generators for binary protocols were first introduced by the
-Hammer~\cite{hammer-parser} parser.
-While previous parser generators could also be used to write grammars
-for binary protocols,\footnote{Theoretically speaking, the alphabet
-over which a grammar is an  abstract set, so most algorithms work
-just as well on an alphabet of $\{0,1\}$.} doing so is practically
-inconvenient. Hammer allows the programmer to specify a grammar in
-terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing are transparently hidden by the library. Hammer
-implements grammars as language-integrated parser combinators, an approach
-popularized by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
-parser combinator style (to our knowledge, first  described
-in~\cite{burge1975recursive}) is a natural way of concisely expressing
-top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
-more background on the history of expressing grammars, see Bryan Ford's
-masters thesis~\cite{ford2002packrat}, which also describes the default
-parsing algorithm used by Hammer.} by composing them from one or multiple
-sub-parsers.  Hammer then constructs a tree of function pointers which
-can be invoked to parse a given input into an abstract syntax tree (AST).
-
-Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
-Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
-Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
-protocols with offset fields or length fields (Hammer has extremely limited support for length
-fields: it can parse arrays immediately preceded by their length).
-
-\paragraph{Application use of parsers.}
-Generated parsers have long been used to parse human input, such as programming languages and
-configuration files. Frequently, such languages are often specified with a formal grammar in an
-executable form.
-
-Unfortunately,  parser frameworks are seldom used  to recognize
-machine-created input. For example, the security-critical and
-well-engineered MIT Kerberos distribution uses parser generators, but
-only for handling configuration files. A notable exception is the Mongrel
-web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
-for HTTP written in the Ragel~\cite{ragel-paper} regular expression
-language. Mongrel was re-written from scratch multiple times to achieve
-better scalability and design, yet the grammar could be  re-used across
-all iterations~\cite{patterson-citation}.
 

### 2014-04-30 paper/osdi14/design.tex
Design and related work

@@
 \paragraph{Semantic Bijecction.}
-Second, this well-defined internal representation allows Nail to invert the
-parser and generate output, while maintaining a \emph{semantic bijection}, that is, when generating
-the corresponding output to a parser input, the two byte streams will have the same meaning, i.e. be
-interpreted the same by Nail, but might not be identical.
-A bijection in the traditional sense often does not make sense when parsing data formats. 
-Consider a grammar for a text language that tolerates white space, or a binary protocol that tolerates
-arbitrarily long padding.\footnote{Say, the physical layer of most communication
-  protocols is a possibly infinite sequence of symbols that are syntactically
-  nil followed by a pre-determined synchronization sequence and the actual
-  contents of the transmission.} Program semantics should be
-independent of the number of padding elements in the input, and Nail therefore
-does not expose that information to the programmer. We call such discarded
-fields \emph{constants}. 
-Similarly, Nail does not preserve the layout of
-objects referred to by their offsets. If the grammar consists of a simple protocol without offset
-fields, constants and the like, there is a normal bijection between the internal model and valid
-parser inputs.
+Second, in order to generate output, Nail maintaining
+a \emph{semantic bijection} between the external format and the internal model. That is, when
+generating the corresponding output to a parser input, the two byte streams will have the same
+meaning, i.e. be interpreted the same by Nail, but might not be identical.
+ A bijection in the traditional sense often does not make sense for data
+formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
+that tolerates arbitrarily long padding.
+%\footnote{Say, the physical layer of most communication
+%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
+%  pre-determined synchronization sequence and the actual contents of the transmission.}
+ Program semantics should be independent of the number of padding elements in the input, and Nail therefore
+does not expose that information to the programmer. We call such discarded fields \emph{constants}.
+Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
+consists of a simple protocol without offset fields, constants and the like, there is a conventional
+bijection between  internal models and valid parser inputs.
 \paragraph{Hide redundant information.}
-Nail's internal model is designed to hide unneeded information from the application. The more
-information Nail automatically deduces, the less the application can produce invalid data.
-Therefore, Nail introduces \emph{dependent fields}, which contain data that can be computed during
-generation and needs to be kept as additional state during parsing.
+Nail's internal model is designed to hide unneeded and redundant information from the application.
+Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
+and needs to be kept as additional state during parsing. Dependent fields are for example used to represent
+lengths, offsets and checksums.
 
@@
 
-We will introduce transformations, which allow the programmer to write arbitrary functions that can
+Nail introduce transformations, which allow the programmer to write arbitrary functions that can
 handle such structures and streams, which are a generic abstraction for input and output data that

### 2014-04-30 paper/osdi14/relwk.tex
Design and related work

@@
 % exploiting another bug feasible.\footnote{For example, by compiling a
-% return-oriented-programming exploit from code fragments discovered on
+% return-oriented-programming exploit from code fragments discoverepd on
 % the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
@@
 
-Proper input recognition has  been shown to be an excellent way of
-eliminating ambiguous inputs. In one case, a  PDF parser implemented in
-Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
-However, manually writing parser code does not scale to the number of
-file formats  and protocols in existence and might result in parser code
-tied to one specific application.
+Proper input recognition has been shown to be an excellent way of eliminating ambiguous inputs. In
+one case, a PDF parser implemented in Coq could eliminate over $95\%$ of known malicious
+PDFs~\cite{Bogk-PDF}. The same project also produced a ``counter-example'' PDF that triggered
+numerous bugs in other PDF parsers. However, manually writing parser code and proving unambiguity
+does not scale to the number of file formats and protocols in existence and might result in parser
+code tied to one specific application.
 

### 2014-05-01 paper/osdi14/design.tex
Initial syntax table

@@
 generating the corresponding output to a parser input, the two byte streams will have the same
-meaning, i.e. be interpreted the same by Nail, but might not be identical.
+meaning, i.e. be interpreted equivalently by Nail, but might not be identical.
  A bijection in the traditional sense often does not make sense for data
@@
 
-\XXX[In the rest of this section, we first present the design of the Nail language and then show how
-Nail's features match these goals.]
+In the rest of this section, we describe the design on the Nail grammar language.
+\input{fig-syntax}
 \subsection{Basics}
-Nail grammars are a set of rules. Each rule assigns a parser, which can be a reference to another
-rule, a trivial parser describing an integer of a certain width or a combinator that combines
-multiple parsers in some way. We provide the most common combinators familiar from other parser
-combinator libraries, such as Parsec\cite{LeijenMeijer:parsec} and Hammer\cite{hammer-parser}.
-However, we extend the sequence and choice combinators normally found in parser combinator libraries
-with \emph{semantic labels} that name fields in the internal models. The syntax described by a Nail
-grammar does not depend on these semantic labels.
+A Nail parser defines both the structure of some external format and a data type to represent that
+format. Parsers are constructed by combinators over simpler parsers, an approach popularised by the Parsec
+framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
+ combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
+ describe a data type. 
+\paragraph{Integers and Constraints}
 
-\paragraph{Integers and Constraints}
-\begin{figure}[tb]
-\begin{tabular}{@{}ll@{}}
-\toprule
-\bf Syntax & \bf Semantics \\
-\midrule
-\cc{int32} & 32-bit signed integer \\
-\multirow{2}{*}{\cc{uint4}}
-  & 4-bit unsigned integer, \\
-  & returned as an 8-bit value \\
-\multirow{2}{*}{\cc{uint8 | 1..3}}
-  & 8-bit unsigned integer, \\
-  & $1\leq x \leq 3$ \\
-\multirow{2}{*}{\cc{uint16 | [1,2]}}
-  & unsigned 16-bit integer, \\
-  & $x \in \{1,2\}$ \\
-\multirow{2}{*}{\cc{int32 | [1,5..255,512]}}
-  & signed 32 bit integer, \\
-  & $x \in \{ 1, 512 \} \vee 5\leq x\leq 255$ \\
-\bottomrule
-\end{tabular}
-\caption{Example Nail grammars for integer values.}
-\label{fig:range}
-\end{figure}
-
-Nail's fundamental parsers recognise signed or unsigned with arbitrary lengths up to 64 bits, an
+Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
 approach popularised by the Hammer framework.\cite{hammer-parser}  
@@
 
-\paragraph{References.}
-To support recursive structures, we include a reference combinator \cc{*}, which is
-syntactically equivalent to directly including the parser, but in the internal model adds a level of
-indirection to the value. This allows recursive structures to be parsed, for example, balanced
-parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren ;uint8=")">}
-
 \paragraph{Choices.}
@@
 usually not ambiguous in this sense.
-For example, Figure~\ref{fig:choice} demonstrates a simple choice
+For example, Figure~\ref{fig:syntax} demonstrates a simple choice
 combinator.
-
-\begin{figure}[h!]
-\smaller[0.5]
-\begin{verbatim}
-  choose {
-    A = uint8 | 1..8
-    B = uint16 | ..256
-  }
-\end{verbatim}
-\caption{A simple choice combinator that parses either an 8-bit unsigned
-integer with a value between 1 and 8 (option A), or a 16-bit unsigned
-integer with a value of at most 256 (option B).}
-\label{fig:choice}
-\end{figure}
 
@@
 
+\paragraph{Rules.}
+A Nail grammar file consists of rules that assign a parser to a name. Rules are written as
+assignments, such as \texttt{ints = many uint8}, which defines a rule called \texttt{ints} with the
+body \texttt{many uint8}.  As we will describe in Sections~\ref{s:dependent}
+and~\ref{s:transforms}, rules can optionally consume parameters. 
+Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
+the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.
+
+Rules allow for recursive grammars. To support recursive data types, we allow introduce the
+reference combinator \cc{*}  that does not change the syntax of the external format described, but
+introduces a layer of indirection, such as a reference or pointer, to the model data type.
+
+% To support recursive structures, we include a reference combinator \cc{*}, which
+%   is syntactically equivalent to directly including the parser, but in the internal model adds a
+%   level of indirection to the value. This allows recursive structures to be parsed, for example,
+%   balanced parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren
+%     ;uint8=")">}
+
 \subsection{Dependent fields}
-
+\label{s:dependent}
 Data formats often contain values that are determined by other values or the layout of information,
@@
 \subsection{Input streams and Transformations.}
-
+\label{s:transforms}
 In order to handle complicated input functions, Nail allows the programmer to

### 2014-05-01 paper/osdi14/eval.tex
Initial syntax table

@@
 challenging features and show that Nail is a good fit for these data formats and some aspects of
-these formats would be very hard to parse with other parser generators. The following table lists
-our example grammars, along with any features that make those protocols hard or impossible to handle
-with existing grammar languages and the size of the grammar and its associated transform functions
-measured in lines of code.
+these formats would be very hard to parse with other parser generators.
+Figure~\ref{fig:eval-protocols} lists various example protocols we have implemented.
 
-\begin{tabular}{l|l|l}
-\textbf{Protocol} & \textbf{LOC} & \textbf{Problematic features} \\ \hline
+\begin{figure}[tb]
+\label{fig:eval-protocols}
+\begin{tabular}{@{}lll@{}}
+\toprule
+\textbf{Protocol} & \textbf{LoC} & \textbf{Problematic features} \\ 
+\midrule
 DNS messages &  48+64 & Label compression\\
@@
 ICMP & 5 & Checksum \\
+\bottomrule
 
 \end{tabular}
-
+\caption{Example Nail grammars, their code size of the grammar plus  the transformations they use
+  and protocol features that cannot be expressed in existing grammar languages. }
+\end{figure}
 \subsubsection{DNS.}
@@
 
-\subsubsection{Ethernet}
-\XXX[TODO]
 \subsubsection{ZIP Files.}
@@
 
-\noindent\begin{tabular}{l | l | l }
+\noindent\begin{tabular}{@{}lll@{}} 
+\toprule
 \textbf{System} & \textbf{Nail LoC}  & \textbf{LoC w/o Nail} \\
-DNS server & 183 + 48 + 64 & 683 (Hammer parser)\\
+\midrule
+DNS server & 295 & 683 (Hammer parser)\\
 &&  ~10.000 (DJBDNS)\\ 
-DNS \texttt{host} tool & 98 + 48 + 64  & \XXX[TODO]\\
-\texttt{unzip} & 50 + 92 + 78 & 1600 (Info-Zip) \\
-
+DNS \texttt{host} tool & 210  & \XXX[TODO]\\
+\texttt{unzip} & 220 & 1600 (Info-Zip) \\
+\bottomrule
 \end{tabular}
@@
 
-\paragraph{Network stack.}
-\XXX[Write this up]
 \subsection{Security.} 
@@
 
-\XXX[Add fuzzing results - no crashes]
+To provide additional assurance that Nail parsers are free of memory corruption attacks, we ran the
+DNS fuzzer provided with the Metasploit\cite{mspframework} framework on our server, which sent
+randomly corrupted DNS queries to our server for 4 hours, during which it did not crash or trigger
+the stack or heap corruption detector.
 \subsection{Performance.}

### 2014-05-01 paper/osdi14/intro.tex
Initial syntax table

@@
 Furthermore, security vulnerability can occur when an application assumes two repetitions of the
-same data to be consistent, such as allocating a buffer based on one size field and copying into
-that buffer based on another.
+same data to be consistent, such as allocating a buffer based on the value of one size field and copying into
+that buffer based on the value of another.
 
@@
 
-Second, this well-defined internal representation allows Nail to generate the external format from
+Second, this well-defined internal representation allows Nail to generate output from
 the internal representation, without requiring the programmer to write additional code.

### 2014-05-01 paper/osdi14/relwk.tex
Initial syntax table

@@
 power of parsing expression grammars (PEGs), but rather implements a
-C-like structure model enhanced with data-dependent length fields and
-constraints.
+C-like structure model enhanced with  length fields and
+constraints. PacketTypes cannot handle complicated encodings such as  compressed data.
 
@@
 A notable exception is the Mongrel
-web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
+web server~\cite{mongrel} which uses a grammar
 for HTTP written in the Ragel~\cite{ragel-paper} regular expression
@@
 inputs.  These mistakes arise in lower-level languages that do not
-provide memory safety guarantees (such as C), and can be partially
-mitigated by a wide range of techniques, such as static analysis,
+provide memory safety guarantees such as C, and can be partially
+mitigated by a wide range of techniques, for example static analysis,
 dynamic instrumentation, and address space layout randomization, that

### 2014-05-01 paper/osdi14/design.tex
Commit

@@
 
-\paragraph{Allow parser extension.}
+\paragraph{Parser Extensions.}
 Real-world protocols contain complicated ways of encoding data. Representing these in an
@@
  describe a data type.
+
+Figure~\ref{fig:syntax} shows the elements of Nail grammars, which we describe in more detail below.
 \paragraph{Integers and Constraints}
@@
 Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
-data offset within the TCP header.  In the internal representation, integer values are represented
-by the nearest usable integer size, for example a 24-bit unsigned integer will be stored in an
-unsigned 32-bit double word.
+data offset within the TCP header.  The internal representation rounds up to the nearest integer
+type, for example a 24-bit unsigned integer will be stored in an unsigned 32-bit double word.
 
-The grammar can also express a constraint on an integer field. The current version of Nail expresses
+The grammar can also constraint the values of an integer. The current version of Nail expresses
 constraints as a set of permissible values or value ranges. Extending the Nail language and
@@
 usually not ambiguous in this sense.
-For example, Figure~\ref{fig:syntax} demonstrates a simple choice
-combinator.
 
@@
 \paragraph{Rules.}
-A Nail grammar file consists of rules that assign a parser to a name. Rules are written as
+A Nail grammar consists of rules that assign a parser to a name. Rules are written as
 assignments, such as \texttt{ints = many uint8}, which defines a rule called \texttt{ints} with the

### 2014-05-01 paper/osdi14/relwk.tex
Commit

@@
 \label{s:relwk}
-
-% \XXX[Move to the intro, into its own section, or kill?]
-
-\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]
 
@@
 
+\paragraph{Workshop paper}
+We presented an early design of Nail at the LangSec workshop~\cite{bangert:nail-langsec} that had
+no support for transformations, as described in Section~\ref{s:transforms}, and much more limited
+dependent fields, as described in Section~\ref{s:dependent}. With our extended and refined design,
+Nail can support much more complicated, real-world data formats.
 %%This is supposed to go into design

### 2014-05-01 paper/osdi14/design.tex
References

@@
 
-Figure~\ref{fig:range} shows several examples.
-
 \paragraph{Repetition.}
@@
 
-For example, \cc{word=<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
+For example, \cc{<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
 lowercase word into an array.
@@
 
-Nail also includes an \cc{optional} combinator, which attempts to recognise a value, but succeeds
+\paragraph{Optional.}Nail also includes an \cc{optional} combinator, which attempts to recognise a value, but succeeds
 without consuming input when it cannot recognise that value. Syntactically, \cc{optional} is
@@
 
-Rules allow for recursive grammars. To support recursive data types, we allow introduce the
+\paragraph{References.}Rules allow for recursive grammars. To support recursive data types, we allow introduce the
 reference combinator \cc{*}  that does not change the syntax of the external format described, but
@@
 
+For example example,   balanced parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren;uint8=")">}
+
 % To support recursive structures, we include a reference combinator \cc{*}, which
 %   is syntactically equivalent to directly including the parser, but in the internal model adds a
-%   level of indirection to the value. This allows recursive structures to be parsed, for example,
-%   balanced parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren
-%     ;uint8=")">}
+%   level of indirection to the value. This allows recursive structures to be parsed, f
 

### 2014-05-01 paper/osdi14/abstract.tex
update abstract

@@
 
-  We present \textit{Nail}, an interface generator that allows programmers to safely parse and
-  generate protocols defined by a Parser-Expression based grammar. Nail uses a richer set of parser
-  combinators that induce an internal representation, obviating the need to write semantic actions.
-  Nail also introduces \textit{dependency fields} and \textit{stream transformations} that capture
-  protocol features such as size and offset fields, check-sums and compressed data that are hard to
-  express in existing protocol languages.
+\textit{Nail} is an interface generator that greatly reduces the
+programmer effort for safely parsing and generating protocols defined
+by a Parser-Expression based grammar.  Nail introduces several key
+ideas to achieve its goal.  First, Nail uses the protocol grammar to
+define not just the data format, but also the internal representation
+of the data.  Second, Nail eliminates the notion of semantic actions,
+used by existing parser generators, which reduces the expressive power
+but allows Nail to both {\em parse} data formats and {\em generate}
+them from internal representations, by maintaining a {\em semantic
+bijection} between data formats and internal representation.  Third,
+Nail introduces {\em dependency fields} and {\em stream transforms} to
+capture protocol features such as size and offset fields, checksums,
+and compressed data, which are impractical to express in existing
+protocol languages.  Using Nail, we implement a complete DNS server in
+C in under 300 lines of code and grammar, and a complete {\tt unzip}
+program in C in 220 lines of code and grammar, demonstrating that Nail
+makes it easy to parse complex real-world data formats.
 

### 2014-05-01 paper/osdi14/code/offset-xform.c
x

@@ -5,7 +5,7 @@ offset_u32_parse(NailArena *tmp,
                  const uint32_t *off)
 {
   /* out_str = suffix of in_current
-               at offset off */
+               at offset *off */
 }
 
 int
@@ -14,6 +14,6 @@ offset_u32_generate(NailArena *tmp,
                     NailStream *out_current,
                     uint32_t *off)
 {
-  /* off = position of out_current */
+  /* *off = position of out_current */
   /* append in_fragment to out_current */
 }

### 2014-05-01 paper/osdi14/eval.tex
Commit
Fixed citations

@@
 implementations disagree on how to find this header in confusing situations, such as when the
-comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
+comment contains the valid magic bytes\cite{wolf:berlinsides-zip}.
 This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
@@
 &&  ~10.000 (DJBDNS)\\ 
-DNS \texttt{host} tool & 210  & \XXX[TODO]\\
+DNS client & 210  & 4,615 (host) \\
 \texttt{unzip} & 220 & 1600 (Info-Zip) \\
@@
 
-The Hammer parser framework\cite{hammer-dns} ships with a toy DNS server that responds to
+The Hammer parser framework\cite{hammer-parser} ships with a toy DNS server that responds to
 any valid DNS query with a CNAME record to the domain ``spargelze.it''. 
@@
 
-We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind8}, one of
+We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind9}, one of
 the oldest and most popular DNS daemons, on a load that resembles the authoritative name server at a

### 2014-05-01 paper/osdi14/relwk.tex
Fixed citations

@@
 \label{s:relwk}
-
-\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
-     but requires explicit encode/decode rules.  actually kind-of like
-     pack/unpack.]
 
@@
 fields: it can parse arrays immediately preceded by their length).
+
+Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports generating output for Ocaml.
+Parsifal structures grammars as an Ocaml type that holds an internal model and functions for parsing
+input and output.
+However, their generator can only produce parsers and generators for simple, fixed-size structures.
+The programmer can then use these when implementing parsers and generators for more
+complicated formats, manually handling offsets, checksums and the like, risking bugs. Nail can
+handle more complicated constructs without the programmer writing code to support them.
 

### 2014-05-01 paper/osdi14/intro.tex

--- 2014-04-30 paper/osdi14/intro.tex
@@
 read and write files, or both receive and send network packets.
-Most parser generators just focus on parsing an input, rather
+Most parser generators focus on just parsing an input, rather
 than producing an output, thus requiring the programmer to manually
@@
 different instances of the same value, an attacker can craft input that causes inconsistencies.
-Furthermore, security vulnerability can occur when an application assumes two repetitions of the
-same data to be consistent, such as allocating a buffer based on one size field and copying into
-that buffer based on another.
+Furthermore, security vulnerabilities can occur when an application assumes two repetitions of the
+same data to be consistent, such as allocating a buffer based on the value of one size field and copying into
+that buffer based on the value of another.
 
@@
 the programmer effort required to use grammars. Nail addresses the above four challenges with
-several key ideas.
+several key ideas, as follows.
 
@@
 
-Second, this well-defined internal representation allows Nail to generate the external format from
-the internal representation, without requiring the programmer to write additional code.
+Second, this well-defined internal representation allows Nail to
+maintain a {\em semantic bijection} between data formats and their
+internal representation.  As a result, this enables Nail to not just
+parse input but also generate output from the internal representation,
+without requiring the programmer to write additional code.
 
@@
 
-We have implemented a prototype of Nail for C\@.  Our experience
-so far suggests that it is a promising approach: we were able to construct
-a succinct grammar for a set of common file formats and protocols that demonstrate Nail's ability to
-handle tricky constructs and build a series of sample applications with these grammars.
+To evaluate whether Nail's design is effective at handling real-world
+data formats, we implemented a prototype of Nail for C\@.  Using our
+prototype, we implemented succinct grammars for parts of an IP network
+stack, for DNS packets, and for Zip files, each in about a hundred lines
+of grammar.  On top of these grammars, we were able to build a fully
+functional DNS server in under 200 lines of C code, and an {\tt unzip}
+utility in about 50 lines of C code.  This suggests both that Nail is
+effective at handling complex real-world data formats, and that Nail
+makes it easy for application developers to handle external data formats.
 

### 2014-05-01 paper/osdi14/intro.tex
xx

@@
 library has had 24 remotely exploitable vulnerabilities from 2007 to
-2013,\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}}
+2013~\cite{cvedetails:libpng},
 and Adobe's PDF and Flash viewers have been notoriously
@@
 library, have had input processing vulnerabilities in the
-past.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}
+past~\cite{cvedetails:zlib}.
 

### 2014-05-01 paper/osdi14/abstract.tex
some intro nits

@@
 ideas to achieve its goal.  First, Nail uses the protocol grammar to
-define not just the data format, but also the internal representation
+define not just the data format, but also the internal object model
 of the data.  Second, Nail eliminates the notion of semantic actions,
@@
 but allows Nail to both {\em parse} data formats and {\em generate}
-them from internal representations, by maintaining a {\em semantic
-bijection} between data formats and internal representation.  Third,
+them from the internal object model, by maintaining a {\em semantic
+bijection} between the data format and the object model.  Third,
 Nail introduces {\em dependency fields} and {\em stream transforms} to

### 2014-05-01 paper/osdi14/intro.tex
some intro nits

@@
 existing parser generators. Those parsers cannot directly deal with length or checksum fields
-directly, so either the programmer has to write potentially unsafe code to deal with such features
+directly, so the programmer has to either write potentially unsafe code to deal with such features,
 or build contrived grammar constructs, such as introducing one grammar rule for each possible value
-of a length fields. Offset fields usually require the programmer to manipulate a parser's internal
-state to re-position its input. More complicated transformations on the external format, such as
-compressed data, cannot be implemented at all.
+of a length field. Offset fields usually require the programmer to manipulate a parser's internal
+state to re-position its input. More complicated transformations, such as
+handling compressed data, cannot be implemented at all.
 
@@
 This removes the semantic actions and type declarations that programmers have to write with existing
-parser generator. While this somewhat reduces the flexibility of the internal model, it forces the programmer
+parser generators. While this somewhat reduces the flexibility of the internal model, it forces the programmer
 to clearly separate syntactic validation and semantic processing.
@@
 maintain a {\em semantic bijection} between data formats and their
-internal representation.  As a result, this enables Nail to not just
+internal object model.  As a result, this enables Nail to not just
 parse input but also generate output from the internal representation,
@@
 effective at handling complex real-world data formats, and that Nail
-makes it easy for application developers to handle external data formats.
+makes it easy for application developers to parse and generate external
+data representations.
 

### 2014-05-01 paper/osdi14/intro.tex
placeholders for citations in intro

@@
 Applications usually do not explicitly check for consistency, and if different applications use
-different instances of the same value, an attacker can craft input that causes inconsistencies.
+different instances of the same value, an attacker can craft input that causes inconsistencies~\cite{cve-python-zip-maybe}.
 Furthermore, security vulnerabilities can occur when an application assumes two repetitions of the
 same data to be consistent, such as allocating a buffer based on the value of one size field and copying into
-that buffer based on the value of another.
+that buffer based on the value of another~\cite{cve-some-example}.
 

### 2014-05-01 paper/osdi14/intro.tex
Typos

--- 2014-05-01 paper/osdi14/intro.tex
@@
 library has had 24 remotely exploitable vulnerabilities from 2007 to
-2013~\cite{cvedetails:libpng},
+2013,\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}}
 and Adobe's PDF and Flash viewers have been notoriously
@@
 library, have had input processing vulnerabilities in the
-past~\cite{cvedetails:zlib}.
+past.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}
 
@@
 Applications usually do not explicitly check for consistency, and if different applications use
-different instances of the same value, an attacker can craft input that causes inconsistencies.
+different instances of the same value, an attacker can craft inputs that cause inconsistencies.
 Furthermore, security vulnerabilities can occur when an application assumes two repetitions of the
@@
 state to re-position its input. More complicated transformations on the external format, such as
-compressed data, cannot be implemented at all.
+compressed data, cannot be represented at all.
 

### 2014-05-01 paper/osdi14/relwk.tex
Typos

@@
 terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing are transparently hidden by the library.
+endianness and bit-packing are handled transparently.
 
@@
 Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
-protocols with offset fields or length fields (Hammer has extremely limited support for length
-fields: it can parse arrays immediately preceded by their length).
+protocols with offset fields or length fields, although Hammer has special facilities for  arrays immediately preceded by their length.
 
@@
 serious security consequences when two systems disagree on the meaning of a network packet or a
-signed message, as in iOS\footnote{The XNU kernel and the user code-signing verifier interpret
+signed message, as in iOS\footnote{The XNU kernel and the user-mode code-signing verifier interpret
   executable metadata differently, so the code signature sees different bytes at a virtual address
@@
 Proper input recognition has been shown to be an excellent way of eliminating ambiguous inputs. In
-one case, a PDF parser implemented in Coq could eliminate over $95\%$ of known malicious
+one case, a PDF parser implemented in Coq rejects over $95\%$ of known malicious
 PDFs~\cite{Bogk-PDF}. The same project also produced a ``counter-example'' PDF that triggered

### 2014-05-01 paper/osdi14/design.tex
Fixed citations
Removed figure 2(DNS example) Re-wrote dependent fields

@@
 last.
-For example, \cc{many uint8} represents an array of 8-bit unsigned
-integers, and \cc{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
-comma-separated lists of decimal numbers.
+%For examle, \cc{many uint8} represents an array of 8-bit unsigned
+%integers, and \cc{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
+%comma-separated lists of decimal numbers.
 
@@
 
-\begin{figure}[tb]
-\smaller[0.5]
-\begin{minipage}{0.45\columnwidth}
-\begin{verbatim}
-dns_header = {
-  id uint16
-  qr uint1
-  opcode uint4
-  aa uint1 
-  tc uint1
-  rd uint1
-  ra uint1
-  uint3 = 0
-  rcode uint4
-}
-\end{verbatim} 
-\end{minipage}
-~
-\begin{minipage}{0.45\columnwidth}
-\begin{verbatim}
-struct dns_header {
-  uint16_t id;
-  uint8_t qr;
-  uint8_t opcode;
-  uint8_t aa;
-  uint8_t tc;
-  uint8_t rd;
-  uint8_t ra;
-  uint8_t rcode;
-};
-\end{verbatim} 
-\end{minipage}
+% \begin{figure}[tb]
+% \smaller[0.5]
+% \begin{minipage}{0.45\columnwidth}
+% \begin{verbatim}
+% dns_header = {
+%   id uint16
+%   qr uint1
+%   opcode uint4
+%   aa uint1 
+%   tc uint1
+%   rd uint1
+%   ra uint1
+%   uint3 = 0
+%   rcode uint4
+% }
+% \end{verbatim} 
+% \end{minipage}
+% ~
+% \begin{minipage}{0.45\columnwidth}
+% \begin{verbatim}
+% struct dns_header {
+%   uint16_t id;
+%   uint8_t qr;
+%   uint8_t opcode;
+%   uint8_t aa;
+%   uint8_t tc;
+%   uint8_t rd;
+%   uint8_t ra;
+%   uint8_t rcode;
+% };
+% \end{verbatim} 
+% \end{minipage}
 
-\caption{Nail grammar (left) and data model (right) for a part of the
-grammar for DNS packets.  The \cc{uint3 = 0} grammar represents
-3 bits of padding (filled with zeroes).}
-\label{fig:dns-struct}
-\end{figure}
+% \caption{Nail grammar (left) and data model (right) for a part of the
+% grammar for DNS packets.  The \cc{uint3 = 0} grammar represents
+% 3 bits of padding (filled with zeroes).}
+% \label{fig:dns-struct}
+% \end{figure}
 
@@
 
-A sample grammar for a DNS header and the corresponding C data type the Nail parser produces, are
-shown in Figure~\ref{fig:dns-struct}.
+% A sample grammar for a DNS header and the corresponding C data type the Nail parser produces, are
+% shown in Figure~\ref{fig:dns-struct}.
 
@@
 As discussed above, choosing to use these combinators on constant parsers
-removes the bijection between byte-strings and our data model, as there are
-multiple byte-strings that correspond to the same internal data structure and
-the generator has to choose one of these representations.
+weakens the bijection between the format and the data type, as there are
+multiple byte-strings that correspond to the same internal representation and
+the generator chooses one of these.
 \paragraph{Wrap combinator.} 
@@
 
-Dependent fields are defined within a structure like normal fields, but their name starts with an \@-sign. 
+Dependent fields are defined within a structure like normal fields, but their name starts with an $@$-sign. 
 A dependent field is in scope and can be referred to by the definition of all subsequent fields in
@@
 
-During parsing, dependent fields are parsed just like other fields, but instead of storing their
-value in a structure that will be returned to the programmer, the parser stores the value in a
-temporary location. Thereafter, the value can be referenced in the grammar.
+Dependent fields are parsed just like other fields in the external format, but they are not stored
+in the internal data type. Instead, the parser stores the value in a
+temporary location the field can be referenced by subsequent parsers.
 
-When generating output, Nail visits a dependent field twice. When generating the fields of a
-structure, it will reserve space for the dependent field in the output, but not write a value. When
-the dependent field goes out of scope, the generator will backtrack to this space and write the
-dependent fields value.
+When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
+structure, the generator reserves space for the dependent field in the output. Once
+the dependent field goes out of scope, the generator  write the
+dependent field's value to this space.
 

### 2014-05-01 paper/osdi14/relwk.tex
checkpoint relwk

--- 2014-05-01 paper/osdi14/relwk.tex
@@
 \XXX[Parsifal at langsec '14]
-\paragraph{Parsers}
 
-Generating parsers and generators from an executable specification
-is the core concept of Interface Generators, e.g. in CORBA or more
-recently~\cite{varda2008}. However, while interface generators work
-very well for existing grammars, they do not allow full control over the
-format of the output, so cannot be used to implement legacy protocols.
-Very related work has been done at Bell Labs with the PacketTypes
-system~\cite{mccann2000packet}, however PacketTypes works only as a
-parser, not as an output generator and does not support the expressive
+\paragraph{Parsers.}
+
+Generating parsers and generators from an executable specification is the
+core concept of interface generators, such as CORBA~\cite{omg:corba},
+XDR~\cite{rfc:1832}, or Protocol Buffers~\cite{varda2008}.  However,
+while interface generators work well for defining new protocols, they
+do not provide full control over the data format, and therefore cannot
+be used to implement existing protocols and data formats.
+
+Closely related work has been done at Bell Labs with the PacketTypes
+system~\cite{mccann2000packet}.  However, PacketTypes works only as a
+parser, not as an output generator, and does not support the expressive
 power of parsing expression grammars (PEGs), but rather implements a
-C-like structure model enhanced with  length fields and
-constraints. PacketTypes cannot handle complicated encodings such as  compressed data.
+C-like structure model enhanced with length fields and constraints.
+PacketTypes also cannot handle complicated encodings such as compressed
+data, which are supported by Nail's stream transforms.
 

### 2014-05-01 paper/osdi14/relwk.tex
checkpoint more relwk

@@
 \label{s:relwk}
-
-\XXX[Parsifal at langsec '14]
 
@@
 Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
-described in~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
-grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For more background on the history of
-  expressing grammars, see Bryan Ford's masters thesis~\cite{ford2002packrat}, which also describes
-  the default parsing algorithm used by Hammer.} by composing them from one or multiple sub-parsers.
+described by Burge~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
+grammars~\cite{Danielsson:2010:TPC:1863543.1863585}
+by composing them from one or multiple sub-parsers.\footnote{For more
+background on the history of expressing grammars, see Bryan Ford's
+masters thesis~\cite{ford2002packrat}, which also describes the
+default parsing algorithm used by Hammer.}
 Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
-an abstract syntax tree (AST).
+an AST\@.
 
-Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
-Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
-Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
-protocols with offset fields or length fields (Hammer has extremely limited support for length
-fields: it can parse arrays immediately preceded by their length).
+Nail improves upon Hammer in three ways.  First, Nail generates output in
+addition to parsing input.  Second, Nail does not require the programmer
+to write potentially insecure semantic actions.  Last, Nail's structural
+dependencies and stream transforms allow it to work with protocols that
+Hammer cannot handle, such as protocols with offset fields, length fields,
+checksums, or compressed data.  (Hammer has limited support for length
+fields: it can parse arrays immediately preceded by their length.)
 
-Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports generating output for Ocaml.
-Parsifal structures grammars as an Ocaml type that holds an internal model and functions for parsing
-input and output.
-However, their generator can only produce parsers and generators for simple, fixed-size structures.
-The programmer can then use these when implementing parsers and generators for more
-complicated formats, manually handling offsets, checksums and the like, risking bugs. Nail can
-handle more complicated constructs without the programmer writing code to support them.
+Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
+generating output for Ocaml.  Parsifal structures grammars as an Ocaml
+type that holds an internal model and functions for parsing input and
+output.  However, their generator can produce parsers and generators
+only for simple, fixed-size structures.  The programmer can then use
+these when implementing parsers and generators for more complicated
+formats, manually handling offsets, checksums, and the like, risking bugs.
+Nail handles more complicated constructs without the programmer manually
+writing code to support them.
 

### 2014-05-01 paper/osdi14/relwk.tex
checkpoint more relwk

@@
 \paragraph{Application use of parsers.}
-Generated parsers have long been used to parse human input, such as programming languages and
-configuration files. Frequently, such languages are often specified with a formal grammar in an
-executable form. Unfortunately,  parser frameworks are seldom used  to recognize
-machine-created input.
 
-A notable exception is the Mongrel
-web server~\cite{mongrel} which uses a grammar
-for HTTP written in the Ragel~\cite{ragel-paper} regular expression
-language. Mongrel was re-written from scratch multiple times to achieve
-better scalability and design, yet the grammar could be  re-used across
-all iterations~\cite{patterson-citation}.
+Generated parsers have long been used to parse human input, such
+as programming languages and configuration files. Frequently, such
+languages are often specified with a formal grammar in an executable
+form. Unfortunately, parser frameworks are seldom used to recognize
+machine-created input; as we demonstrate in \S\ref{s:eval},
+state-of-the-art parser generators are not suitable for parsing or
+generating many real-world data formats.
 
-\paragraph{Other vulnerability mitigations.}
+A notable exception is the Mongrel web server~\cite{mongrel} which
+uses a grammar for HTTP written in the Ragel regular expression
+language~\cite{ragel-paper}.  Mongrel was re-written from scratch
+multiple times to achieve better scalability and design, yet the
+grammar was reused across all iterations~\cite{patterson-citation}.
+We hope that Nail's ideas make it possible to handle a wider range of
+protocols using parser generators, and to build more applications on
+top of grammar-based parsers.
+
+\paragraph{Vulnerability mitigation.}
 

### 2014-05-01 paper/osdi14/relwk.tex
done with relwk pass

@@
 
+We presented an earlier design of Nail at a
+workshop~\cite{bangert:nail-langsec-anon}.  At that stage, Nail had only
+limited support for dependent fields, and did not support stream
+transforms at all, which are crucial for supporting real-world formats
+like DNS and Zip.  The workshop paper also did not provide a detailed
+design discussion or evaluation.
+
 \paragraph{Application use of parsers.}
@@
 
-The second class is logic errors, where application code misinterprets input data. Safe languages
-and exploit mitigation technologies do not help against such vulnerabilities. This can lead to
-serious security consequences when two systems disagree on the meaning of a network packet or a
-signed message, as in iOS\footnote{The XNU kernel and the user code-signing verifier interpret
-  executable metadata differently, so the code signature sees different bytes at a virtual address
-  than the executable that runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing
-and even the X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Logic errors are often
-causing memory corruption bugs as well, for example when a protocol stores the size of a buffer in
-two places and the implementation does not check their consistency. These mistakes are
-highly application-specific, and are difficult to mitigate using existing techniques, and these
-mistakes can occur even in high-level languages that guarantee memory safety. By allowing developers
-to specify their data format just once, Nail avoids logic errors and inconsistencies in parsing and
-output generation.
+The second class is logic errors, where application code misinterprets
+input data. Safe languages and exploit mitigation technologies do not
+help against such vulnerabilities. This can lead to serious security
+consequences when two systems disagree on the meaning of a network
+packet or a signed message.  For example, the XNU kernel in iOS and the
+user code-signing verifier interpret executable metadata differently,
+so the code signature sees different bytes at a virtual address than the
+executable that runs~\cite{evaders6}.  Android's code signing system also
+had a vulnerability due to parsing ambiguities~\cite{saurik-masterkey},
+and parsing ambiguities even contributed to attacks on the X.509 protocol
+underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  Logic errors can also
+cause memory corruption bugs as well, for example when a protocol stores
+the size of a buffer in two places and the implementation does not check
+their consistency. These mistakes are highly application-specific, and
+are difficult to mitigate using existing techniques, and these mistakes
+can occur even in high-level languages that guarantee memory safety. By
+allowing developers to specify their data format just once, Nail avoids
+logic errors and inconsistencies in parsing and output generation.
 
@@
 
-Proper input recognition has been shown to be an excellent way of eliminating ambiguous inputs. In
-one case, a PDF parser implemented in Coq could eliminate over $95\%$ of known malicious
-PDFs~\cite{Bogk-PDF}. The same project also produced a ``counter-example'' PDF that triggered
-numerous bugs in other PDF parsers. However, manually writing parser code and proving unambiguity
-does not scale to the number of file formats and protocols in existence and might result in parser
-code tied to one specific application.
+Proper input recognition has been shown to be an excellent way of
+eliminating ambiguous inputs.  In one case, a PDF parser (implemented
+in Coq) could eliminate over $95\%$ of known malicious PDFs by checking
+whether they conform to the PDF format~\cite{Bogk-PDF}. The same project
+also produced a ``counter-example'' PDF that triggered numerous bugs
+in other PDF parsers. However, manually writing parser code and proving
+unambiguity does not scale to the number of file formats and protocols
+in existence and might result in parser code tied to one specific
+application. By supporting a wide range of real-world file formats,
+Nail helps programmers produce precise parsers for complex grammars.
 
-\paragraph{Workshop paper}
-We presented an early design of Nail at the LangSec workshop~\cite{bangert:nail-langsec} that had
-no support for transformations, as described in Section~\ref{s:transforms}, and much more limited
-dependent fields, as described in Section~\ref{s:dependent}. With our extended and refined design,
-Nail can support much more complicated, real-world data formats.
 %%This is supposed to go into design

### 2014-05-01 paper/osdi14/design.tex
make spell

--- 2014-05-01 paper/osdi14/design.tex
@@
 
-Nail produces the following from a single, descriptive grammmar: 
+Nail produces the following from a single, descriptive grammar: 
 \begin{itemize}
@@
 
-\paragraph{Semantic Bijecction.}
+\paragraph{Semantic Bijection.}
 Second, in order to generate output, Nail maintaining
@@
 A Nail parser defines both the structure of some external format and a data type to represent that
-format. Parsers are constructed by combinators over simpler parsers, an approach popularised by the Parsec
+format. Parsers are constructed by combinators over simpler parsers, an approach popularized by the Parsec
 framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
@@
 Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
-approach popularised by the Hammer framework.\cite{hammer-parser}  
+approach popularized by the Hammer framework.\cite{hammer-parser}  
 Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
@@
 \paragraph{Choices.}
-Our choice combinator takes multiple labelled parsers and tries each of them in turn, backtracking
-to the initial position if an option fails. This behaviour for the choice combinator, the defining
+Our choice combinator takes multiple labeled parsers and tries each of them in turn, backtracking
+to the initial position if an option fails. This behavior for the choice combinator, the defining
 feature of Parser Expression Grammars\cite{ford2002packrat}, resolves any ambiguities explicitly.
@@
 
-\paragraph{Optional.}Nail also includes an \cc{optional} combinator, which attempts to recognise a value, but succeeds
-without consuming input when it cannot recognise that value. Syntactically, \cc{optional} is
+\paragraph{Optional.}Nail also includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
+without consuming input when it cannot recognize that value. Syntactically, \cc{optional} is
 equivalent to a choice between the parser and an empty structure, but in the internal model it is

### 2014-05-01 paper/osdi14/eval.tex
make spell

@@
 awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
-Keeping track of the position all recognised labels would not be enough, as the offset field may
+Keeping track of the position all recognized labels would not be enough, as the offset field may
 refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
@@
 utility\cite{infozip} that is shipped with most Linux distributions. The entire unzip distribution
-is about 46.000 lines of code, which is mostly optimised implementation of various compression
+is about 46.000 lines of code, which is mostly optimized implementation of various compression
 algorithms and other configuration and portability code.
@@
 We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
-analyse a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
+analyze a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
 immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
@@
 queries outstanding at once. The queryperf tool was configured to
-repeat the same, randomised sequence of queries for one minute and the average throughput and
+repeat the same, randomized sequence of queries for one minute and the average throughput and
 response latency were measured.  We repeated each test seven times  with 50.000
@@
 Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
-of optimisation by a factor of 3.  This demonstrates that even though Nail was
+of optimization by a factor of 3.  This demonstrates that even though Nail was
 not developed focusing on performance, applications built on Nail can compete with real-world

### 2014-05-01 paper/osdi14/relwk.tex
make spell

@@
 Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
-generating output for Ocaml.  Parsifal structures grammars as an Ocaml
+generating output for OCaml.  Parsifal structures grammars as an OCaml
 type that holds an internal model and functions for parsing input and

### 2014-05-01 paper/osdi14/relwk.tex

--- 2014-05-01 paper/osdi14/relwk.tex
@@
 
-Nail improves upon Hammer in three ways.  First, Nail generates output in
-addition to parsing input.  Second, Nail does not require the programmer
-to write potentially insecure semantic actions.  Last, Nail's structural
-dependencies and stream transforms allow it to work with protocols that
-Hammer cannot handle, such as protocols with offset fields, length fields,
-checksums, or compressed data.  (Hammer has limited support for length
-fields: it can parse arrays immediately preceded by their length.)
+Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
+Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
+Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
+protocols with offset fields or length fields, although Hammer has special facilities for  arrays immediately preceded by their length.
 
-Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
-generating output for Ocaml.  Parsifal structures grammars as an Ocaml
-type that holds an internal model and functions for parsing input and
-output.  However, their generator can produce parsers and generators
-only for simple, fixed-size structures.  The programmer can then use
-these when implementing parsers and generators for more complicated
-formats, manually handling offsets, checksums, and the like, risking bugs.
-Nail handles more complicated constructs without the programmer manually
-writing code to support them.
+Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports generating output for Ocaml.
+Parsifal structures grammars as an Ocaml type that holds an internal model and functions for parsing
+input and output.
+However, their generator can only produce parsers and generators for simple, fixed-size structures.
+The programmer can then use these when implementing parsers and generators for more
+complicated formats, manually handling offsets, checksums and the like, risking bugs. Nail can
+handle more complicated constructs without the programmer writing code to support them.
 

### 2014-05-01 paper/osdi14/design.tex
Stream writeup

@@
 \label{s:transforms}
-In order to handle complicated input functions, Nail allows the programmer to
-break out of Nail's generated code in a controlled fashion, by manipulating dependent fields and
-streams. 
-By default, Nail implements streams with in-memory buffers, but the programmer can override this implementation.
-While parsing or generating a Nail grammar, there is always an implicit \emph{current stream} and
-referred to as \cc{\$current} in the Nail language.
+Traditional parsers handle input one symbol at a time, beginning to end.
+However, real-world formats often require non-linear parsing. Offset fields require a parser to move
+to a different position in the input, possibly backwards. Size fields require the parser to stop
+processing before the end of input has been reached and perhaps continue executing a parent parser.
+Other cases, such as compressed data require more complicated processing on parts of the input
+before it can be handled.
 
-The programmer can define additional streams by writing a \emph{transformation}, a pair of
-functions, one that generates new streams from existing streams during parsing and one that
-combines streams during output generation. Temporary streams follow the naming and
-scoping convention of dependent fields, but their names start with a \$-sign.
+Nail introduces two concepts to handle this, \emph{streams} and \emph{transformations}. 
+Streams represent a sequence of bytes that contain some external format. The parsers and generators
+Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
+back, reading input or appending output.
 
-The parsing function of the transformation takes any number of streams, either temporary or the current,
-and outputs any number of temporary streams. The parsing function may reposition and read from the
-input streams and read the values of dependent fields, but not change their values. The generating
-function has to be an inverse of the parsing function. It takes the same set of temporary streams
-the parsing function returned and is supposed to change the values of the same set of values the
-generating function produced.
+Grammars can use the \cc{apply} combinator to parse or generate external data on a different stream,
+inserting the result in the data model. Streams can be defined in two ways.
+
+First, a rule can require streams as arguments, which can be passed in by the program that invokes
+the generated parser.Every rule needs toe passed at least the implicit current stream.
+
+Second, the programmer can create temporary streams in the grammars with \emph{transformations},
+which are two arbitrary functions called during parsing and output generation.
+The parsing function  takes any number of stream arguments and dependent field values
+and produces any number of temporary streams. This function may reposition and read from the
+input streams and read the values of dependent fields, but not change their contents and values. 
+The generating function has to be an inverse of the parsing function. It takes the same number of temporary streams
+the parsing function produces and and writes the same number of streams and dependent field
+values the parsing function consumes.
+
+Typically, the top level of most grammars is a rule that only takes a single stream, which may then
+be broken up by various transformations and passed to sub-rules, which eventually parse various linear
+fragment streams. Upon parsing, these fragment streams are generated and then combined by the
+transforms.
 

### 2014-05-01 paper/osdi14/relwk.tex

--- 2014-05-01 paper/osdi14/relwk.tex
@@
 Hammer cannot handle, such as protocols with offset fields, length fields,
-checksums, or compressed data.  (Hammer has limited support for length
-fields: it can parse arrays immediately preceded by their length.)
+checksums, or compressed data, although Hammer has special facilities for
+arrays immediately preceded by their length.
 
@@
 packet or a signed message.  For example, the XNU kernel in iOS and the
-user code-signing verifier interpret executable metadata differently,
+user-mode code-signing verifier interpret executable metadata differently,
 so the code signature sees different bytes at a virtual address than the
@@
 eliminating ambiguous inputs.  In one case, a PDF parser (implemented
-in Coq) could eliminate over $95\%$ of known malicious PDFs by checking
+in Coq) rejects over $95\%$ of known malicious PDFs by checking
 whether they conform to the PDF format~\cite{Bogk-PDF}. The same project

### 2014-05-01 paper/osdi14/design.tex
pass through 3.1

@@
       storing both an explicit length field and an implicit length
-      of a container data structure, to ensure that programmers
-      see a consistent view of the data.
+      of a container data structure, to provide programmers
+      an unambiguous view of the data.
 
@@
 
-\paragraph{Internal Model.}
-Nail grammars describe both the \emph{external format} and an \emph{internal representation} of a
-protocol, as opposed to existing grammar languages describe only the former and leave it up to the programmer to
-construct an internal representation, typically in the form of an abstract syntax tree.
+\paragraph{Internal model.}
 
+Nail grammars describe both the \emph{external format} and an
+\emph{internal representation} of a protocol, as opposed to existing
+grammar languages, which describe only the former and leave it up to
+the programmer to construct an internal representation, typically in
+the form of an abstract syntax tree.
 Nail produces the following from a single, descriptive grammar:
-\begin{itemize}
-\item type declarations for the internal model,
+
+\begin{CompactItemize}
+\item \textit{type declarations} for the internal model,
 \item the \textit{parser}, a function to parse a sequence of bytes into an
@@
 sequence of bytes from an instance of the model.
-\end{itemize}
+\end{CompactItemize}
 
@@
 
-\paragraph{Semantic Bijection.}
-Second, in order to generate output, Nail maintaining
-a \emph{semantic bijection} between the external format and the internal model. That is, when
-generating the corresponding output to a parser input, the two byte streams will have the same
-meaning, i.e. be interpreted equivalently by Nail, but might not be identical.
+\paragraph{Semantic bijection.}
+
+In order to generate output, Nail maintains
+a \emph{semantic bijection} between the external format and the internal model.
+That is, when Nail parses an input into an internal representation, and then
+generates output from that representation, the two byte streams (input and
+output) will have the same meaning (i.e., be interpreted equivalently by Nail).
+However, the byte streams might not be identical.
 A bijection in the traditional sense often does not make sense for data
@@
 Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
-consists of a simple protocol without offset fields, constants and the like, there is a conventional
+consists of a simple protocol without offset fields, constants, and the like, there is a conventional
 bijection between  internal models and valid parser inputs.
+
 \paragraph{Hide redundant information.}
@@
 Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
-and needs to be kept as additional state during parsing. Dependent fields are for example used to represent
-lengths, offsets and checksums.
+and need to be kept as additional state during parsing. Dependent fields are, for example, used to represent
+lengths, offsets, and checksums.
 
-\paragraph{Parser Extensions.}
-Real-world protocols contain complicated ways of encoding data. Representing these in an
-intentionally limited model such as our parser language is complicated. Therefore, Nail introduces
+\paragraph{Parser extensions.}
+Real-world protocols contain complicated ways of encoding data. Fully representing these in an
+intentionally limited model such as our parser language is impractical. Therefore, Nail introduces
 \emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
@@
 functions in a general purpose language that consume streams and define new temporary streams, while
-also reading or writing the value of dependent fields.
+also reading or writing the values of dependent fields.
 
-Initial versions of Nail included an offset combinator, which consumed a dependent field and applied
-a parser at the offset specified therein. However, it proved impossible to foresee all the ways in which a protocol could encode an offset, for example
-some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
-end of the file or a fixed offset. In nested grammars, offsets are also not necessarily from
+Initial versions of Nail's design included a special combinator for handling offset fields,
+which consumed a dependent field and applied
+a parser at the offset specified therein. However, it proved impossible to foresee all the ways in which a protocol could encode an offset;
+for example, some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
+end of the file or a fixed offset. In nested grammars, offsets are also not necessarily computed from
 the beginning of a file or packet.
+Nail's transformations allow the programmer to write arbitrary functions that can
+handle such structures and streams, which are a generic abstraction for input and
+output data that allow the decoded data to be integrated with the rest of the
+generated Nail parser.
 
-Nail introduce transformations, which allow the programmer to write arbitrary functions that can
-handle such structures and streams, which are a generic abstraction for input and output data that
-allow the decoded data to be integrated with the rest of the generated Nail parser.
+\input{fig-syntax}
 
-In the rest of this section, we describe the design on the Nail grammar language.
-\input{fig-syntax}
 \subsection{Basics}
+
 A Nail parser defines both the structure of some external format and a data type to represent that

### 2014-05-01 paper/osdi14/design.tex
pass through 3.2

@@
 format. Parsers are constructed by combinators over simpler parsers, an approach popularized by the Parsec
-framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
- combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
+framework~\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
+ combinator libraries, such as Parsec and Hammer~\cite{hammer-parser} and extend them so they also
  describe a data type. 
+Figure~\ref{fig:syntax} shows the elements of Nail grammars, which we describe in more detail below.
 
-Figure~\ref{fig:syntax} shows the elements of Nail grammars, which we describe in more detail below.
-\paragraph{Integers and Constraints}
+\paragraph{Integers and constraints.}
 
-Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
-approach popularized by the Hammer framework.\cite{hammer-parser}  
-Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
+Nail's fundamental parsers represent signed or unsigned
+integers with arbitrary lengths up to 64 bits, an
+approach popularized by the Hammer framework~\cite{hammer-parser}.
+Note that is possible to define parsers for sub-byte lengths, e.g.,
+to parse the 4-bit
 data offset within the TCP header.  The internal representation rounds up to the nearest integer
-type, for example a 24-bit unsigned integer will be stored in an unsigned 32-bit double word.
+type; for example, a 24-bit unsigned integer will be stored in an unsigned 32-bit integer.
 
-The grammar can also constraint the values of an integer. The current version of Nail expresses
+The grammar can also constrain the values of an integer.  Nail expresses
 constraints as a set of permissible values or value ranges. Extending the Nail language and
@@
 The \emph{many} combinator takes a parser and applies it repeatedly
-until it fails, returning an array of the inner parsers results. The
+until it fails, returning an array of the inner parser's results. The
 \emph{sepBy} combinator
@@
 two values, but not before parsing the first value or after parsing the
-last.
+last.  This is useful for parsing an array of items delimited by a separator.
 %For examle, \cc{many uint8} represents an array of 8-bit unsigned
@@
 parser that describes the contents of that field. Other field types will be described below.
- The syntax of the structure combinator is inspired by the Go language\cite{golang}, with field names preceding their definition.
+ The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field names preceding their definition.
 In our opinion, this syntax is easier to read than the widespread C-style structure and type
@@
 Nail therefore allows repeated constants, such as \cc{many (uint8=' ')}, which parses any number of space characters, or
-\cc{|| uint8 = 0x90 || uint16 = 0x1F0F}, which parses two of the many representations for X86
+\cc{|| uint8 = 0x90 || uint16 = 0x1F0F}, which parses two of the many representations for x86
 NOP instructions, which are used as padding between basic blocks in an executable.
@@
 the generator chooses one of these.
+
 \paragraph{Wrap combinator.} 
-When implementing real protocols with Nail, we often found that
-structures that consist of many constant parsers and only one named field. This pattern is
+When implementing real protocols with Nail, we often found
+structures that consist of many constants and only one named field. This pattern is
 common in binary protocols which use fixed headers to denote the type of data
@@
 (and thus application code) more concise.
-
 For example, \cc{<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
@@
 to the initial position if an option fails. This behavior for the choice combinator, the defining
-feature of Parser Expression Grammars\cite{ford2002packrat}, resolves any ambiguities explicitly.
+feature of Parser Expression Grammars~\cite{ford2002packrat}, resolves any ambiguities explicitly.
 Internally, the choice combinator represents choices as tagged unions.
@@
 usually not ambiguous in this sense.
+\XXX[nz][I don't understand this paragraph (esp the first sentence).]
 
-\paragraph{Optional.}Nail also includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
+\paragraph{Optional.}
+Nail includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
 without consuming input when it cannot recognize that value. Syntactically, \cc{optional} is
@@
 A Nail grammar consists of rules that assign a parser to a name. Rules are written as
-assignments, such as \texttt{ints = many uint8}, which defines a rule called \texttt{ints} with the
-body \texttt{many uint8}.  As we will describe in Sections~\ref{s:dependent}
-and~\ref{s:transforms}, rules can optionally consume parameters. 
+assignments, such as \cc{ints = many uint8}, which defines a rule called \cc{ints} with the
+body \cc{many uint8}.  As we will describe in \S\ref{s:dependent}
+and \S\ref{s:transforms}, rules can optionally consume parameters. 
 Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
@@
 
-\paragraph{References.}Rules allow for recursive grammars. To support recursive data types, we allow introduce the
+\paragraph{References.}
+Rules allow for recursive grammars. To support recursive data types, we allow introduce the
 reference combinator \cc{*}  that does not change the syntax of the external format described, but
 introduces a layer of indirection, such as a reference or pointer, to the model data type.
-
-For example example,   balanced parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren;uint8=")">}
+For example example,   balanced parentheses could be parsed as follows: \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
 

### 2014-05-01 paper/osdi14/design.tex
pass through 3.3

@@
 Data formats often contain values that are determined by other values or the layout of information,
-such as checksums, duplicated information  or offset and  length fields.
-We represent such as values with \emph{dependent fields} and handle them transparently during
+such as checksums, duplicated information, or offset and  length fields.
+We represent such values using \emph{dependent fields} and handle them transparently during
 parsing and generation without exposing them to the internal model.
 
-Dependent fields are defined within a structure like normal fields, but their name starts with an $@$-sign. 
+Dependent fields are defined within a structure like normal fields, but their name starts with an \cc{@} symbol.
 A dependent field is in scope and can be referred to by the definition of all subsequent fields in
-the same structure. Dependency fields can be passed to rule invocations as parameters.
+the same structure. Dependent fields can be passed to rule invocations as parameters.
 
@@
 in the internal data type. Instead, the parser stores the value in a
-temporary location the field can be referenced by subsequent parsers.
+temporary location, so that the field can be referenced by subsequent parsers.
 
@@
 structure, the generator reserves space for the dependent field in the output. Once
-the dependent field goes out of scope, the generator  write the
+the dependent field goes out of scope, the generator  writes the
 dependent field's value to this space.
 
-Nail only provides one built-in combinator, \cc{n_of} that uses dependency fields, which is used
+Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which is used
 for count fields. For example, DNS labels are encoded as Pascal-style strings: a length followed by
 the value of the label. The Nail grammar for DNS labels can represent
-this pattern as follows:
+this pattern as:
 
@@
 
-Other dependencies, such as offset fields or checksums are not handled directly by combinators, but
-through  transformations.
+Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
+through  transformations, as we describe next.
 

### 2014-05-01 paper/osdi14/design.tex
pass over 3.4

@@
 
-\subsection{Input streams and Transformations.}
+\subsection{Input streams and transformations}
 \label{s:transforms}
-Traditional parsers handle input one symbol at a time, beginning to end.
+
+Traditional parsers handle input one symbol at a time, from beginning to end.
 However, real-world formats often require non-linear parsing. Offset fields require a parser to move
 to a different position in the input, possibly backwards. Size fields require the parser to stop
-processing before the end of input has been reached and perhaps continue executing a parent parser.
-Other cases, such as compressed data require more complicated processing on parts of the input
+processing before the end of input has been reached, and perhaps resume executing a parent parser.
+Other cases, such as compressed data, require more complicated processing on parts of the input
 before it can be handled.
@@
 Streams represent a sequence of bytes that contain some external format. The parsers and generators
-Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
+that Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
 back, reading input or appending output.
+Grammars can use the \cc{apply} combinator to parse or generate external data on a different stream,
+inserting the result in the data model.
 
-Grammars can use the \cc{apply} combinator to parse or generate external data on a different stream,
-inserting the result in the data model. Streams can be defined in two ways.
+Streams can be defined in two ways.
+First, a rule can require streams as arguments, which can be passed in by the program that invokes
+the generated parser. Every rule needs to be passed at least the implicit current stream.
 
-First, a rule can require streams as arguments, which can be passed in by the program that invokes
-the generated parser.Every rule needs toe passed at least the implicit current stream.
-
-Second, the programmer can create temporary streams in the grammars with \emph{transformations},
-which are two arbitrary functions called during parsing and output generation.
-The parsing function  takes any number of stream arguments and dependent field values
+Second, the programmer can create temporary streams in the grammar with \emph{transformations},
+which are two arbitrary functions called during parsing and output generation respectively.
+The parsing function  takes any number of stream arguments and dependent field values,
 and produces any number of temporary streams. This function may reposition and read from the
 input streams and read the values of dependent fields, but not change their contents and values. 
-The generating function has to be an inverse of the parsing function. It takes the same number of temporary streams
-the parsing function produces and and writes the same number of streams and dependent field
-values the parsing function consumes.
+The generating function has to be an inverse of the parsing function. It takes the same number of
+temporary streams that the parsing function produces, and writes the same number of streams and
+dependent field values that the parsing function consumes.
 
-Typically, the top level of most grammars is a rule that only takes a single stream, which may then
+Typically, the top level of most grammars is a rule that takes only a single stream, which may then
 be broken up by various transformations and passed to sub-rules, which eventually parse various linear
@@
 reduce both programmer effort and the risk of unsafe operations, we provide implementations of
-transformations for many common features, such as checksums and size and offset fields. Furthermore,
+transformations for many common features, such as checksums, size, and offset fields. Furthermore,
 we provide library functions that can be used to safely operate on streams, such as splitting and
@@
 \paragraph{Offsets.}
-We provide a transformation for handling offset fields that is invoked as follows:
+We provide a transformation for handling offset fields, which is invoked as follows:
 \cc{\$fragment transform offset\_u32(\$file\_stream, \@offset)}. 
-They are implemented by two functions, as shown in Figure~\ref{fig:xform-sig}.
+This transformation corresponds to two functions for parsing and
+generation, as shown in Figure~\ref{fig:xform-sig}.
 
@@
 \paragraph{Compressed data.}
-Encoded, compressed or encrypted data can be handled transparently by writing a custom
+Encoded, compressed, or encrypted data can be handled transparently by writing a custom
 transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
@@
 the checksum being set to some particular value. Because the functions implementing a transformation
-are passed a pointer to any dependent fields, the checksum function can set the checksums initial
+are passed a pointer to any dependent fields, the checksum function can set the checksum's initial
 value before calculating the checksum over the entire buffer, including the checksum.
 
-\noindent A real-world example with many different transforms is explained in Section~\ref{s:eval-format-zip}.
+\paragraph{}
+A real-world example with many different transforms, used to support
+the Zip file format, is described in \S\ref{s:eval-format-zip}.
 

### 2014-05-01 paper/osdi14/relwk.tex
x

@@
 eliminating ambiguous inputs.  In one case, a PDF parser (implemented
-in Coq) rejects over $95\%$ of known malicious PDFs by checking
+in Coq) was used to reject over $95\%$ of known malicious PDFs by checking
 whether they conform to the PDF format~\cite{Bogk-PDF}. The same project

### 2014-05-01 paper/osdi14/impl.tex
do not mention the url due to anonymous submission
pass over impl

@@
 The current prototype of the Nail parser generator supports the C programming
-language and top-down parsers. The implementation parses Nail grammars with Nail itself, using a 130
-line Nail grammar feeding into a 2000 line C++ program that implements the parser. Bootstrapping is
+language and top-down parsers. The implementation parses Nail grammars with Nail itself, using a
+130-line Nail grammar feeding into a 2,000-line C++ program that
+implements the parser. Bootstrapping is
 performed with a subset of the grammar implemented using conventional grammars. 
@@
 
-The source code of our implementation, together with the examples described in
-\S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
+% The source code of our implementation, together with the examples described in
+% \S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
 
@@
 Thus, for each of those combinators, we store the minimum amount of information
-required to reconstruct the syntactic structure of the input. The trace is an
+required to reconstruct the syntactic structure of the input.
+
+The trace is an
 array of integers.
-Whenever the parser encounters a choice, it appends two integers to the trace. After it
-successfully parses a choice, the parser writes the number of that choice and
+Whenever the parser encounters a choice, it appends two integers to the trace:
+the number of that choice and
 the length of the trace when it began parsing that choice. When backtracking in
 the input, the parser does not backtrack in the trace. This means that offsets
-within the trace can be used for a Packrat hash-table to memoize backtrack-heavy
+within the trace can be used for a Packrat hash table to memoize backtrack-heavy
 parsers.
-
 When encountering a repetition combinator, the parser records the number of

### 2014-05-01 paper/osdi14/eval.tex
checkpoint eval (partway through 5.1)

@@
 
-In our evaluation of Nail, we try to answer four questions:
+In our evaluation of Nail, we answer four questions:
 
@@
 \label{s:eval-formats}
-To answer the first question, we implemented grammars for four protocols with particularly
-challenging features and show that Nail is a good fit for these data formats and some aspects of
-these formats would be very hard to parse with other parser generators.
-Figure~\ref{fig:eval-protocols} lists various example protocols we have implemented.
+
+To answer the first question, we used Nail to implement grammars
+for seven protocols with a range of challenging features.
+Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
+of code for their Nail grammars, and the challenging features that make
+the protocols difficult to parse with state-of-the-art parser generators.
+We find that despite the challenging aspects of these protocols, Nail is
+able to capture the protocols, by relying on its novel features: dependent
+fields, streams, and transforms.  In contrast, state-of-the-art parser
+generators would be unable to fully handle 5 out of the 7 data formats.
 
 \begin{figure}[tb]
-\label{fig:eval-protocols}
-\begin{tabular}{@{}lll@{}}
+\centering
+\begin{tabular}{lrl}
 \toprule
-\textbf{Protocol} & \textbf{LoC} & \textbf{Problematic features} \\ 
+\textbf{Protocol} & \textbf{LoC} & \textbf{Challenging features} \\ 
 \midrule
-DNS messages &  48+64 & Label compression\\
-& & Count fields \\
-ZIP archives & 92 + 78 & Checksums, \\ 
-& & compression, offsets, \\
-& &  variable length trailer\\ 
-Ethernet frames & 16 & \\
-ARP &  10  & \\
-IP &  25 & Total Length field, options \\
-UDP & 7 & Checksum, Length field \\
+DNS packets & 48+64 & Label compression,\\
+  & & count fields \\
+ZIP archives & 92+78 & Checksums, offsets, \\ 
+  & & variable length trailer, \\
+  & & compression \\
+Ethernet & 16 & --- \\
+ARP & 10 & --- \\
+IP & 25 & Total length field, options \\
+UDP & 7 & Checksum, length field \\
 ICMP & 5 & Checksum \\
 \bottomrule
+\end{tabular}
 
-\end{tabular}
-\caption{Example Nail grammars, their code size of the grammar plus  the transformations they use
-  and protocol features that cannot be expressed in existing grammar languages. }
+\caption{Protocols, sizes of their Nail grammars, and challenging aspects
+of the protocol that cannot be expressed in existing grammar languages.
+A + symbol counts lines of Nail grammar code (before the +) and lines of
+C code for protocol-specific transforms (after the +).}
+\label{fig:eval-protocols}
 \end{figure}
+
 \subsubsection{DNS.}

### 2014-05-01 paper/osdi14/eval.tex
x

@@
   & & compression \\
-Ethernet & 16 & --- \\
-ARP & 10 & --- \\
-IP & 25 & Total length field, options \\
-UDP & 7 & Checksum, length field \\
-ICMP & 5 & Checksum \\
+Ethernet  & 16+0\phantom{0} & --- \\
+ARP       & 10+0\phantom{0} & --- \\
+IP        & 25+0\phantom{0} & Total length field, options \\
+UDP       &  7+0\phantom{0} & Checksum, length field \\
+ICMP      &  5+0\phantom{0} & Checksum \\
 \bottomrule

### 2014-05-01 paper/osdi14/eval.tex
pass over the DNS text in 5.1; fix the DNS grammar

@@
 generators would be unable to fully handle 5 out of the 7 data formats.
+In the rest of this subsection, we describe the DNS and Zip grammars in
+more detail, focusing on how Nail's features enable us to support these
+formats.
 
@@
 
-\subsubsection{DNS.}
+\paragraph{DNS.}
+
+Figure~\ref{fig:dns-xform} presents the Nail grammar for DNS packets.
+The grammar corresponds almost directly to the diagrams in RFC 1035,
+which defines DNS~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
+header, a set of question records, and a set of answer records. Domain
+names in both queries and answers are encoded as a sequence of labels,
+terminated by a zero byte.  Labels are Pascal-style strings, consisting
+of a length field followed by that many bytes comprising the label.
+
+One challenging aspect of DNS packets lies in the count fields (\cc{qc},
+\cc{ac}, \cc{ns}, and \cc{ar}), which represent the number of questions
+or answers in another part of the packet.  Nail's \cc{n_of} combinator
+handles this situation easily, which would have been difficult to handle
+for other parsers.
+
+Another challenging aspect of DNS is label
+compression~\cite[\S4.1.4]{RFC:1035}.  Label compression is used to reduce
+the size overhead of including each domain name multiple times in a DNS
+reply (once in the question section, and at least once in the response
+section).  If a domain name suffix is repeated, instead of repeating that
+suffix, the DNS packet may write a two-bit marker sequence followed by
+a 14-bit offset into the packet, indicating the position of where that
+suffix was previously encoded.
 
 \input{fig-dnsfull}
+
+Handling label compression in existing tools, such as Bison or Hammer,
+would at best be very awkward, because some ad-hoc trick would have to
+be used to re-position the parser's input stream.  Keeping track of the
+position of all recognized labels would not be enough, as the offset field
+may refer to any byte within the packet, not just the beginning of labels.
+For this reason, the DNS server used as the example for Hammer does not
+support compression.
 
@@
 
-Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
-corresponds to the diagrams in Section 4 of that document. Each DNS packet consists of a header, a
-set of question records and a set of answer records. Domain names in both queries and answers are
-encoded as a sequence of Pascal-style strings. Each label is preceded by a length field, and the
-domain name is terminated by a $0$ octet. The client's query packet contains one question
-record to the server for each DNS look-up it wants to perform. In addition to the appropriate
-response records, the server's reply contains a duplicate of the client's question section, which
-allows stateless DNS clients and caches.
+In contrast, Nail is able to handle label compression, by using a stream
+transform; the signatures of the two transform functions are shown
+in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
+decompresses the DNS label stream by following the offset pointers.
+When generating a packet, this transform receives the current suffix as
+an input, and scans the packet so far for previous occurrences, which
+implements compression.
 
-To reduce the size overhead of including each domain name twice in a DNS reply, once in the
-question section, at least once in the response record, DNS includes a label compression scheme. If
-a domain name suffix is repeated, instead of repeating that suffix, the DNS packet may write a
-two-bit marker sequence followed by a 14-bit offset into the packet where that suffix was previously
-encoded.
+\paragraph{ZIP files.}
 
-Processing these encodings in existing tools, such as Bison or Hammer, would at best be very
-awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
-Keeping track of the position all recognized labels would not be enough, as the offset field may
-refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
-server used as a tutorial example for Hammer does not support compression.
-
-Nail handles compression with a stream transform. When parsing, this transform decompresses the DNS
-label stream by following the offset pointers. When generating, this transform receives the current
-suffix as an input and can scan the packet so far for previous occurrences.
-
-\subsubsection{ZIP Files.}
 \input{fig-zip}

### 2014-05-01 paper/osdi14/code/zip-eod-xform.tex
pass through 5.1

\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
\PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}end\PYZus{}of\PYZus{}directory\PYZus{}parse}\PY{p}{(}
  \PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,} \PY{n}{NailStream} \PY{o}{*}\PY{n}{out\PYZus{}files}\PY{p}{,}
  \PY{n}{NailStream} \PY{o}{*}\PY{n}{out\PYZus{}dir}\PY{p}{,} \PY{n}{NailStream} \PY{o}{*}\PY{n}{in\PYZus{}current}\PY{p}{)}\PY{p}{;}
\PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}end\PYZus{}of\PYZus{}directory\PYZus{}generate}\PY{p}{(}
  \PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,} \PY{n}{NailStream} \PY{o}{*}\PY{n}{in\PYZus{}files}\PY{p}{,}
  \PY{n}{NailStream} \PY{o}{*}\PY{n}{in\PYZus{}dir}\PY{p}{,} \PY{n}{NailStream} \PY{o}{*}\PY{n}{out\PYZus{}current}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

### 2014-05-01 paper/osdi14/design.tex
Fixing typos. Explaining diff
more typos
Acked-by: Julian Bangert <jbangert@acm.org>
xx note
nit
pass through 5.1

@@
 
+\XXX[nz][Do you need \cc{*} here?  It seems like \cc{optional paren} would
+already have a C representation of \cc{paren*} according to our big figure.]
+
 % To support recursive structures, we include a reference combinator \cc{*}, which
@@
 We provide a transformation for handling offset fields, which is invoked as follows:
-\cc{\$fragment transform offset\_u32(\$file\_stream, \@offset)}. 
+\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
 This transformation corresponds to two functions for parsing and
-generation, as shown in Figure~\ref{fig:xform-sig}.
+generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
+be used to parse data at the offset contained in \cc{@offset}, by
+using \cc{apply \$fragment some_parser}.
 
@@
 A real-world example with many different transforms, used to support
-the Zip file format, is described in \S\ref{s:eval-format-zip}.
+the ZIP file format, is described in \S\ref{s:eval-formats}.
 

### 2014-05-01 paper/osdi14/eval.tex
pass through 5.1

@@
 
+An especially tricky data format is the ZIP compressed archive
+format~\cite{pkzip}.  ZIP files are normally parsed end-to-beginning. At
+the end of each ZIP file is an \emph{end-of-directory header}. This header
+contains a variable-length comment, so it has to be located by scanning
+backwards from the end of the file until a magic number and a valid
+length field is found. Many ZIP implementations disagree on how to find
+this header in confusing situations, such as when the comment contains
+the magic number~\cite{wolf:berlinsides-zip}.  This end-of-directory
+header contains the offset and size of the \emph{ZIP directory}, which
+is an array of \emph{directory entry headers}, one for every file
+in the archive.  Each entry stores file metadata, such as file name,
+compressed and uncompressed size, and a checksum, in addition to the
+offset of a \emph{local file header}. The local file header duplicates
+most information from the directory entry header and the compressed file
+contents follow it immediately.
+
 \input{fig-zip}
-\label{s:eval-format-zip}
-Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
-ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
-header}. This header contains a variable length comment, so it has to be located by scanning
-backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
-implementations disagree on how to find this header in confusing situations, such as when the
-comment contains the valid magic bytes\cite{wolf:berlinsides-zip}.
-This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
-array of one \emph{directory entry header} per file in the archive.
-Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
-in addition to the offset of a \emph{local file header}. The local file header duplicates most
-information from the directory entry header and the compressed file contents follow it immediately.
 
-Duplicating information made sense when ZIP files were
-stored on floppy disks with slow seek times and high fault rates and memory constraints made it
-impossible to keep the ZIP directory in memory or the archive was split across multiple disks.
-However, care must be taken that the meta-data is consistent. For example, vulnerabilities could occur
-if the length in the central directory is used to allocate memory and the length in the local
-directory is used to extract without checking that they are equal first.
+Duplicating information made sense when ZIP files were stored on floppy
+disks with slow seek times and high fault rates, and memory constraints
+made it impossible to keep the ZIP directory in memory or the archive
+was split across multiple disks.  However, care must be taken that the
+metadata is consistent. For example, vulnerabilities could occur if
+the length in the central directory is used to allocate memory and the
+length in the local directory is used to extract without checking that
+they are equal first, as was the case in the Python ZIP
+library~\cite{cve-python-zip-maybe}.
 
-We have implemented a grammar for ZIP. A strongly abbreviated version is shown in
-Figure~\ref{fig:zip-extract}.
+\begin{figure}[tb]
+\smaller[0.5]
+\input{code/zip-eod-xform}
+\caption{Signatures of stream transform functions for handling the
+end-to-beginning structure of ZIP files.}
+\label{fig:zip-eod-xform}
+\end{figure}
 
-Our ZIP grammar is a good example of transformations can capture complicated syntax in real-world
-file format. The \cc{zip\_file} grammar first splits the entire file stream into two streams
-based on the \cc{zip\_end\_of\_directory} transform.
-\cc{zip\_end\_of\_directory_parse(NailStream *files, NailStream *dir, NailStream (current))}
-finds the end of directory header as described above by scanning the file backwards and splits the
-file into two streams, one containing the end-of-directory header and one containing the file
-contents. The header stream is then parsed. The file stream, to which all the offsets in the
-metadata refer, is passed as an argument to the \cc{end\_of\_directory} parser.
-The directory header is then parsed and the offset and size transforms provided with Nail are used
-to isolate the actual directory from the file stream. Each directory entry in turn points to a local
-file header.
+Figure~\ref{fig:zip-extract} shows an abbreviated version of our ZIP
+file grammar.  The ZIP grammar is a good example of how transformations
+capture complicated syntax in a real-world file format; existing parser
+languages cannot handle a file format of this complexity.
 
-The \cc{file} grammar for a zip file entry is particularly interesting. It receives the file
-size, checksum and compression method of the file from the directory header. However, this same
-information is duplicated in the file entry, so the grammar uses a standard \cc{depend}
-transform to check whether the two values are equal. Unlike most other transforms, \cc{depend}
-does not consume or produce strings, it only checks that two dependent fields are equal when parsing
-and assigns the value of the second field to the first when generating. Immediately following the
-file entry is the compressed data. Because most compression algorithms operate on unbounded streams
-of data, Nail decompresses data in two steps.
-First, it isolates the compressed data from the rest of the stream by using the size transform. In
-this case, size operates on the current stream, which means it will consume data starting at the
-current position of the parser in the input.
-Second, Nail invokes a custom transform that implements the appropriate compression and decompression
-functions based on the specified compression method. These functions are otherwise oblivious about
-the layout or metadata of the file.
+The \cc{zip\_file} grammar first splits the entire file stream into two
+streams based on the \cc{zip\_end\_of\_directory} transform (whose two
+C function signatures are shown in Figure~\ref{fig:zip-eod-xform}).
+\cc{zip\_end\_of\_directory_parse} finds the end-of-directory header
+as described above, by scanning the file backwards, and splits the
+file into two streams, one containing the end-of-directory header and
+one containing the file contents.  The header stream is then parsed,
+according to the \cc{apply} rule, using the \cc{end\_of\_directory}
+parser.  The file stream, to which all the offsets in the metadata refer,
+is passed as an argument to that parser.  The directory header is then
+parsed, and the offset and size transforms provided by Nail are used
+to isolate the actual directory from the file stream.  Each directory
+entry in turn points to a local file header.
 
-Existing parser languages are not able to handle a file format of this complexity.
+The \cc{file} grammar for a ZIP file entry is particularly interesting.
+It receives as arguments the file size, checksum, and compression method
+of the file from the directory header.  However, this same information
+is duplicated in the file entry, so the grammar uses the Nail-supplied
+\cc{depend} transform to check whether the two values are equal. Unlike
+most other transforms, \cc{depend} does not consume or produce strings; it
+only checks that two dependent fields are equal when parsing, and assigns
+the value of the second field to the first when generating.  This ensures
+that the programmer does not have to worry about inconsistencies when
+handling the internal representation of a ZIP file.
 
-\noindent The results suggest that Nail is a good fit for these data formats.
+Immediately following the file entry is the compressed data.  Because
+most compression algorithms operate on unbounded streams of data, Nail
+decompresses data in two steps.  First, it isolates the compressed
+data from the rest of the stream by using the \cc{size} transform,
+which operates on the current stream, meaning it will consume data
+starting at the current position of the parser in the input.  Second,
+Nail invokes a custom \cc{zip\_compression} transform that implements
+the appropriate compression and decompression functions based on the
+specified compression method. These functions are otherwise oblivious
+to the layout or metadata of the file.
 

### 2014-05-01 paper/osdi14/intro.tex
pass through 5.1

--- 2014-05-01 paper/osdi14/intro.tex
@@
 state to re-position its input. More complicated transformations, such as
-handling compressed data, cannot be implemented at all.
+handling compressed data, cannot be represented at all.
 
@@
 prototype, we implemented succinct grammars for parts of an IP network
-stack, for DNS packets, and for Zip files, each in about a hundred lines
+stack, for DNS packets, and for ZIP files, each in about a hundred lines
 of grammar.  On top of these grammars, we were able to build a fully

### 2014-05-01 paper/osdi14/relwk.tex
pass through 5.1

@@
 transforms at all, which are crucial for supporting real-world formats
-like DNS and Zip.  The workshop paper also did not provide a detailed
+like DNS and ZIP\@.  The workshop paper also did not provide a detailed
 design discussion or evaluation.

### 2014-05-01 paper/osdi14/eval.tex
checkpoint some 5.2

@@
 \subsection{Programmer effort}
+\label{s:eval-effort}
 
@@
 %2821 lines
-\label{s:eval-effort}
-To answer the second question, we implemented small example applications based on the above
-grammars and compared code size with comparable applications that process data manually. We will
-also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
-example for the Hammer parser generator. All code size measurements for C were performed by
-SlocCount\cite{sloccount}.
 
-\noindent We implemented the following applications and compare them to applications written without
-Nail:
+To evaluate how much programmer effort is required to build an
+application that uses Nail, we implemented three applications---a
+DNS server, a replacement for the \cc{host} DNS client tool, and an
+\cc{unzip} program---based on the above grammars, and compared code
+size with comparable applications that process data manually, using
+SlocCount~\cite{sloccount}.  We also compare the code size of our
+DNS server to a DNS server written using the Hammer parsing framework,
+although it does not fully support DNS (e.g., it lacks label compression,
+among other things).  Figure~\ref{fig:effort} summarizes the results.
 
-\noindent\begin{tabular}{@{}lll@{}} 
+\XXX[nz][should we kill \cc{host}, since our replacement is nowhere
+near as full-featured as the real \cc{host} or \cc{dig}?]
+
+\begin{figure}[tb]
+\centering
+\begin{tabular}{llr@{~}l}
 \toprule
-\textbf{System} & \textbf{Nail LoC}  & \textbf{LoC w/o Nail} \\
+\textbf{System}
+  & \textbf{LoC w/ Nail}
+  & \multicolumn{2}{c}{\textbf{LoC w/o Nail}} \\
 \midrule
-DNS server & 295 & 683 (Hammer parser)\\
-&&  ~10.000 (DJBDNS)\\ 
-DNS client & 210  & 4,615 (host) \\
-\texttt{unzip} & 220 & 1600 (Info-Zip) \\
+DNS server
+  & 295
+  & 683
+  & (Hammer parser) \\
+
+  &
+  & 10,000
+  & (DJBDNS) \\
+
+DNS client
+  & 210
+  & 4,615
+  & (host) \\
+
+\cc{unzip}
+  & 220
+  & 1600
+  & (Info-Zip) \\
 \bottomrule
 \end{tabular}
-%%DNS: 
+\caption{Comparison of code size for three applications written in Nail,
+         and several existing alternative implementations.}
+\label{fig:effort}
+\end{figure}
+
 \paragraph{DNS.}
 
-Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full}, consists of a 48 line Nail
-grammar and 64 lines of C implementing DNS label compression, whose signatures are shown in
-Figure~\ref{fig:dns-xform}.
-The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
-(16 lines) supporting A, NS, MX and CNAME records.
+Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full},
+consists of a 48 line Nail grammar and 64 lines of C
+implementing DNS label compression, whose signatures are shown in
+Figure~\ref{fig:dns-xform}.  The grammar describes both the structure
+of DNS packets (36 lines) and of a simple zone-file format (16 lines)
+supporting A, NS, MX and CNAME records.
 

### 2014-05-01 paper/osdi14/eval.tex
x

@@
 To evaluate how much programmer effort is required to build an
-application that uses Nail, we implemented three applications---a
-DNS server, a replacement for the \cc{host} DNS client tool, and an
+application that uses Nail, we implemented two applications---a
+DNS server, and an
 \cc{unzip} program---based on the above grammars, and compared code
 size with comparable applications that process data manually, using
-SlocCount~\cite{sloccount}.  We also compare the code size of our
+\cc{sloccount}~\cite{sloccount}.  We also compare the code size of our
 DNS server to a DNS server written using the Hammer parsing framework,
@@
 
-\XXX[nz][should we kill \cc{host}, since our replacement is nowhere
-near as full-featured as the real \cc{host} or \cc{dig}?]
-
 \begin{figure}[tb]
 \centering
-\begin{tabular}{llr@{~}l}
+\begin{tabular}{@{~}lrr@{~}l@{~}}
 \toprule
-\textbf{System}
+\textbf{Application}
   & \textbf{LoC w/ Nail}
@@
   & (DJBDNS) \\
-
-DNS client
-  & 210
-  & 4,615
-  & (host) \\
 
@@
 
-Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full},
-consists of a 48 line Nail grammar and 64 lines of C
-implementing DNS label compression, whose signatures are shown in
-Figure~\ref{fig:dns-xform}.  The grammar describes both the structure
-of DNS packets (36 lines) and of a simple zone-file format (16 lines)
-supporting A, NS, MX and CNAME records.
-
-With this grammar, we built a simple authoritative DNS server, which parses a zone file, listens to incoming DNS
-requests, parses them and generates appropriate responses, is implemented in 183 lines of C. The
-same grammar is used, together with 98 lines of C, to implement a minimal clone of the
-\cc{host} command-line tool. Most of this code consists of a custom hashtable and system
-interface code, such as listening to sockets and reading the zonefile.
-
-The Hammer parser framework\cite{hammer-parser} ships with a toy DNS server that responds to
-any valid DNS query with a CNAME record to the domain ``spargelze.it''. 
-The server consists of 683 lines of C, mostly custom validators, semantic actions,
-and data structure definitions, with only 52 lines of code defining the
-grammar with Hammer's combinators.
+Our DNS server parses a zone file, listens to incoming DNS requests,
+parses them, and generates appropriate responses.  The DNS server is
+implemented in 183 lines of C, together with 48 lines of Nail grammar
+and 64 lines of C code implementing stream transforms for DNS label
+compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
+DNS server that responds to any valid DNS query with a CNAME record to the
+domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
+custom validators, semantic actions, and data structure definitions, with
+only 52 lines of code defining the grammar with Hammer's combinators.
+Their DNS server does not implement label compression, zone files, etc.
+From this, we conclude that Nail leads to much more compact code for
+dealing with DNS packet formats.
 
 It is hard to compare the programming effort required to implement
-our DNS server to that of a real world DNS server, since we implement less functionality.
- In particular, we do not send the additional glue records real-world DNS servers send and
-we did not implement any configuration options.
-
-However, the closest in functionality and intent is Dan Bernstein's
-djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be a minimalist, highly secure,
-authoritative-only DNS server. The latest release of djbdns, including various support tools, is
-about 10,000 lines of C as measured by \cc{sloccount}. DJ
+our DNS server to that of a real world DNS server, since we implement
+less functionality: we do not send the additional glue records and
+we did not implement any configuration options.  However, the closest
+in functionality and intent is Dan Bernstein's djbdns~\cite{djbdns},
+which aims to be a minimalist, highly secure, authoritative-only DNS
+server. The latest release of djbdns, including various support tools,
+is about 10,000 lines of C\@.  \XXX[nz][so what does this mean?]
 

### 2014-05-01 paper/osdi14/eval.tex
checkpoint through 5.2

@@
 
-  &
-  & 10,000
-  & (DJBDNS) \\
+%  &
+%  & 10,000
+%  & (DJBDNS) \\
 
@@
   & 220
-  & 1600
+  & 1,600
   & (Info-Zip) \\
@@
 
-It is hard to compare the programming effort required to implement
-our DNS server to that of a real world DNS server, since we implement
-less functionality: we do not send the additional glue records and
-we did not implement any configuration options.  However, the closest
-in functionality and intent is Dan Bernstein's djbdns~\cite{djbdns},
-which aims to be a minimalist, highly secure, authoritative-only DNS
-server. The latest release of djbdns, including various support tools,
-is about 10,000 lines of C\@.  \XXX[nz][so what does this mean?]
+% It is hard to compare the programming effort required to implement
+% our DNS server to that of a real world DNS server, since we implement
+% less functionality: we do not send the additional glue records and
+% we did not implement any configuration options.  However, the closest
+% in functionality and intent is Dan Bernstein's djbdns~\cite{djbdns},
+% which aims to be a minimalist, highly secure, authoritative-only DNS
+% server. The latest release of djbdns, including various support tools,
+% is about 10,000 lines of C\@.  \XXX[nz][so what does this mean?]
 
@@
 
-We wrote a grammar for ZIP, which consists of 92 lines of Nail and 78 lines of C implementing two
-stream transforms, one for the DEFLATE compression algorithm with the help of the zlib library and
-one for finding the end-of-directory header.
+We implemented a ZIP file extractor in 50 lines of C code, together with
+92 lines of Nail grammar and 78 lines of C code implementing two stream
+transforms (one for the DEFLATE compression algorithm with the help of
+the zlib library, and one for finding the end-of-directory header).
 
-Our grammar is also very efficient to use. Using our grammar, we build a ZIP file extractor in 50
-lines of C. Because more recent versions of ZIP have added more features, such as large file support
-and encryption, the closest existing tool in functionality is the historic version 5.4 of the Info-Zip unzip
-utility\cite{infozip} that is shipped with most Linux distributions. The entire unzip distribution
-is about 46.000 lines of code, which is mostly optimized implementation of various compression
-algorithms and other configuration and portability code.
+Because more recent versions of ZIP have added more features, such
+as large file support and encryption, the closest existing tool
+in functionality is the historic version 5.4 of the Info-Zip unzip
+utility~\cite{infozip} that is shipped with most Linux distributions.
+The entire unzip distribution is about 46,000 lines of code, which is
+mostly optimized implementations of various compression algorithms and
+other configuration and portability code.  However, unzip isolates the
+equivalent of our Nail tool in the file \cc{extract.c}, which parses the
+ZIP metadata and calls various decompression routines in other files. This
+file measures over 1,600 lines of C, which suggests that Nail is highly
+effective at reducing manual input parsing code, even for the complex
+ZIP file format.
 
-However, unzip isolates the equivalent of our Nail tool in the file extract.c, which parses the ZIP
-metadata and calls various decompression routines in other files. This file measures over 1,600
-lines of C.
+\subsection{Security}
 
-\subsection{Security.} 
 We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
@@
 the stack or heap corruption detector.
-\subsection{Performance.}
-\input{fig-perf-dns}
+
+\subsection{Performance}
+
 To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
 benchmarked our DNS server against BIND.
+
+\input{fig-perf-dns}
 

### 2014-05-01 paper/osdi14/motivation.tex
add an explicit "motivation section"
current CVE vulnerability analysis from the security eval should move there.

--- 2014-05-01 paper/osdi14/relwk.tex
@@
-\section{Related work}
-\label{s:relwk}
-
-\paragraph{Parsers.}
-
-Generating parsers and generators from an executable specification is the
-core concept of interface generators, such as CORBA~\cite{omg:corba},
-XDR~\cite{rfc:1832}, or Protocol Buffers~\cite{varda2008}.  However,
-while interface generators work well for defining new protocols, they
-do not provide full control over the data format, and therefore cannot
-be used to implement existing protocols and data formats.
-
-Closely related work has been done at Bell Labs with the PacketTypes
-system~\cite{mccann2000packet}.  However, PacketTypes works only as a
-parser, not as an output generator, and does not support the expressive
-power of parsing expression grammars (PEGs), but rather implements a
-C-like structure model enhanced with length fields and constraints.
-PacketTypes also cannot handle complicated encodings such as compressed
-data, which are supported by Nail's stream transforms.
-
-Parser generators for binary protocols were first introduced by the
-Hammer~\cite{hammer-parser} parser.
-While previous parser generators could also be used to write grammars
-for binary protocols,\footnote{Theoretically speaking, the alphabet
-over which a grammar is an abstract set, so most algorithms work
-just as well on an alphabet of $\{0,1\}$.} doing so is practically
-inconvenient. Hammer allows the programmer to specify a grammar in
-terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing are handled transparently.
-
-Hammer implements grammars as language-integrated parser combinators, an approach popularized by
-Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
-described by Burge~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
-grammars~\cite{Danielsson:2010:TPC:1863543.1863585}
-by composing them from one or multiple sub-parsers.\footnote{For more
-background on the history of expressing grammars, see Bryan Ford's
-masters thesis~\cite{ford2002packrat}, which also describes the
-default parsing algorithm used by Hammer.}
-Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
-an AST\@.
-
-Nail improves upon Hammer in three ways.  First, Nail generates output in
-addition to parsing input.  Second, Nail does not require the programmer
-to write potentially insecure semantic actions.  Last, Nail's structural
-dependencies and stream transforms allow it to work with protocols that
-Hammer cannot handle, such as protocols with offset fields, length fields,
-checksums, or compressed data, although Hammer has special facilities for
-arrays immediately preceded by their length.
-
-Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
-generating output for OCaml.  Parsifal structures grammars as an OCaml
-type that holds an internal model and functions for parsing input and
-output.  However, their generator can produce parsers and generators
-only for simple, fixed-size structures.  The programmer can then use
-these when implementing parsers and generators for more complicated
-formats, manually handling offsets, checksums, and the like, risking bugs.
-Nail handles more complicated constructs without the programmer manually
-writing code to support them.
-
-We presented an earlier design of Nail at a
-workshop~\cite{bangert:nail-langsec-anon}.  At that stage, Nail had only
-limited support for dependent fields, and did not support stream
-transforms at all, which are crucial for supporting real-world formats
-like DNS and Zip.  The workshop paper also did not provide a detailed
-design discussion or evaluation.
-
-\paragraph{Application use of parsers.}
-
-Generated parsers have long been used to parse human input, such
-as programming languages and configuration files. Frequently, such
-languages are often specified with a formal grammar in an executable
-form. Unfortunately, parser frameworks are seldom used to recognize
-machine-created input; as we demonstrate in \S\ref{s:eval},
-state-of-the-art parser generators are not suitable for parsing or
-generating many real-world data formats.
-
-A notable exception is the Mongrel web server~\cite{mongrel} which
-uses a grammar for HTTP written in the Ragel regular expression
-language~\cite{ragel-paper}.  Mongrel was re-written from scratch
-multiple times to achieve better scalability and design, yet the
-grammar was reused across all iterations~\cite{patterson-citation}.
-We hope that Nail's ideas make it possible to handle a wider range of
-protocols using parser generators, and to build more applications on
-top of grammar-based parsers.
+\section{Motivation}
+\label{s:motivation}
 

### 2014-05-01 paper/osdi14/relwk.tex
add an explicit "motivation section"
current CVE vulnerability analysis from the security eval should move there.

@@
 top of grammar-based parsers.
-
-\paragraph{Vulnerability mitigation.}
-
-Input processing vulnerabilities fall into two broad classes.  The first
-class is memory safety bugs, such as buffer overflows, which allow an
-adversary to corrupt the application's memory using specially crafted
-inputs.  These mistakes arise in lower-level languages that do not
-provide memory safety guarantees such as C, and can be partially
-mitigated by a wide range of techniques, for example static analysis,
-dynamic instrumentation, and address space layout randomization, that
-make it more difficult for an adversary to exploit these bugs.
-Nail helps developers of lower-level languages avoid these bugs
-in the first place.
-
-The second class is logic errors, where application code misinterprets
-input data. Safe languages and exploit mitigation technologies do not
-help against such vulnerabilities. This can lead to serious security
-consequences when two systems disagree on the meaning of a network
-packet or a signed message.  For example, the XNU kernel in iOS and the
-user-mode code-signing verifier interpret executable metadata differently,
-so the code signature sees different bytes at a virtual address than the
-executable that runs~\cite{evaders6}.  Android's code signing system also
-had a vulnerability due to parsing ambiguities~\cite{saurik-masterkey},
-and parsing ambiguities even contributed to attacks on the X.509 protocol
-underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  Logic errors can also
-cause memory corruption bugs as well, for example when a protocol stores
-the size of a buffer in two places and the implementation does not check
-their consistency. These mistakes are highly application-specific, and
-are difficult to mitigate using existing techniques, and these mistakes
-can occur even in high-level languages that guarantee memory safety. By
-allowing developers to specify their data format just once, Nail avoids
-logic errors and inconsistencies in parsing and output generation.
-
-% A subclass of logic errors are so-called \textit{weird machines}, where
-% implementation side effects or under-specified parser behavior leads to a
-% protocol or data format inadvertently becoming a Turing-complete execution
-% environment, even though the original grammar did not require it.
-% Frequently, this execution environment can either
-% then directly manipulate data in unwanted ways or be used to make
-% exploiting another bug feasible.\footnote{For example, by compiling a
-% return-oriented-programming exploit from code fragments discoverepd on
-% the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
-% ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
-% research community, this has been generalized into treating a program
-% as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
-% an input, analogous to a virtual machine operating on bytecode.
-% Nail avoids these problems by having the parser precisely match the
-% specified grammar, eliminating under-specified behavior.
-
-% With currently prevalent software design methods, it is hard to isolate
-% input-handling code from the rest of the program, as user input is
-% typically passed through the program and processed in bits and pieces as a
-% ``shotgun parser''~\cite{shotgun-parser}. Thus, partially invalid input might be partly 
-% processed by the program until the invalid sections are encountered.
-
-Proper input recognition has been shown to be an excellent way of
-eliminating ambiguous inputs.  In one case, a PDF parser (implemented
-in Coq) was used to reject over $95\%$ of known malicious PDFs by checking
-whether they conform to the PDF format~\cite{Bogk-PDF}. The same project
-also produced a ``counter-example'' PDF that triggered numerous bugs
-in other PDF parsers. However, manually writing parser code and proving
-unambiguity does not scale to the number of file formats and protocols
-in existence and might result in parser code tied to one specific
-application. By supporting a wide range of real-world file formats,
-Nail helps programmers produce precise parsers for complex grammars.
-
-%%This is supposed to go into design
-% While it is possible to express short transformations on the input
-% entirely as semantic actions\footnote{This is in fact the design
-% rationale, to perform computation on the fly as the parser walks the
-% parse tree.}, more complicated programs usually construct an internal
-% representation, which contains all relevant information from the input
-% in a format native to the programming language used. For example, a C
-% programmer ideally wants to deal with structs and NULL-terminated arrays,
-% whereas a C++ programmer might expect STL containers, a Java programmer
-% interfaces, a Haskell programmer records and a LISP programmer property
-% lists.  The structure of this internal representation usually resembles
-% the structure of the grammar.

### 2014-05-01 paper/osdi14/eval.tex
sec eval pass

@@
 
-We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
-analyze a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
-immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
-server.
+We use a twofold approach to evaluate the security of applications
+implemented with Nail. First, we analyze a list of CVE's related to the
+ZIP file format and argue how our ZIP tools based on Nail are immune
+against those vulnerability classes. Second, we present the results of
+fuzz-testing our DNS server.
 
-The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
-security vulnerabilities in most implementations of ZIP and other archive formats that are directly
+\paragraph{ZIP analysis.}
+
+The ZIP format has been associated with many vulnerabilities, and the
+PROTOS Genome project found numerous security vulnerabilities in most
+implementations of ZIP and other archive formats that are directly
 related to input handling.
 
-Between May 2011 and May 2014, the CVE database\cite{cve-database} lists 83 vulnerabilities when searching for the
-string ``ZIP''. Of these, 42 were completely unrelated to the archive format, for example reporting
-vulnerabilities revealing Zip codes of other users. 26 vulnerabilities were related to ZIP archives,
-but not to input processing, such as trusting arbitrary update ZIP files sent to a server or
-vulnerabilities that can be accessed through manipulating files in a ZIP container. This includes
-several vulnerabilities allowing another files to replaced while they are extracted from ZIP
-files because insufficiently secure temporary directories are used. Of the 15 remaining
-vulnerabilities pertaining to ZIP files, 14 vulnerabilities were related to input processing. The
- other vulnerability, which falls outside of Nails purview, concerns a ZIP packer choosing a weak
- legacy cryptosystem even when configured to use AES.
+Between May 2011 and May 2014, the CVE database~\cite{cve-database} lists
+83 vulnerabilities when searching for the string ``ZIP''. Of these, 42
+were completely unrelated to the archive format, for example reporting
+vulnerabilities revealing ZIP codes of other users. 26 vulnerabilities
+were related to ZIP archives, but not to input processing, such as
+trusting arbitrary update ZIP files sent to a server or vulnerabilities
+that can be accessed through manipulating files in a ZIP container. This
+includes several vulnerabilities allowing another files to replaced while
+they are extracted from ZIP files because insufficiently secure temporary
+directories are used. Of the 15 remaining vulnerabilities pertaining
+to ZIP files, 14 vulnerabilities were related to input processing. The
+other vulnerability, which falls outside of Nail's purview, concerns a
+ZIP packer choosing a weak legacy cryptosystem even when configured to
+use AES.
 
- Nail parsers are immune by design to vulnerabilities like these 14. Ten of those vulnerabilities
- are memory corruption attacks, that could also be avoided by using safer languages or better
- verification tools. Most of the buffer overflows seem to result from inconsistent offset fields
- being manipulated without checking. Because Nail's generated code checks offsets before reading and
- does not expose any untrusted pointers to the application, it protects against memory corruption
- attacks by design.
+Nail parsers are immune by design to vulnerabilities like these 14.
+Ten of those vulnerabilities are memory corruption attacks, that could
+also be avoided by using safer languages.  Most of the buffer overflows
+seem to result from inconsistent offset fields being manipulated
+without checking.  Because Nail's generated code checks offsets before
+reading and does not expose any untrusted pointers to the application,
+it protects against memory corruption attacks by design.
 
-The remaining four vulnerabilities do not involve memory corruption, but rather parser ambiguities.
-Two of them relate to implementations mis-parsing ZIP files, particularly in virus scanners. One
-other vulnerability is in a Python ZIP library, which uses both the size field in the file entry and
-in the header, without checking for their consistency. This results in a denial-of-service through
-endless loops or incorrectly extracted ZIP files.
+The remaining four vulnerabilities do not involve memory corruption,
+but rather parser ambiguities.  Two of them relate to implementations
+mis-parsing ZIP files, particularly in virus scanners. One other
+vulnerability is in a Python ZIP library, which uses both the size
+field in the file entry and in the header, without checking for their
+consistency.  This results in a denial-of-service through endless loops
+or incorrectly extracted ZIP files.
 
-A similar attack was used by the fourth vulnerability, the infamous Android master key bug
-\cite{saurik-masterkey}, that completely by-passed Android security multiple times through parser
-differentials between the ZIP handler that checks signatures for privileged applications and the ZIP
-implementations that ultimately extracts those files. Thus, valid  application bundles could be
-modified to include malicious applications without breaking their signatures. Instead of both
-components containing an ad-hoc archive implementation, a single Nail grammar would be reusable
+A similar attack was used by the fourth vulnerability, the infamous
+Android master key bug~\cite{saurik-masterkey} that completely
+bypassed Android security through parser inconsistencies between the
+ZIP handler that checks signatures for privileged applications and the
+ZIP implementations that ultimately extract those files. Thus, valid
+application bundles could be modified to include malicious applications
+without breaking their signatures. Instead of both components containing
+an ad-hoc archive implementation, a single Nail grammar would be reusable
 across both implementations.
 
-To provide additional assurance that Nail parsers are free of memory corruption attacks, we ran the
-DNS fuzzer provided with the Metasploit\cite{mspframework} framework on our server, which sent
-randomly corrupted DNS queries to our server for 4 hours, during which it did not crash or trigger
-the stack or heap corruption detector.
+\paragraph{DNS fuzzing.}
+
+To provide additional assurance that Nail parsers are free of memory
+corruption attacks, we ran the DNS fuzzer provided with the Metasploit
+framework~\cite{mspframework} on our DNS server, which sent randomly
+corrupted DNS queries to our server for 4 hours, during which it did
+not crash or trigger the stack or heap corruption detector.
 

### 2014-05-01 paper/osdi14/concl.tex
do not mention the url due to anonymous submission
done with a pass

@@
 
-This paper presented the initial design and implementation of
-\textit{Nail}, an interface generator for data formats.  Nail helps
-programmers to avoid memory corruption and ambiguity vulnerabilities
-while reducing effort in parsing and generating real-world protocols
-and file formats.  Nail achieves this by reducing the expressive
-power of the grammar, maintaining a \emph{semantic bijection} between
-data formats and internal representations, and allowing programmers to
-specify \emph{structural dependencies} in the data format.  Preliminary
-experience with implementing a DNS server using Nail suggests that this
-is a promising approach.
+This paper presented the design and implementation of \textit{Nail},
+an interface generator for data formats.  Nail helps programmers avoid
+memory corruption and ambiguity vulnerabilities while reducing effort
+in parsing and generating real-world protocols and file formats.
+Nail achieves this by reducing the expressive power of the grammar,
+maintaining a \emph{semantic bijection} between data formats and internal
+representations.  Nail captures complex data formats by introducing
+\emph{dependent fields}, \emph{streams}, and \emph{transforms}.
+Using these techniques, Nail is able to support DNS packet and ZIP file
+formats, and enables applications to handle these data formats in many
+fewer lines of code.
 
-The source code for our prototype of Nail is available at
-\url{https://github.com/jbangert/nail}.
+% The source code for our prototype of Nail is available at
+% \url{https://github.com/jbangert/nail}.

### 2014-05-01 paper/osdi14/eval.tex
done with a pass

@@
 
-To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
-benchmarked our DNS server against BIND.
+To evaluate whether Nail-based parsers are compatible with good
+performance, we compare the performance of our DNS server to that of ISC
+BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
+We simulate a load resembling that of an authoritative name server.
+First, we generate domain names consisting of one or two labels randomly
+selected from an English dictionary, and one label that is one of
+three popular top-level domains (\cc{com}, \cc{net}, and \cc{org}).
+Second, we randomly selected 90\% of these domains and created a zone
+file that mapped these domain names to \cc{127.0.0.1}.  Finally, we used
+the \cc{queryperf} tool provided with BIND to query each domain between
+zero and three times, using a DNS server running on the local machine.
+We used a single core of an Intel i7-3610QM system with 12GB of RAM.
+The benchmark tool kept at most 20 queries outstanding at once, and
+was configured to repeat the same randomized sequence of queries for
+one minute.  We repeated each test seven times with 50,000 domain names,
+restarting each daemon in between; we also repeated the tests with 1
+million domain names, and found similar results.  We also performed one
+initial dry run to warm the file system cache for the zone file.
+
+The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
+our Nail-based DNS server can achieve higher performance and lower latency
+than BIND\@.  Although BIND is a much more sophisticated DNS server,
+and implements many features that are not present in our Nail-based DNS
+server, we believe our results demonstrate that Nail's parsers are not
+a barrier to achieving acceptable performance.
 
 \input{fig-perf-dns}
-
-We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind9}, one of
-the oldest and most popular DNS daemons, on a load that resembles the authoritative name server at a
-hosting company. First, we generated domain names consisting of one or
-two labels randomly selected from an English dictionary and one label that is one of three popular
-Top Level Domains: ``com'', ``net'' or ``org''. Second, we randomly selected 90\% of these domains and
-created a zone file that mapped these domain names to the local IP address.
-
-Finally, we used the \cc{queryperf} tool provided with BIND to query each domain between zero
-and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
-core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
-queries outstanding at once. The queryperf tool was configured to
-repeat the same, randomized sequence of queries for one minute and the average throughput and
-response latency were measured.  We repeated each test seven times  with 50.000
-domain names, restarting each daemon in between. We also performed one initial dry run to warm the
-file system cache for the zone file. We repeated the tests with 1 million domain names and found
-that the servers performed almost identically.
-
-Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
-of optimization by a factor of 3.  This demonstrates that even though Nail was
-not developed focusing on performance, applications built on Nail can compete with real-world
-hand-written parsers.

### 2014-05-01 paper/osdi14/future.tex
Initial syntax table
done with a pass

@@
 \label{s:future}
-
-The Nail parser generator is currently a work-in-progress, and we would
-appreciate feedback on our initial design and prototype implementation.
-Short-term next steps include improving the scoping of dependency fields
-and adding support for offset fields.
 
@@
 
-Finally, we would like to demonstrate the capabilities of Nail by
-implementing various binary formats ``notorious'' for their insecurity
-in Nail. Nail was designed with the idioms of formats such as PDF and
-PNG in mind. Ultimately, we want to provide examples of successful Nail
-parsers throughout a network stack, from a user-space TCP stack to a
-PNG de-compressor.
+% Finally, we would like to demonstrate the capabilities of Nail by
+% implementing various binary formats ``notorious'' for their insecurity
+% in Nail. Nail was designed with the idioms of formats such as PDF and
+% PNG in mind. Ultimately, we want to provide examples of successful Nail
+% parsers throughout a network stack, from a user-space TCP stack to a
+% PNG de-compressor.

### 2014-05-01 paper/osdi14/eval.tex
move text to the motivation section

@@
 
-The ZIP format has been associated with many vulnerabilities, and the
-PROTOS Genome project found numerous security vulnerabilities in most
-implementations of ZIP and other archive formats that are directly
-related to input handling.
-
-Between May 2011 and May 2014, the CVE database~\cite{cve-database} lists
-83 vulnerabilities when searching for the string ``ZIP''. Of these, 42
-were completely unrelated to the archive format, for example reporting
-vulnerabilities revealing ZIP codes of other users. 26 vulnerabilities
-were related to ZIP archives, but not to input processing, such as
-trusting arbitrary update ZIP files sent to a server or vulnerabilities
-that can be accessed through manipulating files in a ZIP container. This
-includes several vulnerabilities allowing another files to replaced while
-they are extracted from ZIP files because insufficiently secure temporary
-directories are used. Of the 15 remaining vulnerabilities pertaining
-to ZIP files, 14 vulnerabilities were related to input processing. The
-other vulnerability, which falls outside of Nail's purview, concerns a
-ZIP packer choosing a weak legacy cryptosystem even when configured to
-use AES.
+XXX refer back to \S\ref{s:motivation}.
 
 Nail parsers are immune by design to vulnerabilities like these 14.
-Ten of those vulnerabilities are memory corruption attacks, that could
-also be avoided by using safer languages.  Most of the buffer overflows
-seem to result from inconsistent offset fields being manipulated
-without checking.  Because Nail's generated code checks offsets before
-reading and does not expose any untrusted pointers to the application,
-it protects against memory corruption attacks by design.
+Because Nail's generated code checks offsets before reading and does not
+expose any untrusted pointers to the application, it protects against
+memory corruption attacks by design.
 
-The remaining four vulnerabilities do not involve memory corruption,
-but rather parser ambiguities.  Two of them relate to implementations
-mis-parsing ZIP files, particularly in virus scanners. One other
-vulnerability is in a Python ZIP library, which uses both the size
-field in the file entry and in the header, without checking for their
-consistency.  This results in a denial-of-service through endless loops
-or incorrectly extracted ZIP files.
+Nail also protects against parser ambiguity vulnerabilities.
 
-A similar attack was used by the fourth vulnerability, the infamous
-Android master key bug~\cite{saurik-masterkey} that completely
-bypassed Android security through parser inconsistencies between the
-ZIP handler that checks signatures for privileged applications and the
-ZIP implementations that ultimately extract those files. Thus, valid
-application bundles could be modified to include malicious applications
-without breaking their signatures. Instead of both components containing
+Instead of both components containing
 an ad-hoc archive implementation, a single Nail grammar would be reusable

### 2014-05-01 paper/osdi14/motivation.tex
move text to the motivation section

@@
 \label{s:motivation}
+
+To motivate the need for Nail, this section presents a case study of
+vulnerabilities due to ad-hoc input parsing and output generation.
+
+The ZIP format has been associated with many vulnerabilities, and the
+PROTOS Genome project found numerous security vulnerabilities in most
+implementations of ZIP and other archive formats that are directly
+related to input handling.
+\XXX[what is the PROTOS Genome project?]
+
+Between May 2011 and May 2014, the CVE database~\cite{cve-database} lists
+83 vulnerabilities when searching for the string ``ZIP''. Of these, 42
+were completely unrelated to the archive format, for example reporting
+vulnerabilities revealing ZIP codes of other users. 26 vulnerabilities
+were related to ZIP archives, but not to input processing, such as
+trusting arbitrary update ZIP files sent to a server or vulnerabilities
+that can be accessed through manipulating files in a ZIP container. This
+includes several vulnerabilities allowing another files to replaced while
+they are extracted from ZIP files because insufficiently secure temporary
+directories are used. Of the 15 remaining vulnerabilities pertaining
+to ZIP files, 14 vulnerabilities were related to input processing. The
+other vulnerability, which falls outside of Nail's purview, concerns a
+ZIP packer choosing a weak legacy cryptosystem even when configured to
+use AES.
+
+Out of the 14 input processing vulnerabilities, ten are memory corruption
+attacks, that could also be avoided by using safer languages.  Most of
+the buffer overflows seem to result from inconsistent offset fields
+being manipulated without checking.
+
+The remaining four vulnerabilities do not involve memory corruption,
+but rather parser ambiguities.  Two of them relate to implementations
+mis-parsing ZIP files, particularly in virus scanners.  One other
+vulnerability is in a Python ZIP library, which uses both the size
+field in the file entry and in the header, without checking for their
+consistency.  This results in a denial-of-service through endless loops
+or incorrectly extracted ZIP files.
+
+A similar attack was used by the fourth vulnerability, the infamous
+Android master key bug~\cite{saurik-masterkey} that completely
+bypassed Android security through parser inconsistencies between the
+ZIP handler that checks signatures for privileged applications and the
+ZIP implementations that ultimately extract those files. Thus, valid
+application bundles could be modified to include malicious applications
+without breaking their signatures.
 
@@
 
-Proper input recognition has been shown to be an excellent way of
-eliminating ambiguous inputs.  In one case, a PDF parser (implemented
-in Coq) was used to reject over $95\%$ of known malicious PDFs by checking
-whether they conform to the PDF format~\cite{Bogk-PDF}. The same project
-also produced a ``counter-example'' PDF that triggered numerous bugs
-in other PDF parsers. However, manually writing parser code and proving
-unambiguity does not scale to the number of file formats and protocols
-in existence and might result in parser code tied to one specific
-application. By supporting a wide range of real-world file formats,
-Nail helps programmers produce precise parsers for complex grammars.
+% Proper input recognition has been shown to be an excellent way of
+% eliminating ambiguous inputs.  In one case, a PDF parser (implemented
+% in Coq) was used to reject over $95\%$ of known malicious PDFs by checking
+% whether they conform to the PDF format~\cite{Bogk-PDF}. The same project
+% also produced a ``counter-example'' PDF that triggered numerous bugs
+% in other PDF parsers. However, manually writing parser code and proving
+% unambiguity does not scale to the number of file formats and protocols
+% in existence and might result in parser code tied to one specific
+% application. By supporting a wide range of real-world file formats,
+% Nail helps programmers produce precise parsers for complex grammars.
 

### 2014-05-01 paper/osdi14/abstract.tex
checkpoint comments from frans through section 2

@@
 programmer effort for safely parsing and generating protocols defined
-by a Parser-Expression based grammar.  Nail introduces several key
+by a Parser-Expression based grammar.
+\XXX[what is an interface generator?]
+\XXX[what is a Parser-Expression based grammar?]
+Nail introduces several key
 ideas to achieve its goal.  First, Nail uses the protocol grammar to
@@
 program in C in 220 lines of code and grammar, demonstrating that Nail
-makes it easy to parse complex real-world data formats.
+makes it easy to parse complex real-world data formats.  Performance
+experiments show that a Nail-based DNS server can outperform the widely
+used BIND DNS server, demonstrating that Nail parsers can achieve good
+performance.
 

### 2014-05-01 paper/osdi14/intro.tex
add an explicit "motivation section"
current CVE vulnerability analysis from the security eval should move there.
checkpoint comments from frans through section 2

@@
 as the parser generator is bug-free, the application will be safe from
-input processing vulnerabilities.  Unfortunately, applying this approach
-in practice, using state-of-the-art parser generators, still requires
-too much manual programmer effort, making it error-prone, as we describe
-next.
+input processing vulnerabilities.
+
+Unfortunately, applying this approach in practice, using state-of-the-art
+parser generators, still requires too much manual programmer effort,
+and is still error-prone, for four reasons:
 
@@
 much like other input processing code, and mistakes can result in memory
-corruption bugs or misinterpreted inputs.  Writing these semantic actions
+corruption bugs or misinterpreted inputs.
+\XXX[do we have any evidence to cite for this claim?]
+Writing these semantic actions
 also requires the programmer to describe the structure of the input three
@@
 than producing an output, thus requiring the programmer to manually
-construct outputs, which is error-prone.  Some parser generators, such as
+construct outputs, which is error-prone.
+\XXX[do we have any evidence to cite for this claim?]
+Some parser generators, such as
 Boost.Spirit~\cite{boost-spirit},
@@
 
-Third, many data formats contain redundancies, such as repeating information in multiple structures.
-Applications usually do not explicitly check for consistency, and if different applications use
-different instances of the same value, an attacker can craft input that causes inconsistencies~\cite{cve-python-zip-maybe}.
-Furthermore, security vulnerabilities can occur when an application assumes two repetitions of the
-same data to be consistent, such as allocating a buffer based on the value of one size field and copying into
-that buffer based on the value of another~\cite{cve-some-example}.
+Third, many data formats contain redundancies, such as repeating
+information in multiple structures.  Applications usually do not
+explicitly check for consistency, and if different applications use
+different instances of the same value, an attacker can craft input that
+causes inconsistencies~\cite{cve-python-zip-maybe}.  Furthermore, security
+vulnerabilities can occur when an application assumes two repetitions
+of the same data to be consistent, such as allocating a buffer based on
+the value of one size field and copying into that buffer based on the
+value of another~\cite{cve-some-example}.
 
-Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
-existing parser generators. Those parsers cannot directly deal with length or checksum fields
-directly, so the programmer has to either write potentially unsafe code to deal with such features,
-or build contrived grammar constructs, such as introducing one grammar rule for each possible value
-of a length field. Offset fields usually require the programmer to manipulate a parser's internal
-state to re-position its input. More complicated transformations, such as
-handling compressed data, cannot be represented at all.
+Finally, real-world data formats, for example PNG or PDF, are hard to
+represent with existing parser generators.  Those parsers cannot directly
+deal with length or checksum fields directly, so the programmer has to
+either write potentially unsafe code to deal with such features, or build
+contrived grammar constructs, such as introducing one grammar rule for
+each possible value of a length field.  Offset fields, which specify
+the position at which some data structure is located, usually require
+the programmer to manipulate a parser's internal state to re-position
+its input.  More complicated transformations, such as handling compressed
+data, cannot be represented at all.
 
@@
 
-Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{transformations}, to
-elegantly handle problematic structures, such as offset fields or checksums. Dependent fields are
-fields in the protocol whose value depends in some way on the value or layout of other parts of the
-format. Transformations allow the programmer to escape the generated code to modify the raw data and
-interact with dependent fields in a controlled manner.
+Third, Nail introduces two abstractions, \emph{dependent fields} and
+\emph{transformations}, to elegantly handle problematic structures,
+such as offset fields or checksums.  Dependent fields capture fields
+in a protocol whose value depends in some way on the value or layout
+of other parts of the format; for example, offset or length fields,
+which specify the position or length of another data structure, fall
+in this category.  Transformations allow the programmer to escape the
+generated code to modify the raw data and interact with dependent fields
+in a controlled manner.
 
@@
 makes it easy for application developers to parse and generate external
-data representations.
+data representations.  Performance results show that the Nail-based DNS
+server outperforms the widely used BIND DNS server, demonstrating that
+Nail-based parsers and generators can achieve good performance.
 
 The rest of this paper is organized as follows.  \S\ref{s:relwk}
-puts Nail in the context of related work.  \S\ref{s:design} describes
+puts Nail in the context of related work.
+\S\ref{s:motivation} motivates the need for Nail by examining
+past data format vulnerabilities.
+\S\ref{s:design} describes
 Nail's design.  \S\ref{s:impl} discusses our initial implementation

### 2014-05-01 paper/osdi14/relwk.tex
checkpoint comments from frans through section 2

@@
 core concept of interface generators, such as CORBA~\cite{omg:corba},
-XDR~\cite{rfc:1832}, or Protocol Buffers~\cite{varda2008}.  However,
-while interface generators work well for defining new protocols, they
-do not provide full control over the data format, and therefore cannot
-be used to implement existing protocols and data formats.
+XDR~\cite{rfc:1832}, and Protocol Buffers~\cite{varda2008}.  However,
+interface generators do not allow the programmer to specify the byte-level
+data format; instead, they define their own data encoding that is
+specific to a particular interface generator.  For instance, XDR-based
+protocols are incompatible with Protocol Buffers.  Moreover, this means
+that interface generators cannot be used to interact with existing protocols
+that were not defined using that interface generator in the first place.
+As a result, interface generators cannot be used to parse or generate
+widely used formats such as DNS or ZIP, which is a goal for Nail.
 
@@
 type that holds an internal model and functions for parsing input and
-output.  However, their generator can produce parsers and generators
+output.  However, Parsifal can produce parsers and generators
 only for simple, fixed-size structures.  The programmer can then use

### 2014-05-01 paper/osdi14/eval.tex
depend vs u16_depend

@@
 is duplicated in the file entry, so the grammar uses the Nail-supplied
-\cc{depend} transform to check whether the two values are equal. Unlike
-most other transforms, \cc{depend} does not consume or produce strings; it
+\cc{u16\_depend} transform to check whether the two values are equal. Unlike
+most other transforms, \cc{u16\_depend} does not consume or produce strings; it
 only checks that two dependent fields are equal when parsing, and assigns

### 2014-05-01 paper/osdi14/eval.tex
Eval-security

@@
 
-XXX refer back to \S\ref{s:motivation}.
+In \S\ref{s:motivation}, we found 14 input handling vulnerabilities related to ZIP files.
 
-Nail parsers are immune by design to vulnerabilities like these 14.
+Ten of these vulnerabilities involved memory corruption during input handling.
 Because Nail's generated code checks offsets before reading and does not
-expose any untrusted pointers to the application, it protects against
+expose any untrusted pointers to the application, it is immune against
 memory corruption attacks by design.
 
-Nail also protects against parser ambiguity vulnerabilities.
-
-Instead of both components containing
-an ad-hoc archive implementation, a single Nail grammar would be reusable
-across both implementations.
+Nail also protects against parsing ambiguity vulnerabilities like the four others we studied. First,
+Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
+that caused a vulnerability in the Python ZIP library. Second, the other three vulnerabilities exist
+because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
+parsers are not very reuseable, as they build application specific data structures and are tightly
+coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
+avoiding protocol misunderstandings.
 

### 2014-05-01 paper/osdi14/motivation.tex
Eval-security

@@
 The ZIP format has been associated with many vulnerabilities, and the
-PROTOS Genome project found numerous security vulnerabilities in most
+PROTOS Genome project~\cite{protoc-c10-archive} found numerous security vulnerabilities in most
 implementations of ZIP and other archive formats that are directly
 related to input handling.
-\XXX[what is the PROTOS Genome project?]
 
@@
 
-\paragraph{Vulnerability mitigation.}
+\paragraph{Vulnerability classes.}
 
@@
 so the code signature sees different bytes at a virtual address than the
-executable that runs~\cite{evaders6}.  Android's code signing system also
+executable that runs~\cite{evaders6}. 
+Android's code signing system also
 had a vulnerability due to parsing ambiguities~\cite{saurik-masterkey},

### 2014-05-01 paper/osdi14/abstract.tex
Abstract

--- 2014-05-01 paper/osdi14/abstract.tex
@@
 
-\textit{Nail} is an interface generator that greatly reduces the
-programmer effort for safely parsing and generating protocols defined
-by a Parser-Expression based grammar.  Nail introduces several key
+\textit{Nail} is an tools that greatly reduces the
+programmer effort for safely parsing and generating data formats defined
+by a grammar.
+Nail introduces several key
 ideas to achieve its goal.  First, Nail uses the protocol grammar to
@@
 and compressed data, which are impractical to express in existing
-protocol languages.  Using Nail, we implement a complete DNS server in
-C in under 300 lines of code and grammar, and a complete {\tt unzip}
+protocol languages.  Using Nail, we implement an authoritative  DNS server in
+C in under 300 lines of code and grammar, and an {\tt unzip}
 program in C in 220 lines of code and grammar, demonstrating that Nail
-makes it easy to parse complex real-world data formats.
+makes it easy to parse complex real-world data formats.  Performance
+experiments show that a Nail-based DNS server can outperform the widely
+used BIND DNS server, demonstrating that Nail parsers can achieve good
+performance.
 

### 2014-05-01 paper/osdi14/design.tex
More examples

@@
 Nail's fundamental parsers represent signed or unsigned
-integers with arbitrary lengths up to 64 bits, an
-approach popularized by the Hammer framework~\cite{hammer-parser}.
+integers with arbitrary lengths up to 64 bits.
 Note that is possible to define parsers for sub-byte lengths, e.g.,
@@
  The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field names preceding their definition.
-In our opinion, this syntax is easier to read than the widespread C-style structure and type
-declarations.
 
@@
 more concisely represented as a reference that is null when the parser fails.
-
+For example, the grammar for  Ethernet headers uses \cc{optional vlan\_header} to parse the VLAN
+header that only appears in Ethernet packets transmitted to a non-default VLAN.
 \paragraph{Rules.}
@@
 introduces a layer of indirection, such as a reference or pointer, to the model data type.
-For example example,   balanced parentheses could be parsed as follows: \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
+The reference combinator does not need to used when another combinator, such as \cc{optional} or
+\cc{many} already introduces indirection in the data type.
+For example example,   arithmetic expressions consisting of sums and products can be parsed as follows:
+
+\begin{verbatim}
+expr = choose {
+  PAREN = <uint8='('; *expr; uint8=')'>
+  PRODUCT = sepBy1 uint8='*' expr
+  SUM = sepBy1 uint8='+' expr
+  INTEGER = many1 uint8 | '0' .. '9'
+}
+\end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
 
@@
 
-The programmer has to guarantee memory safety and unambiguity of the stream transformations. To
-reduce both programmer effort and the risk of unsafe operations, we provide implementations of
+ To reduce both programmer effort and the risk of unsafe operations, Nails provides implementations of
 transformations for many common features, such as checksums, size, and offset fields. Furthermore,
-we provide library functions that can be used to safely operate on streams, such as splitting and
+Nail provides library functions that can be used to safely operate on streams, such as splitting and
 concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
@@
 
+Transformations need to be carefully written, because they can escape Nail's security and introduce
+bugs. However, as we will show in \S\ref{s:eval-effort}, Nail transformations are much shorter than
+hand-written parsers and many formats can be represented with just the provided transforms
+
 Transformations can handle a wide variety of patterns in data formats, including the following: 
 \paragraph{Offsets.}
-We provide a transformation for handling offset fields, which is invoked as follows:
+A built-in transformation for handling offset fields, which is invoked as follows:
 \cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
@@
 transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
-versa. Usually, this involves calling an existing library.
+versa. This transformation must be carefully written to not have bugs.
 

### 2014-05-01 paper/osdi14/intro.tex
More examples

@@
 update the application's internal representation of the data based on
-each AST node.  Writing these semantic actions by hand is error-prone,
-much like other input processing code, and mistakes can result in memory
-corruption bugs or misinterpreted inputs.
-\XXX[do we have any evidence to cite for this claim?]
-Writing these semantic actions
+each AST node.  Not only does the programmer allocate memory and manipulate pointers,
+ operations that are error-prone, but
+writing these semantic actions
 also requires the programmer to describe the structure of the input three
@@
 
-To evaluate whether Nail's design is effective at handling real-world
-data formats, we implemented a prototype of Nail for C\@.  Using our
-prototype, we implemented succinct grammars for parts of an IP network
-stack, for DNS packets, and for ZIP files, each in about a hundred lines
-of grammar.  On top of these grammars, we were able to build a fully
-functional DNS server in under 200 lines of C code, and an {\tt unzip}
-utility in about 50 lines of C code.  This suggests both that Nail is
-effective at handling complex real-world data formats, and that Nail
-makes it easy for application developers to parse and generate external
-data representations.  Performance results show that the Nail-based DNS
-server outperforms the widely used BIND DNS server, demonstrating that
-Nail-based parsers and generators can achieve good performance.
+To evaluate whether Nail's design is effective at handling real-world data
+formats, we implemented a prototype of Nail for C\@. Using our prototype, we
+implemented succinct grammars for parts of an IP network stack, for DNS packets,
+and for ZIP files, each in about a hundred lines of grammar. On top of these
+grammars, we were able to build a DNS server in under 200 lines of C code, and
+an {\tt unzip} utility in about 50 lines of C code, with performance comparable
+to or exceeding existing implementations.This suggests both that Nail is
+effective at handling complex real-world data formats, and that Nail makes it
+easy for application developers to parse and generate external data
+representations. Performance results show that the Nail-based DNS server
+outperforms the widely used BIND DNS server, demonstrating that Nail-based
+parsers and generators can achieve good performance.
 

### 2014-05-01 paper/osdi14/relwk.tex
More examples

@@
 Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
-described by Burge~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
+described by Burge~\cite{burge1975recursive}) is a natural way of concisely e               xpressing top-down
 grammars~\cite{Danielsson:2010:TPC:1863543.1863585}

### 2014-05-01 paper/osdi14/design.tex
Design

@@
 
-Nail introduces two concepts to handle this, \emph{streams} and \emph{transformations}. 
+Nail introduces two concepts to handle these challenges, \emph{streams} and \emph{transformations}. 
 Streams represent a sequence of bytes that contain some external format. The parsers and generators
@@
 
-Streams can be defined in two ways.
-First, a rule can require streams as arguments, which can be passed in by the program that invokes
-the generated parser. Every rule needs to be passed at least the implicit current stream.
+Streams are passed as arguments to a rule or defined within the grammar through \emph{transformations}.
+The current stream is always passed as an implicit parameter.
 
-Second, the programmer can create temporary streams in the grammar with \emph{transformations},
-which are two arbitrary functions called during parsing and output generation respectively.
+Transformations are two arbitrary functions called during parsing and output generation.
 The parsing function  takes any number of stream arguments and dependent field values,

### 2014-05-01 paper/osdi14/design.tex
DNS with line numbers

@@
  describe a data type.
-Figure~\ref{fig:syntax} shows the elements of Nail grammars, which we describe in more detail below.
 
+We present both a systematic overview of Nail's syntax with short examples in
+Figure~\ref{fig:syntax} shows the elements of Nail grammars and explain our design in more detail
+below, using a grammar for the well-known DNS protocol as a guiding example.
+
+\input{fig-dnsfull}
+\paragraph{Rules.}
+A Nail grammar consists of rules that assign a parser to a name. Rules are written as assignments,
+such as \cc{ints = /*parser definition*/}, which defines a rule called \cc{ints}. As we will
+describe later in \S\ref{s:dependent} and \S\ref{s:transforms}, rules can optionally consume parameters.
+Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
+the body of the rule had been substituted in the code. If parameters appear, they are passed by
+reference.
 \paragraph{Integers and constraints.}
@@
 integers with arbitrary lengths up to 64 bits.
-Note that is possible to define parsers for sub-byte lengths, e.g.,
-to parse the 4-bit
-data offset within the TCP header.  The internal representation rounds up to the nearest integer
-type; for example, a 24-bit unsigned integer will be stored in an unsigned 32-bit integer.
+Note that is possible to define parsers for sub-byte lengths, for example, the flag bits in the DNS
+message header, in lines 4 through 8.
 
@@
 The \emph{many} combinator takes a parser and applies it repeatedly
-until it fails, returning an array of the inner parser's results. The
-\emph{sepBy} combinator
+until it fails, returning an array of the inner parser's results. In line 39 of the DNS grammar, a
+sequence of labels is parsed by parsing as many labels as possible, that is, until an invalid length
+field is encountered.
+The \emph{sepBy} combinator
 additionally takes a constant parser, which it applies in between parsing
@@
 
-% \begin{figure}[tb]
-% \smaller[0.5]
-% \begin{minipage}{0.45\columnwidth}
-% \begin{verbatim}
-% dns_header = {
-%   id uint16
-%   qr uint1
-%   opcode uint4
-%   aa uint1 
-%   tc uint1
-%   rd uint1
-%   ra uint1
-%   uint3 = 0
-%   rcode uint4
-% }
-% \end{verbatim} 
-% \end{minipage}
-% ~
-% \begin{minipage}{0.45\columnwidth}
-% \begin{verbatim}
-% struct dns_header {
-%   uint16_t id;
-%   uint8_t qr;
-%   uint8_t opcode;
-%   uint8_t aa;
-%   uint8_t tc;
-%   uint8_t rd;
-%   uint8_t ra;
-%   uint8_t rcode;
-% };
-% \end{verbatim} 
-% \end{minipage}
-
-% \caption{Nail grammar (left) and data model (right) for a part of the
-% grammar for DNS packets.  The \cc{uint3 = 0} grammar represents
-% 3 bits of padding (filled with zeroes).}
-% \label{fig:dns-struct}
-% \end{figure}
-
 Nail provides a structure combinator with semantic labels instead of the sequence combinator that
@@
 The structure combinator consists of a sequence of fields, typically consisting of a label and a
-parser that describes the contents of that field. Other field types will be described below.
- The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field names preceding their definition.
+parser that describes the contents of that field, surrounded by curly braces. Other field types will be described below.
+ The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field
+ names preceding their definition.
 
@@
 parsing and generated during output.  Constants can either have integer values, such as
-\cc{uint3 = 0} or string values for text-based protocols, e.g. \cc{many uint8 = ``Foo''}.
+in line 9 of the DNS grammar, or string values for text-based protocols, e.g. \cc{many uint8 = ``Foo''}.
 
@@
 we introduced the wrap combinator, which takes a sequence of parsers containing
-exactly one non-constant parser. The parser and generator act as though the wrap
-combinator was a sequence of parsers, but the data model does not wrap the
-single value in another structure, making the application-visible representation
+exactly one non-constant parser. The external format is defined as though the wrap combinator were a
+structure, but the data model does not introduce a structure with just one element, making the application-visible representation
 (and thus application code) more concise.
-For example, \cc{<uint8='"'; many int8 | 'a'..'z'; uint8='"'>} parses a quoted
-lowercase word into an array.
+Line 39 of the DNS grammar uses the wrap combinator to hide the terminating NUL-byte of a sequence
+of labels.
 
@@
 header that only appears in Ethernet packets transmitted to a non-default VLAN.
-\paragraph{Rules.}
-A Nail grammar consists of rules that assign a parser to a name. Rules are written as
-assignments, such as \cc{ints = many uint8}, which defines a rule called \cc{ints} with the
-body \cc{many uint8}.  As we will describe in \S\ref{s:dependent}
-and \S\ref{s:transforms}, rules can optionally consume parameters. 
-Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
-the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.
 

### 2014-05-01 paper/osdi14/eval.tex
Eval
DNS with line numbers

@@
 
-Figure~\ref{fig:dns-xform} presents the Nail grammar for DNS packets.
+In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
+for DNS packets, shown in Figure~\ref{fig:dns-xform}.
 The grammar corresponds almost directly to the diagrams in RFC 1035,
@@
 suffix was previously encoded.
-
-\input{fig-dnsfull}
 
@@
 this header in confusing situations, such as when the comment contains
-the magic number~\cite{wolf:berlinsides-zip}.  This end-of-directory
+the magic number~\cite{wolf:berlinsides-zip}.
+
+ This end-of-directory
 header contains the offset and size of the \emph{ZIP directory}, which
@@
 offset of a \emph{local file header}. The local file header duplicates
-most information from the directory entry header and the compressed file
-contents follow it immediately.
+most information from the directory entry header. The compressed file
+contents follow the header immediately.
 
@@
 
-The \cc{zip\_file} grammar first splits the entire file stream into two
-streams based on the \cc{zip\_end\_of\_directory} transform (whose two
-C function signatures are shown in Figure~\ref{fig:zip-eod-xform}).
-\cc{zip\_end\_of\_directory_parse} finds the end-of-directory header
-as described above, by scanning the file backwards, and splits the
-file into two streams, one containing the end-of-directory header and
-one containing the file contents.  The header stream is then parsed,
-according to the \cc{apply} rule, using the \cc{end\_of\_directory}
-parser.  The file stream, to which all the offsets in the metadata refer,
-is passed as an argument to that parser.  The directory header is then
-parsed, and the offset and size transforms provided by Nail are used
-to isolate the actual directory from the file stream.  Each directory
-entry in turn points to a local file header.
+The \cc{zip\_file} grammarfirst splits the entire file stream into two
+streams based on the \cc{zip\_end\_of\_directory} transform in line 2
+(whose two C function signatures are shown in
+Figure~\ref{fig:zip-eod-xform}). \cc{zip\_end\_of\_directory_parse}
+finds the end-of-directory header as described above, by scanning the
+file backwards, and splits the file into two streams, one containing
+the end-of-directory header and one containing the file contents. The
+\cc{end\_of\_directory} rule is then applied to the header stream in
+line 4. All offsets in the ZIP file refer to the beginning of the
+file, so the stream \cc{\$file} which contains the file contents
+without the header is passed as an argument to all parsers from
+hereon.
 
-The \cc{file} grammar for a ZIP file entry is particularly interesting.
-It receives as arguments the file size, checksum, and compression method
-of the file from the directory header.  However, this same information
-is duplicated in the file entry, so the grammar uses the Nail-supplied
-\cc{u16\_depend} transform to check whether the two values are equal. Unlike
-most other transforms, \cc{u16\_depend} does not consume or produce strings; it
-only checks that two dependent fields are equal when parsing, and assigns
-the value of the second field to the first when generating.  This ensures
-that the programmer does not have to worry about inconsistencies when
-handling the internal representation of a ZIP file.
+The directory header contains the offset and size of the ZIP
+directory (lines 9 and 10). The \cc{offset} and \cc{size}
+transformations extract a stream containing just the directory from
+the file contents. This stream is then parsed as an array of directory
+entries in line 17.
+Each directory entry in turn points to a local file header, which is
+similarly extracted and parsed with the \cc{file} rule.
+
+The \cc{file} rule starting at line 41, describing a ZIP file entry takes dependent field
+parameters containing file metadata information from the directory
+header. However, this same information is duplicated in the file
+entry, so the grammar uses the Nail-supplied \cc{u16\_depend}
+transform to check whether the two values are equal. Unlike most other
+transforms, \cc{u16\_depend} does not consume or produce strings; it
+only checks that two dependent fields are equal when parsing, and
+assigns the value of the second field to the first when generating.
+This ensures that the programmer does not have to worry about
+inconsistencies when handling the internal representation of a ZIP
+file.
 

### 2014-05-01 paper/osdi14/impl.tex
Implementation

@@
 The current prototype of the Nail parser generator supports the C programming
-language and top-down parsers. The implementation parses Nail grammars with Nail itself, using a
+language. The implementation parses Nail grammars with Nail itself, using a
 130-line Nail grammar feeding into a 2,000-line C++ program that
-implements the parser. Bootstrapping is
+emits the parser and generator code. Bootstrapping is
 performed with a subset of the grammar implemented using conventional grammars. 
-Options for C++ STL data models and emitting
-Packrat parsers~\cite{packrat-parsing:icfp02} are under development. In
+And option for C++ STL data models  is in development. In
 this section, we will discuss some particular features of our parser
@@
 validate and recognize the input, and the second to bind this data to the internal
-model. Currently the parser is a straightforward top-down parser, although
-preparations have been made to add Packrat parsing to achieve linear parsing
-times.
+model. Currently the parser uses a straightforward top-down algorithm, which can perform poorly on
+grammars that backtrack heavily. However, preparations have been made to add Packrat
+parsing~\cite{packrat-parsing:icfp02} that achieve linear time even in the worst case.
 
-\paragraph{Memory allocation.}
+\paragraph{Defense-in-Depth.}
 
-Many security vulnerabilities can occur when heap allocations are done improperly. Therefore, just
-like Hammer, Nail avoids using the heap as much as possible, using a custom arena allocator and
-allocating only fixed-size blocks from the system allocator (malloc). However, Nail extends upon
-Hammer's approach and uses two arenas for each parsed input. One arena is used for intermediate
-structures, including dependency fields and any temporary streams, and is released (and zeroed)
-after parsing completes, whereas the other arena is used only for allocating the result, and has to
-be freed by the programmer.
+Security exploits often rely on raw inputs being present in memory, for example to include
+shell-code or crafted stack frames for ROP-attacks in padding fields\cite{shotgun-parser} or the
+application executing a controlled sequence of heap allocations and de-allocations to place
+specific data at predictable addresses\cite{jp-advanced, vudo-malloc}. Because the rest of the
+application or even Nail's generated code may contain memory corruption bug, Nail
+carefully handles memory allocations as defense-in-depth to make exploiting such vulnerabilities harder.
 
-\paragraph{Intermediate representation.}
+When parsing input, Nail uses two separate memory arenas. These arenas allocate memory from the
+system allocator in large, fixed size blocks. Allocations are handled linearly and all data in the
+arena is zeroed and freed at the same time. Nail uses one arena for data used only during parsing,
+including dependent fields and temporary streams and is released before the Nail's parser returns.
+The other arena is used to allocate the internal data type returned and is freed by the application
+once it is done processing an input.
 
-Most parser generators, such as Bison, do not have to dynamically allocate temporary data on the
-heap, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
-computation as possible before the input has been validated, and we do not want to mix temporary
-objects with the results of our parse, we use an append-only trace to store intermediate parser
-results.
+Furthermore, the internal representation does not include any references to the input stream, which
+can therefore be zeroed immediately after the parser succeeds, so an attacker has to write an
+exploit that works without referencing data in the raw input.
+% \paragraph{Intermediate representation.}
 
-Hammer solves this problem by storing a full abstract syntax tree. However, this
-abstract syntax tree is at least an order of magnitude larger than the input,
-because it stores a large tree node structure for each input byte and for each
-rule reduced. This allows Hammer semantic actions to get all of the necessary
-information without ever seeing the raw input stream. However, because we also
-automatically generate our second pass, which corresponds to Hammer's semantic
-actions, we can trust it as much as we trust the parser, and thus can expose it
-to the raw input stream.
+% Most parser generators, such as Bison, do not have to dynamically allocate temporary data on the
+% heap, as they evaluate a semantic action on every rule. However, as our goal is to perform as little
+% computation as possible before the input has been validated, and we do not want to mix temporary
+% objects with the results of our parse, we use an append-only trace to store intermediate parser
+% results.
 
-Under this premise, the actions need limited information from the
-recognizer to correctly handle the input stream. In particular, the parser's
-control flow branches only at the choice, repetition, and constant combinators.
-Thus, for each of those combinators, we store the minimum amount of information
-required to reconstruct the syntactic structure of the input.
+% Hammer solves this problem by storing a full abstract syntax tree. However, this
+% abstract syntax tree is at least an order of magnitude larger than the input,
+% because it stores a large tree node structure for each input byte and for each
+% rule reduced. This allows Hammer semantic actions to get all of the necessary
+% information without ever seeing the raw input stream. However, because we also
+% automatically generate our second pass, which corresponds to Hammer's semantic
+% actions, we can trust it as much as we trust the parser, and thus can expose it
+% to the raw input stream.
 
-The trace is an
-array of integers.
-Whenever the parser encounters a choice, it appends two integers to the trace:
-the number of that choice and
-the length of the trace when it began parsing that choice. When backtracking in
-the input, the parser does not backtrack in the trace. This means that offsets
-within the trace can be used for a Packrat hash table to memoize backtrack-heavy
-parsers.
-When encountering a repetition combinator, the parser records the number of
-times the inner parser succeeded, and when encountering a constant parser of
-variable size, it records how much input was consumed by the constant parser.
+% Under this premise, the actions need limited information from the
+% recognizer to correctly handle the input stream. In particular, the parser's
+% control flow branches only at the choice, repetition, and constant combinators.
+% Thus, for each of those combinators, we store the minimum amount of information
+% required to reconstruct the syntactic structure of the input.
 
-In a second pass, the parser then allocates the internal representation from an
-arena allocator and binds the fields to values from the input, while following
-the trace to determine how many array fields to parse and which choices to pick.
+% The trace is an
+% array of integers.
+% Whenever the parser encounters a choice, it appends two integers to the trace:
+% the number of that choice and
+% the length of the trace when it began parsing that choice. When backtracking in
+% the input, the parser does not backtrack in the trace. This means that offsets
+% within the trace can be used for a Packrat hash table to memoize backtrack-heavy
+% parsers.
+% When encountering a repetition combinator, the parser records the number of
+% times the inner parser succeeded, and when encountering a constant parser of
+% variable size, it records how much input was consumed by the constant parser.
+
+% In a second pass, the parser then allocates the internal representation from an
+% arena allocator and binds the fields to values from the input, while following
+% the trace to determine how many array fields to parse and which choices to pick.
 

### 2014-05-01 paper/osdi14/design.tex

@@
 
-We present both a systematic overview of Nail's syntax with short examples in
-Figure~\ref{fig:syntax} shows the elements of Nail grammars and explain our design in more detail
-below, using a grammar for the well-known DNS protocol as a guiding example.
+We present both a systematic overview of Nail's syntax with short examples
+in Figure~\ref{fig:syntax}, and explain our design in more detail below,
+using a grammar for the well-known DNS protocol as a running example
+(shown in Figure~\ref{fig:dns-full}).
 

### 2014-05-02 paper/osdi14/eval.tex
Motivation figure

@@
 \end{tabular}
-\caption{Comparison of code size for three applications written in Nail,
-         and several existing alternative implementations.}
+\caption{Comparison of code size for three applications written in Nail,and several existing alternative implementations.}
 \label{fig:effort}
@@
 custom validators, semantic actions, and data structure definitions, with
-only 52 lines of code defining the grammar with Hammer's combinators.
+ 52 lines of code defining the grammar with Hammer's combinators.
 Their DNS server does not implement label compression, zone files, etc.

### 2014-05-02 paper/osdi14/fig-motivation.tex
Motivation figure

\begin{figure}
\begin{tabular}{@{}llll @{}}
 \bf Classification & \bf Example CVE & \bf Example description & \bf Count\\
  Memory corruption\footnote{We classified the following vulnerabilities as memory corruption
    attacks based on their description: 2013-5660, - 0742, -0138, 2012-4987, -1163,-1162, 2011-2265,
    2010-4535, -1657, -1336, -1218} & 2013-5660 & Buffer overflow & 11\\
  Parser ambiguities & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\
  Semantic misunderstanding & 2014-2319 & Compressor uses weak cryptography, even if the user
  selects AES & 1\\
  \multicolumn{3}{l}{\bf Total vulnerabilities related .zip processing }& \bf 16  \\  
  Mishandling archive contents &  2012-4389 & A web application overwrites .htaccess files and
  executable files by extracting uploaded .ZIP files to a publicly accessible folder  & 25\\
   Unrelated to .zip &2011-3371& Vulnerability leaking postal Zip codes & 42\\
  \multicolumn{3}{l}{\bf Total vulnerabilities} & \bf 83\\

\end{tabular}
\caption{Classification of known vulnerabilities in the CVE database between 2010 and 2014 related to the search term ``ZIP''}
\label{fig:motivation} 
\end{figure}

### 2014-05-02 paper/osdi14/motivation.tex
Motivation figure

--- 2014-05-01 paper/osdi14/motivation.tex
@@
 
-\paragraph{Vulnerability mitigation.}
+To motivate the need for Nail, this section presents a case study of
+vulnerabilities due to ad-hoc input parsing and output generation.
 
-Input processing vulnerabilities fall into two broad classes.  The first
-class is memory safety bugs, such as buffer overflows, which allow an
+\paragraph{Parsing bugs} Two recent high profile security vulnerabilities in mobile operating systems are due to improper
+input recognition, without using any memory corruption. In both cases, when the vulnerabilities were
+fixed, a similar flaw was exposed immediately afterwards, showing the need for a different approach
+to input handling that eliminates those vulnerabilities by design.
+
+The Evasi0n jailbreak for iOS 6~\cite{evaders6} relies on the XNU kernel and
+user-mode code-signing verifier interpreting executable metadata differently, so the code signature
+sees different bytes at a virtual address than kernel maps into the process. The next version of iOS
+added an explicit check for this particular metadata inconsistency. However, because parsing and
+processing of the input data is still mixed, the jailbreakers could set a flag that re-introduced
+the ambiguity after the check, but before signatures are verified~\cite{geohot-evasion}, which
+allowed iOS 7 to be jailbroken.
+
+Implementations of the X.509 format for SSL certificates were also shown to parse various ASN.1
+structures, such as length fields, strings containing invalid characters and repeated fields with
+conflicting values different ~\cite{DBLP:conf/fc/KaminskyPS10}. For example, certificates could be
+crafted where the X.509 parsers used certificate authorities and web browsers disagree on the
+subject of the certificate, allowing an attacker to get certificates for domains they do not control.
+
+Similarly, the infamous Android master key bug~\cite{saurik-masterkey} completely bypassed Android
+security exploited parser inconsistencies
+between the ZIP handler that checks signatures for privileged applications and the ZIP
+implementation that ultimately extract those files. Thus, privileged application bundles could be
+modified to include malicious code without breaking their signatures. Google quickly fixed this
+particular parser inconsistency, but another vulnerability, based on a different inconsistency
+between the parsers, was quickly disclosed~\cite{saurik-masterkey2}.
+
+Android application bundles are based on ZIP files, so we decided to conduct a systematic studies of
+vulnerabilities related to ZIP parsing.  
+The ZIP format has been associated with many vulnerabilities, and the
+PROTOS Genome project~\cite{protoc-c10-archive} found numerous security vulnerabilities in most
+implementations of ZIP and other archive formats that are directly
+related to input handling. We extend this study by looking at the CVE database.
+
+\input{fig-motivation}
+Figure~\ref{fig:motivation} shows  the 83 vulnerabilities the CVE database~\cite{cve-database} lists
+ for the search string ``ZIP''. 16 of these vulnerabilities are related to processing .ZIP archives
+ themselves; the rest are either totally unrelated to .ZIP archives or involve applications insecurely using
+ the contents of untrusted ZIP files.
+
+These input processing vulnerabilities fall into two broad classes.  The first
+class, of occurred 11 times, is memory safety bugs, such as buffer overflows, which allow an
 adversary to corrupt the application's memory using specially crafted
@@
 
-The second class is logic errors, where application code misinterprets
-input data. Safe languages and exploit mitigation technologies do not
-help against such vulnerabilities. This can lead to serious security
-consequences when two systems disagree on the meaning of a network
-packet or a signed message.  For example, the XNU kernel in iOS and the
-user-mode code-signing verifier interpret executable metadata differently,
-so the code signature sees different bytes at a virtual address than the
-executable that runs~\cite{evaders6}.  Android's code signing system also
-had a vulnerability due to parsing ambiguities~\cite{saurik-masterkey},
-and parsing ambiguities even contributed to attacks on the X.509 protocol
-underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  Logic errors can also
-cause memory corruption bugs as well, for example when a protocol stores
-the size of a buffer in two places and the implementation does not check
-their consistency. These mistakes are highly application-specific, and
+The second class, which occurred four times in our study is logic errors, where application code
+misinterprets input data. Safe languages and exploit mitigation technologies do not help against
+such vulnerabilities. This can lead to serious security consequences when two systems disagree on
+the meaning of a network packet or a signed message, as shown by the vulnerabilities we described
+before. CVE-2013-0211 shows that logic errors can be the underlying cause of memory corruption, when
+one part of a parser interprets a size field as a signed integer and another interprets it as an
+unsigned integer. CVE-2013-7338 is a logic error that allows an attacker to craft ZIP files that are
+incorrectly extracted or result in applications hangs with applications using a Python ZIP library,
+because this library does not check that two fields that contain the size of a file contain the same value.
+
+ These mistakes are highly application-specific, and
 are difficult to mitigate using existing techniques, and these mistakes
@@
 logic errors and inconsistencies in parsing and output generation.
+e through endless loops
+or incorrectly extracted ZIP files.
+
+The fourth vulnerability (CVE-\XXX) is the Android master key bug mentioned above, which results
+from one parser interpreting a size field as an unsigned integer, another interpreting the same
+field as a signed integer.
 

### 2014-05-02 paper/osdi14/fig-motivation.tex
Figure

@@
-\begin{figure}
+\begin{figure*}
 \begin{tabular}{@{}llll @{}}
@@
   selects AES & 1\\
-  \multicolumn{3}{l}{\bf Total vulnerabilities related .zip processing }& \bf 16  \\  
-  Mishandling archive contents &  2012-4389 & A web application overwrites .htaccess files and
-  executable files by extracting uploaded .ZIP files to a publicly accessible folder  & 25\\
+  \multicolumn{3}{l}{\bf Total vulnerabilities related to .zip processing }& \bf 16  \\  
+  Mishandling archive contents &  2012-4389 & 
+\begin{minipage}{7cm}
+A web application overwrites .htaccess files and
+  executable files by extracting uploaded .ZIP files to a publicly accessible folder  
+\end{minipage}& 25\\
    Unrelated to .zip &2011-3371& Vulnerability leaking postal Zip codes & 42\\
@@
 \label{fig:motivation} 
-\end{figure}
+\end{figure*}

### 2014-05-02 paper/osdi14/fig-motivation.tex
x

@@
 \begin{figure*}
-\begin{tabular}{@{}llll @{}}
+\centering
+\begin{tabular}{llll}
+\toprule
  \bf Classification & \bf Example CVE & \bf Example description & \bf Count\\
+\midrule
   Memory corruption\footnote{We classified the following vulnerabilities as memory corruption
@@
   Parser ambiguities & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\
-  Semantic misunderstanding & 2014-2319 & Compressor uses weak cryptography, even if the user
-  selects AES & 1\\
+  Semantic misunderstanding & 2014-2319 & Weak cryptography used even if user selects AES & 1\\
   \multicolumn{3}{l}{\bf Total vulnerabilities related to .zip processing }& \bf 16  \\
-  Mishandling archive contents &  2012-4389 & 
-\begin{minipage}{7cm}
-A web application overwrites .htaccess files and
-  executable files by extracting uploaded .ZIP files to a publicly accessible folder  
-\end{minipage}& 25\\
-   Unrelated to .zip &2011-3371& Vulnerability leaking postal Zip codes & 42\\
-  \multicolumn{3}{l}{\bf Total vulnerabilities} & \bf 83\\
+\bottomrule
 
 \end{tabular}
-\caption{Classification of known vulnerabilities in the CVE database between 2010 and 2014 related to the search term ``ZIP''}
+\caption{Classification of known vulnerabilities in the CVE database
+between 2010 and 2014 related to the search term ``ZIP'' and involving
+the ZIP file format.}
 \label{fig:motivation} 

### 2014-05-02 paper/osdi14/motivation.tex
x

@@
 vulnerabilities due to ad-hoc input parsing and output generation.
+Broadly speaking, parsing vulnerabilities can lead to two kinds of
+problems---memory corruption and logic errors---and as we show, both
+are prevalent in software and lead to significant security problems.
 
-\paragraph{Parsing bugs} Two recent high profile security vulnerabilities in mobile operating systems are due to improper
-input recognition, without using any memory corruption. In both cases, when the vulnerabilities were
-fixed, a similar flaw was exposed immediately afterwards, showing the need for a different approach
-to input handling that eliminates those vulnerabilities by design.
+\paragraph{Widely exploited parsing errors.}
+
+Three recent high profile security vulnerabilities are due to logic
+errors in input processing.  In all cases, when the vulnerabilities were
+fixed, a similar flaw was exposed immediately afterwards, showing the
+need for a different approach to input handling that eliminates those
+vulnerabilities by design.
 
@@
 structures, such as length fields, strings containing invalid characters and repeated fields with
-conflicting values different ~\cite{DBLP:conf/fc/KaminskyPS10}. For example, certificates could be
+conflicting values different~\cite{DBLP:conf/fc/KaminskyPS10}. For example, certificates could be
 crafted where the X.509 parsers used certificate authorities and web browsers disagree on the
@@
 between the parsers, was quickly disclosed~\cite{saurik-masterkey2}.
+
+\paragraph{Case study: ZIP file handling.}
 
@@
 or incorrectly extracted ZIP files.
-
-The fourth vulnerability (CVE-\XXX) is the Android master key bug mentioned above, which results
-from one parser interpreting a size field as an unsigned integer, another interpreting the same
-field as a signed integer.
 

### 2014-05-02 paper/osdi14/motivation.tex
typo

--- 2014-05-02 paper/osdi14/motivation.tex
@@
 These input processing vulnerabilities fall into two broad classes.  The first
-class, of occurred 11 times, is memory safety bugs, such as buffer overflows, which allow an
+class, which occurred 11 times, is memory safety bugs, such as buffer overflows, which allow an
 adversary to corrupt the application's memory using specially crafted

### 2014-05-02 paper/osdi14/motivation.tex
Motivation

@@
 
-Implementations of the X.509 format for SSL certificates were also shown to parse various ASN.1
-structures, such as length fields, strings containing invalid characters and repeated fields with
-conflicting values different~\cite{DBLP:conf/fc/KaminskyPS10}. For example, certificates could be
-crafted where the X.509 parsers used certificate authorities and web browsers disagree on the
-subject of the certificate, allowing an attacker to get certificates for domains they do not control.
+Similarly, vulnerabilities in X.509 parsers for SSL certificates allowed attackers to get
+certificates for domains they do not control. First, Moxie Marlinspike discovered that the X.509
+parsers in popular browsers handle NUL-bytes in certificates incorrectly~\cite{Marlinspike:SSL}. After this
+vulnerability was fixed, Dan Kaminsky discovered~\cite{DBLP:conf/fc/KaminskyPS10} that other
+structures, such as length fields and duplicated data, were also handled incorrectly.
 

### 2014-05-02 paper/osdi14/design.tex
Goals motivated by bugs

@@
 
-Nail's goals are to reduce programmer effort required to safely interact
-with data formats.  In particular, this means:
+Nail's goals are to reduce programmer effort required to safely interact with data formats and
+prevent vulnerabilities like those described in \S\ref{s:motivation}. In particular, this means:
 
@@
 
-\item Using the grammar to define both the \emph{external format}
-      and the \emph{internal representation}.
+\item Using the grammar to define both the \emph{external format} and the \emph{internal
+    representation}, allowing the same grammar to be re-used in multiple applications, avoiding
+  vulnerabilities like the Android Master Key.
 
@@
       generating outputs from internal representations, without
-      requiring the programmer to write any semantic actions.
+      requiring the programmer to write any semantic actions. This prevents
+      vulnerabilities such as the XNU bug, where format recognition and data processing
+      are mixed and interact in unexpected ways.
 
@@
       of a container data structure, to provide programmers
-      an unambiguous view of the data.
+      an unambiguous view of the data, to avoid bugs such as in the Python ZIP library.
 

### 2014-05-02 paper/osdi14/intro.tex
Added CVE bugs

@@
 different instances of the same value, an attacker can craft input that
-causes inconsistencies~\cite{cve-python-zip-maybe}.  Furthermore, security
+causes inconsistencies~\cite{evaders6}.  Furthermore, security
 vulnerabilities can occur when an application assumes two repetitions
@@
 the value of one size field and copying into that buffer based on the
-value of another~\cite{cve-some-example}.
+value of another~\cite{cve-2012-7338}.
 

### 2014-05-02 paper/osdi14/abstract.tex
Fixed citations

@@
 bijection} between the data format and the object model.  Third,
-Nail introduces {\em dependency fields} and {\em stream transforms} to
+Nail introduces {\em dependent fields} and {\em stream transforms} to
 capture protocol features such as size and offset fields, checksums,

### 2014-05-02 paper/osdi14/eval.tex
Fixed citations

@@
 they are equal first, as was the case in the Python ZIP
-library~\cite{cve-python-zip-maybe}.
+library~\cite{cve-2012-7338}.
 

### 2014-05-02 paper/osdi14/intro.tex
Fixed citations

@@
 Most parser generators focus on just parsing an input, rather
-than producing an output, thus requiring the programmer to manually
-construct outputs, which is error-prone.
-\XXX[do we have any evidence to cite for this claim?]
+than producing also an output, thus requiring the programmer to manually
+construct outputs, which is work-intensive and more code that could contain errors.
 Some parser generators, such as

### 2014-05-02 paper/osdi14/motivation.tex
Fixed citations

@@
 The ZIP format has been associated with many vulnerabilities, and the
-PROTOS Genome project~\cite{protoc-c10-archive} found numerous security vulnerabilities in most
+PROTOS Genome project~\cite{protos-c10-archive} found numerous security vulnerabilities in most
 implementations of ZIP and other archive formats that are directly

### 2014-05-02 paper/osdi14/design.tex
Design XXX

@@
 \paragraph{Choices.}
-Our choice combinator takes multiple labeled parsers and tries each of them in turn, backtracking
-to the initial position if an option fails. This behavior for the choice combinator, the defining
-feature of Parser Expression Grammars~\cite{ford2002packrat}, resolves any ambiguities explicitly.
-Internally, the choice combinator represents choices as tagged unions.
+If multiple structures can appear at a given position in a format, the programmer lists the options
+along with a label for each in the \cc{choose} combinator. 
+During parsing, Nail remembers the current input position and attempts each option in the order they
+appear in the grammar. If an option fails, the parser backtracks to the initial position. If no
+options succeed, the entire combinator fails. In the data model, choices are represented as tagged
+unions.  The programmer has to be careful when options overlap, because if the programmer meant to
+generate output for a choice, but the external representation is also valid for an earlier,
+higher-priority option, the parser will interpret it as such. However, real data formats normally do
+not have this overlap and we did not encounter it in the grammars we wrote.
 
-If two options the same string could be parser by two options, generated output for the latter
-option is not necessarily understood identically by the parser. However, actual data formats are
-usually not ambiguous in this sense.
-\XXX[nz][I don't understand this paragraph (esp the first sentence).]
+% If two options the same string could be parsed by two options, generated output for the latter
+% option is not necessarily understood identically by the parser. However, actual data formats are
+% usually not ambiguous in this sense.
+% \XXX[nz][I don't understand this paragraph (esp the first sentence).]
 
@@
 \end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
-
-\XXX[nz][Do you need \cc{*} here?  It seems like \cc{optional paren} would
-already have a C representation of \cc{paren*} according to our big figure.]
 

### 2014-05-02 paper/osdi14/eval.tex
Shortened code

@@
 library~\cite{cve-2012-7338}.
-
+\begin{comment}
 \begin{figure}[tb]
@@
 \end{figure}
-
+\end{comment}
 Figure~\ref{fig:zip-extract} shows an abbreviated version of our ZIP
@@
 streams based on the \cc{zip\_end\_of\_directory} transform in line 2
-(whose two C function signatures are shown in
-Figure~\ref{fig:zip-eod-xform}). \cc{zip\_end\_of\_directory_parse}
+%(whose two C function signatures are shown in Figure~\ref{fig:zip-eod-xform}). 
+\cc{zip\_end\_of\_directory_parse}
 finds the end-of-directory header as described above, by scanning the

### 2014-05-02 paper/osdi14/code/dns-xform.c

@@ -1,10 +1,9 @@
-int
-dnscompress_parse(NailArena *tmp,
+
+int dnscompress_parse(NailArena *tmp,
                   NailStream *out_decomp,
                   NailStream *in_current);
 
-int
-dnscompress_generate(NailArena *tmp,
+int dnscompress_generate(NailArena *tmp,
                      NailStream *in_decomp,
                      NailStream *out_current);
 

### 2014-05-02 paper/osdi14/code/offset-xform.c

@@ -1,19 +1,15 @@
-int
-offset_u32_parse(NailArena *tmp,
+int offset_u32_parse(NailArena *tmp,
                  NailStream *out_str,
                  NailStream *in_current,
-                 const uint32_t *off)
-{
+                 const uint32_t *off){
   /* out_str = suffix of in_current
                at offset *off */
 }
 
-int
-offset_u32_generate(NailArena *tmp,
+int offset_u32_generate(NailArena *tmp,
                     NailStream *in_fragment,
                     NailStream *out_current,
-                    uint32_t *off)
-{
+                    uint32_t *off){
   /* *off = position of out_current */
   /* append in_fragment to out_current */
 }

### 2014-05-02 paper/osdi14/abstract.tex
Change title and abstract

@@
 
-\textit{Nail} is an tools that greatly reduces the
+\textit{Nail} is a tool that greatly reduces the
 programmer effort for safely parsing and generating data formats defined
@@
 experiments show that a Nail-based DNS server can outperform the widely
-used BIND DNS server, demonstrating that Nail parsers can achieve good
+used BIND DNS server, demonstrating that systems built with Nail can achieve good
 performance.

### 2014-05-02 paper/osdi14/motivation.tex
typo

@@
 unsigned integer. CVE-2013-7338 is a logic error that allows an attacker to craft ZIP files that are
-incorrectly extracted or result in applications hangs with applications using a Python ZIP library,
+incorrectly extracted or result in application hangs with applications using a Python ZIP library,
 because this library does not check that two fields that contain the size of a file contain the same value.
@@
 logic errors and inconsistencies in parsing and output generation.
-e through endless loops
-or incorrectly extracted ZIP files.
 

### 2014-05-02 paper/osdi14/design.tex
Rewrote n_of. one paragraph over

@@
 not have this overlap and we did not encounter it in the grammars we wrote.
+An example is provided in Figure~\ref{fig:grammar-arith}.
 
@@
 The reference combinator does not need to used when another combinator, such as \cc{optional} or
-\cc{many} already introduces indirection in the data type.
-For example example,   arithmetic expressions consisting of sums and products can be parsed as follows:
+\cc{many} already introduces indirection in the data type. An example is shown in Figure~\ref{fig:grammar-arith}.
 
+\begin{figure}
 \begin{verbatim}
@@
 \end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
-
+\label{fig:grammar-arith}
+\caption{Grammar for sums and products of integers.}
+\end{figure}
 % To support recursive structures, we include a reference combinator \cc{*}, which
@@
 
-Dependent fields are parsed just like other fields in the external format, but they are not stored
-in the internal data type. Instead, the parser stores the value in a
-temporary location, so that the field can be referenced by subsequent parsers.
-
+Dependent fields are handled like other fields when parsing input, but their values are not stored
+in the internal data type. Instead the value can be referenced by subsequent parsers and it
+discarded when the field goes out of scope.
 When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
@@
 
-Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which is used
-for count fields. For example, DNS labels are encoded as Pascal-style strings: a length followed by
-the value of the label. The Nail grammar for DNS labels can represent
-this pattern as:
-
-{
-\smaller[0.5]
-\begin{verbatim}
-  {
-    @length uint8 | 1..64
-    label n_of @length uint8
-  }
-\end{verbatim}
-}
-
+Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
+the \cc{many} combinator, except it represents an exact number, specified in the dependent field, of
+repetitions, as opposed as many repetitions as possible. For example, DNS labels, which
+are encoded as a length followed by
+a value are described in line 38 of the DNS grammar.
 Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but

### 2014-05-02 paper/osdi14/design.tex
typo

@@
 \item Using the grammar to define both the \emph{external format} and the \emph{internal
-    representation}, allowing the same grammar to be re-used in multiple applications, avoiding
+    representation}, allowing the same grammar to be re-used in multiple programs, avoiding
   vulnerabilities like the Android Master Key.
@@
       requiring the programmer to write any semantic actions. This prevents
-      vulnerabilities such as the XNU bug, where format recognition and data processing
+      vulnerabilities such as the XNU bug, where format recognition and semantics
       are mixed and interact in unexpected ways.

### 2014-05-02 paper/osdi14/impl.tex
another type

@@
 specific data at predictable addresses\cite{jp-advanced, vudo-malloc}. Because the rest of the
-application or even Nail's generated code may contain memory corruption bug, Nail
+application or even Nail's generated code may contain memory corruption bugs, Nail
 carefully handles memory allocations as defense-in-depth to make exploiting such vulnerabilities harder.

### 2014-05-02 paper/osdi14/intro.tex
Future work removed
intro and relwk nits

@@
 an {\tt unzip} utility in about 50 lines of C code, with performance comparable
-to or exceeding existing implementations.This suggests both that Nail is
+to or exceeding existing implementations. This suggests both that Nail is
 effective at handling complex real-world data formats, and that Nail makes it
@@
 \S\ref{s:design} describes
-Nail's design.  \S\ref{s:impl} discusses our initial implementation
-of Nail.  \S\ref{s:eval} provides evaluation results.
-\S\ref{s:future} suggests several directions for future work, and
+Nail's design.  \S\ref{s:impl} discusses our implementation
+of Nail.  \S\ref{s:eval} provides evaluation results, and
 \S\ref{s:concl} concludes.

### 2014-05-02 paper/osdi14/relwk.tex
intro and relwk nits

--- 2014-05-01 paper/osdi14/relwk.tex
@@
 terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing are handled transparently.
+endianness and bit-packing, are handled transparently.
 
@@
 as programming languages and configuration files. Frequently, such
-languages are often specified with a formal grammar in an executable
+languages are specified with a formal grammar in an executable
 form. Unfortunately, parser frameworks are seldom used to recognize

### 2014-05-02 paper/osdi14/intro.tex
x

@@
 formats, we implemented a prototype of Nail for C\@. Using our prototype, we
-implemented succinct grammars for parts of an IP network stack, for DNS packets,
+implemented grammars for parts of an IP network stack, for DNS packets,
 and for ZIP files, each in about a hundred lines of grammar. On top of these

### 2014-05-02 paper/osdi14/motivation.tex
motivation nits

@@
 Similarly, the infamous Android master key bug~\cite{saurik-masterkey} completely bypassed Android
-security exploited parser inconsistencies
+security by exploiting parser inconsistencies
 between the ZIP handler that checks signatures for privileged applications and the ZIP
-implementation that ultimately extract those files. Thus, privileged application bundles could be
+implementation that ultimately extracts those files. Thus, privileged application bundles could be
 modified to include malicious code without breaking their signatures. Google quickly fixed this
@@
 
-Android application bundles are based on ZIP files, so we decided to conduct a systematic studies of
-vulnerabilities related to ZIP parsing.  
+To understand the impact of parsing mistakes in real-world software,
+we conducted a systematic study of
+vulnerabilities related to ZIP file parsing.  
 The ZIP format has been associated with many vulnerabilities, and the
@@
 \input{fig-motivation}
-Figure~\ref{fig:motivation} shows  the 83 vulnerabilities the CVE database~\cite{cve-database} lists
- for the search string ``ZIP''. 16 of these vulnerabilities are related to processing .ZIP archives
- themselves; the rest are either totally unrelated to .ZIP archives or involve applications insecurely using
- the contents of untrusted ZIP files.
+
+We found 83 vulnerabilities in the CVE database~\cite{cve-database} that
+that mention the search string ``ZIP.''  Just 16 of these vulnerabilities
+were related to processing ZIP archives; the rest were unrelated to ZIP
+archives or involved applications insecurely using the contents of untrusted
+ZIP files.  Figure~\ref{fig:motivation} summarizes the 16 ZIP-related
+vulnerabilities.
 
@@
 inputs.  These mistakes arise in lower-level languages that do not
-provide memory safety guarantees such as C, and can be partially
+provide memory safety guarantees, such as C, and can be partially
 mitigated by a wide range of techniques, for example static analysis,
@@
 make it more difficult for an adversary to exploit these bugs.
-Nail helps developers of lower-level languages avoid these bugs
+Nail helps developers using lower-level languages to avoid these bugs
 in the first place.
 
-The second class, which occurred four times in our study is logic errors, where application code
+The second class, which occurred four times in our study,
+is logic errors, where application code
 misinterprets input data. Safe languages and exploit mitigation technologies do not help against
@@
 because this library does not check that two fields that contain the size of a file contain the same value.
+The Android ZIP file signature verification bug that we described
+earlier was also among these 4 vulnerabilities.
 

### 2014-05-02 paper/osdi14/fig-motivation.tex
right-align numbers
footnote moved

@@
 \centering
-\begin{tabular}{llll}
+\begin{tabular}{lllr}
 \toprule
@@
 \midrule
-  Memory corruption\footnote{We classified the following vulnerabilities as memory corruption
-    attacks based on their description: 2013-5660, - 0742, -0138, 2012-4987, -1163,-1162, 2011-2265,
-    2010-4535, -1657, -1336, -1218} & 2013-5660 & Buffer overflow & 11\\
+  Memory corruption & 2013-5660 & Buffer overflow & 11\\
   Parser ambiguities & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\

### 2014-05-02 paper/osdi14/motivation.tex
footnote moved

@@
 These input processing vulnerabilities fall into two broad classes.  The first
-class, which occurred 11 times, is memory safety bugs, such as buffer overflows, which allow an
+class, which occurred 11 times\footnote{We classified the following vulnerabilities as memory corruption
+    attacks based on their description: 2013-5660, - 0742, -0138, 2012-4987, -1163,-1162, 2011-2265,
+    2010-4535, -1657, -1336, -1218}, is memory safety bugs, such as buffer overflows, which allow an
 adversary to corrupt the application's memory using specially crafted

### 2014-05-02 paper/osdi14/code/offset-xform.c
x

@@ -1,15 +1,15 @@
 int offset_u32_parse(NailArena *tmp,
-                 NailStream *out_str,
-                 NailStream *in_current,
-                 const uint32_t *off){
+  NailStream *out_str, NailStream *in_current,
+  const uint32_t *off)
+{
   /* out_str = suffix of in_current
                at offset *off */
 }
 
 int offset_u32_generate(NailArena *tmp,
-                    NailStream *in_fragment,
-                    NailStream *out_current,
-                    uint32_t *off){
+  NailStream *in_fragment,
+  NailStream *out_current, uint32_t *off)
+{
   /* *off = position of out_current */
   /* append in_fragment to out_current */
 }

### 2014-05-02 paper/osdi14/impl.tex
x

@@
 Security exploits often rely on raw inputs being present in memory, for example to include
-shell-code or crafted stack frames for ROP-attacks in padding fields\cite{shotgun-parser} or the
+shell-code or crafted stack frames for ROP-attacks in padding fields~\cite{shotgun-parser} or the
 application executing a controlled sequence of heap allocations and de-allocations to place
-specific data at predictable addresses\cite{jp-advanced, vudo-malloc}. Because the rest of the
+specific data at predictable addresses~\cite{jp-advanced, vudo-malloc}. Because the rest of the
 application or even Nail's generated code may contain memory corruption bugs, Nail

### 2014-05-02 paper/osdi14/code/dns-xform.c

@@ -1,9 +1,7 @@
-
 int dnscompress_parse(NailArena *tmp,
-                  NailStream *out_decomp,
-                  NailStream *in_current);
+  NailStream *out_decomp,
+  NailStream *in_current);
 
 int dnscompress_generate(NailArena *tmp,
-                     NailStream *in_decomp,
-                     NailStream *out_current);
-
+  NailStream *in_decomp,
+  NailStream *out_current);

### 2014-05-02 paper/osdi14/eval.tex
typo
eval nits

@@
 
-The \cc{zip\_file} grammarfirst splits the entire file stream into two
-streams based on the \cc{zip\_end\_of\_directory} transform in line 2
+The \cc{zip\_file} grammar first splits the entire file stream into two
+streams based on the \cc{zip\_end\_of\_directory} transform on line 2.
 %(whose two C function signatures are shown in Figure~\ref{fig:zip-eod-xform}). 
@@
 
-The \cc{file} rule starting at line 41, describing a ZIP file entry takes dependent field
+The \cc{file} rule starting at line 41, describing a ZIP file entry,
+takes dependent field
 parameters containing file metadata information from the directory
@@
 \end{tabular}
-\caption{Comparison of code size for three applications written in Nail,and several existing alternative implementations.}
+\caption{Comparison of code size for three applications written in
+  Nail,and a comparable existing implementation without Nail.}
 \label{fig:effort}
@@
 
-In \S\ref{s:motivation}, we found 14 input handling vulnerabilities related to ZIP files.
+In \S\ref{s:motivation}, we found 15 input handling vulnerabilities
+related to ZIP files.
 
-Ten of these vulnerabilities involved memory corruption during input handling.
+11 of these vulnerabilities involved memory corruption during input handling.
 Because Nail's generated code checks offsets before reading and does not

### 2014-05-02 paper/osdi14/concl.tex
done with pass

@@
 This paper presented the design and implementation of \textit{Nail},
-an interface generator for data formats.  Nail helps programmers avoid
+a tool for parsing and generating complex data formats based on a
+precise grammar.  Nail helps programmers avoid
 memory corruption and ambiguity vulnerabilities while reducing effort
@@
 fewer lines of code.
+Nail and all of the applications and grammars developed in this paper
+are released as open-source software.
 

### 2014-05-02 paper/osdi14/eval.tex
done with pass

@@
 our Nail-based DNS server can achieve higher performance and lower latency
-than BIND\@.  Although BIND is a much more sophisticated DNS server,
+than BIND\@.  Although BIND is a more sophisticated DNS server,
 and implements many features that are not present in our Nail-based DNS
 server, we believe our results demonstrate that Nail's parsers are not
-a barrier to achieving acceptable performance.
+a barrier to achieving good performance.
 

### 2014-05-02 paper/osdi14/eval.tex
Typo

--- 2014-05-02 paper/osdi14/eval.tex
@@
 
-The \cc{zip\_file} grammarfirst splits the entire file stream into two
+The \cc{zip\_file} grammar first splits the entire file stream into two
 streams based on the \cc{zip\_end\_of\_directory} transform in line 2

### 2014-05-02 paper/osdi14/design.tex
x
make spell

@@
 For example, the grammar for  Ethernet headers uses \cc{optional vlan\_header} to parse the VLAN
-header that only appears in Ethernet packets transmitted to a non-default VLAN.
+header that only appears in Ethernet packets transmitted to a non-default VLAN\@.
 
@@
 The reference combinator does not need to used when another combinator, such as \cc{optional} or
-\cc{many} already introduces indirection in the data type. An example is shown in Figure~\ref{fig:grammar-arith}.
+\cc{many}, already introduces indirection in the data type. An example is shown in Figure~\ref{fig:grammar-arith}.
 
@@
 \end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
+\caption{Grammar for sums and products of integers.}
 \label{fig:grammar-arith}
-\caption{Grammar for sums and products of integers.}
 \end{figure}

### 2014-05-02 paper/osdi14/eval.tex
make spell

--- 2014-05-02 paper/osdi14/eval.tex
@@
 because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
-parsers are not very reuseable, as they build application specific data structures and are tightly
+parsers are not very reusable, as they build application specific data structures and are tightly
 coupled to the rest of the code. Nail grammars, however, can be re-used between applications,

### 2014-05-02 paper/osdi14/motivation.tex
make spell

@@
 
-We found 83 vulnerabilities in the CVE database~\cite{cve-database} that
+We found 83 vulnerabilities in the CVE database~\cite{cve-database}
 that mention the search string ``ZIP.''  Just 16 of these vulnerabilities

### 2014-08-14 paper/osdi14/design.tex
cite data languages

@@
 bugs. However, as we will show in \S\ref{s:eval-effort}, Nail transformations are much shorter than
-hand-written parsers and many formats can be represented with just the provided transforms
+hand-written parsers and many formats can be represented with just the transformations in the
+standard library.
 

### 2014-08-14 paper/osdi14/impl.tex
cite data languages

--- 2014-05-02 paper/osdi14/impl.tex
@@
 
-% The source code of our implementation, together with the examples described in
-% \S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
+The source code of our implementation, together with the examples described in
+\S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
 
@@
 
+\XXX[Add API here]
 \paragraph{Defense-in-Depth.}
 
-Security exploits often rely on raw inputs being present in memory, for example to include
-shell-code or crafted stack frames for ROP-attacks in padding fields\cite{shotgun-parser} or the
+Security exploits often rely on raw inputs being present in memory~\cite{shotgun-parser}, for example to include
+shell-code or crafted stack frames for ROP~\cite{phrack58:4-nergal}-attacks in padding fields or the
 application executing a controlled sequence of heap allocations and de-allocations to place
-specific data at predictable addresses\cite{jp-advanced, vudo-malloc}. Because the rest of the
+specific data at predictable addresses~\cite{jp-advanced, vudo-malloc}. Because the rest of the
 application or even Nail's generated code may contain memory corruption bugs, Nail

### 2014-08-14 paper/osdi14/relwk.tex
cite data languages

@@
 
-Closely related work has been done at Bell Labs with the PacketTypes
-system~\cite{mccann2000packet}.  However, PacketTypes works only as a
-parser, not as an output generator, and does not support the expressive
-power of parsing expression grammars (PEGs), but rather implements a
-C-like structure model enhanced with length fields and constraints.
-PacketTypes also cannot handle complicated encodings such as compressed
-data, which are supported by Nail's stream transforms.
+Closely related work has been done in the field of data description languages, for example
+PacketTypes~\cite{mccann2000packet} and DataScript~\cite{DataScript}. However, PacketTypes works
+only as a parser, not as an output generator. Data description languages typically are not as
+expressive as parsing expression grammars (PEGs), but rather implement a C-like structure model
+enhanced with length fields and constraints. These systems also cannot handle complicated encodings
+such as compressed data, which are supported by Nail's stream transforms.
 

### 2014-08-14 paper/osdi14/relwk.tex
Related work (de-anonymized citation)
Cite DataScript

@@
 PacketTypes~\cite{mccann2000packet} and DataScript~\cite{DataScript}. However, PacketTypes works
-only as a parser, not as an output generator. Data description languages typically are not as
-expressive as parsing expression grammars (PEGs), but rather implement a C-like structure model
-enhanced with length fields and constraints. These systems also cannot handle complicated encodings
-such as compressed data, which are supported by Nail's stream transforms.
+only as a parser, not as an output generator and implement a C-like structure model enhanced with
+length fields and constraints. DataScript adds output generation and built-in support for offset
+fields. However, these systems cannot handle complicated encodings such as compressed data, which
+are supported by Nail's stream transforms. Many built-in features of these date description
+languages, such as the various ways of encoding array size or length and offsets, can be expressed
+by Nail transformations, keeping the core language itself smaller and increasing flexibility.
 
@@
 We presented an earlier design of Nail at a
-workshop~\cite{bangert:nail-langsec-anon}.  At that stage, Nail had only
+workshop~\cite{bangert:nail-langsec}.  At that stage, Nail had only
 limited support for dependent fields, and did not support stream

### 2014-08-14 paper/osdi14/relwk.tex
Cite Fisher, DDL

@@
 Closely related work has been done in the field of data description languages, for example
-PacketTypes~\cite{mccann2000packet} and DataScript~\cite{DataScript}. However, PacketTypes works
+PacketTypes~\cite{mccann2000packet} and DataScript~\cite{DataScript}. A broader overview of data
+description languages can be found in~\cite{Fisher:700DDL} However, PacketTypes works
 only as a parser, not as an output generator and implement a C-like structure model enhanced with

### 2014-08-21 paper/osdi14/abstract.tex
Explain BIND difference

@@
 experiments show that a Nail-based DNS server can outperform the widely
-used BIND DNS server, demonstrating that systems built with Nail can achieve good
+used BIND DNS server on an authoritative workload,
+ demonstrating that systems built with Nail can achieve good
 performance.

### 2014-08-21 paper/osdi14/eval.tex
Explain BIND difference

@@
 and implements many features that are not present in our Nail-based DNS
-server, we believe our results demonstrate that Nail's parsers are not
+server and allow it to be used in more complicated configurations, 
+we believe our results demonstrate that Nail's parsers are not
 a barrier to achieving good performance.

### 2014-08-21 paper/osdi14/intro.tex
Intro

@@
 
-A promising approach to avoid such vulnerabilities is to specify
+Typically, input parsers, particularly for binary formats, are often
+written by hand. Not only is such code plagued by the bugs mentioned above, but
+two separate hand-written implementations of the same protocol might disagree on
+the meaning of a particular input, which can lead to security
+vulnerabilities~\cite{saurik-masterkey}.
+
+A promising approach to avoid these problems is to specify
 a precise grammar for the input data format, and to use a parser
@@
 as the parser generator is bug-free, the application will be safe from
-input processing vulnerabilities.
+input processing vulnerabilities. Grammars can be re-used between
+applications, further reducing effort and eliminating ambiguities.
 

### 2014-09-09 paper/osdi14/relwk.tex
Typos
fix some undefined citations and other compile errors

--- 2014-08-14 paper/osdi14/relwk.tex
@@
 core concept of interface generators, such as CORBA~\cite{omg:corba},
-XDR~\cite{rfc:1832}, and Protocol Buffers~\cite{varda2008}.  However,
+XDR~\cite{RFC:1832}, and Protocol Buffers~\cite{varda2008}.  However,
 interface generators do not allow the programmer to specify the byte-level
@@
 Closely related work has been done in the field of data description languages, for example
-PacketTypes~\cite{mccann2000packet} and DataScript~\cite{DataScript}. However, PacketTypes works
+PacketTypes~\cite{mccann2000packet} and DataScript~\cite{back2002datascript}.
+A broader overview of data
+description languages can be found in~\cite{Fisher:700DDL}. However, PacketTypes works
 only as a parser, not as an output generator and implement a C-like structure model enhanced with

### 2014-09-09 paper/osdi14/eval.tex
whew, lots of bib crap

@@
 they are equal first, as was the case in the Python ZIP
-library~\cite{cve-2012-7338}.
+library~\cite{python-bug:20078}.
 \begin{comment}

### 2014-09-09 paper/osdi14/intro.tex
whew, lots of bib crap

@@
 the value of one size field and copying into that buffer based on the
-value of another~\cite{cve-2012-7338}.
+value of another~\cite{python-bug:20078}.
 

### 2014-09-10 paper/osdi14/motivation.tex
make spell

@@
 We found 83 vulnerabilities in the CVE database~\cite{cve-database}
-that mention the search string ``ZIP.''  Just 16 of these vulnerabilities
+that mention the search string ``ZIP\@.''  Just 16 of these vulnerabilities
 were related to processing ZIP archives; the rest were unrelated to ZIP

### 2014-09-10 paper/osdi14/design.tex
fix some undefined citations and other compile errors
Many:
-eliminated review_todo
-More details on why transformations are safe
-Bib fixes

@@
 }
-\end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
+\end{verbatim}
+% \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
 \caption{Grammar for sums and products of integers.}
@@
 standard library. 
+For example, our Zip transformations are 78 lines of code, compared to 1600 lines of code for a
+   hand-written parser. Additionally, Nail provides convenient and safe interfaces for allocating
+   memory and accessing streams that address the most common occurences of buffer overflow
+   vulnerabilities.
 

### 2014-09-10 paper/osdi14/impl.tex
impl.tex: fix an obvious typo pointed out by Reviewer C
Many:
-eliminated review_todo
-More details on why transformations are safe
-Bib fixes

@@
 performed with a subset of the grammar implemented using conventional grammars. 
-And option for C++ STL data models  is in development. In
+An option for C++ STL data models is in development.
+Figure~\ref{fig:dns-api} shows the API for the Nail-generated code.  In
 this section, we will discuss some particular features of our parser
@@
 
+\begin{figure}[h]
+\smaller[0.5]
+% \input{code/dns-api}
+\caption{API for the parser and generator Nail produces.}
+\label{fig:dns-api}
+\end{figure}
 \paragraph{Parsing.}
@@
 
-\XXX[Add API here]
 \paragraph{Defense-in-Depth.}
@@
 exploit that works without referencing data in the raw input.
+
+Finally, Nail only supports error handling in a special debug configuration and will only print
+error messages about the input to standard error. Besides complicating the parser, advanced error handling
+invites programmers to attempt and fix malformed input, such as adding reasonable defaults for a
+missing field. Such error-fixing not only introduces parser differential, but also might allow an
+attacker to sneak inconsistent input past a parser.
 % \paragraph{Intermediate representation.}

### 2014-09-10 paper/osdi14/intro.tex
some intro tweaks

@@
 data or parsing a file, is error-prone and is often exploited by
-attackers.  This is in part because attackers have precise control
-over the inputs to that code, and can craft inputs that trigger
-subtle corner cases in input processing.  For example, the libpng image decompression
-library has had 24 remotely exploitable vulnerabilities from 2007 to
-2013~\cite{cvedetails:libpng},
-and Adobe's PDF and Flash viewers have been notoriously
-plagued by input processing vulnerabilities.  Even relatively
-simple formats, such as those used by the zlib compression
-library, have had input processing vulnerabilities in the
-past~\cite{cvedetails:zlib}.
-
-Typically, input parsers, particularly for binary formats, are often
-written by hand. Not only is such code plagued by the bugs mentioned above, but
-two separate hand-written implementations of the same protocol might disagree on
-the meaning of a particular input, which can lead to security
-vulnerabilities~\cite{saurik-masterkey}.
+attackers.
+%   This is in part because attackers have precise control
+% over the inputs to that code, and can craft inputs that trigger
+% subtle corner cases in input processing.
+For example, the libpng image decompression library had 24 remotely
+exploitable vulnerabilities from 2007 to 2013~\cite{cvedetails:libpng},
+Adobe's PDF and Flash viewers have been notoriously plagued by input
+processing vulnerabilities, and even the zlib compression library had
+input processing vulnerabilities in the past~\cite{cvedetails:zlib}.
+With a memory-unsafe language like C, mistakes in input processing
+code can lead to memory errors like buffer overflows, but even with
+a memory-safe language like Java, inconsistencies between different
+parsers can lead to security issues~\cite{saurik-masterkey}.
 
@@
 update the application's internal representation of the data based on
-each AST node.  Not only does the programmer allocate memory and manipulate pointers,
- operations that are error-prone, but
-writing these semantic actions
-also requires the programmer to describe the structure of the input three
+each AST node.  
+Writing these semantic actions requires the programmer to describe the
+structure of the input three
 times---once to describe the grammar, once to describe the internal data
 structure, and once again in the semantic actions that translate the
-grammar into the data structure---leading to another potential source
-of bugs and inconsistencies.
+grammar into the data structure---leading to potential bugs and
+inconsistencies.
+Furthermore, in a memory-unsafe language like C, these semantic
+actions often involve error-prone manual memory allocation and
+pointer manipulation.
 
@@
 Most parser generators focus on just parsing an input, rather
-than producing also an output, thus requiring the programmer to manually
+than producing an output, thus requiring the programmer to manually
 construct outputs, which is work-intensive and more code that could contain errors.
@@
 semantic actions to be written, transforming the internal representation
-into an AST\@.
+back into an AST\@.
 

### 2014-09-10 paper/osdi14/code/dns-api.c
DNS api code

@@ -1,0 +1,6 @@
+struct dnspacket*parse_dnspacket(NailArena *arena, 
+  const uint8_t *data, 
+  size_t size);
+int gen_dnspacket(NailArena *tmp_arena,
+  NailStream *out,
+  struct dnspacket * val) ;

### 2014-09-10 paper/osdi14/relwk.tex
Add PADS citation

@@
 length fields and constraints. DataScript adds output generation and built-in support for offset
-fields. However, these systems cannot handle complicated encodings such as compressed data, which
-are supported by Nail's stream transforms. Many built-in features of these date description
-languages, such as the various ways of encoding array size or length and offsets, can be expressed
-by Nail transformations, keeping the core language itself smaller and increasing flexibility.
+fields. A
+particularly sophisticated data description language, PADS~\cite{Fisher:2005:PDL:1064978.1065046},
+which is targeted more towards offline analysis, even features built-in support for XML and
+automatic grammar inference.
 
-Parser generators for binary protocols were first introduced by the
-Hammer~\cite{hammer-parser} parser.
-While previous parser generators could also be used to write grammars
-for binary protocols,\footnote{Theoretically speaking, the alphabet
-over which a grammar is an abstract set, so most algorithms work
-just as well on an alphabet of $\{0,1\}$.} doing so is practically
-inconvenient. Hammer allows the programmer to specify a grammar in
-terms of bits and bytes instead of characters. Common concerns, such as
-endianness and bit-packing, are handled transparently.
+However, these systems cannot easily handle complicated encodings such as compressed data, which are
+supported by Nail's stream transforms. While sophisticated languages like PADS allow for handling
+particular variations of offset fields, compressed data or even XML entities, each of these
+features has to be implemented in the data description language - and all associated tools. Nail's
+transformations keep the core language practicably small, while enabling the wide range of
+features real-world protocols require.
 
+Recently, the Hammer project~\cite{hammer-parser} introduced a security-focused parser framework for
+binary protocols.
 Hammer implements grammars as language-integrated parser combinators, an approach popularized by

### 2014-09-10 paper/osdi14/design.tex

@@
    hand-written parser. Additionally, Nail provides convenient and safe interfaces for allocating
-   memory and accessing streams that address the most common occurences of buffer overflow
+   memory and accessing streams that address the most common occurrences of buffer overflow
    vulnerabilities.

### 2014-09-10 paper/osdi14/intro.tex
abstr and intro nits

@@
 With a memory-unsafe language like C, mistakes in input processing
-code can lead to memory errors like buffer overflows, but even with
+code can lead to memory errors like buffer overflows, and even with
 a memory-safe language like Java, inconsistencies between different
@@
 as the parser generator is bug-free, the application will be safe from
-input processing vulnerabilities. Grammars can be re-used between
-applications, further reducing effort and eliminating ambiguities.
+input processing vulnerabilities. Grammars can also be re-used between
+applications, further reducing effort and eliminating inconsistencies.
 
@@
 than producing an output, thus requiring the programmer to manually
-construct outputs, which is work-intensive and more code that could contain errors.
+construct outputs, which is work-intensive and leads to
+more code that could contain errors.
 Some parser generators, such as
@@
 explicitly check for consistency, and if different applications use
-different instances of the same value, an attacker can craft input that
+different instances of the same value, an attacker can craft an input that
 causes inconsistencies~\cite{evaders6}.  Furthermore, security
@@
 
-Finally, real-world data formats, for example PNG or PDF, are hard to
+Finally, real-world data formats, such as PNG or PDF, are hard to
 represent with existing parser generators.  Those parsers cannot directly
-deal with length or checksum fields directly, so the programmer has to
+deal with length or checksum fields, so the programmer has to
 either write potentially unsafe code to deal with such features, or build
@@
 This paper presents the design and implementation of Nail, a parser generator that greatly reduces
-the programmer effort required to use grammars. Nail addresses the above four challenges with
+the programmer effort required to use grammars. Nail addresses the above challenges with
 several key ideas, as follows.
@@
 Second, this well-defined internal representation allows Nail to
-maintain a {\em semantic bijection} between data formats and their
+establish a {\em semantic bijection} between data formats and their
 internal object model.  As a result, this enables Nail to not just

### 2014-09-10 paper/osdi14/concl.tex
include github URL
replace ambiguity with inconsistency
i think inconsistency more precisely captures what we mean

@@
 precise grammar.  Nail helps programmers avoid
-memory corruption and ambiguity vulnerabilities while reducing effort
+memory corruption and inconsistency vulnerabilities while reducing effort
 in parsing and generating real-world protocols and file formats.

### 2014-09-10 paper/osdi14/design.tex
replace ambiguity with inconsistency
i think inconsistency more precisely captures what we mean

@@
       of a container data structure, to provide programmers
-      an unambiguous view of the data, to avoid bugs such as in the Python ZIP library.
+      a consistent, unambiguous view of the data,
+      to avoid bugs such as in the Python ZIP library.
 

### 2014-09-10 paper/osdi14/eval.tex
make spell
replace ambiguity with inconsistency
i think inconsistency more precisely captures what we mean

@@
 
-Nail also protects against parsing ambiguity vulnerabilities like the four others we studied. First,
+Nail also protects against parsing inconsistency vulnerabilities like the four others we studied. First,
 Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
@@
 zero and three times, using a DNS server running on the local machine.
-We used a single core of an Intel i7-3610QM system with 12GB of RAM.
+We used a single core of an Intel i7-3610QM system with 12GB of RAM\@.
 The benchmark tool kept at most 20 queries outstanding at once, and

### 2014-09-10 paper/osdi14/fig-motivation.tex
replace ambiguity with inconsistency
i think inconsistency more precisely captures what we mean

@@
   Memory corruption & 2013-5660 & Buffer overflow & 11\\
-  Parser ambiguities & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\
+  Parser inconsistencies & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\
   Semantic misunderstanding & 2014-2319 & Weak cryptography used even if user selects AES & 1\\

### 2014-09-10 paper/osdi14/motivation.tex
replace ambiguity with inconsistency
i think inconsistency more precisely captures what we mean

@@
 processing of the input data is still mixed, the jailbreakers could set a flag that re-introduced
-the ambiguity after the check, but before signatures are verified~\cite{geohot-evasion}, which
+the inconsistency after the check, but before signatures are verified~\cite{geohot-evasion}, which
 allowed iOS 7 to be jailbroken.

### 2014-09-10 paper/osdi14/design.tex
typo
New figure

@@
 \end{CompactItemize}
+
+For example, Figure~\ref{fig:dns-full} shows an example grammar for DNS packets and the
+corresponding data type and functions that Nail produces.
+
+\input{fig-dnsfull}
 
@@
 
-\input{fig-dnsfull}
+\paragraph{Parsing.}
 \paragraph{Rules.}
@@
 introduces a layer of indirection, such as a reference or pointer, to the model data type.
-The reference combinator does not need to used when another combinator, such as \cc{optional} or
+The reference combinator does not need to be used when another combinator, such as \cc{optional} or
 \cc{many}, already introduces indirection in the data type. An example is shown in Figure~\ref{fig:grammar-arith}.

### 2014-09-10 paper/osdi14/code/dns-api.c

@@ -1,4 +1,15 @@
-struct dnspacket*parse_dnspacket(NailArena *arena, 
+struct dnspacket {
+    uint16_t id;
+    uint8_t qr;
+    /*  abbreviated */
+    struct {
+        question*elem;
+        size_t count;
+    } questions;
+}; 
+/* Other structures */
+struct dnspacket*parse_dnspacket(
+  NailArena *arena, 
   const uint8_t *data, 
   size_t size);
 int gen_dnspacket(NailArena *tmp_arena,

### 2014-09-10 paper/osdi14/design.tex
Example for dependent field. Grammar example.

@@
 \paragraph{Semantic bijection.}
+Parsing inputs and generating outputs form a bijection. However,
+a bijection in the traditional sense often does not make sense for data
+formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
+that tolerates arbitrarily long padding. Program semantics should be independent of the number of padding elements in the input, and Nail therefore
+does not expose that information to the programmer. We call such discarded fields \emph{constants}. 
+Similarly, programs should not necessarily preserve the layout of objects referred to by their offsets.
 
-In order to generate output, Nail maintains
-a \emph{semantic bijection} between the external format and the internal model.
+Therefore, Nail  maintains only a  \emph{semantic bijection} between the external format and the internal model.
 That is, when Nail parses an input into an internal representation, and then
@@
 output) will have the same meaning (i.e., be interpreted equivalently by Nail).
-However, the byte streams might not be identical.
-A bijection in the traditional sense often does not make sense for data
-formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
-that tolerates arbitrarily long padding.
+However, the byte streams might not be identical. If the grammar
+consists of a simple protocol without offset fields, constants, and the like, there is a
+conventional bijection between  internal models and valid parser inputs.
 %\footnote{Say, the physical layer of most communication
@@
 %  pre-determined synchronization sequence and the actual contents of the transmission.}
- Program semantics should be independent of the number of padding elements in the input, and Nail therefore
-does not expose that information to the programmer. We call such discarded fields \emph{constants}.
-Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
-consists of a simple protocol without offset fields, constants, and the like, there is a conventional
-bijection between  internal models and valid parser inputs.
 
@@
 and need to be kept as additional state during parsing. Dependent fields are, for example, used to represent
-lengths, offsets, and checksums.
+lengths, offsets, and checksums. If dependent fields were exposed in the internal model, information
+would be duplicated and inconsistent internal data structures could be produced when data is
+modified. For example when using Nail to handle UDP packets, without dependent fields, programmers
+might forget to update checksum fields when they modify the payload data.
 

### 2014-09-10 paper/osdi14/fig-motivation.tex
motivation nits

--- 2014-05-02 paper/osdi14/fig-motivation.tex
@@
 \midrule
-  Memory corruption & 2013-5660 & Buffer overflow & 11\\
-  Parser ambiguities & 2013-1462 & Multiple virus scanners interpret ZIP files incorrectly.& 4\\
-  Semantic misunderstanding & 2014-2319 & Weak cryptography used even if user selects AES & 1\\
-  \multicolumn{3}{l}{\bf Total vulnerabilities related to .zip processing }& \bf 16  \\
+  Memory corruption & CVE-2013-5660 & Buffer overflow & 11\\
+  Parsing inconsistency & CVE-2013-1462 & Multiple virus scanners interpret ZIP files incorrectly & 4\\
+  Semantic misunderstanding & CVE-2014-2319 & Weak cryptography used even if user selects AES & 1\\
+\midrule
+  \multicolumn{3}{l}{\bf Total of all vulnerabilities related to .zip processing }& \bf 16  \\
 \bottomrule
@@
 \end{tabular}
-\caption{Classification of known vulnerabilities in the CVE database
-between 2010 and 2014 related to the search term ``ZIP'' and involving
+\caption{Classification of vulnerabilities in the CVE database
+from 2010 to May 2014 containing the term ``ZIP'' and involving
 the ZIP file format.}

### 2014-09-10 paper/osdi14/motivation.tex
motivation nits

@@
 
-Three recent high profile security vulnerabilities are due to logic
+Three recent high-profile security vulnerabilities are due to logic
 errors in input processing.  In all cases, when the vulnerabilities were
@@
 user-mode code-signing verifier interpreting executable metadata differently, so the code signature
-sees different bytes at a virtual address than kernel maps into the process. The next version of iOS
+checker sees different bytes at a virtual address than what the kernel maps into the process. The next version of iOS
 added an explicit check for this particular metadata inconsistency. However, because parsing and
@@
 The ZIP format has been associated with many vulnerabilities, and the
-PROTOS Genome project~\cite{protos-c10-archive} found numerous security vulnerabilities in most
-implementations of ZIP and other archive formats that are directly
-related to input handling. We extend this study by looking at the CVE database.
+PROTOS Genome project~\cite{protos-c10-archive} found numerous security vulnerabilities
+related to input handling
+in most implementations of ZIP and other archive formats.
+We extend this study by looking at the CVE database.
 
@@
 
-These input processing vulnerabilities fall into two broad classes.  The first
-class, which occurred 11 times\footnote{We classified the following vulnerabilities as memory corruption
-    attacks based on their description: 2013-5660, - 0742, -0138, 2012-4987, -1163,-1162, 2011-2265,
-    2010-4535, -1657, -1336, -1218}, is memory safety bugs, such as buffer overflows, which allow an
+These input-processing vulnerabilities fall into two broad classes.  The first
+class, which occurred 11 times,\footnote{We classified the following vulnerabilities as memory corruption
+    attacks based on their description: CVE-2013-5660, -0742, -0138, CVE-2012-4987, -1163, -1162, CVE-2011-2265,
+    CVE-2010-4535, -1657, -1336, and -1218.} is memory safety bugs, such as buffer overflows, which allow an
 adversary to corrupt the application's memory using specially crafted

### 2014-09-10 paper/osdi14/design.tex
design (section 4.0) nits

@@
 
-\item Using the grammar to define both the \emph{external format} and the \emph{internal
-    representation}, allowing the same grammar to be re-used in multiple programs, avoiding
-  vulnerabilities like the Android Master Key.
+\item Using a single grammar to define both the \emph{external format} and the
+  \emph{internal representation}.  This allows the same grammar to be re-used
+  in multiple programs, and helps avoid vulnerabilities like the Android Master
+  Key bug.
 
-\item Both parsing inputs into internal representations, as well as
+\item Parsing inputs into internal representations, as well as
       generating outputs from internal representations, without
       requiring the programmer to write any semantic actions. This prevents
-      vulnerabilities such as the XNU bug, where format recognition and semantics
+      vulnerabilities such as the iOS XNU bug, where format recognition and semantics
       are mixed and interact in unexpected ways.
@@
       of a container data structure, to provide programmers
-      a consistent, unambiguous view of the data,
-      to avoid bugs such as in the Python ZIP library.
+      a consistent, unambiguous view of the data.
+      This helps avoid bugs such as the one discovered in the
+      Python ZIP library~\cite{python-bug:20078}.
 
@@
       data formats through well-defined extensibility mechanisms. 
+      This helps prevent programmers from falling back on manual
+      parsing when encountering a complex data format.
 

### 2014-09-10 paper/osdi14/code/dns-api.c
tweak DNS grammar figure

@@ -1,17 +1,7 @@
-struct dnspacket {
-    uint16_t id;
-    uint8_t qr;
-    /*  abbreviated */
-    struct {
-        question*elem;
-        size_t count;
-    } questions;
-}; 
-/* Other structures */
-struct dnspacket*parse_dnspacket(
-  NailArena *arena, 
-  const uint8_t *data, 
+struct dnspacket *parse_dnspacket(NailArena *arena,
+  const uint8_t *data,
   size_t size);
+
 int gen_dnspacket(NailArena *tmp_arena,
   NailStream *out,
-  struct dnspacket * val) ;
+  struct dnspacket *val);

### 2014-09-10 paper/osdi14/design.tex
some minor typesetting
4.1 nits

@@
 Nail grammars describe both the \emph{external format} and an
-\emph{internal representation} of a protocol, as opposed to existing
-grammar languages, which describe only the former and leave it up to
-the programmer to construct an internal representation, typically in
-the form of an abstract syntax tree.
+\emph{internal representation} of a protocol.
 Nail produces the following from a single, descriptive grammar:
 
-\begin{CompactItemize}
-\item \textit{type declarations} for the internal model,
-\item the \textit{parser}, a function to parse a sequence of bytes into an
-instance of the model, and
-\item the \textit{generator}, a function to create a
-sequence of bytes from an instance of the model.
-\end{CompactItemize}
+\begin{itemize}
+\item \textit{Type declarations} for the internal model, which
+the application should use to represent data items in memory.
 
-For example, Figure~\ref{fig:dns-full} shows an example grammar for DNS packets and the
-corresponding data type and functions that Nail produces.
+\item The \textit{parser}, a function that the application
+should invoke to parse a sequence of bytes into an instance
+of the above model.
+
+\item The \textit{generator}, a function that the application
+should invoke to create a sequence of bytes from an instance
+of the model.
+\end{itemize}
+
+For example, Figure~\ref{fig:dns-full} shows a Nail grammar for DNS
+packets.  For this grammar, Nail produces the type declarations
+shown in Figure~\ref{fig:dns-full-struct}, and the parser and
+generator functions shown in Figure~\ref{fig:dns-full-api}.
 
@@
 \paragraph{Semantic bijection.}
-Parsing inputs and generating outputs form a bijection. However,
+Parsing inputs and generating outputs suggests a bijection
+between external data and its internal representation. However,
 a bijection in the traditional sense often does not make sense for data
@@
 
-Therefore, Nail  maintains only a  \emph{semantic bijection} between the external format and the internal model.
+Therefore, Nail  establishes only a  \emph{semantic bijection} between the external format and the internal model.
 That is, when Nail parses an input into an internal representation, and then
@@
 
-\paragraph{Hide redundant information.}
+\paragraph{Hiding redundant information.}
 Nail's internal model is designed to hide unneeded and redundant information from the application.
@@
 would be duplicated and inconsistent internal data structures could be produced when data is
-modified. For example when using Nail to handle UDP packets, without dependent fields, programmers
+modified. For example, when using Nail to handle UDP packets, without dependent fields, programmers
 might forget to update checksum fields when they modify the payload data.
@@
 generator. Nail parsers and generators interact with data through an abstract stream, which allows
-reading and writing of bits as re-positioning. Transformations allow the programmer to write
-functions in a general purpose language that consume streams and define new temporary streams, while
+reading and writing of bits and re-positioning. Transformations allow the programmer to write
+functions in a general-purpose language that consume streams and define new temporary streams, while
 also reading or writing the values of dependent fields.
@@
 for example, some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
-end of the file or a fixed offset. In nested grammars, offsets are also not necessarily computed from
+end of the file or at a fixed offset. In nested grammars, offsets are also not necessarily computed from
 the beginning of a file or packet.
@@
 \begin{figure}
+\smaller[0.5]
 \begin{verbatim}

### 2014-09-10 paper/osdi14/design.tex
4.2 nits

@@
 
+\input{fig-syntax}
+
 Initial versions of Nail's design included a special combinator for handling offset fields,
@@
 generated Nail parser.
-
-\input{fig-syntax}
 
@@
 
-\paragraph{Parsing.}
 \paragraph{Rules.}
@@
 Note that is possible to define parsers for sub-byte lengths, for example, the flag bits in the DNS
-message header, in lines 4 through 8.
+message header, in lines 5 through 8.
 
@@
 
-The \emph{many} combinator takes a parser and applies it repeatedly
+The \cc{many} combinator takes a parser and applies it repeatedly
 until it fails, returning an array of the inner parser's results. In line 39 of the DNS grammar, a
@@
 field is encountered.
-The \emph{sepBy} combinator
+The \cc{sepBy} combinator
 additionally takes a constant parser, which it applies in between parsing
@@
 For example, the grammar for  Ethernet headers uses \cc{optional vlan\_header} to parse the VLAN
-header that only appears in Ethernet packets transmitted to a non-default VLAN\@.
+header that appears only in Ethernet packets transmitted to a non-default VLAN\@.
 
 \paragraph{References.}
-Rules allow for recursive grammars. To support recursive data types, we allow introduce the
+Rules allow for recursive grammars. To support recursive data types, we introduce the
 reference combinator \cc{*}  that does not change the syntax of the external format described, but

### 2014-09-10 paper/osdi14/design.tex
4.3 nits

@@
 the \cc{many} combinator, except it represents an exact number, specified in the dependent field, of
-repetitions, as opposed as many repetitions as possible. For example, DNS labels, which
+repetitions, as opposed to as many repetitions as possible. For example, DNS labels, which
 are encoded as a length followed by
-a value are described in line 38 of the DNS grammar.
+a value, are described in line 38 of the DNS grammar.
 Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but

### 2014-09-10 paper/osdi14/design.tex
done with pass over design

@@
 
-Transformations need to be carefully written, because they can escape Nail's security and introduce
+Transformations need to be carefully written, because they can violate Nail's safety properties
+and introduce
 bugs. However, as we will show in \S\ref{s:eval-effort}, Nail transformations are much shorter than
-hand-written parsers and many formats can be represented with just the transformations in the
+hand-written parsers, and many formats can be represented with just the transformations in Nail's
 standard library.

### 2014-09-11 paper/osdi14/concl.tex
eval and concl nits

@@
 Nail achieves this by reducing the expressive power of the grammar,
-maintaining a \emph{semantic bijection} between data formats and internal
+establishing a \emph{semantic bijection} between data formats and internal
 representations.  Nail captures complex data formats by introducing
@@
 Nail and all of the applications and grammars developed in this paper
-are released as open-source software.
-
-% The source code for our prototype of Nail is available at
-% \url{https://github.com/jbangert/nail}.
+are released as open-source software, available at
+\url{https://github.com/jbangert/nail}.

### 2014-09-11 paper/osdi14/design.tex
design

--- 2014-09-10 paper/osdi14/design.tex
@@
 Nail grammars describe both the \emph{external format} and an
-\emph{internal representation} of a protocol.
+\emph{internal representation} of a protocol, as opposed to existing
+grammar languages, which describe only the former and leave it up to
+the programmer to construct an internal representation, typically in
+the form of an abstract syntax tree.
 Nail produces the following from a single, descriptive grammar:
 
-\begin{itemize}
-\item \textit{Type declarations} for the internal model, which
-the application should use to represent data items in memory.
+\begin{CompactItemize}
+\item \textit{type declarations} for the internal model,
+\item the \textit{parser}, a function to parse a sequence of bytes into an
+instance of the model, and
+\item the \textit{generator}, a function to create a
+sequence of bytes from an instance of the model.
+\end{CompactItemize}
 
-\item The \textit{parser}, a function that the application
-should invoke to parse a sequence of bytes into an instance
-of the above model.
-
-\item The \textit{generator}, a function that the application
-should invoke to create a sequence of bytes from an instance
-of the model.
-\end{itemize}
-
-For example, Figure~\ref{fig:dns-full} shows a Nail grammar for DNS
-packets.  For this grammar, Nail produces the type declarations
-shown in Figure~\ref{fig:dns-full-struct}, and the parser and
-generator functions shown in Figure~\ref{fig:dns-full-api}.
+For example, Figure~\ref{fig:dns-full} shows an example grammar for DNS packets. Nail produces two
+functions shown in Figure~\ref{fig:dns-full-api} that parse a DNS packet into the structures
+partially shown in
+Figure~\ref{fig:dns-full-struct} and vice versa.
 
@@
 \paragraph{Semantic bijection.}
-Parsing inputs and generating outputs suggests a bijection
-between external data and its internal representation. However,
+Parsing inputs and generating outputs form a bijection. However,
 a bijection in the traditional sense often does not make sense for data
@@
 
-Therefore, Nail  establishes only a  \emph{semantic bijection} between the external format and the internal model.
+Therefore, Nail  maintains only a  \emph{semantic bijection} between the external format and the internal model.
 That is, when Nail parses an input into an internal representation, and then
@@
 
-\paragraph{Hiding redundant information.}
+\paragraph{Hide redundant information.}
 Nail's internal model is designed to hide unneeded and redundant information from the application.
@@
 would be duplicated and inconsistent internal data structures could be produced when data is
-modified. For example, when using Nail to handle UDP packets, without dependent fields, programmers
+modified. For example when using Nail to handle UDP packets, without dependent fields, programmers
 might forget to update checksum fields when they modify the payload data.
@@
 generator. Nail parsers and generators interact with data through an abstract stream, which allows
-reading and writing of bits and re-positioning. Transformations allow the programmer to write
-functions in a general-purpose language that consume streams and define new temporary streams, while
+reading and writing of bits as re-positioning. Transformations allow the programmer to write
+functions in a general purpose language that consume streams and define new temporary streams, while
 also reading or writing the values of dependent fields.
-
-\input{fig-syntax}
 
@@
 for example, some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
-end of the file or at a fixed offset. In nested grammars, offsets are also not necessarily computed from
+end of the file or a fixed offset. In nested grammars, offsets are also not necessarily computed from
 the beginning of a file or packet.
@@
 generated Nail parser.
+
+\input{fig-syntax}
 
@@
 
+\paragraph{Parsing.}
 \paragraph{Rules.}
@@
 Note that is possible to define parsers for sub-byte lengths, for example, the flag bits in the DNS
-message header, in lines 5 through 8.
+message header, in lines 4 through 8.
 
@@
 
-The \cc{many} combinator takes a parser and applies it repeatedly
+The \emph{many} combinator takes a parser and applies it repeatedly
 until it fails, returning an array of the inner parser's results. In line 39 of the DNS grammar, a
@@
 field is encountered.
-The \cc{sepBy} combinator
+The \emph{sepBy} combinator
 additionally takes a constant parser, which it applies in between parsing
@@
 For example, the grammar for  Ethernet headers uses \cc{optional vlan\_header} to parse the VLAN
-header that appears only in Ethernet packets transmitted to a non-default VLAN\@.
+header that only appears in Ethernet packets transmitted to a non-default VLAN\@.
 
 \paragraph{References.}
-Rules allow for recursive grammars. To support recursive data types, we introduce the
+Rules allow for recursive grammars. To support recursive data types, we allow introduce the
 reference combinator \cc{*}  that does not change the syntax of the external format described, but

### 2014-09-17 paper/osdi14/code/zip.nail.tex
Poster first shot

\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
\PY{n}{zip\PYZus{}file} \PY{o}{=} \PY{p}{\PYZob{}} 
 \PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{,} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{transform} \PY{n}{zip\PYZus{}end\PYZus{}of\PYZus{}directory} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current}\PY{p}{)}
 \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
\PY{p}{\PYZcb{}}
\PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}} 
 \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x06054b50}
 \PY{n}{disks} \PY{n}{uint16} \PY{o}{|} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
 \PY{n}{directory\PYZus{}disk} \PY{n}{uint16} \PY{o}{|} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
 \PY{err}{@}\PY{n}{this\PYZus{}records} \PY{n}{uint16}
 \PY{err}{@}\PY{n}{total\PYZus{}records} \PY{n}{uint16} 
 \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{this\PYZus{}records} \PY{err}{@}\PY{n}{total\PYZus{}records}\PY{p}{)}
 \PY{err}{@}\PY{n}{directory\PYZus{}size} \PY{n}{uint32} 
 \PY{err}{@}\PY{n}{directory\PYZus{}start} \PY{n}{uint32}
 \PY{err}{\PYZdl{}}\PY{n}{dirstr1} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{directory\PYZus{}start}\PY{p}{)} 
 \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{transform} \PY{n}{size\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{dirstr1} \PY{err}{@}\PY{n}{directory\PYZus{}size}\PY{p}{)}
 \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint16}
 \PY{n}{comment} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint8}
 \PY{n}{files} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{total\PYZus{}directory\PYZus{}records} \PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
\PY{p}{\PYZcb{}}
\PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}}  \PY{c+cm}{/*cut for brevity*/}
 \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x02014b50}\PY{c+cm}{/*...*/}
 \PY{err}{@}\PY{n}{compression\PYZus{}method} \PY{n}{uint16} \PY{c+cm}{/*...*/}
 \PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{c+cm}{/*...*/}
 \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint16}\PY{c+cm}{/*...*/}
 \PY{err}{@}\PY{n}{off} \PY{n}{uint32}
 \PY{n}{filename} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint8}\PY{c+cm}{/*...*/}
 \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{off}\PY{p}{)}
 \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compression\PYZus{}method}\PY{c+cm}{/*...*/}\PY{p}{)}
\PY{p}{\PYZcb{}}
\PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{p}{,}\PY{err}{@}\PY{n}{method} \PY{n}{uint16}\PY{c+cm}{/*...*/}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}}
 \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x04034b50}\PY{c+cm}{/*...*/}
 \PY{err}{@}\PY{n}{method\PYZus{}local} \PY{n}{uint16}\PY{c+cm}{/*...*/}
 \PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{n}{transform} \PY{n}{size\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current} \PY{err}{@}\PY{n}{compressed\PYZus{}size}\PY{p}{)}
 \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{transform} \PY{n}{zip\PYZus{}compression} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{err}{@}\PY{n}{method}\PY{c+cm}{/*...*/}\PY{p}{)}
 \PY{n}{transform} \PY{n}{crc\PYZus{}32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{err}{@}\PY{n}{crc32}\PY{p}{)}
 \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{many} \PY{n}{uint8}

 \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{method\PYZus{}local} \PY{err}{@}\PY{n}{method}\PY{p}{)}\PY{c+cm}{/*...*/}
\PY{p}{\PYZcb{}}
\end{Verbatim}

### 2014-09-17 paper/osdi14/fig-motivation.tex
some minor typesetting
Poster first shot

@@
 \centering
-\begin{tabular}{lllr}
-\toprule
- \bf Classification & \bf Example CVE & \bf Example description & \bf Count\\
-\midrule
-  Memory corruption & CVE-2013-5660 & Buffer overflow & 11\\
-  Parsing inconsistency & CVE-2013-1462 & Multiple virus scanners interpret ZIP files incorrectly & 4\\
-  Semantic misunderstanding & CVE-2014-2319 & Weak cryptography used even if user selects AES & 1\\
-\midrule
-  \multicolumn{3}{l}{\bf Total of all vulnerabilities related to .zip processing }& \bf 16  \\
-\bottomrule
-
-\end{tabular}
+\smaller[0.5]
+\input{table-motivation.tex}
 \caption{Classification of vulnerabilities in the CVE database

### 2014-09-17 paper/osdi14/poster.tex
Poster first shot

\documentclass[25pt, a0paper, portrait, margin=0mm, innermargin=15mm,
blockverticalspace=10mm, colspace=15mm, subcolspace=8mm]{tikzposter}

\usepackage{times}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{color}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{relsize}
\usepackage{fancyvrb}
\usepackage{pgfplots}

\title{Nail: A practical tool for parsing and generating data formats}
\author{Julian Bangert and Nickolai Zeldovich }
\institute{MIT CSAIL}
%\titlegraphic{Logo}
\usetheme{Envelope}

\input{code/fmt}

\begin{document}
\maketitle
\begin{columns}
\column{.5}
\block{Parsing vulnerabilities}{
Current software typically uses hand-implemented parsers, which are not only prone to bugs, but two
separate parser implementations can disagree about the meaning of an input. This has been the source
of many recent high-profile vulnerabilities:
\begin{itemize}
\item Evasi0n jailbreaks on iOS~\cite{evaders6}.
\item PKI layer cake~\cite{Marlinspike:SSL},\cite{DBLP:conf/fc/KaminskyPS10}.
\item Android master key ~\cite{saurik-masterkey}
\end{itemize}

{\bf Case study of ZIP vulnerabilities in the CVE database:}

\begin{tabular}{lllr}
\toprule
 \bf Classification  & \bf Example description & \bf Count\\
\midrule
  Memory corruption  & Buffer overflow & 11\\
  Parsing inconsistency & Virus scanners interpret ZIP files incorrectly & 4\\
  Semantic misunderstanding & Weak cryptography used even if user selects AES & 1\\
\midrule
  \multicolumn{2}{l}{\bf Total of all vulnerabilities related to .zip processing }& \bf 16  \\
\bottomrule

\end{tabular}
}
\block{Solution: Grammar-based Parsers}{   
Grammar-based parsers, such as \texttt{bison} or Hammer~\cite{hammer-parser}, are compact, reuseable
and safer. However, traditional parsers are inconvenient to use, because the programmer has to write all the
yellow-colored parts:

\includegraphics{StateOfTheArt.png}

Nail grammars describe both the format and a data type to represent it.

\includegraphics{Nail.png}

}
\block{Complicated real-world example: ZIP }{
  \includegraphics[width=0.4\columnwidth]{ZipHorizontal.png}
  \begin{itemize}
    \item Variable-length end-of-file header
    \item Length and offset fields
    \item Duplicated information
    \item Checksums, compression and encryption
  \end{itemize}
}
\block{Design Goals}{
  \begin{itemize}
    \item Reduce programmer effort
  \begin{itemize}
      \item Semantic combinators describes data type and format.
      \item Generate output.
  \end{itemize}
     \item Handle complex formats
  \begin{itemize}
       \item Dependency Fields: hide redundant information.
       \item Transforms: Extensibility.
       \end{itemize}
  \end{itemize}
}
\block{Semantic Bijection}{
  Traditional bijections don't make sense for some data formats. For example, offsets are and should
  be discarded during parsing. Nail parsers and generators therefore form a looser 'semantic
  bijection', in which parser(generator(x)) = x, but not necessarily vice versa.

}
\column{.5}
\block{Nail Grammar for DNS}{
\small
\input{code/zip.nail}
}
\block{Dependent Fields}{
Duplicated data and layout information can cause parser consistency bugs. Nail represents them with
Dependent Fields, which are automatically verified during parsing and created during 
generation. 
}
\block{Streams and Transformations}{
  Existing parsers are linear, consuming input front to back; Offset fields require kludges.
  Nail grammars feature multiple streams, and introduce \emph{Transformations} to create
  new streams. Transformations are a pair of functions operating on streams and dependent fields.
  A standard library of Transformations is provided, but for complex formats, programmers can write
  their own Transformations, for which they have to guarantee security and semantic bijection.
 For example, in the ZIP grammar above, the programmer has to write two functions to find and
 generate the end-of-file header, with the following prototype:
\input{code/dns-xform}
}
\block{Related Work and Bibliography}{
  \begin{itemize}
    \item Interface generators such XDR~\cite{RFC:1832} and ProtoBuf~\cite{varda2008} can parse and
      generate, but cannot interoperate with legacy formats.
    \item Data Description Languages, such as PacketTypes~\cite{mccann2000packet},
      DataScript~\cite{back2002datascript} and PADS~\cite{Fisher:2005:PDL:1064978.1065046}, can
      parse and generate, but are not extensible, so they cannot support complicated encodings in
      formats like ZIP.
    \item See our paper for a detailed discussion, including application use of parsers.
  \end{itemize}
\begin{minipage}{\textwidth}
\tiny \bibliography{n-str,main,locasto,langsec,n,n-conf}
\bibliographystyle{alpha} 
\end{minipage}
}

\end{columns}
\end{document}

### 2014-09-17 paper/osdi14/code/zip.nail.tex
Poster

@@
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
-\PY{n}{zip\PYZus{}file} \PY{o}{=} \PY{p}{\PYZob{}} 
- \PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{,} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{transform} \PY{n}{zip\PYZus{}end\PYZus{}of\PYZus{}directory} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current}\PY{p}{)}
+\PY{n}{zip} \PY{o}{=} \PY{p}{\PYZob{}} 
+ \PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{,} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{transform} \PY{n}{zip\PYZus{}eod} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current}\PY{p}{)}
  \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
@@
  \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x02014b50}\PY{c+cm}{/*...*/}
- \PY{err}{@}\PY{n}{compression\PYZus{}method} \PY{n}{uint16} \PY{c+cm}{/*...*/}
+ \PY{err}{@}\PY{n}{compressed\PYZus{}size} \PY{n}{uint16} \PY{c+cm}{/*...*/}
  \PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{c+cm}{/*...*/}
@@
  \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{off}\PY{p}{)}
- \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compression\PYZus{}method}\PY{c+cm}{/*...*/}\PY{p}{)}
+ \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compressed\PYZus{}size} \PY{c+cm}{/*...*/}\PY{p}{)}
 \PY{p}{\PYZcb{}}
-\PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{p}{,}\PY{err}{@}\PY{n}{method} \PY{n}{uint16}\PY{c+cm}{/*...*/}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}}
+\PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{p}{,}\PY{err}{@}\PY{n}{size} \PY{n}{uint32}\PY{c+cm}{/*...*/}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}}
  \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x04034b50}\PY{c+cm}{/*...*/}
- \PY{err}{@}\PY{n}{method\PYZus{}local} \PY{n}{uint16}\PY{c+cm}{/*...*/}
- \PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{n}{transform} \PY{n}{size\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current} \PY{err}{@}\PY{n}{compressed\PYZus{}size}\PY{p}{)}
- \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{transform} \PY{n}{zip\PYZus{}compression} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{err}{@}\PY{n}{method}\PY{c+cm}{/*...*/}\PY{p}{)}
+ \PY{err}{@}\PY{n}{crc\PYZus{}local} \PY{n}{uint32}\PY{c+cm}{/*...*/}
+ \PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{n}{transform} \PY{n}{size\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current} \PY{err}{@}\PY{n}{size}\PY{p}{)}
+ \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{transform} \PY{n}{zip\PYZus{}compression} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{err}{@}\PY{n}{size}\PY{c+cm}{/*...*/}\PY{p}{)}
  \PY{n}{transform} \PY{n}{crc\PYZus{}32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{err}{@}\PY{n}{crc32}\PY{p}{)}
@@
 
- \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{method\PYZus{}local} \PY{err}{@}\PY{n}{method}\PY{p}{)}\PY{c+cm}{/*...*/}
+ \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{crc\PYZus{}local} \PY{err}{@}\PY{n}{crc32}\PY{p}{)}\PY{c+cm}{/*...*/}
 \PY{p}{\PYZcb{}}

### 2014-09-17 paper/osdi14/poster.tex
Poster

@@
 \usepackage{pgfplots}
-
+\usepackage{adjustbox}
 \title{Nail: A practical tool for parsing and generating data formats}
@@
 \begin{itemize}
-\item Evasi0n jailbreaks on iOS~\cite{evaders6}.
-\item PKI layer cake~\cite{Marlinspike:SSL},\cite{DBLP:conf/fc/KaminskyPS10}.
-\item Android master key ~\cite{saurik-masterkey}
+\item Evasi0n jailbreaks on iOS.
+\item PKI layer cake.
+\item Android master key.
 \end{itemize}
@@
 }
-\block{Solution: Grammar-based Parsers}{   
-Grammar-based parsers, such as \texttt{bison} or Hammer~\cite{hammer-parser}, are compact, reuseable
+
+\block{State of the art}{   
+Grammar-based parsers, such as \texttt{bison} or Hammer, are compact, reuseable
 and safer. However, traditional parsers are inconvenient to use, because the programmer has to write all the
@@
 
-Nail grammars describe both the format and a data type to represent it.
+Nail grammars describe both the format and a data type to represent it, an idea introduced by data
+description languages such as PacketTypes.
 
@@
 
-}
-\block{Complicated real-world example: ZIP }{
+However, existing solutions cannot handle complicated file formats like ZIP, which feature
+duplicated data, offsets and format-specific complications, such as a variable-length end-of-file header:
+
   \includegraphics[width=0.4\columnwidth]{ZipHorizontal.png}
-  \begin{itemize}
-    \item Variable-length end-of-file header
-    \item Length and offset fields
-    \item Duplicated information
-    \item Checksums, compression and encryption
-  \end{itemize}
 }
@@
       \item Semantic combinators describes data type and format.
-      \item Generate output.
+      \item Generate output. Semantic bijection, preserves meaning, but can change representation.
   \end{itemize}
@@
   \begin{itemize}
-       \item Dependency Fields: hide redundant information.
-       \item Transforms: Extensibility.
+       \item Dependency Fields: hide redundant and structural information.
+       \item Transforms: Extensibility mechanism to capture many features without complicating the
+         grammar language.
        \end{itemize}
   \end{itemize}
-}
-\block{Semantic Bijection}{
-  Traditional bijections don't make sense for some data formats. For example, offsets are and should
-  be discarded during parsing. Nail parsers and generators therefore form a looser 'semantic
-  bijection', in which parser(generator(x)) = x, but not necessarily vice versa.
-
-}
-\column{.5}
-\block{Nail Grammar for DNS}{
-\small
-\input{code/zip.nail}
 }
@@
 \block{Streams and Transformations}{
-  Existing parsers are linear, consuming input front to back; Offset fields require kludges.
-  Nail grammars feature multiple streams, and introduce \emph{Transformations} to create
-  new streams. Transformations are a pair of functions operating on streams and dependent fields.
-  A standard library of Transformations is provided, but for complex formats, programmers can write
-  their own Transformations, for which they have to guarantee security and semantic bijection.
- For example, in the ZIP grammar above, the programmer has to write two functions to find and
- generate the end-of-file header, with the following prototype:
-\input{code/dns-xform}
+  Existing parsers are linear, consuming input front to back. Complicated formats require inelegant hacks. 
+  Nail grammars feature multiple streams and introduce \emph{Transformations} to create
+  them. Transformations are a pair of functions operating on streams and dependent fields.
+  A standard library of Transformations is provided, but  programmers can extend
+  Nail with custom transformations, which are security critical.
+
+%  For example, in the ZIP grammar above, the programmer has to write two functions to find and
+%  generate the end-of-file header, with the following prototype:
+% \input{code/dns-xform}
 }
-\block{Related Work and Bibliography}{
-  \begin{itemize}
-    \item Interface generators such XDR~\cite{RFC:1832} and ProtoBuf~\cite{varda2008} can parse and
-      generate, but cannot interoperate with legacy formats.
-    \item Data Description Languages, such as PacketTypes~\cite{mccann2000packet},
-      DataScript~\cite{back2002datascript} and PADS~\cite{Fisher:2005:PDL:1064978.1065046}, can
-      parse and generate, but are not extensible, so they cannot support complicated encodings in
-      formats like ZIP.
-    \item See our paper for a detailed discussion, including application use of parsers.
-  \end{itemize}
-\begin{minipage}{\textwidth}
-\tiny \bibliography{n-str,main,locasto,langsec,n,n-conf}
-\bibliographystyle{alpha} 
+% \block{Semantic Bijection}{
+%   Traditional bijections don't make sense for some data formats. For example, offsets are and should
+%   be discarded during parsing. Nail parsers and generators therefore form a looser 'semantic
+%   bijection', in which parser(generator(x)) = x, but not necessarily vice versa.
+
+% }
+\column{.5}
+\block{Nail Grammar for ZIP}{
+{\small
+\input{code/zip.nail}}
+An excerpt of Nail's output for this example:
+\input{code/zip-out}
+}
+\block{Evaluation}{
+\begin{tabular}{p{0.43\linewidth}p{0.57\linewidth}}
+\begin{minipage}[t]{0.43\linewidth}
+\vspace{0pt}\begin{tabular}{lrr@{~}l}
+\toprule
+\multicolumn{4}{l}{\textbf{Comparing total lines of code}}\\
+
+\midrule
+\textbf{Application}
+  & \textbf{Nail}
+  & \multicolumn{2}{c}{\textbf{Alternative}} \\
+\midrule
+DNS server
+  & 295
+  & 683
+  & (Hammer) \\
+
+%  &
+%  & 10,000
+%  & (DJBDNS) \\
+
+\texttt{unzip}
+  & 220
+  & 1,600
+  & (Info-Zip) \\
+\bottomrule
+\end{tabular}
+\end{minipage}&
+\vspace{0pt}\begin{minipage}{0.57\linewidth}
+
+\begin{tabular}[t]{lrl}
+\toprule
+\textbf{Protocol} & \textbf{LoC} & \textbf{Challenging features} \\ 
+\midrule
+DNS packets & 48+64 & Label compression,\\
+  & & count fields \\
+ZIP archives & 92+78 & Checksums, offsets, \\ 
+  & & variable length trailer, \\
+  & & compression \\
+Ethernet  & 16+0\phantom{0} & --- \\
+ARP       & 10+0\phantom{0} & --- \\
+IP        & 25+0\phantom{0} & Total length field, options \\
+UDP       &  7+0\phantom{0} & Checksum, length field \\
+ICMP      &  5+0\phantom{0} & Checksum \\
+\bottomrule
+\end{tabular}
 \end{minipage}
+\end{tabular}
+
+Nail and our example applications are available at \url{https://github.com/jbangert/nail}.
 }
@@
 \end{columns}
+
 \end{document}

### 2014-09-17 paper/osdi14/code/zip-out.c

@@ -1,0 +1,28 @@
+struct zip{
+  end_of_directory contents;
+};
+struct end_of_directory{
+  uint16_t disks;
+  uint16_t directory_disk;/*...*/
+  struct {
+    dir_fileheader* elem;
+    size_t count;
+  } files;
+};
+struct dir_fileheader{
+  struct {
+    uint8_t*elem;
+    size_t count;
+  } filename;
+  fileentry contents;
+};/*..*/
+//The programmer calls these generated functions.
+int gen_zip(NailArena *tmp_arena,NailStream *out,zip * val);
+zip*parse_zip(NailArena *arena, 
+              const uint8_t *data, size_t size);
+//The programmers implements these two transformation
+//functions and two similar ones for output.
+extern int zip_eod_parse(NailArena *tmp,
+ NailStream *filestream, NailStream *,NailStream *current);
+extern int zip_compression_parse(NailArena *tmp,
+ NailStream *uncomp,NailStream *compressed,uint32_t* size);

### 2014-09-17 paper/osdi14/code/zip.nail.c

@@ -1,0 +1,39 @@
+zip = { 
+ $filestream, $end_directory transform zip_eod ($current)
+ contents apply $end_directory end_of_directory($filestream)
+}
+end_of_directory($filestream) = { 
+ uint32 = 0x06054b50
+ disks uint16 | [0]
+ directory_disk uint16 | [0]
+ @this_records uint16
+ @total_records uint16 
+ transform uint16_depend (@this_records @total_records)
+ @directory_size uint32 
+ @directory_start uint32
+ $dirstr1 transform offset_u32 ($filestream @directory_start) 
+ $directory_stream transform size_u32 ($dirstr1 @directory_size)
+ @comment_length uint16
+ comment n_of @comment_length uint8
+ files apply $directory_stream n_of @total_directory_records dir_fileheader($filestream)
+}
+dir_fileheader($filestream) = {  /*cut for brevity*/
+ uint32 = 0x02014b50/*...*/
+ @compressed_size uint16 /*...*/
+ @crc32 uint32/*...*/
+ @file_name_len uint16/*...*/
+ @off uint32
+ filename n_of @file_name_len uint8/*...*/
+ $contentstream transform offset_u32 ($filestream @off)
+ contents apply $contentstream fileentry(@crc32,@compressed_size /*...*/)
+}
+fileentry(@crc32 uint32,@size uint32/*...*/) = {
+ uint32 = 0x04034b50/*...*/
+ @crc_local uint32/*...*/
+ $compressed transform size_u32 ($current @size)
+ $uncompressed transform zip_compression ($compressed @size/*...*/)
+ transform crc_32 ($uncompressed @crc32)
+ contents apply $uncompressed many uint8
+ 
+ transform uint16_depend (@crc_local @crc32)/*...*/
+}

### 2014-09-26 paper/osdi14/code/zip.nail.tex
Fixed margin

@@
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
-\PY{n}{zip} \PY{o}{=} \PY{p}{\PYZob{}} 
+\PY{n}{zip} \PY{o}{=} \PY{p}{\PYZob{}}  \PY{c+cm}{/*cut for brevity*/}
  \PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{,} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{transform} \PY{n}{zip\PYZus{}eod} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current}\PY{p}{)}
@@
      \PY{n}{comment} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint8}
- \PY{n}{files} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{total\PYZus{}directory\PYZus{}records} \PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
+     \PY{n}{files} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{total\PYZus{}directory\PYZus{}records} 
+                 \PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
 \PY{p}{\PYZcb{}}
-\PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}}  \PY{c+cm}{/*cut for brevity*/}
+\PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}} 
      \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x02014b50}\PY{c+cm}{/*...*/}
@@
      \PY{n}{filename} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint8}\PY{c+cm}{/*...*/}
- \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{off}\PY{p}{)}
- \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{contentstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compressed\PYZus{}size} \PY{c+cm}{/*...*/}\PY{p}{)}
+     \PY{err}{\PYZdl{}}\PY{n}{cstream} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{off}\PY{p}{)}
+     \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{cstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compressed\PYZus{}size}\PY{c+cm}{/*...*/}\PY{p}{)}
 \PY{p}{\PYZcb{}}
@@
      \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{many} \PY{n}{uint8}
-
      \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{crc\PYZus{}local} \PY{err}{@}\PY{n}{crc32}\PY{p}{)}\PY{c+cm}{/*...*/}

### 2014-09-26 paper/osdi14/poster.tex
Fixed margin

@@
 However, existing solutions cannot handle complicated file formats like ZIP, which feature
-duplicated data, offsets and format-specific complications, such as a variable-length end-of-file header:
+duplicated data, offsets and format-specific complications, such as a variable-length end-of-file
+header:
 
-  \includegraphics[width=0.4\columnwidth]{ZipHorizontal.png}
+\begin{centering}
+  \includegraphics[width=0.35\columnwidth]{ZipHorizontal.png}
+\end{centering}
 }
@@
 \block{Dependent Fields}{
-Duplicated data and layout information can cause parser consistency bugs. Nail represents them with
-Dependent Fields, which are automatically verified during parsing and created during 
-generation. 
+Duplicated data and layout information can confuse the application, for
+example if a length field is not correct or multiple copies of the same data are inconsistent. 
+Nail grammars can use Dependent Fields, which are automatically verified during parsing and created during 
+generation, to hide such information from the application.
 }
@@
   A standard library of Transformations is provided, but  programmers can extend
-  Nail with custom transformations, which are security critical.
-
+  Nail with custom transformations, which contain arbitrary code and should be carefully written.
 %  For example, in the ZIP grammar above, the programmer has to write two functions to find and
@@
 \vspace{0pt}\begin{tabular}{lrr@{~}l}
+%
+\multicolumn{4}{l}{\textrm{Comparing total lines of code}}\\
 \toprule
-\multicolumn{4}{l}{\textbf{Comparing total lines of code}}\\
-
-\midrule
+%\midrule
 \textbf{Application}
@@
 \vspace{0pt}\begin{minipage}{0.57\linewidth}
-
 \begin{tabular}[t]{lrl}
+\multicolumn{3}{l}{LoC for Nail grammar+transformations}\\
 \toprule

### 2014-09-26 paper/osdi14/code/zip.nail.c

@@ -1,39 +1,39 @@
-zip = { 
+zip = {  /*cut for brevity*/
  $filestream, $end_directory transform zip_eod ($current)
  contents apply $end_directory end_of_directory($filestream)
 }
 end_of_directory($filestream) = { 
- uint32 = 0x06054b50
- disks uint16 | [0]
- directory_disk uint16 | [0]
- @this_records uint16
- @total_records uint16 
- transform uint16_depend (@this_records @total_records)
- @directory_size uint32 
- @directory_start uint32
- $dirstr1 transform offset_u32 ($filestream @directory_start) 
- $directory_stream transform size_u32 ($dirstr1 @directory_size)
- @comment_length uint16
- comment n_of @comment_length uint8
- files apply $directory_stream n_of @total_directory_records dir_fileheader($filestream)
+     uint32 = 0x06054b50
+     disks uint16 | [0]
+     directory_disk uint16 | [0]
+     @this_records uint16
+     @total_records uint16 
+     transform uint16_depend (@this_records @total_records)
+     @directory_size uint32 
+     @directory_start uint32
+     $dirstr1 transform offset_u32 ($filestream @directory_start) 
+     $directory_stream transform size_u32 ($dirstr1 @directory_size)
+     @comment_length uint16
+     comment n_of @comment_length uint8
+     files apply $directory_stream n_of @total_directory_records 
+                 dir_fileheader($filestream)
 }
-dir_fileheader($filestream) = {  /*cut for brevity*/
- uint32 = 0x02014b50/*...*/
- @compressed_size uint16 /*...*/
- @crc32 uint32/*...*/
- @file_name_len uint16/*...*/
- @off uint32
- filename n_of @file_name_len uint8/*...*/
- $contentstream transform offset_u32 ($filestream @off)
- contents apply $contentstream fileentry(@crc32,@compressed_size /*...*/)
+dir_fileheader($filestream) = { 
+     uint32 = 0x02014b50/*...*/
+     @compressed_size uint16 /*...*/
+     @crc32 uint32/*...*/
+     @file_name_len uint16/*...*/
+     @off uint32
+     filename n_of @file_name_len uint8/*...*/
+     $cstream transform offset_u32 ($filestream @off)
+     contents apply $cstream fileentry(@crc32,@compressed_size/*...*/)
 }
 fileentry(@crc32 uint32,@size uint32/*...*/) = {
- uint32 = 0x04034b50/*...*/
[…]

### 2014-11-14 paper/login/abstract.tex
Login first rough draft

--- 2014-08-21 paper/osdi14/abstract.tex
@@
 Nail introduces several key
-ideas to achieve its goal.  First, Nail uses the protocol grammar to
+ideas to achieve its goal.  First, Nail uses a protocol grammar to
 define not just the data format, but also the internal object model
@@
 but allows Nail to both {\em parse} data formats and {\em generate}
-them from the internal object model, by maintaining a {\em semantic
+them from the internal object model, by establishing a {\em semantic
 bijection} between the data format and the object model.  Third,

### 2014-11-14 paper/login/code/dns-api.tex
Login first rough draft

\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
\PY{k}{struct} \PY{n}{dnspacket} \PY{o}{*}\PY{n+nf}{parse\PYZus{}dnspacket}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{arena}\PY{p}{,}
  \PY{k}{const} \PY{k+kt}{uint8\PYZus{}t} \PY{o}{*}\PY{n}{data}\PY{p}{,}
  \PY{k+kt}{size\PYZus{}t} \PY{n}{size}\PY{p}{)}\PY{p}{;}

\PY{k+kt}{int} \PY{n+nf}{gen\PYZus{}dnspacket}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp\PYZus{}arena}\PY{p}{,}
  \PY{n}{NailStream} \PY{o}{*}\PY{n}{out}\PY{p}{,}
  \PY{k}{struct} \PY{n}{dnspacket} \PY{o}{*}\PY{n}{val}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

### 2014-11-14 paper/login/concl.tex
Login first rough draft

--- 2014-09-11 paper/osdi14/concl.tex
@@
 
-This paper presented the design and implementation of \textit{Nail},
+ We presented the design and implementation of \textit{Nail},
 a tool for parsing and generating complex data formats based on a
-precise grammar.  Nail helps programmers avoid
+precise grammar. This helps programmers avoid
 memory corruption and inconsistency vulnerabilities while reducing effort
 in parsing and generating real-world protocols and file formats.
-Nail achieves this by reducing the expressive power of the grammar,
-establishing a \emph{semantic bijection} between data formats and internal
-representations.  Nail captures complex data formats by introducing
+ Nail captures complex data formats by introducing
 \emph{dependent fields}, \emph{streams}, and \emph{transforms}.
@@
 fewer lines of code.
+
 Nail and all of the applications and grammars developed in this paper
 are released as open-source software, available at
-\url{https://github.com/jbangert/nail}.
+\url{https://github.com/jbangert/nail}. A more detailed discussion of our design and our results is
+available in~\cite{bangert:osdi-nail14}.

### 2014-11-14 paper/login/design.tex
Login first rough draft

--- 2014-09-10 paper/osdi14/design.tex
@@
-\section{Design}
+\section{Design By Example}
+In this section, we will explain how to handle basic data formats in Nail, how to handle
+redundancies in the format with dependent fields and how Nail parsers can be extended with
+transformations.
+
+We will use DNS packets as a leading example, as defined in RFC 1035~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
+header, a set of question records, and a set of answer records. Domain
+names in both queries and answers are encoded as a sequence of labels,
+terminated by a zero byte.  Labels are Pascal-style strings, consisting
+of a length field followed by that many bytes comprising the label.
 \label{s:design}
+\subsection{Basic Data formats}
+\input{fig-dnsfull}
+Let's stepping through a simplified Nail grammar for non-compressed
+packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
+shown in Figure~\ref{fig:dns-full-struct}, and the parser and
+generator functions shown in Figure~\ref{fig:dns-full-api}. A more concise reference of the syntax
+can be found in~\ref{fig:syntax}.
+Nail grammars are reuseable between applications and we will use this grammar to implement
+both a DNS server and client, which previously would have had two separate hand-written parsers,
+leading to bugs such as  the Android Master Key.
 
-Nail's goals are to reduce programmer effort required to safely interact with data formats and
-prevent vulnerabilities like those described in \S\ref{s:motivation}. In particular, this means:
+A Nail grammar file consists of rule definitions, for example l. 1-20,  which assign a name
+(\texttt{dnspacket}) to a grammar production (2-20). If you are not familiar with
+other parsers, you can imagine rules as C type declarations on steroids
+(although our syntax is inspired by Go).
 
-\begin{itemize}
+Just like C supports various constructs to build up types, such as structures, unions, pointers and
+elemental types, Nail supports a integers as elemental parsers and a plethora of \emph{combinators}
+to represent features of a file or protocol.
 
-\item Using a single grammar to define both the \emph{external format} and the
-  \emph{internal representation}.  This allows the same grammar to be re-used
-  in multiple programs, and helps avoid vulnerabilities like the Android Master
-  Key bug.
-
-\item Parsing inputs into internal representations, as well as
-      generating outputs from internal representations, without
-      requiring the programmer to write any semantic actions. This prevents
-      vulnerabilities such as the iOS XNU bug, where format recognition and semantics
-      are mixed and interact in unexpected ways.
-
-\item Eliminating redundancy in internal representations, such as
-      storing both an explicit length field and an implicit length
-      of a container data structure, to provide programmers
-      a consistent, unambiguous view of the data.
-      This helps avoid bugs such as the one discovered in the
-      Python ZIP library~\cite{python-bug:20078}.
-
-\item Allow programmers to define grammars for complex real-world
-      data formats through well-defined extensibility mechanisms. 
-      This helps prevent programmers from falling back on manual
-      parsing when encountering a complex data format.
-
-\end{itemize}
-
-\subsection{Overview}
-
-\paragraph{Internal model.}
-
-Nail grammars describe both the \emph{external format} and an
-\emph{internal representation} of a protocol.
-Nail produces the following from a single, descriptive grammar:
-
-\begin{itemize}
-\item \textit{Type declarations} for the internal model, which
-the application should use to represent data items in memory.
-
-\item The \textit{parser}, a function that the application
-should invoke to parse a sequence of bytes into an instance
-of the above model.
-
-\item The \textit{generator}, a function that the application
-should invoke to create a sequence of bytes from an instance
-of the model.
-\end{itemize}
-
-For example, Figure~\ref{fig:dns-full} shows a Nail grammar for DNS
-packets.  For this grammar, Nail produces the type declarations
-shown in Figure~\ref{fig:dns-full-struct}, and the parser and
-generator functions shown in Figure~\ref{fig:dns-full-api}.
-
-\input{fig-dnsfull}
-
-\noindent
-
-\paragraph{Semantic bijection.}
-Parsing inputs and generating outputs suggests a bijection
-between external data and its internal representation. However,
-a bijection in the traditional sense often does not make sense for data
-formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
-that tolerates arbitrarily long padding. Program semantics should be independent of the number of padding elements in the input, and Nail therefore
-does not expose that information to the programmer. We call such discarded fields \emph{constants}. 
-Similarly, programs should not necessarily preserve the layout of objects referred to by their offsets.
-
-Therefore, Nail  establishes only a  \emph{semantic bijection} between the external format and the internal model.
-That is, when Nail parses an input into an internal representation, and then
-generates output from that representation, the two byte streams (input and
-output) will have the same meaning (i.e., be interpreted equivalently by Nail).
-However, the byte streams might not be identical. If the grammar
-consists of a simple protocol without offset fields, constants, and the like, there is a
-conventional bijection between  internal models and valid parser inputs.
-%\footnote{Say, the physical layer of most communication
-%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
-%  pre-determined synchronization sequence and the actual contents of the transmission.}
-
-\paragraph{Hiding redundant information.}
-Nail's internal model is designed to hide unneeded and redundant information from the application.
-Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
-and need to be kept as additional state during parsing. Dependent fields are, for example, used to represent
-lengths, offsets, and checksums. If dependent fields were exposed in the internal model, information
-would be duplicated and inconsistent internal data structures could be produced when data is
-modified. For example, when using Nail to handle UDP packets, without dependent fields, programmers
-might forget to update checksum fields when they modify the payload data.
-
-\paragraph{Parser extensions.}
-Real-world protocols contain complicated ways of encoding data. Fully representing these in an
-intentionally limited model such as our parser language is impractical. Therefore, Nail introduces
-\emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
-generator. Nail parsers and generators interact with data through an abstract stream, which allows
-reading and writing of bits and re-positioning. Transformations allow the programmer to write
-functions in a general-purpose language that consume streams and define new temporary streams, while
-also reading or writing the values of dependent fields.
-
-\input{fig-syntax}
-
-Initial versions of Nail's design included a special combinator for handling offset fields,
-which consumed a dependent field and applied
-a parser at the offset specified therein. However, it proved impossible to foresee all the ways in which a protocol could encode an offset;
-for example, some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
-end of the file or at a fixed offset. In nested grammars, offsets are also not necessarily computed from
-the beginning of a file or packet.
-Nail's transformations allow the programmer to write arbitrary functions that can
-handle such structures and streams, which are a generic abstraction for input and
-output data that allow the decoded data to be integrated with the rest of the
-generated Nail parser.
-
-\subsection{Basics}
-
-A Nail parser defines both the structure of some external format and a data type to represent that
-format. Parsers are constructed by combinators over simpler parsers, an approach popularized by the Parsec
-framework~\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
- combinator libraries, such as Parsec and Hammer~\cite{hammer-parser} and extend them so they also
- describe a data type.
-
-We present both a systematic overview of Nail's syntax with short examples
-in Figure~\ref{fig:syntax}, and explain our design in more detail below,
-using a grammar for the well-known DNS protocol as a running example
-(shown in Figure~\ref{fig:dns-full}).
-
-\paragraph{Rules.}
-A Nail grammar consists of rules that assign a parser to a name. Rules are written as assignments,
-such as \cc{ints = /*parser definition*/}, which defines a rule called \cc{ints}. As we will
-describe later in \S\ref{s:dependent} and \S\ref{s:transforms}, rules can optionally consume parameters.
-Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
-the body of the rule had been substituted in the code. If parameters appear, they are passed by
-reference.
-\paragraph{Integers and constraints.}
-
-Nail's fundamental parsers represent signed or unsigned
-integers with arbitrary lengths up to 64 bits.
-Note that is possible to define parsers for sub-byte lengths, for example, the flag bits in the DNS
-message header, in lines 5 through 8.
-
-The grammar can also constrain the values of an integer.  Nail expresses
-constraints as a set of permissible values or value ranges. Extending the Nail language and
-implementation to support richer constraints languages would be relatively trivial, however we have
-found that the current syntax covers permissible values within existing protocols correctly and
-concisely.
-
-\paragraph{Repetition.}
-
-The \cc{many} combinator takes a parser and applies it repeatedly
-until it fails, returning an array of the inner parser's results. In line 39 of the DNS grammar, a
-sequence of labels is parsed by parsing as many labels as possible, that is, until an invalid length
-field is encountered.
-The \cc{sepBy} combinator
-additionally takes a constant parser, which it applies in between parsing
-two values, but not before parsing the first value or after parsing the
-last.  This is useful for parsing an array of items delimited by a separator.
-%For examle, \cc{many uint8} represents an array of 8-bit unsigned
-%integers, and \cc{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
-%comma-separated lists of decimal numbers.
+\paragraph{Integers and Constraints}.
+Because Nail is designed to cope with binary formats,
+it handles not only common integer types (such as \texttt{uint16} on l. 16), but bit fields of any
+length, such as \texttt{uint1}. These integers are exposed to the programmer as an appropriately
+sized machine integer, e.g. \texttt{uint8\_t}. Nail also supports constraints on integer values,
+limiting the values to either a range (l. 23, \texttt{|1..16}), which can optionally be half-open,
+or a fixed set (l. 24, \texttt{|[1,255]}). Both types of constraint can be combined, e.g. \texttt{|
+  [1..16,255]}. Constant values are also supported, e.g. l. 10: \texttt{uint3=0} represents
+three reserved bits that must be 0. Because constant values carry no information, they are not
+represented in the data type.
 
@@
 
-Nail provides a structure combinator with semantic labels instead of the sequence combinator that
-other parser combinator libraries use to capture structures in data formats. 
-The structure combinator consists of a sequence of fields, typically consisting of a label and a
-parser that describes the contents of that field, surrounded by curly braces. Other field types will be described below.
- The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field
- names preceding their definition.
-
-% A sample grammar for a DNS header and the corresponding C data type the Nail parser produces, are
-% shown in Figure~\ref{fig:dns-struct}.
-
-\paragraph{Constants.}
-In some cases, not all bytes in a structure actually contain information, such as magic numbers or
-reserved fields. Those fields can be represented in Nail grammars by constant fields in structures.
-Constant fields do not correspond to a field in the internal model, but they are validated during
-parsing and generated during output.  Constants can either have integer values, such as
-in line 9 of the DNS grammar, or string values for text-based protocols, e.g. \cc{many uint8 = ``Foo''}.
-
-In some protocols, there might be many ways to represent the same constant field
-and there is no semantic difference between the different syntactic representations.
-Nail therefore allows repeated constants, such as \cc{many (uint8=' ')}, which parses any number of space characters, or
-\cc{|| uint8 = 0x90 || uint16 = 0x1F0F}, which parses two of the many representations for x86
-NOP instructions, which are used as padding between basic blocks in an executable.
-
-As discussed above, choosing to use these combinators on constant parsers
-weakens the bijection between the format and the data type, as there are
-multiple byte-strings that correspond to the same internal representation and
-the generator chooses one of these.
+The body of the \texttt{dnspacket} rule is a structure, which contains any number of fields enclosed
+between  curly braces. Each field in the structure is parsed in sequence, and represented as a
+structure to the programmer.  Contrary to other programming languages, Nail does not have a special
+keyword for structs. We also reverse the usual structure-field syntax: \texttt{id uint1} is a field
+called \texttt{id} with type \texttt{uint1}\footnote{This is a feature from Go, and makes nested
+  types more readable and the parser for Nail itself easier to write.}.
 
@@
 An example is provided in Figure~\ref{fig:grammar-arith}.
+\paragraph{Repetition.}
 
-% If two options the same string could be parsed by two options, generated output for the latter
-% option is not necessarily understood identically by the parser. However, actual data formats are
-% usually not ambiguous in this sense.
-% \XXX[nz][I don't understand this paragraph (esp the first sentence).]
-
+The \cc{many} combinator takes a parser and applies it repeatedly
+until it fails, returning an array of the inner parser's results. In line 39 of the DNS grammar, a
+sequence of labels is parsed by parsing as many labels as possible, that is, until an invalid length
+field is encountered.
+The \cc{sepBy} combinator
+additionally takes a constant parser, which it applies in between parsing
+two values, but not before parsing the first value or after parsing the
+last.  This is useful for parsing an array of items delimited by a separator.
+%For examle, \cc{many uint8} represents an array of 8-bit unsigned
 \paragraph{Optional.}
@@
 \paragraph{References.}
-Rules allow for recursive grammars. To support recursive data types, we introduce the
+Rules allow for recursive grammars, but some programming languages, such as C, do not support
+recursive structures.  We introduce the
 reference combinator \cc{*}  that does not change the syntax of the external format described, but
-introduces a layer of indirection, such as a reference or pointer, to the model data type.
+introduces a reference or pointer to the internal data type.
 The reference combinator does not need to be used when another combinator, such as \cc{optional} or
@@
 
-\subsection{Dependent fields}
+\subsection{Redundant Data}
 \label{s:dependent}
 Data formats often contain values that are determined by other values or the layout of information,
-such as checksums, duplicated information, or offset and  length fields.
-We represent such values using \emph{dependent fields} and handle them transparently during
-parsing and generation without exposing them to the internal model.
+such as checksums, duplicated information, or offset and  length fields. Exposing such values risks
+inconsistencies that could trick the program into unsafe behaviour. Therefore, we represent such values using \emph{dependent fields} and handle them transparently during
+parsing and generation without exposing them to the application.
+
+In DNS packets, the packet header contains count fields (\cc{qc},
+\cc{ac}, \cc{ns}, and \cc{ar}), which contain the number of questions and answers that follow the
+header, which we represent by dependent fields (line 12-15).
 
@@
 
-Dependent fields are handled like other fields when parsing input, but their values are not stored
-in the internal data type. Instead the value can be referenced by subsequent parsers and it
-discarded when the field goes out of scope.
+Dependent fields are handled like other fields when parsing input, but only stored temporarily
+instead of in the data type. Their value can be referenced by other parsers until it goes out of scope.
 When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
@@
 Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
-the \cc{many} combinator, except it represents an exact number, specified in the dependent field, of
-repetitions, as opposed to as many repetitions as possible. For example, DNS labels, which
-are encoded as a length followed by
-a value, are described in line 38 of the DNS grammar.
+the \cc{many} combinator, except it parses an exact number of repetitions specified in the dependent
+field. Lines 16-19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
+answer records in a DNS packet.
 Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
@@
 back, reading input or appending output.
-Grammars can use the \cc{apply} combinator to parse or generate external data on a different stream,
-inserting the result in the data model.
+Grammars can use the \cc{apply} combinator to parse a stream or generate output into it.
 
@@
 Transformations can handle a wide variety of patterns in data formats, including the following:
+
+\paragraph{Compressed data.}
+Encoded, compressed, or encrypted data can be handled transparently by writing a custom
+transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
+versa. This transformation must be carefully written to not have bugs.
+
+For example, DNS uses label compression to reduce the overhead of including each domain name
+multiple times in a DNS reply. If a domain name suffix is repeated, it may be replaced by a two-bit
+marker followed by a  14-bit offset into the packet, indicating the position where that suffix was
+previously encoded.  We use a transform in line 35
+of our DNS grammar to handle DNS label compression; the signatures of the two transform functions are shown
+in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
+decompresses the DNS label stream.
+When generating a packet, this transform receives the current suffix as
+an input, and scans the packet so far for previous occurrences and inserts the offset.
+
+Handling label compression in existing tools, such as Bison or Hammer,
+would be awkward at best, because some ad-hoc trick would have to
+be used to re-position the parser's input stream.  Keeping track of the
+position of all recognized labels would not be enough, although still awkward, as the offset field
+may refer to any byte within the packet.
+For this reason, the sample DNS grammar in Hammer does not
+support compression.
+
 \paragraph{Offsets.}
@@
 
-\begin{figure}[h]
-\smaller[0.5]
-\input{code/offset-xform}
-\caption{Pseudocode for two functions that implement the offset transform.}
-\label{fig:xform-sig}
-\end{figure}
-
 \paragraph{Sizes.}
@@
 
-\paragraph{Compressed data.}
-Encoded, compressed, or encrypted data can be handled transparently by writing a custom
-transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
-versa. This transformation must be carefully written to not have bugs.
-
 \paragraph{Checksums.}
@@
 field. In some cases, a checksum is calculated over a buffer that contains the checksum itself, with
-the checksum being set to some particular value. Because the functions implementing a transformation
-are passed a pointer to any dependent fields, the checksum function can set the checksum's initial
-value before calculating the checksum over the entire buffer, including the checksum.
+the checksum being set to some particular value.
 
@@
 A real-world example with many different transforms, used to support
-the ZIP file format, is described in \S\ref{s:eval-formats}.
+the ZIP file format, is described in our full paper.%\S\ref{s:eval-formats}.
 

### 2014-11-14 paper/login/eval.tex
Login first rough draft

--- 2014-08-21 paper/osdi14/eval.tex
@@
 
-\input{fig-ethernet}
+% In our evaluation of Nail, we answer four questions:
 
-In our evaluation of Nail, we answer four questions:
+% \begin{itemize}
 
-\begin{itemize}
+% \item Can Nail grammars support real-world data formats, and
+%       are Nail's techniques critical to handling these formats?
 
-\item Can Nail grammars support real-world data formats, and
-      are Nail's techniques critical to handling these formats?
+% \item How much programmer effort is required to build an
+%       application that uses Nail for data input and output?
 
-\item How much programmer effort is required to build an
-      application that uses Nail for data input and output?
+% \item Does using Nail for handling input and output improve
+%       application security?
 
-\item Does using Nail for handling input and output improve
-      application security?
+% \item Does Nail achieve acceptable performance?
 
-\item Does Nail achieve acceptable performance?
+% \end{itemize}
 
-\end{itemize}
-
-\subsection{Data formats}
+\subsection{Real-World Formats}
 \label{s:eval-formats}
 
-To answer the first question, we used Nail to implement grammars
+First, we ask whether Nail can handle real-world data formats and whether its techniques are
+critical to that. We used Nail to implement grammars
 for seven protocols with a range of challenging features.
-Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
-of code for their Nail grammars, and the challenging features that make
-the protocols difficult to parse with state-of-the-art parser generators.
+Figure~\ref{fig:eval-protocols} summarizes our results.
 We find that despite the challenging aspects of these protocols, Nail is
@@
 
-\paragraph{DNS.}
+\paragraph{DNS} In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
+for DNS packets, shown in Figure~\ref{fig:dns-full}.
+The grammar corresponds almost directly to the diagrams in RFC 1035,
+which defines DNS~\cite[\S4]{RFC:1035}. Nail's dependent fields handle DNS's count fields, and
+transformations represent label compression. Both of these features are awkward at best to handle
+with existing tools.
 
-In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
-for DNS packets, shown in Figure~\ref{fig:dns-xform}.
-The grammar corresponds almost directly to the diagrams in RFC 1035,
-which defines DNS~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
-header, a set of question records, and a set of answer records. Domain
-names in both queries and answers are encoded as a sequence of labels,
-terminated by a zero byte.  Labels are Pascal-style strings, consisting
-of a length field followed by that many bytes comprising the label.
+\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format~\cite{pkzip}.
+ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an
+\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
+located by scanning backwards from the end of the file until a magic number and a valid length field
+is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
+contains the magic number~\cite{wolf:berlinsides-zip}.
 
-One challenging aspect of DNS packets lies in the count fields (\cc{qc},
-\cc{ac}, \cc{ns}, and \cc{ar}), which represent the number of questions
-or answers in another part of the packet.  Nail's \cc{n_of} combinator
-handles this situation easily, which would have been difficult to handle
-for other parsers.
-
-Another challenging aspect of DNS is label
-compression~\cite[\S4.1.4]{RFC:1035}.  Label compression is used to reduce
-the size overhead of including each domain name multiple times in a DNS
-reply (once in the question section, and at least once in the response
-section).  If a domain name suffix is repeated, instead of repeating that
-suffix, the DNS packet may write a two-bit marker sequence followed by
-a 14-bit offset into the packet, indicating the position of where that
-suffix was previously encoded.
-
-Handling label compression in existing tools, such as Bison or Hammer,
-would at best be very awkward, because some ad-hoc trick would have to
-be used to re-position the parser's input stream.  Keeping track of the
-position of all recognized labels would not be enough, as the offset field
-may refer to any byte within the packet, not just the beginning of labels.
-For this reason, the DNS server used as the example for Hammer does not
-support compression.
-
-\begin{figure}
-\smaller[0.5]
-\input{code/dns-xform}
-\caption{Signatures of stream transform functions for handling DNS label compression.}
-\label{fig:dns-xform}
-\end{figure}
-
-In contrast, Nail is able to handle label compression, by using a stream
-transform; the signatures of the two transform functions are shown
-in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
-decompresses the DNS label stream by following the offset pointers.
-When generating a packet, this transform receives the current suffix as
-an input, and scans the packet so far for previous occurrences, which
-implements compression.
-
-\paragraph{ZIP files.}
-
-An especially tricky data format is the ZIP compressed archive
-format~\cite{pkzip}.  ZIP files are normally parsed end-to-beginning. At
-the end of each ZIP file is an \emph{end-of-directory header}. This header
-contains a variable-length comment, so it has to be located by scanning
-backwards from the end of the file until a magic number and a valid
-length field is found. Many ZIP implementations disagree on how to find
-this header in confusing situations, such as when the comment contains
-the magic number~\cite{wolf:berlinsides-zip}.
-
- This end-of-directory
-header contains the offset and size of the \emph{ZIP directory}, which
-is an array of \emph{directory entry headers}, one for every file
-in the archive.  Each entry stores file metadata, such as file name,
-compressed and uncompressed size, and a checksum, in addition to the
+This header  contains points to the \emph{ZIP directory}, contains an entry every file
+in the archive.  Each entry stores file metadata in addition to the
 offset of a \emph{local file header}. The local file header duplicates
-most information from the directory entry header. The compressed file
-contents follow the header immediately.
-
-\input{fig-zip}
-
+most information from the directory entry header and is followed immediately by the compressed
+archive entry.
 Duplicating information made sense when ZIP files were stored on floppy
 disks with slow seek times and high fault rates, and memory constraints
-made it impossible to keep the ZIP directory in memory or the archive
-was split across multiple disks.  However, care must be taken that the
-metadata is consistent. For example, vulnerabilities could occur if
+made it impossible to keep the ZIP directory in memory.
+
+ However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
 the length in the central directory is used to allocate memory and the
@@
 they are equal first, as was the case in the Python ZIP
-library~\cite{cve-2012-7338}.
-\begin{comment}
-\begin{figure}[tb]
-\smaller[0.5]
-\input{code/zip-eod-xform}
-\caption{Signatures of stream transform functions for handling the
-end-to-beginning structure of ZIP files.}
-\label{fig:zip-eod-xform}
-\end{figure}
-\end{comment}
-Figure~\ref{fig:zip-extract} shows an abbreviated version of our ZIP
-file grammar.  The ZIP grammar is a good example of how transformations
-capture complicated syntax in a real-world file format; existing parser
-languages cannot handle a file format of this complexity.
+library~\cite{python-bug:20078}.
 
-The \cc{zip\_file} grammar first splits the entire file stream into two
-streams based on the \cc{zip\_end\_of\_directory} transform on line 2.
-%(whose two C function signatures are shown in Figure~\ref{fig:zip-eod-xform}). 
-\cc{zip\_end\_of\_directory_parse}
-finds the end-of-directory header as described above, by scanning the
-file backwards, and splits the file into two streams, one containing
-the end-of-directory header and one containing the file contents. The
-\cc{end\_of\_directory} rule is then applied to the header stream in
-line 4. All offsets in the ZIP file refer to the beginning of the
-file, so the stream \cc{\$file} which contains the file contents
-without the header is passed as an argument to all parsers from
-hereon.
+Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
+files transparently with transformations, and the programmer
+is presented with a data structure containing the archive contents. To handle formats based on ZIP,
+such as office documents, our ZIP grammar could be extended to apply additional grammars to the
+files within an archive.
 
-The directory header contains the offset and size of the ZIP
-directory (lines 9 and 10). The \cc{offset} and \cc{size}
-transformations extract a stream containing just the directory from
-the file contents. This stream is then parsed as an array of directory
-entries in line 17.
-Each directory entry in turn points to a local file header, which is
-similarly extracted and parsed with the \cc{file} rule.
-
-The \cc{file} rule starting at line 41, describing a ZIP file entry,
-takes dependent field
-parameters containing file metadata information from the directory
-header. However, this same information is duplicated in the file
-entry, so the grammar uses the Nail-supplied \cc{u16\_depend}
-transform to check whether the two values are equal. Unlike most other
-transforms, \cc{u16\_depend} does not consume or produce strings; it
-only checks that two dependent fields are equal when parsing, and
-assigns the value of the second field to the first when generating.
-This ensures that the programmer does not have to worry about
-inconsistencies when handling the internal representation of a ZIP
-file.
-
-Immediately following the file entry is the compressed data.  Because
-most compression algorithms operate on unbounded streams of data, Nail
-decompresses data in two steps.  First, it isolates the compressed
-data from the rest of the stream by using the \cc{size} transform,
-which operates on the current stream, meaning it will consume data
-starting at the current position of the parser in the input.  Second,
-Nail invokes a custom \cc{zip\_compression} transform that implements
-the appropriate compression and decompression functions based on the
-specified compression method. These functions are otherwise oblivious
-to the layout or metadata of the file.
-
-\subsection{Programmer effort}
+\subsection{Applications}
 \label{s:eval-effort}
@@
 %2821 lines
-
-To evaluate how much programmer effort is required to build an
-application that uses Nail, we implemented two applications---a
-DNS server, and an
-\cc{unzip} program---based on the above grammars, and compared code
-size with comparable applications that process data manually, using
-\cc{sloccount}~\cite{sloccount}.  We also compare the code size of our
-DNS server to a DNS server written using the Hammer parsing framework,
-although it does not fully support DNS (e.g., it lacks label compression,
-among other things).  Figure~\ref{fig:effort} summarizes the results.
-
+We implemented two applications---a
+DNS server and an
+\cc{unzip} program---based on the above grammars, and will compare the effort involved and the
+resulting security to similar applications with hand-written parsers and with other parser generators.
+We will use lines of code as a proxy for programmer effort. To evaluate security, we will argue how
+our design avoids classes of vulnerabilities and fuzz-test one of our applications.
 \begin{figure}[tb]
 \centering
-\begin{tabular}{@{~}lrr@{~}l@{~}}
+\smaller[0.5]
+\begin{tabular}{lrr@{~}l}
 \toprule
@@
 
-% It is hard to compare the programming effort required to implement
-% our DNS server to that of a real world DNS server, since we implement
-% less functionality: we do not send the additional glue records and
-% we did not implement any configuration options.  However, the closest
-% in functionality and intent is Dan Bernstein's djbdns~\cite{djbdns},
-% which aims to be a minimalist, highly secure, authoritative-only DNS
-% server. The latest release of djbdns, including various support tools,
-% is about 10,000 lines of C\@.  \XXX[nz][so what does this mean?]
+We ran the DNS fuzzer provided with the Metasploit
+framework~\cite{mspframework} on our DNS server, which sent randomly
+corrupted DNS queries to our server for 4 hours, during which it did
+not crash or trigger the stack or heap corruption detector.
 
+\paragraph{Performance}
+
+To evaluate whether Nail-based parsers are compatible with good
+performance, we compare the performance of our DNS server to that of ISC
+BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
+We simulate a load resembling that of an authoritative name server, generating a random zone file
+and a random sequence of queries, with 10\% non-existent domains. We repeated this sequence of
+queries for one minute against both DNS servers. The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
+our Nail-based DNS server can achieve higher performance and lower latency
+than BIND\@.  Although BIND is a more sophisticated DNS server,
+and implements many features that are not present in our Nail-based DNS
+server and that allow it to be used in more complicated configurations, 
+we believe our results demonstrate that Nail's parsers are not
+a barrier to achieving good performance.
 \paragraph{ZIP.}
@@
 the zlib library, and one for finding the end-of-directory header).
-
-Because more recent versions of ZIP have added more features, such
-as large file support and encryption, the closest existing tool
-in functionality is the historic version 5.4 of the Info-Zip unzip
-utility~\cite{infozip} that is shipped with most Linux distributions.
-The entire unzip distribution is about 46,000 lines of code, which is
-mostly optimized implementations of various compression algorithms and
-other configuration and portability code.  However, unzip isolates the
-equivalent of our Nail tool in the file \cc{extract.c}, which parses the
+ The \cc{unzip} utility contains a file \cc{extract.c}, which parses
 ZIP metadata and calls various decompression routines in other files. This
@@
 
-\subsection{Security}
-
-We use a twofold approach to evaluate the security of applications
-implemented with Nail. First, we analyze a list of CVE's related to the
-ZIP file format and argue how our ZIP tools based on Nail are immune
-against those vulnerability classes. Second, we present the results of
-fuzz-testing our DNS server.
-
-\paragraph{ZIP analysis.}
-
-In \S\ref{s:motivation}, we found 15 input handling vulnerabilities
-related to ZIP files.
-
-11 of these vulnerabilities involved memory corruption during input handling.
-Because Nail's generated code checks offsets before reading and does not
-expose any untrusted pointers to the application, it is immune against
+In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
+11 of these vulnerabilities involved memory corruption during input handling, which generated code is immune to
 memory corruption attacks by design.
-
-Nail also protects against parsing ambiguity vulnerabilities like the four others we studied. First,
+More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
 Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
-that caused a vulnerability in the Python ZIP library. Second, the other three vulnerabilities exist
+that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
 because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
-parsers are not very reusable, as they build application specific data structures and are tightly
+parsers are not very reusable, as they build application-specific data structures and are tightly
 coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
 avoiding protocol misunderstandings.
-
-\paragraph{DNS fuzzing.}
-
-To provide additional assurance that Nail parsers are free of memory
-corruption attacks, we ran the DNS fuzzer provided with the Metasploit
-framework~\cite{mspframework} on our DNS server, which sent randomly
-corrupted DNS queries to our server for 4 hours, during which it did
-not crash or trigger the stack or heap corruption detector.
-
-\subsection{Performance}
-
-To evaluate whether Nail-based parsers are compatible with good
-performance, we compare the performance of our DNS server to that of ISC
-BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
-We simulate a load resembling that of an authoritative name server.
-First, we generate domain names consisting of one or two labels randomly
-selected from an English dictionary, and one label that is one of
-three popular top-level domains (\cc{com}, \cc{net}, and \cc{org}).
-Second, we randomly selected 90\% of these domains and created a zone
-file that mapped these domain names to \cc{127.0.0.1}.  Finally, we used
-the \cc{queryperf} tool provided with BIND to query each domain between
-zero and three times, using a DNS server running on the local machine.
-We used a single core of an Intel i7-3610QM system with 12GB of RAM.
-The benchmark tool kept at most 20 queries outstanding at once, and
-was configured to repeat the same randomized sequence of queries for
-one minute.  We repeated each test seven times with 50,000 domain names,
-restarting each daemon in between; we also repeated the tests with 1
-million domain names, and found similar results.  We also performed one
-initial dry run to warm the file system cache for the zone file.
-
-The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
-our Nail-based DNS server can achieve higher performance and lower latency
-than BIND\@.  Although BIND is a more sophisticated DNS server,
-and implements many features that are not present in our Nail-based DNS
-server and allow it to be used in more complicated configurations, 
-we believe our results demonstrate that Nail's parsers are not
-a barrier to achieving good performance.
-
-\input{fig-perf-dns}

### 2014-11-14 paper/login/impl.tex
Login first rough draft

--- 2014-05-02 paper/osdi14/impl.tex
@@
 performed with a subset of the grammar implemented using conventional grammars. 
-And option for C++ STL data models  is in development. In
+An option for C++ STL data models is in development.  In
 this section, we will discuss some particular features of our parser
 implementation.
-
-% The source code of our implementation, together with the examples described in
-% \S\ref{s:eval} are available on GitHub at \url{https://github.com/jbangert/nail}.
-
-\paragraph{Parsing.}
 
@@
 
-\paragraph{Defense-in-Depth.}
+\paragraph{Defense-in-depth.}
 
-Security exploits often rely on raw inputs being present in memory, for example to include
-shell-code or crafted stack frames for ROP-attacks in padding fields~\cite{shotgun-parser} or the
+Security exploits often rely on raw inputs being present in memory~\cite{shotgun-parser}, for example to include
+shell-code or crafted stack frames for ROP~\cite{phrack58:4-nergal} attacks in padding fields or the
 application executing a controlled sequence of heap allocations and de-allocations to place
@@
 When parsing input, Nail uses two separate memory arenas. These arenas allocate memory from the
-system allocator in large, fixed size blocks. Allocations are handled linearly and all data in the
+system allocator in large, fixed-size blocks. Allocations are handled linearly and all data in the
 arena is zeroed and freed at the same time. Nail uses one arena for data used only during parsing,
-including dependent fields and temporary streams and is released before the Nail's parser returns.
+including dependent fields and temporary streams; this arena is released before the parser returns.
 The other arena is used to allocate the internal data type returned and is freed by the application
@@
 can therefore be zeroed immediately after the parser succeeds, so an attacker has to write an
-exploit that works without referencing data in the raw input.
+exploit that works without referencing data from the raw input.
+
+Finally, Nail performs sophisticated error handling only in a special debug configuration and will print
+error messages about the input only to \cc{stderr}. Besides complicating the parser, advanced error handling
+invites programmers to attempt to fix malformed input, such as adding reasonable defaults for a
+missing field. Such error-fixing not only introduces parser inconsistencies, but also might allow an
+attacker to sneak inconsistent input past a parser.
 % \paragraph{Intermediate representation.}

### 2014-11-14 paper/login/intro.tex
Login first rough draft

\begin{quote}
  I am ZIP, file of files.
  Parse me, ye mighty and despair.

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013~\cite{cvedetails:libpng},
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past~\cite{cvedetails:zlib}.
Most of these attacks involve memory corruption - therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

However, as memory-safe languages and exploit mitigation  tricks are becoming more prevalent,
attackers are moving to a new class of attack - parser differentials. Many applications use
hand-written input processing code, which is often mixed with the rest of the application, e.g. by
passing a pointer to the raw input through the application. This (anti-)pattern makes it impossible
to figure out whether two implementations of the same format or protocol are identical, and input
handling code can't be easily reused between applications. As a result, different applications often
disagree in the corner cases of a protocol, which can have fatal security consequences. For example,
Android has two parsers for ZIP archives involved in securely installing applications. First, a Java
program checks the signatures of files contained within an app archive and then another tool
extracts them to the filesystem. Because the two ZIP parsers disagree in multiple places, attackers
can modify a valid file so that the verifier will see the original contents, but attacker-controlled
files will be extracted to the filesystem, bypassing Android's code signing. Similar issues showed
up on iOS \cite{evasion} and SSL \cite{layercake}.

Instead of attempting to parse inputs by hand and failing, 
a promising approach is to specify
a precise grammar for the input data format, and automatically generate parsing code from that with tools like {\tt yacc}. As long
as the parser generator is bug-free, the application will be safe from many
input processing vulnerabilities. Grammars can also be re-used between
applications, further reducing effort and eliminating inconsistencies.

This approach is typical in compiler design and in other applications handling text-based inputs, but not
common for binary inputs. The Hammer framework~\cite{hammer} and data description languages such as
PADS~\cite{pads} have been developing generated parsers for binary protocols.

However, existing tools are not powerful enough to handle some complixities formats such as PDF
and ZIP, so programmers still have to hand-write code to handle the most complicated and therefore
most error-prone formats. For example, existing parser generators cannot conveniently represent size or offset fields and more complex features, such as data compression or checksums cannot be expressed at all.

Furthermore, some parser generators are cumbersome to use when parsing binary data for several reasons. First, parser generators typically parse inputs into an abstract syntax tree that corresponds to the grammar. Application developers must then write explicit {\em semantic actions} to construct a data structure the rest of the program can process. Therefore, programmers must describe the format up to three times -- in the grammar, the data structure and the semantic actions. Second, many applications also need to produce outputs in the same formats they parse - say when saving files or responding to network requests. However, most parser generators only address parsing inputs, so  the programmer must manually construct outputs. Some parser generators, such as 
Boost.Spirit~\cite{boost-spirit} allow generating output, but require the programmer to write another set of semantic actions.

We address these challenges with Nail, a new parser generator  with several features geared towards
binary formats.
First, Nail grammars describe not only a format, but also a data type to represent it within the program.
This removes the semantic actions and type declarations that programmers have to write with existing
parser generators. This forces the programmer
to clearly separate syntactic validation and semantic processing.

Second, Nail does not just
parse input into an internal representation, but also generates output from the internal representation,
without requiring the programmer to write additional code.

Third, Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow the programmer to escape the
generated code to modify the raw data and interact with dependent fields
in a controlled manner.

In the rest of this article, we will show some tricky features of real world formats and how to
handle them with Nail.

%% To evaluate whether Nail's design is effective at handling real-world data
%% formats, we implemented a prototype of Nail for C\@. Using our prototype, we
%% implemented grammars for parts of an IP network stack, for DNS packets,
%% and for ZIP files, each in about a hundred lines of grammar. On top of these
%% grammars, we were able to build a DNS server in under 200 lines of C code, and
%% an {\tt unzip} utility in about 50 lines of C code, with performance comparable
%% to or exceeding existing implementations. This suggests both that Nail is
%% effective at handling complex real-world data formats, and that Nail makes it
%% easy for application developers to parse and generate external data
%% representations. Performance results show that the Nail-based DNS server
%% outperforms the widely used BIND DNS server, demonstrating that Nail-based
%% parsers and generators can achieve good performance.

%% The rest of this paper is organized as follows.  \S\ref{s:relwk}
%% puts Nail in the context of related work.
%% \S\ref{s:motivation} motivates the need for Nail by examining
%% past data format vulnerabilities.
%% \S\ref{s:design} describes
%% Nail's design.  \S\ref{s:impl} discusses our implementation
%% of Nail.  \S\ref{s:eval} provides evaluation results, and
%% \S\ref{s:concl} concludes.

### 2014-11-14 paper/login/table-motivation.tex
Login first rough draft

\begin{tabular}{lllr}
\toprule
 \bf Classification & \bf Example CVE & \bf Example description & \bf Count\\
\midrule
  Memory corruption & CVE-2013-5660 & Buffer overflow & 11\\
  Parsing inconsistency & CVE-2013-1462 & Multiple virus scanners interpret ZIP files incorrectly & 4\\
  Semantic misunderstanding & CVE-2014-2319 & Weak cryptography used even if user selects AES & 1\\
\midrule
  \multicolumn{3}{l}{\bf Total of all vulnerabilities related to .zip processing }& \bf 16  \\
\bottomrule

\end{tabular}

### 2014-11-14 paper/login/code/dns-api.c

@@ -1,0 +1,7 @@
+struct dnspacket *parse_dnspacket(NailArena *arena,
+  const uint8_t *data,
+  size_t size);
+
+int gen_dnspacket(NailArena *tmp_arena,
+  NailStream *out,
+  struct dnspacket *val);

### 2014-11-14 paper/login/code/dns-struct.c

@@ -1,0 +1,9 @@
+struct dnspacket {
+  uint16_t id;
+  uint8_t qr;
+  /* ... */
+  struct {
+    struct question *elem;
+    size_t count;
+  } questions;
+};

### 2014-11-14 paper/login/code/dns.nail.c

@@ -1,0 +1,39 @@
+labels = <many {@length  uint8 | 1..64
+              label n_of @length uint8 }
+          uint8 = 0 > 
+compressed={
+   $decompressed transform dnscompress($current)
+   labels apply $decompressed label
+}     
+question={
+  labels compressed
+  qtype uint16 | 1..16 
+  qclass uint16 | [1,255]
+}
+answer={
+  labels compressed
+  rtype uint16 | 1..16
+  class uint16 | [1]
+  ttl uint32
+  @rlength uint16 
+  rdata n_of @rlength uint8 
+}
+dnspacket = {
+  id uint16
+  qr uint1
+  opcode uint4
+  aa uint1 
+  tc uint1
+  rd uint1
+  ra uint1
+  uint3 = 0
+  rcode uint4
+  @questioncount uint16
+  @answercount uint16
+  @authoritycount uint16   // Ignored
+  @additionalcount uint16  //Ignored       
+  questions n_of @questioncount question         
+  responses n_of @answercount answer
+  authority n_of @authoritycount answer
+  additional n_of @additionalcount answer
+}

### 2014-11-14 paper/login/code/zip-eod-xform.c

@@ -1,0 +1,6 @@
+int zip_end_of_directory_parse(
+  NailArena *tmp, NailStream *out_files,
+  NailStream *out_dir, NailStream *in_current);
+int zip_end_of_directory_generate(
+  NailArena *tmp, NailStream *in_files,
+  NailStream *in_dir, NailStream *out_current);

### 2014-11-14 paper/login/code/zip-out.c

@@ -1,0 +1,28 @@
+struct zip{
+  end_of_directory contents;
+};
+struct end_of_directory{
+  uint16_t disks;
+  uint16_t directory_disk;/*...*/
+  struct {
+    dir_fileheader* elem;
+    size_t count;
+  } files;
+};
+struct dir_fileheader{
+  struct {
+    uint8_t*elem;
+    size_t count;
+  } filename;
+  fileentry contents;
+};/*..*/
+//The programmer calls these generated functions.
+int gen_zip(NailArena *tmp_arena,NailStream *out,zip *val);
+zip*parse_zip(NailArena *arena, 
+              const uint8_t *data, size_t size);
+//The programmers implements these two transformation
+//functions and two similar ones for output.
+extern int zip_eod_parse(NailArena *tmp,
+ NailStream *filestream, NailStream *,NailStream *current);
+extern int zip_compression_parse(NailArena *tmp,
+ NailStream *uncomp,NailStream *compressed,uint32_t* size);

### 2014-11-14 paper/login/code/zip.nail.c

@@ -1,0 +1,36 @@
+zip = {   /*simplified and cut for brevity*/
+    /* Call zip_eod transform to isolate end_directory and contents streams*/
+     $contents, $end_directory transform zip_eod ($current)
+     /* Parse end_directory stream. The end_of_directory parser takes the 
+      content stream as a parameter */
+     contents apply $end_directory end_of_directory($contents)
+}
+end_of_directory($filestream) = { /*Grammar rules can have parameters*/
+     uint32 = 0x06054b50 /* constant*/
+     disks uint16 | [0]  /* constraint*/
+     directory_disk uint16 | [0]
+     @this_records uint16 /*dependent field*/
+     @total_records uint16 
+     /*The following transform ensures these two fields are always equal*/
+     transform uint16_depend (@this_records @total_records)
+     @directory_size uint32 /*These two dependent fields are used by the  */
+     @directory_start uint32/* transformations below to find the directory*/
+     /*dirstr1 is the the suffic of filestream starting @directory_start*/
+     $dirstr1 transform offset_u32 ($filestream @directory_start) 
+     /*Another stream with @directory_size bytes starting at that offset*/
+     $directory_stream transform size_u32 ($dirstr1 @directory_size)
+     @comment_length uint16 /*Dependent field used with the built in n_of*/
+     comment n_of @comment_length uint8 /* Variable-length comment*/
+     files apply $directory_stream n_of @total_records 
+        dir_fileheader($filestream) /*Array of directory entries*/
+}
+dir_fileheader($filestream) = { 
+     uint32 = 0x02014b50
+     @compressed_size uint16 
+     @crc32 uint32 
+     @file_name_len uint16
+     @off uint32
+     filename n_of @file_name_len uint8
+     $cstream transform offset_u32 ($filestream @off)
+     contents apply $cstream fileentry(@crc32,@compressed_size)
+}

### 2014-11-14 paper/login/design.tex
Login

@@
 \subsection{Basic Data formats}
+
 \input{fig-dnsfull}
@@
 shown in Figure~\ref{fig:dns-full-struct}, and the parser and
-generator functions shown in Figure~\ref{fig:dns-full-api}. A more concise reference of the syntax
-can be found in~\ref{fig:syntax}.
+generator functions shown in Figure~\ref{fig:dns-full-api}.
 Nail grammars are reuseable between applications and we will use this grammar to implement
@@
 elemental types, Nail supports a integers as elemental parsers and a plethora of \emph{combinators}
-to represent features of a file or protocol.
+to represent features of a file or protocol. We will present the features we used in implementing
+DNS. A more complete reference can be found in~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.
 
@@
 called \texttt{id} with type \texttt{uint1}\footnote{This is a feature from Go, and makes nested
-  types more readable and the parser for Nail itself easier to write.}.
+  types more readable and the parser for Nail itself easier to write.}. Often, Nail grammars have
+structures with just one non-constant field, for example when parsing a fixed header. To prevent
+this from cluttering the data type, we support an alternative form of the structure combinator,
+using angle brackets, that has just one unnamed, non-constant field, which is represented directly
+in the data type.
 
-\paragraph{Wrap combinator.} 
-When implementing real protocols with Nail, we often found
-structures that consist of many constants and only one named field. This pattern is
-common in binary protocols which use fixed headers to denote the type of data
-structure to be parsed.  In order to keep the internal representation cleaner,
-we introduced the wrap combinator, which takes a sequence of parsers containing
-exactly one non-constant parser. The external format is defined as though the wrap combinator were a
-structure, but the data model does not introduce a structure with just one element, making the application-visible representation
-(and thus application code) more concise.
-Line 39 of the DNS grammar uses the wrap combinator to hide the terminating NUL-byte of a sequence
-of labels.
-
-\paragraph{Choices.}
-If multiple structures can appear at a given position in a format, the programmer lists the options
-along with a label for each in the \cc{choose} combinator. 
-During parsing, Nail remembers the current input position and attempts each option in the order they
-appear in the grammar. If an option fails, the parser backtracks to the initial position. If no
-options succeed, the entire combinator fails. In the data model, choices are represented as tagged
-unions.  The programmer has to be careful when options overlap, because if the programmer meant to
-generate output for a choice, but the external representation is also valid for an earlier,
-higher-priority option, the parser will interpret it as such. However, real data formats normally do
-not have this overlap and we did not encounter it in the grammars we wrote.
-An example is provided in Figure~\ref{fig:grammar-arith}.
-\paragraph{Repetition.}
-
-The \cc{many} combinator takes a parser and applies it repeatedly
-until it fails, returning an array of the inner parser's results. In line 39 of the DNS grammar, a
-sequence of labels is parsed by parsing as many labels as possible, that is, until an invalid length
-field is encountered.
-The \cc{sepBy} combinator
-additionally takes a constant parser, which it applies in between parsing
-two values, but not before parsing the first value or after parsing the
-last.  This is useful for parsing an array of items delimited by a separator.
-%For examle, \cc{many uint8} represents an array of 8-bit unsigned
-\paragraph{Optional.}
-Nail includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
-without consuming input when it cannot recognize that value. Syntactically, \cc{optional} is
-equivalent to a choice between the parser and an empty structure, but in the internal model it is
-more concisely represented as a reference that is null when the parser fails.
-For example, the grammar for  Ethernet headers uses \cc{optional vlan\_header} to parse the VLAN
-header that appears only in Ethernet packets transmitted to a non-default VLAN\@.
-
-\paragraph{References.}
-Rules allow for recursive grammars, but some programming languages, such as C, do not support
-recursive structures.  We introduce the
-reference combinator \cc{*}  that does not change the syntax of the external format described, but
-introduces a reference or pointer to the internal data type.
-The reference combinator does not need to be used when another combinator, such as \cc{optional} or
-\cc{many}, already introduces indirection in the data type. An example is shown in Figure~\ref{fig:grammar-arith}.
-
-\begin{figure}
-\smaller[0.5]
-\begin{verbatim}
-expr = choose {
-  PAREN = <uint8='('; *expr; uint8=')'>
-  PRODUCT = sepBy1 uint8='*' expr
-  SUM = sepBy1 uint8='+' expr
-  INTEGER = many1 uint8 | '0' .. '9'
-}
-\end{verbatim}
-% \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
-\caption{Grammar for sums and products of integers.}
-\label{fig:grammar-arith}
-\end{figure}
 % To support recursive structures, we include a reference combinator \cc{*}, which
@@
 %   level of indirection to the value. This allows recursive structures to be parsed, f
+
+\input{fig-syntax}
 
@@
 parsing and generation without exposing them to the application. 
+Dependent fields are handled like other fields when parsing input, but only stored temporarily
+instead of in the data type. Their value can be referenced by other parsers until it goes out of scope.
+When generating output, Nail inserts the correct value.
 
@@
 header, which we represent by dependent fields (line 12-15).
-
 Dependent fields are defined within a structure like normal fields, but their name starts with an \cc{@} symbol.
@@
 the same structure. Dependent fields can be passed to rule invocations as parameters.
-
-Dependent fields are handled like other fields when parsing input, but only stored temporarily
-instead of in the data type. Their value can be referenced by other parsers until it goes out of scope.
-When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
-structure, the generator reserves space for the dependent field in the output. Once
-the dependent field goes out of scope, the generator  writes the
-dependent field's value to this space.
 
@@
 
+\begin{figure}
+\smaller[0.5]
+\input{code/dns-xform}
+\caption{Signatures of stream transform functions for handling DNS label compression.}
+\label{fig:dns-xform}
+\end{figure}
 Handling label compression in existing tools, such as Bison or Hammer,
@@
 using \cc{apply \$fragment some_parser}.
+
+\begin{figure}[h]
+\smaller[0.5]
+\input{code/offset-xform}
+\caption{Pseudocode for two functions that implement the offset transform.}
+\label{fig:xform-sig}
+\end{figure}
 

### 2014-11-14 paper/login/intro.tex
Login

@@
 files will be extracted to the filesystem, bypassing Android's code signing. Similar issues showed
-up on iOS \cite{evasion} and SSL \cite{layercake}.
+up on iOS \cite{geohot-evasion} and SSL \cite{DBLP:conf/fc/KaminskyPS10}.
 
@@
 This approach is typical in compiler design and in other applications handling text-based inputs, but not
-common for binary inputs. The Hammer framework~\cite{hammer} and data description languages such as
-PADS~\cite{pads} have been developing generated parsers for binary protocols.
+common for binary inputs. The Hammer framework~\cite{hammer-parser} and data description languages such as
+PADS~\cite{Fisher:2005:PDL:1064978.1065046} have been developing generated parsers for binary protocols.
 
@@
 
-Furthermore, some parser generators are cumbersome to use when parsing binary data for several reasons.
-First, parser generators typically parse inputs into an abstract syntax tree that corresponds to the grammar.
-Application developers must then write explicit {\em semantic actions} to construct a data structure the rest of the program can process.
-Therefore, programmers must describe the format up to three times -- in the grammar, the data structure and the semantic actions.
-Second, many applications also need to produce outputs in the same formats they parse - say when saving files or responding to network requests.
-However, most parser generators only address parsing inputs, so  the programmer must manually construct outputs.
-Some parser generators, such as
+Furthermore, some parser generators are cumbersome to use when parsing binary data for several
+reasons. First, many parser generators don't directly data structures, but call semantic actions
+that the programmer writes to construct a data structure. Therefore, programmers must
+describe the format up to three times -- in the grammar, the data structure and the semantic
+actions. Second, most parser generators only address parsing inputs, so programmers have to manually
+construct outputs. Some parser generators, such as 
 Boost.Spirit~\cite{boost-spirit} allow generating output, but require the programmer to write another set of semantic actions.
 
-We address these challenges with Nail, a new parser generator  with several features geared towards
-binary formats.
+We address these challenges with Nail, a new parser generator for binary formats.
 First, Nail grammars describe not only a format, but also a data type to represent it within the program.
 This removes the semantic actions and type declarations that programmers have to write with existing
-parser generators. This forces the programmer
-to clearly separate syntactic validation and semantic processing.
-
+parser generators. It also forces the programmer to clearly separate syntactic validation and semantic processing.
 Second, Nail does not just

### 2014-11-24 paper/login/design.tex
login

@@
 keyword for structs. We also reverse the usual structure-field syntax: \texttt{id uint1} is a field
-called \texttt{id} with type \texttt{uint1}\footnote{This is a feature from Go, and makes nested
-  types more readable and the parser for Nail itself easier to write.}. Often, Nail grammars have
-structures with just one non-constant field, for example when parsing a fixed header. To prevent
-this from cluttering the data type, we support an alternative form of the structure combinator,
-using angle brackets, that has just one unnamed, non-constant field, which is represented directly
-in the data type.
+called \texttt{id} with type \texttt{uint1}. Often, Nail grammars have
+structures with just one non-constant field, for example when parsing a fixed header. Nail supports
+this with an alternative form of the structure combinator,
+using angle brackets, that contains one unnamed, non-constant field, which is represented directly
+in the datatype, without introducing another layer of indirection, as shown on line 40.
 
@@
 \label{s:transforms}
+Formats have
 
@@
 before it can be handled.
-
-Nail introduces two concepts to handle these challenges, \emph{streams} and \emph{transformations}. 
 Streams represent a sequence of bytes that contain some external format. The parsers and generators
@@
 Grammars can use the \cc{apply} combinator to parse a stream or generate output into it. 
-
 Streams are passed as arguments to a rule or defined within the grammar through \emph{transformations}.
 The current stream is always passed as an implicit parameter.
+
+There are many ways conceptual streams could be encoded in a data format --
+A tool like Nail needs to support all of them. This makes data description languages like
+PADS~\cite{Fisher:2005:PDL:1064978.1065046} contain not just a kitchen sink, but a kitchen store
+full of features -- and a language that can handle all possible format oddities will be a general
+purpose language.  Nail instead of allows you to plug in your own code as \emph{transformations},
+with a small, but powerful camping kitchen included that will handle most formats.
 
@@
 
-Typically, the top level of most grammars is a rule that takes only a single stream, which may then
+Typically, grammars start out with a single stream, which may then
 be broken up by various transformations and passed to sub-rules, which eventually parse various linear
@@
 
- To reduce both programmer effort and the risk of unsafe operations, Nails provides implementations of
+ To make your life easier and your code safer, Nails provides implementations of
 transformations for many common features, such as checksums, size, and offset fields. Furthermore,
@@
 \paragraph{Compressed data.}
-Encoded, compressed, or encrypted data can be handled transparently by writing a custom
-transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
-versa. This transformation must be carefully written to not have bugs.
-
 For example, DNS uses label compression to reduce the overhead of including each domain name
@@
 \paragraph{Offsets.}
-A built-in transformation for handling offset fields, which is invoked as follows:
+We provide a  transformation for handling offset fields, which is invoked as follows:
 \cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 

### 2014-11-24 paper/login/eval.tex
Login
login

@@
 
-\paragraph{Performance}
-
 To evaluate whether Nail-based parsers are compatible with good
@@
 and a random sequence of queries, with 10\% non-existent domains. We repeated this sequence of
-queries for one minute against both DNS servers. The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
-our Nail-based DNS server can achieve higher performance and lower latency
-than BIND\@.  Although BIND is a more sophisticated DNS server,
+queries for one minute against both DNS servers. We found that our DNS server is approximately three
+times faster than BIND. Although BIND is a more sophisticated DNS server,
 and implements many features that are not present in our Nail-based DNS
@@
 In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
-11 of these vulnerabilities involved memory corruption during input handling, which generated code is immune to
-memory corruption attacks by design.
+11 of these vulnerabilities involved memory corruption during input handling, which generated code
+is immune to by design.
 More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.

### 2014-11-24 paper/login/intro.tex
login

@@
   I am ZIP, file of files.
-  Parse me, ye mighty and despair.
+  Parse me, ye mighty and drop a shell.
 
@@
 
-However, existing tools are not powerful enough to handle some complixities formats such as PDF
-and ZIP, so programmers still have to hand-write code to handle the most complicated and therefore
-most error-prone formats.
+However, if you wanted to use existing tools to parse PDF
+or ZIP, you would soon find that they cannot handle the complicated -- and therefore most
+error-prone-- aspects of such formats-- so you'd still have to
+hand-write the riskiest bits of code.
 For example, existing parser generators cannot conveniently represent size or offset fields and more complex features, such as data compression or checksums cannot be expressed at all.
@@
 reasons. First, many parser generators don't directly data structures, but call semantic actions
-that the programmer writes to construct a data structure. Therefore, programmers must
+that the you write to construct a data structure. Therefore,you must
 describe the format up to three times -- in the grammar, the data structure and the semantic
-actions. Second, most parser generators only address parsing inputs, so programmers have to manually
+actions. Second, most parser generators only address parsing inputs, so you'd have to manually
 construct outputs. Some parser generators, such as 
-Boost.Spirit~\cite{boost-spirit} allow generating output, but require the programmer to write another set of semantic actions.
+Boost.Spirit~\cite{boost-spirit} allow generating output, but require you to write another set of semantic actions.
 
@@
 First, Nail grammars describe not only a format, but also a data type to represent it within the program.
-This removes the semantic actions and type declarations that programmers have to write with existing
-parser generators. It also forces the programmer to clearly separate syntactic validation and semantic processing.
-Second, Nail does not just
-parse input into an internal representation, but also generates output from the internal representation,
-without requiring the programmer to write additional code.
+Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
+Second, Nail will also take care of generating output from the same data type, without requiring you
+to write more risky code or think about the format again.
 
@@
 which specify the position or length of another data structure, fall
-in this category.  Transformations allow the programmer to escape the
+in this category.  Transformations allow you to escape the
 generated code to modify the raw data and interact with dependent fields

### 2014-11-25 paper/login/design.tex
Login

@@
 
-We will use DNS packets as a leading example, as defined in RFC 1035~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
+We will use DNS packets as a leading example, as defined in RFC 1035.  Each DNS packet consists of a
 header, a set of question records, and a set of answer records. Domain
@@
 \input{fig-dnsfull}
-Let's stepping through a simplified Nail grammar for non-compressed
+Let's step through a simplified Nail grammar DNS
 packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
@@
 
-Just like C supports various constructs to build up types, such as structures, unions, pointers and
-elemental types, Nail supports a integers as elemental parsers and a plethora of \emph{combinators}
-to represent features of a file or protocol. We will present the features we used in implementing
+Just like C supports various constructs to build up types, such as structures and unions from pointers and
+elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
 DNS. A more complete reference can be found in~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.
@@
 structures with just one non-constant field, for example when parsing a fixed header. Nail supports
-this with an alternative form of the structure combinator,
+this with an alternative form of structures,
 using angle brackets, that contains one unnamed, non-constant field, which is represented directly
 in the datatype, without introducing another layer of indirection, as shown on line 40.
+
+\paragraph{Arrays.}
+
+Nail supports various forms of arrays. Line 40 shows how to parse a domain in a DNS packet with
+\cc{many}, which keeps repeating the \cc{label} rule until it fails.  In the next section, we will
+explain how to handle count fields, and our full paper describes how to handle various array
+representations (such as delimiters or non-empty arrays).
 
@@
 
-Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
-the \cc{many} combinator, except it parses an exact number of repetitions specified in the dependent
-field. Lines 16-19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
+Nail allows handling count fields with \cc{n\_of}, which parses an exact number of repetitions of a rule. Lines 16-19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
 answer records in a DNS packet.
@@
 \label{s:transforms}
-Formats have
-
-Traditional parsers handle input one symbol at a time, from beginning to end.
+So far, we have described a parser that consumes input a byte at a time from beginning to end.
 However, real-world formats often require non-linear parsing. Offset fields require a parser to move
 to a different position in the input, possibly backwards. Size fields require the parser to stop
-processing before the end of input has been reached, and perhaps resume executing a parent parser.
-Other cases, such as compressed data, require more complicated processing on parts of the input
+processing before the end of input has been reached.
+Other cases, such as compressed data and checksums, require more complicated processing on parts of the input
 before it can be handled.
-Streams represent a sequence of bytes that contain some external format. The parsers and generators
-that Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
-back, reading input or appending output.
-Grammars can use the \cc{apply} combinator to parse a stream or generate output into it. 
-Streams are passed as arguments to a rule or defined within the grammar through \emph{transformations}.
-The current stream is always passed as an implicit parameter.
 
-There are many ways conceptual streams could be encoded in a data format --
-A tool like Nail needs to support all of them. This makes data description languages like
+For a parser to be useful, it needs to support all these ways of structuring a format. This makes data description languages like
 PADS~\cite{Fisher:2005:PDL:1064978.1065046} contain not just a kitchen sink, but a kitchen store
-full of features -- and a language that can handle all possible format oddities will be a general
-purpose language.  Nail instead of allows you to plug in your own code as \emph{transformations},
-with a small, but powerful camping kitchen included that will handle most formats.
+full of features -- and a language that can handle all possible formats will be a general
+purpose programming language.  Instead, we keep Nail itself small, and introduce an
+interface that allows complicated format structures to be handled by  plug-in \emph{transformations} 
+in a general purpose
+language. Of course, we ship Nail with a handy library of common transformations to handle common
+format features, such as offsets,sizes and checksums.
+
+These \emph{transformations} consume and produce
+\emph{streams} -- sequences of bytes -- which can be further passed to other transformations and
+eventually parsed by a Nail rule. Transformations can also access values in dependent fields. 
+Streams can be subsets of other streams, for example the substream starting at an
+offset given in a dependent field to handle pointer fields, or computed at runtime
+, such as by decompressing another stream with zlib.
 
 Transformations are two arbitrary functions called during parsing and output generation.
-The parsing function  takes any number of stream arguments and dependent field values,
+The parsing function consumes any number of streams and dependent field values,
 and produces any number of temporary streams. This function may reposition and read from the
 input streams and read the values of dependent fields, but not change their contents and values. 
-The generating function has to be an inverse of the parsing function. It takes the same number of
-temporary streams that the parsing function produces, and writes the same number of streams and
-dependent field values that the parsing function consumes.
+The generating function has to be an inverse of the parsing function, consuming streams and
+producing dependent field values and other streams.
 
-Typically, grammars start out with a single stream, which may then
-be broken up by various transformations and passed to sub-rules, which eventually parse various linear
-fragment streams. Upon parsing, these fragment streams are generated and then combined by the
-transforms.
+\input{fig-zip}
 
- To make your life easier and your code safer, Nails provides implementations of
-transformations for many common features, such as checksums, size, and offset fields. Furthermore,
-Nail provides library functions that can be used to safely operate on streams, such as splitting and
-concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
-efficiently duplicated and split.
+As a concrete example, we will show a grammar for ProtoZIP, a very simple archive format inspired by ZIP in
+Figure~\ref{fig:zip-extract}. ProtoZIP consists of a variable-length end-of-file directory, which
+is a magic number followed by an array of filenames and pointers to compressed files. A grammar for the real ZIP format, which
+has more layers of indirection, is presented in the full paper.
+
+ The grammar first calls the zipdir transform on line 2, which finds the magic number and splits the file into
+ two streams, one containing the compressed files, the other the directory.  Streams are referred to
+ with \cc{\$identifiers}, similar to dependent fields. A C prototype of the \cc{zipdir} transform is
+ shown in\ref{fig:zip-transform}. 
+\begin{figure}[tb]
+\smaller[0.5]
+\input{code/zip-eod-xform}
+\caption{Signatures of stream transform functions for handling the
+end-to-beginning structure of ProtoZIP files.}
+\label{fig:zip-transform}
+\end{figure}
+When parsing input, this will call \cc{zipdir\_parse}, which takes \cc{\$current} -- an implicit
+identifier always refering to the stream currently being handled -- and returns \cc{\$files} and
+\cc{\$header}. When generating output, this will call \cc{zipdir\_generate}, which appends
+\cc{\$files} and \cc{\$header} to \cc{\$current}.
+
+Line 3 then applies the \cc{dir} rule to  the \cc{\$header} stream, passing it the \cc{\$files}
+stream. Within \cc{dir}, \cc{\$current} is now \cc{\$header} and input is parsed from and output
+generated to that stream. The \cc{dir} rule in turn describes the structure of the directory -- a
+magic number and a count field, followed by that many file descripors. Each file descriptor, is then parsed with two
+transformations, the standard-library \cc{slice} which describes an offset and a size within another
+stream, and the custom \cc{zlib}, which compresses a stream using zlib. Finally, we apply a trivial
+grammar (line 14) to the contents.
+
+In a more complicated example, such as an Office document, we could now specify grammars for each
+entry within an archive.
 
@@
    vulnerabilities.
-
-Transformations can handle a wide variety of patterns in data formats, including the following:
-
-\paragraph{Compressed data.}
-For example, DNS uses label compression to reduce the overhead of including each domain name
-multiple times in a DNS reply. If a domain name suffix is repeated, it may be replaced by a two-bit
-marker followed by a  14-bit offset into the packet, indicating the position where that suffix was
-previously encoded.  We use a transform in line 35
-of our DNS grammar to handle DNS label compression; the signatures of the two transform functions are shown
-in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
-decompresses the DNS label stream.
-When generating a packet, this transform receives the current suffix as
-an input, and scans the packet so far for previous occurrences and inserts the offset.
-
-\begin{figure}
-\smaller[0.5]
-\input{code/dns-xform}
-\caption{Signatures of stream transform functions for handling DNS label compression.}
-\label{fig:dns-xform}
-\end{figure}
-Handling label compression in existing tools, such as Bison or Hammer,
-would be awkward at best, because some ad-hoc trick would have to
-be used to re-position the parser's input stream.  Keeping track of the
-position of all recognized labels would not be enough, although still awkward, as the offset field
-may refer to any byte within the packet.
-For this reason, the sample DNS grammar in Hammer does not
-support compression.
-
-\paragraph{Offsets.}
-We provide a  transformation for handling offset fields, which is invoked as follows:
-\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
-This transformation corresponds to two functions for parsing and
-generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
-be used to parse data at the offset contained in \cc{@offset}, by
-using \cc{apply \$fragment some_parser}.
-
-\begin{figure}[h]
-\smaller[0.5]
-\input{code/offset-xform}
-\caption{Pseudocode for two functions that implement the offset transform.}
-\label{fig:xform-sig}
-\end{figure}
-
-\paragraph{Sizes.}
-A similar transformation handles size fields. Just like the offset transform, it takes two parameters, a
-stream and a dependent field, but instead of returning the suffix of the current stream after an
-offset, it returns a slice of the given size from the current stream starting at its current
-position. When generating, it appends the fragment stream to the current stream and writes the size
-of the fragment to the dependent field.
-
-\paragraph{Checksums.}
-Checksums can be verified and computed in a transformation that takes a stream and a dependent
-field. In some cases, a checksum is calculated over a buffer that contains the checksum itself, with
-the checksum being set to some particular value.
-
-\paragraph{}
-A real-world example with many different transforms, used to support
-the ZIP file format, is described in our full paper.%\S\ref{s:eval-formats}.
 

### 2014-11-25 paper/login/eval.tex
Login

@@
 The grammar corresponds almost directly to the diagrams in RFC 1035,
-which defines DNS~\cite[\S4]{RFC:1035}. Nail's dependent fields handle DNS's count fields, and
+which defines DNS. Nail's dependent fields handle DNS's count fields, and
 transformations represent label compression. Both of these features are awkward at best to handle
@@
 
-\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format~\cite{pkzip}.
+\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
+by PKWare.
 ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an
@@
 length in the local directory is used to extract without checking that
-they are equal first, as was the case in the Python ZIP
-library~\cite{python-bug:20078}.
+they are equal first, as occured in Python bug 20078.
 
@@
 We ran the DNS fuzzer provided with the Metasploit
-framework~\cite{mspframework} on our DNS server, which sent randomly
+framework on our DNS server, which sent randomly
 corrupted DNS queries to our server for 4 hours, during which it did
@@
 performance, we compare the performance of our DNS server to that of ISC
-BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
+BIND 9 release 9.9.5, a mature and widely used DNS server.
 We simulate a load resembling that of an authoritative name server, generating a random zone file

### 2014-11-25 paper/login/intro.tex
Login

@@
 The libpng image decompression library had 24 remotely
-exploitable vulnerabilities from 2007 to 2013~\cite{cvedetails:libpng},
+exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
 Adobe's PDF and Flash viewers have been notoriously plagued by input
 processing vulnerabilities, and even the zlib compression library had
-input processing vulnerabilities in the past~\cite{cvedetails:zlib}.
+input processing vulnerabilities in the past.
 Most of these attacks involve memory corruption - therefore, it is easy to assume that 
@@
 construct outputs. Some parser generators, such as 
-Boost.Spirit~\cite{boost-spirit} allow generating output, but require you to write another set of semantic actions.
+Boost.Spirit  allow generating output, but require you to write another set of semantic actions.
 

### 2014-11-25 paper/osdi14/code/zip.nail.tex
Poster stuff?

@@
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
-\PY{n}{zip} \PY{o}{=} \PY{p}{\PYZob{}}  \PY{c+cm}{/*cut for brevity*/}
- \PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{,} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{transform} \PY{n}{zip\PYZus{}eod} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current}\PY{p}{)}
- \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
+\PY{n}{zip} \PY{o}{=} \PY{p}{\PYZob{}}   \PY{c+cm}{/*simplified and cut for brevity*/}
+    \PY{c+cm}{/* Call zip\PYZus{}eod transform to isolate end\PYZus{}directory and contents streams*/}
+     \PY{err}{\PYZdl{}}\PY{n}{contents}\PY{p}{,} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{transform} \PY{n}{zip\PYZus{}eod} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current}\PY{p}{)}
+     \PY{c+cm}{/* Parse end\PYZus{}directory stream. The end\PYZus{}of\PYZus{}directory parser takes the }
+\PY{c+cm}{      content stream as a parameter */}
+     \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{end\PYZus{}directory} \PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{contents}\PY{p}{)}
 \PY{p}{\PYZcb{}}
-\PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}} 
-     \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x06054b50}
-     \PY{n}{disks} \PY{n}{uint16} \PY{o}{|} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
+\PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}} \PY{c+cm}{/*Grammar rules can have parameters*/}
+     \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x06054b50} \PY{c+cm}{/* constant*/}
+     \PY{n}{disks} \PY{n}{uint16} \PY{o}{|} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}  \PY{c+cm}{/* constraint*/}
      \PY{n}{directory\PYZus{}disk} \PY{n}{uint16} \PY{o}{|} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
-     \PY{err}{@}\PY{n}{this\PYZus{}records} \PY{n}{uint16}
+     \PY{err}{@}\PY{n}{this\PYZus{}records} \PY{n}{uint16} \PY{c+cm}{/*dependent field*/}
      \PY{err}{@}\PY{n}{total\PYZus{}records} \PY{n}{uint16} 
+     \PY{c+cm}{/*The following transform ensures these two fields are always equal*/}
      \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{this\PYZus{}records} \PY{err}{@}\PY{n}{total\PYZus{}records}\PY{p}{)}
-     \PY{err}{@}\PY{n}{directory\PYZus{}size} \PY{n}{uint32} 
-     \PY{err}{@}\PY{n}{directory\PYZus{}start} \PY{n}{uint32}
+     \PY{err}{@}\PY{n}{directory\PYZus{}size} \PY{n}{uint32} \PY{c+cm}{/*These two dependent fields are used by the  */}
+     \PY{err}{@}\PY{n}{directory\PYZus{}start} \PY{n}{uint32}\PY{c+cm}{/* transformations below to find the directory*/}
+     \PY{c+cm}{/*dirstr1 is the the suffic of filestream starting @directory\PYZus{}start*/}
      \PY{err}{\PYZdl{}}\PY{n}{dirstr1} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{directory\PYZus{}start}\PY{p}{)} 
+     \PY{c+cm}{/*Another stream with @directory\PYZus{}size bytes starting at that offset*/}
      \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{transform} \PY{n}{size\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{dirstr1} \PY{err}{@}\PY{n}{directory\PYZus{}size}\PY{p}{)}
-     \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint16}
-     \PY{n}{comment} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint8}
-     \PY{n}{files} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{total\PYZus{}directory\PYZus{}records} 
-                 \PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)}
+     \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint16} \PY{c+cm}{/*Dependent field used with the built in n\PYZus{}of*/}
+     \PY{n}{comment} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{comment\PYZus{}length} \PY{n}{uint8} \PY{c+cm}{/* Variable\PYZhy{}length comment*/}
+     \PY{n}{files} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{directory\PYZus{}stream} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{total\PYZus{}records} 
+        \PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{c+cm}{/*Array of directory entries*/}
 \PY{p}{\PYZcb{}}
 \PY{n}{dir\PYZus{}fileheader}\PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}} 
-     \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x02014b50}\PY{c+cm}{/*...*/}
-     \PY{err}{@}\PY{n}{compressed\PYZus{}size} \PY{n}{uint16} \PY{c+cm}{/*...*/}
-     \PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{c+cm}{/*...*/}
-     \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint16}\PY{c+cm}{/*...*/}
+     \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x02014b50}
+     \PY{err}{@}\PY{n}{compressed\PYZus{}size} \PY{n}{uint16} 
+     \PY{err}{@}\PY{n}{crc32} \PY{n}{uint32} 
+     \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint16}
      \PY{err}{@}\PY{n}{off} \PY{n}{uint32}
-     \PY{n}{filename} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint8}\PY{c+cm}{/*...*/}
+     \PY{n}{filename} \PY{n}{n\PYZus{}of} \PY{err}{@}\PY{n}{file\PYZus{}name\PYZus{}len} \PY{n}{uint8}
      \PY{err}{\PYZdl{}}\PY{n}{cstream} \PY{n}{transform} \PY{n}{offset\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{filestream} \PY{err}{@}\PY{n}{off}\PY{p}{)}
-     \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{cstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compressed\PYZus{}size}\PY{c+cm}{/*...*/}\PY{p}{)}
-\PY{p}{\PYZcb{}}
-\PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32} \PY{n}{uint32}\PY{p}{,}\PY{err}{@}\PY{n}{size} \PY{n}{uint32}\PY{c+cm}{/*...*/}\PY{p}{)} \PY{o}{=} \PY{p}{\PYZob{}}
-     \PY{n}{uint32} \PY{o}{=} \PY{l+m+mh}{0x04034b50}\PY{c+cm}{/*...*/}
-     \PY{err}{@}\PY{n}{crc\PYZus{}local} \PY{n}{uint32}\PY{c+cm}{/*...*/}
-     \PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{n}{transform} \PY{n}{size\PYZus{}u32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{current} \PY{err}{@}\PY{n}{size}\PY{p}{)}
-     \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{transform} \PY{n}{zip\PYZus{}compression} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{compressed} \PY{err}{@}\PY{n}{size}\PY{c+cm}{/*...*/}\PY{p}{)}
-     \PY{n}{transform} \PY{n}{crc\PYZus{}32} \PY{p}{(}\PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{err}{@}\PY{n}{crc32}\PY{p}{)}
-     \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{uncompressed} \PY{n}{many} \PY{n}{uint8}
-     \PY{n}{transform} \PY{n}{uint16\PYZus{}depend} \PY{p}{(}\PY{err}{@}\PY{n}{crc\PYZus{}local} \PY{err}{@}\PY{n}{crc32}\PY{p}{)}\PY{c+cm}{/*...*/}
+     \PY{n}{contents} \PY{n}{apply} \PY{err}{\PYZdl{}}\PY{n}{cstream} \PY{n}{fileentry}\PY{p}{(}\PY{err}{@}\PY{n}{crc32}\PY{p}{,}\PY{err}{@}\PY{n}{compressed\PYZus{}size}\PY{p}{)}
 \PY{p}{\PYZcb{}}

### 2014-11-25 paper/osdi14/poster.tex
Poster stuff?

--- 2014-09-17 paper/osdi14/poster.tex
@@
 \column{.5}
+
 \block{Parsing vulnerabilities}{
-Current software typically uses hand-implemented parsers, which are not only prone to bugs, but two
-separate parser implementations can disagree about the meaning of an input. This has been the source
-of many recent high-profile vulnerabilities:
+\begin{itemize}
+\item Hand-written parsers introduce bugs.
+\item Different parsers not equivalent
 \begin{itemize}
@@
 \end{itemize}
+\end{itemize}
 
-{\bf Case study of ZIP vulnerabilities in the CVE database:}
+\innerblock{ZIP: A tricky file format}{
+  \includegraphics[width=0.43\columnwidth]{ZipHorizontal.png}
+}
+
+\innerblock{Case study of ZIP vulnerabilities in the CVE database:}{
 
@@
 }
+}
 
-\block{State of the art}{   
-Grammar-based parsers, such as \texttt{bison} or Hammer, are compact, reuseable
-and safer. However, traditional parsers are inconvenient to use, because the programmer has to write all the
+\block{Existing parsers}{   
+Automatically generated parsers, such as  \texttt{bison} or Hammer are
+\begin{itemize}
+\item Immune to classes of bugs (e.g. memory corruption)\
+\item Easier to use
+\item Re-useable, eliminating parser ambiguities
+\end{itemize} 
+However, traditional parsers are inconvenient to use, because the programmer has to write all the
 yellow-colored parts:
@@
 
+}
+\block{Nail}{
 Nail grammars describe both the format and a data type to represent it, an idea introduced by data
@@
 
-However, existing solutions cannot handle complicated file formats like ZIP, which feature
-duplicated data, offsets and format-specific complications, such as a variable-length end-of-file header:
-
-  \includegraphics[width=0.4\columnwidth]{ZipHorizontal.png}
-}
-\block{Design Goals}{
+\innerblock{Dependent Fields}{
 \begin{itemize}
-    \item Reduce programmer effort
+\item Redundant Information in protocol
 \begin{itemize}
-      \item Semantic combinators describes data type and format.
-      \item Generate output. Semantic bijection, preserves meaning, but can change representation.
+\item Length field
+\item Repeated information
 \end{itemize}
-     \item Handle complex formats
+\item Confuses application, e.g.: 
 \begin{itemize}
-       \item Dependency Fields: hide redundant and structural information.
-       \item Transforms: Extensibility mechanism to capture many features without complicating the
-         grammar language.
+\item Android master key: Uses local name length for extracting, directory for signature
+\item PKI layer cake: Mix implicit (terminator) string length and length field.
 \end{itemize}
 \end{itemize}
+Dependent Fields \textbf{hide redundant information}and automatically
+verify and generate it.
 }
-\block{Dependent Fields}{
-Duplicated data and layout information can cause parser consistency bugs. Nail represents them with
-Dependent Fields, which are automatically verified during parsing and created during 
-generation. 
+\innerblock{Streams and Transformations}{
+\begin{itemize}
+ \item Existing parsers are linear, consuming input front to back.
+\item  Nail grammars feature multiple streams.
+\item \emph{Transformations} modify and create streams.
+\item Pair of functions operating on streams and dependent fields.
+\item Standard library of Transformations enough for many formats.
+\item Programmer can write their own for complex formats.
+\end{itemize}
 }
-\block{Streams and Transformations}{
-  Existing parsers are linear, consuming input front to back. Complicated formats require inelegant hacks. 
-  Nail grammars feature multiple streams and introduce \emph{Transformations} to create
-  them. Transformations are a pair of functions operating on streams and dependent fields.
-  A standard library of Transformations is provided, but  programmers can extend
-  Nail with custom transformations, which are security critical.
-
-%  For example, in the ZIP grammar above, the programmer has to write two functions to find and
-%  generate the end-of-file header, with the following prototype:
-% \input{code/dns-xform}
 }
+% \block{Design Goals}{
+%   \begin{itemize}
+%     \item Reduce programmer effort
+%   \begin{itemize}
+%       \item Semantic combinators describes data type and format.
+%       \item Generate output. Semantic bijection, preserves meaning, but can change representation.
+%   \end{itemize}
+%      \item Handle complex formats
+%   \begin{itemize}
+%        \item Dependency Fields: hide redundant and structural information.
+%        \item Transforms: Extensibility mechanism to capture many features without complicating the
+%          grammar language.
+%        \end{itemize}
+%   \end{itemize}
+%}
 % \block{Semantic Bijection}{
@@
 \column{.5}
-\block{Nail Grammar for ZIP}{
-{\small
-\input{code/zip.nail}}
-An excerpt of Nail's output for this example:
+\block{Nail Grammar for .ZIP }{
+\innerblock{Nail Grammar}{
+\small
+\input{code/zip.nail}
+}
+\innerblock{Excerpt of the generated API}{
 \input{code/zip-out}
 }
+}
 \block{Evaluation}{
-\begin{tabular}{p{0.43\linewidth}p{0.57\linewidth}}
-\begin{minipage}[t]{0.43\linewidth}
-\vspace{0pt}\begin{tabular}{lrr@{~}l}
+\begin{tabular}{lllllll}
 \toprule
-\multicolumn{4}{l}{\textbf{Comparing total lines of code}}\\
+\textbf{Protocol}&\textbf{Grammar} & \textbf{Transform}& \textbf{Application} & \textbf{App LoC} &
+\textbf{Total LoC } &
+\textbf{Alternative}\\
+\midrule
+DNS & 48&64 & Server & 186 & 298 & 683 (Hammer) \\
+\multicolumn{3}{l}{} & Resolver &  97 & 209 & >3000 (libc) \\
+ZIP & 92&78 & Extractor & 49 & 219 & 1600(Info-ZIP)\\
+Ethernet & 16 & 0 & \multicolumn{4}{l}{}\\
+ARP & 10 & 0 & \multicolumn{4}{l}{}\\
+IP & 25 & 0 & \multicolumn{4}{l}{}\\
+UDP & 7 & 0 & \multicolumn{4}{l}{}\\
+ICMP & 5 & 0 & \multicolumn{4}{l}{}\\
 
-\midrule
-\textbf{Application}
-  & \textbf{Nail}
-  & \multicolumn{2}{c}{\textbf{Alternative}} \\
-\midrule
-DNS server
-  & 295
-  & 683
-  & (Hammer) \\
-
-%  &
-%  & 10,000
-%  & (DJBDNS) \\
-
-\texttt{unzip}
-  & 220
-  & 1,600
-  & (Info-Zip) \\
-\bottomrule
-\end{tabular}
-\end{minipage}&
-\vspace{0pt}\begin{minipage}{0.57\linewidth}
-
-\begin{tabular}[t]{lrl}
-\toprule
-\textbf{Protocol} & \textbf{LoC} & \textbf{Challenging features} \\ 
-\midrule
-DNS packets & 48+64 & Label compression,\\
-  & & count fields \\
-ZIP archives & 92+78 & Checksums, offsets, \\ 
-  & & variable length trailer, \\
-  & & compression \\
-Ethernet  & 16+0\phantom{0} & --- \\
-ARP       & 10+0\phantom{0} & --- \\
-IP        & 25+0\phantom{0} & Total length field, options \\
-UDP       &  7+0\phantom{0} & Checksum, length field \\
-ICMP      &  5+0\phantom{0} & Checksum \\
-\bottomrule
-\end{tabular}
-\end{minipage}
 \end{tabular}
 
-Nail and our example applications are available at \url{https://github.com/jbangert/nail}.
+Nail and our example applications are available at
+
+\url{https://github.com/jbangert/nail}.
 }

### 2014-11-25 paper/osdi14/code/zip-out.c

@@ -17,7 +17,7 @@ struct dir_fileheader{
   fileentry contents;
 };/*..*/
 //The programmer calls these generated functions.
-int gen_zip(NailArena *tmp_arena,NailStream *out,zip * val);
+int gen_zip(NailArena *tmp_arena,NailStream *out,zip *val);
 zip*parse_zip(NailArena *arena, 
               const uint8_t *data, size_t size);
 //The programmers implements these two transformation

### 2014-11-25 paper/osdi14/code/zip.nail.c

@@ -1,39 +1,36 @@
-zip = {  /*cut for brevity*/
- $filestream, $end_directory transform zip_eod ($current)
- contents apply $end_directory end_of_directory($filestream)
+zip = {   /*simplified and cut for brevity*/
+    /* Call zip_eod transform to isolate end_directory and contents streams*/
+     $contents, $end_directory transform zip_eod ($current)
+     /* Parse end_directory stream. The end_of_directory parser takes the 
+      content stream as a parameter */
+     contents apply $end_directory end_of_directory($contents)
 }
-end_of_directory($filestream) = { 
-     uint32 = 0x06054b50
-     disks uint16 | [0]
+end_of_directory($filestream) = { /*Grammar rules can have parameters*/
+     uint32 = 0x06054b50 /* constant*/
+     disks uint16 | [0]  /* constraint*/
      directory_disk uint16 | [0]
-     @this_records uint16
+     @this_records uint16 /*dependent field*/
      @total_records uint16 
+     /*The following transform ensures these two fields are always equal*/
      transform uint16_depend (@this_records @total_records)
-     @directory_size uint32 
-     @directory_start uint32
+     @directory_size uint32 /*These two dependent fields are used by the  */
+     @directory_start uint32/* transformations below to find the directory*/
+     /*dirstr1 is the the suffic of filestream starting @directory_start*/
      $dirstr1 transform offset_u32 ($filestream @directory_start) 
+     /*Another stream with @directory_size bytes starting at that offset*/
      $directory_stream transform size_u32 ($dirstr1 @directory_size)
-     @comment_length uint16
-     comment n_of @comment_length uint8
-     files apply $directory_stream n_of @total_directory_records 
-                 dir_fileheader($filestream)
+     @comment_length uint16 /*Dependent field used with the built in n_of*/
+     comment n_of @comment_length uint8 /* Variable-length comment*/
+     files apply $directory_stream n_of @total_records 
+        dir_fileheader($filestream) /*Array of directory entries*/
 }
 dir_fileheader($filestream) = { 
-     uint32 = 0x02014b50/*...*/
-     @compressed_size uint16 /*...*/
-     @crc32 uint32/*...*/
-     @file_name_len uint16/*...*/
+     uint32 = 0x02014b50
+     @compressed_size uint16 
+     @crc32 uint32 
+     @file_name_len uint16
      @off uint32
[…]

### 2014-11-25 paper/osdi14/code/zip-out.tex
Missing files

\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`\$=3\catcode`\^=7\catcode`\_=8}]
\PY{k}{struct} \PY{n}{zip}\PY{p}{\PYZob{}}
  \PY{n}{end\PYZus{}of\PYZus{}directory} \PY{n}{contents}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{k}{struct} \PY{n}{end\PYZus{}of\PYZus{}directory}\PY{p}{\PYZob{}}
  \PY{k+kt}{uint16\PYZus{}t} \PY{n}{disks}\PY{p}{;}
  \PY{k+kt}{uint16\PYZus{}t} \PY{n}{directory\PYZus{}disk}\PY{p}{;}\PY{c+cm}{/*...*/}
  \PY{k}{struct} \PY{p}{\PYZob{}}
    \PY{n}{dir\PYZus{}fileheader}\PY{o}{*} \PY{n}{elem}\PY{p}{;}
    \PY{k+kt}{size\PYZus{}t} \PY{n}{count}\PY{p}{;}
  \PY{p}{\PYZcb{}} \PY{n}{files}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{k}{struct} \PY{n}{dir\PYZus{}fileheader}\PY{p}{\PYZob{}}
  \PY{k}{struct} \PY{p}{\PYZob{}}
    \PY{k+kt}{uint8\PYZus{}t}\PY{o}{*}\PY{n}{elem}\PY{p}{;}
    \PY{k+kt}{size\PYZus{}t} \PY{n}{count}\PY{p}{;}
  \PY{p}{\PYZcb{}} \PY{n}{filename}\PY{p}{;}
  \PY{n}{fileentry} \PY{n}{contents}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;}\PY{c+cm}{/*..*/}
\PY{c+c1}{//The programmer calls these generated functions.}
\PY{k+kt}{int} \PY{n+nf}{gen\PYZus{}zip}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp\PYZus{}arena}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{out}\PY{p}{,}\PY{n}{zip} \PY{o}{*}\PY{n}{val}\PY{p}{)}\PY{p}{;}
\PY{n}{zip}\PY{o}{*}\PY{n+nf}{parse\PYZus{}zip}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{arena}\PY{p}{,} 
              \PY{k}{const} \PY{k+kt}{uint8\PYZus{}t} \PY{o}{*}\PY{n}{data}\PY{p}{,} \PY{k+kt}{size\PYZus{}t} \PY{n}{size}\PY{p}{)}\PY{p}{;}
\PY{c+c1}{//The programmers implements these two transformation}
\PY{c+c1}{//functions and two similar ones for output.}
\PY{k}{extern} \PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}eod\PYZus{}parse}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,}
 \PY{n}{NailStream} \PY{o}{*}\PY{n}{filestream}\PY{p}{,} \PY{n}{NailStream} \PY{o}{*}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{current}\PY{p}{)}\PY{p}{;}
\PY{k}{extern} \PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}compression\PYZus{}parse}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,}
 \PY{n}{NailStream} \PY{o}{*}\PY{n}{uncomp}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{compressed}\PY{p}{,}\PY{k+kt}{uint32\PYZus{}t}\PY{o}{*} \PY{n}{size}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

### 2014-11-25 paper/login/design.tex
More typowrangling

@@
 
-We will use DNS packets as a leading example, as defined in RFC 1035.  Each DNS packet consists of a
+As a motivating example, we will parse DNS packets, as defined in RFC 1035. Each DNS packet consists of a
 header, a set of question records, and a set of answer records. Domain
@@
 \input{fig-dnsfull}
-Let's step through a simplified Nail grammar DNS
+Let's step through a simplified Nail grammar for DNS
 packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations

### 2014-11-25 paper/login/intro.tex
More typowrangling

@@
 Furthermore, some parser generators are cumbersome to use when parsing binary data for several
-reasons. First, many parser generators don't directly data structures, but call semantic actions
-that the you write to construct a data structure. Therefore,you must
+reasons. First, many parser generators don't produce convenient data structures, but call
+\emph{semantic actions} that you have to write to build up a data structure your program can use. Therefore,you must
 describe the format up to three times -- in the grammar, the data structure and the semantic
-actions. Second, most parser generators only address parsing inputs, so you'd have to manually
+actions. Second, most parser generators only address parsing inputs, so you have to manually
 construct outputs. Some parser generators, such as 
@@
 Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
-Second, Nail will also take care of generating output from the same data type, without requiring you
-to write more risky code or think about the format again.
+Second, Nail will also generate output from this data type, without requiring you
+to write more risky code or giving you a chance to introduce ambiguity.
 
@@
 which specify the position or length of another data structure, fall
-in this category.  Transformations allow you to escape the
-generated code to modify the raw data and interact with dependent fields
-in a controlled manner.
+in this category.  Transformations allow you to write plug-ins, allowing your programs to handle complicated structures, while
+keeping Nail itself small, yet flexible.
 

### 2014-11-25 paper/login/eval.tex
pass

@@
 
-First, we ask whether Nail can handle real-world data formats and whether its techniques are
-critical to that. We used Nail to implement grammars
+First, we'll demonstrate how Nail's features help process real-world protocols. We used Nail to implement grammars
 for seven protocols with a range of challenging features.
@@
 generators would be unable to fully handle 5 out of the 7 data formats.
-In the rest of this subsection, we describe the DNS and Zip grammars in
-more detail, focusing on how Nail's features enable us to support these
-formats.
 
@@
 \paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
-by PKWare.
-ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an
+by PKWare. At the end of each ZIP file is an
 \emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
@@
 contains the magic number~\cite{wolf:berlinsides-zip}.
-
 This header  contains points to the \emph{ZIP directory}, contains an entry every file
@@
 Duplicating information made sense when ZIP files were stored on floppy
-disks with slow seek times and high fault rates, and memory constraints
-made it impossible to keep the ZIP directory in memory.
+disks with slow seek times and high fault rates, but nowadays lead to parsers being confused, such
+as in the recent Android Master Key bug.
 
- However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
-the length in the central directory is used to allocate memory and the
-length in the local directory is used to extract without checking that
-they are equal first, as occured in Python bug 20078.
+%  However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
+% the length in the central directory is used to allocate memory and the
+% length in the local directory is used to extract without checking that
+% they are equal first, as occured in Python bug 20078.
 
@@
 files transparently with transformations, and the programmer
-is presented with a data structure containing the archive contents. To handle formats based on ZIP,
-such as office documents, our ZIP grammar could be extended to apply additional grammars to the
-files within an archive.
+is presented with a data structure containing the archive contents. Nail even allows parsing the
+contents of an archive file -- allowing formats based on ZIP, such as Microsoft Office documents, to
+be handled with one grammar.
 
@@
 Their DNS server does not implement label compression, zone files, etc.
-From this, we conclude that Nail leads to much more compact code for
-dealing with DNS packet formats.
-
-We ran the DNS fuzzer provided with the Metasploit
-framework on our DNS server, which sent randomly
-corrupted DNS queries to our server for 4 hours, during which it did
-not crash or trigger the stack or heap corruption detector.
 
@@
 11 of these vulnerabilities involved memory corruption during input handling, which generated code
-is immune to by design.
+is immune to by design. We also fuzz-tested our DNS server.
 More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.

### 2014-11-26 paper/login/eval.tex
Nail article

@@
-\section{Evaluation}
+\section{Using Nail}
 \label{s:eval}
@@
 \label{s:eval-formats}
-
-First, we'll demonstrate how Nail's features help process real-world protocols. We used Nail to implement grammars
+ We used Nail to implement grammars
 for seven protocols with a range of challenging features.
 Figure~\ref{fig:eval-protocols} summarizes our results.
-We find that despite the challenging aspects of these protocols, Nail is
-able to capture the protocols, by relying on its novel features: dependent
+Despite the challenging aspects of these protocols, Nail is
+able to capture them by relying on its novel features: dependent
 fields, streams, and transforms.  In contrast, state-of-the-art parser
@@
 
-\paragraph{DNS} In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
-for DNS packets, shown in Figure~\ref{fig:dns-full}.
-The grammar corresponds almost directly to the diagrams in RFC 1035,
+\paragraph{DNS} Previously, we used a grammar for DNS packets shown in Figure~\ref{fig:dns-full} to
+show how to write Nail grammars.
+This example grammar corresponds almost directly to the diagrams in RFC 1035,
 which defines DNS. Nail's dependent fields handle DNS's count fields, and
@@
 
-Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
-files transparently with transformations, and the programmer
-is presented with a data structure containing the archive contents. Nail even allows parsing the
+Nail captures these redundancies with dependent fields, eliminating the ambiguities. It also  decompresses
+archive contents transparently with transformations, which allows parsing the
 contents of an archive file -- allowing formats based on ZIP, such as Microsoft Office documents, to

### 2014-11-26 paper/login/intro.tex
Nail article

@@
+\begin{quote}
+Julian Bangert is a second year Ph.D. student working on computer security at MIT. When he is not
+building parsers for complicated formats, he is interested is building exploit mitigation
+techniques, side-channel resistant cryptography and finding turing-complete weird machines in
+unexpected places, such as your processors virtual memory system.
+
+Nickolai Zeldovich is an Associate Professor at MIT's department of Electrical Engineering and Computer Science, and a member of the Computer Science and Artificial Intelligence Laboratory.
+His research interests are in building practical secure systems, from operating systems and hardware to programming languages and security analysis tools.
+He received his PhD from Stanford University in 2008, where he developed HiStar, an operating system designed to minimize the amount of trusted code by controlling information flow.
+In 2005, he co-founded MokaFive, a company focused on improving desktop management and mobility using x86 virtualization.
+Prof.
+Zeldovich received a Sloan fellowship (2010), an NSF CAREER award (2011), the MIT EECS Spira teaching award (2013), and the MIT Edgerton faculty achievement award (2014).
+\end{quote}
 \begin{quote}

### 2014-11-26 paper/login/design.tex
pass
typos

@@
 When parsing input, this will call \cc{zipdir\_parse}, which takes \cc{\$current} -- an implicit
-identifier always refering to the stream currently being handled -- and returns \cc{\$files} and
+identifier always referring to the stream currently being handled -- and returns \cc{\$files} and
 \cc{\$header}. When generating output, this will call \cc{zipdir\_generate}, which appends
@@
 generated to that stream. The \cc{dir} rule in turn describes the structure of the directory -- a
-magic number and a count field, followed by that many file descripors. Each file descriptor, is then parsed with two
+magic number and a count field, followed by that many file descriptors. Each file descriptor, is then parsed with two
 transformations, the standard-library \cc{slice} which describes an offset and a size within another
@@
    vulnerabilities.
-
-% For example, we imagine the following grammar could be used to represent
-% a sequence of bytes \texttt{data} followed by its CRC32 checksum:
-
-% \begin{verbatim}
-% data many uint8; @checksum uint32
-% raw_depend @checksum data crc32
-% \end{verbatim}
-
-% \noindent
-% where \texttt{crc32} is a function supplied by the application, with
-% the following signature:
-
-% \begin{verbatim}
-% bool crc32(uint32_t *out, uint8_t *in);
-% \end{verbatim}
-
-% Because this feature compromises Nail's security guarantees, it should
-% only be used in limited circumstances and with carefully prepared checksum
-% functions.  This feature is not implemented in the current prototype.

### 2014-12-03 paper/login/design.tex
pass through 2.1
one \XXX where i have no idea what the sentence meant

--- 2014-11-25 paper/login/design.tex
@@
 In this section, we will explain how to handle basic data formats in Nail, how to handle
-redundancies in the format with dependent fields and how Nail parsers can be extended with
+redundancies in the format with dependent fields, and how Nail parsers can be extended with
 transformations.
@@
 \label{s:design}
-\subsection{Basic Data formats}
+\subsection{Basic data formats}
 
@@
 generator functions shown in Figure~\ref{fig:dns-full-api}.
-Nail grammars are reuseable between applications and we will use this grammar to implement
-both a DNS server and client, which previously would have had two separate hand-written parsers,
-leading to bugs such as  the Android Master Key.
+Nail grammars are reusable between applications and we will use this grammar to implement
+both a DNS server and a DNS client, which previously would have had two separate hand-written parsers,
+leading to bugs such as the Android ZIP file vulnerability.
 
-A Nail grammar file consists of rule definitions, for example l. 1-20,  which assign a name
-(\texttt{dnspacket}) to a grammar production (2-20). If you are not familiar with
+A Nail grammar file consists of rule definitions, for example lines 1--20,  which assign a name
+(\texttt{dnspacket}) to a grammar production (lines 2--20). If you are not familiar with
 other parsers, you can imagine rules as C type declarations on steroids
@@
 elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
-DNS. A more complete reference can be found in~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.
+DNS. A more complete reference can be found in Figure~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.
 
-\paragraph{Integers and Constraints}.
+\paragraph{Integers and Constraints.}
 Because Nail is designed to cope with binary formats,
-it handles not only common integer types (such as \texttt{uint16} on l. 16), but bit fields of any
+it handles not only common integer types (such as \texttt{uint16} on line 16), but bit fields of any
 length, such as \texttt{uint1}. These integers are exposed to the programmer as an appropriately
 sized machine integer, e.g. \texttt{uint8\_t}. Nail also supports constraints on integer values,
-limiting the values to either a range (l. 23, \texttt{|1..16}), which can optionally be half-open,
-or a fixed set (l. 24, \texttt{|[1,255]}). Both types of constraint can be combined, e.g. \texttt{|
-  [1..16,255]}. Constant values are also supported, e.g. l. 10: \texttt{uint3=0} represents
+limiting the values to either a range (line 23, \texttt{|1..16}), which can optionally be half-open,
+or a fixed set (line 24, \texttt{|[1,255]}). Both types of constraint can be combined, e.g. \texttt{|
+  [1..16,255]}. Constant values are also supported; for instance, on line 10, \texttt{uint3=0} represents
 three reserved bits that must be 0. Because constant values carry no information, they are not
@@
 \cc{many}, which keeps repeating the \cc{label} rule until it fails.  In the next section, we will
-explain how to handle count fields, and our full paper describes how to handle various array
+explain how to handle count fields, and our full paper~\cite{bangert:nail-osdi14} describes how to handle various array
 representations (such as delimiters or non-empty arrays).

### 2014-12-03 paper/login/intro.tex
pass through 2.1
one \XXX where i have no idea what the sentence meant

--- 2014-11-25 paper/login/intro.tex
@@
 input processing vulnerabilities in the past.
-Most of these attacks involve memory corruption - therefore, it is easy to assume that 
+Most of these attacks involve memory corruption; therefore, it is easy to assume that 
 solving memory corruption will end all our woes when handling untrusted inputs.
@@
 However, as memory-safe languages and exploit mitigation  tricks are becoming more prevalent,
-attackers are moving to a new class of attack - parser differentials. Many applications use
+attackers are moving to a new class of attack called {\em parser differentials}. Many applications use
 hand-written input processing code, which is often mixed with the rest of the application, e.g. by
@@
 files will be extracted to the filesystem, bypassing Android's code signing. Similar issues showed
-up on iOS \cite{geohot-evasion} and SSL \cite{DBLP:conf/fc/KaminskyPS10}.
+up on iOS~\cite{geohot-evasion} and SSL~\cite{DBLP:conf/fc/KaminskyPS10}.
 
@@
 However, if you wanted to use existing tools to parse PDF
-or ZIP, you would soon find that they cannot handle the complicated -- and therefore most
-error-prone-- aspects of such formats-- so you'd still have to
+or ZIP files, you would soon find that they cannot handle the complicated---and therefore most
+error-prone---aspects of such formats, so you'd still have to
 hand-write the riskiest bits of code.
-For example, existing parser generators cannot conveniently represent size or offset fields and more complex features, such as data compression or checksums cannot be expressed at all.
+For example, existing parser generators cannot conveniently represent size or offset fields, and more complex features, such as data compression or checksums, cannot be expressed at all.
 
 Furthermore, some parser generators are cumbersome to use when parsing binary data for several
-reasons. First, many parser generators don't directly data structures, but call semantic actions
-that the you write to construct a data structure. Therefore,you must
-describe the format up to three times -- in the grammar, the data structure and the semantic
-actions. Second, most parser generators only address parsing inputs, so you'd have to manually
+reasons. First, many parser generators don't produce convenient data structures, and instead call
+but call \emph{semantic actions} that the developer has to write to build up a data structure that
+the rest of the program can use. Therefore, the developer must
+describe the format up to three times: in the grammar, in the data structure, and in the semantic
+actions. Second, most parser generators only address parsing inputs, so the developer would have to manually
 construct outputs. Some parser generators, such as 
-Boost.Spirit  allow generating output, but require you to write another set of semantic actions.
+Boost.Spirit, allow generating output, but require the developer to write another set of semantic actions.
 
 We address these challenges with Nail, a new parser generator for binary formats.
-First, Nail grammars describe not only a format, but also a data type to represent it within the program.
-Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
-Second, Nail will also take care of generating output from the same data type, without requiring you
-to write more risky code or think about the format again.
+
+First, Nail grammars describe not only a format, but also a data type to represent the format within the program.
+Therefore the developer doesn't have to write semantic actions and type declarations.
+\XXX[what did this mean?  and can no longer syntactic validation and semantic processing.]
+
+Second, Nail will also generate output from this data type, without requiring the developer
+to write more risky code or giving the developer a chance to introduce ambiguity.
 
@@
 which specify the position or length of another data structure, fall
-in this category.  Transformations allow you to escape the
-generated code to modify the raw data and interact with dependent fields
-in a controlled manner.
+in this category.  Transformations allow the developer to write plug-ins,
+allowing programs to handle complicated structures, while
+keeping Nail itself small, yet flexible.
 

### 2014-12-03 paper/login/design.tex
sync n*.bib

@@
 elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
-DNS. A more complete reference can be found in Figure~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.
+DNS. A more complete reference can be found in Figure~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail}.
 
@@
 \cc{many}, which keeps repeating the \cc{label} rule until it fails.  In the next section, we will
-explain how to handle count fields, and our full paper~\cite{bangert:nail-osdi14} describes how to handle various array
+explain how to handle count fields, and our full paper~\cite{bangert:nail} describes how to handle various array
 representations (such as delimiters or non-empty arrays).

### 2014-12-03 paper/login/design.tex
make spell

@@
 elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
-DNS. A more complete reference can be found in Figure~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail}.
+DNS\@. A more complete reference can be found in Figure~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail}.
 
@@
 using angle brackets, that contains one unnamed, non-constant field, which is represented directly
-in the datatype, without introducing another layer of indirection, as shown on line 40.
+in the data type, without introducing another layer of indirection, as shown on line 40.
 
@@
 such as checksums, duplicated information, or offset and  length fields. Exposing such values risks
-inconsistencies that could trick the program into unsafe behaviour. Therefore, we represent such values using \emph{dependent fields} and handle them transparently during
+inconsistencies that could trick the program into unsafe behavior. Therefore, we represent such values using \emph{dependent fields} and handle them transparently during
 parsing and generation without exposing them to the application. 

### 2014-12-03 paper/login/eval.tex
sync n*.bib
make spell

--- 2014-11-25 paper/login/eval.tex
@@
 The grammar corresponds almost directly to the diagrams in RFC 1035,
-which defines DNS. Nail's dependent fields handle DNS's count fields, and
+which defines DNS\@. Nail's dependent fields handle DNS's count fields, and
 transformations represent label compression. Both of these features are awkward at best to handle
@@
 
-\paragraph{DNS.}
+\paragraph{DNS\@.}
 
@@
 queries for one minute against both DNS servers. We found that our DNS server is approximately three
-times faster than BIND. Although BIND is a more sophisticated DNS server,
+times faster than BIND\@. Although BIND is a more sophisticated DNS server,
 and implements many features that are not present in our Nail-based DNS
@@
 a barrier to achieving good performance.
-\paragraph{ZIP.}
+\paragraph{ZIP\@.}
 
@@
 
-In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
+In our full paper~\cite{bangert:nail}, we present a study of 15 ZIP parsing bugs.
 11 of these vulnerabilities involved memory corruption during input handling, which generated code

### 2014-12-03 paper/login/design.tex
pass thru design

@@
 
-\subsection{Redundant Data}
+\subsection{Redundant data}
 \label{s:dependent}
@@
 \cc{ac}, \cc{ns}, and \cc{ar}), which contain the number of questions and answers that follow the
-header, which we represent by dependent fields (line 12-15).
-Dependent fields are defined within a structure like normal fields, but their name starts with an \cc{@} symbol.
+header, which we represent by dependent fields (lines 12--15).
+Dependent fields are defined within a structure like normal fields, but their names start with an \cc{@} symbol.
 A dependent field is in scope and can be referred to by the definition of all subsequent fields in
@@
 
-Nail allows handling count fields with \cc{n\_of}, which parses an exact number of repetitions of a rule. Lines 16-19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
+Nail allows handling count fields with \cc{n\_of}, which parses an exact number of repetitions of a rule. Lines 16--19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
 answer records in a DNS packet.
@@
 PADS~\cite{Fisher:2005:PDL:1064978.1065046} contain not just a kitchen sink, but a kitchen store
-full of features -- and a language that can handle all possible formats will be a general
-purpose programming language.  Instead, we keep Nail itself small, and introduce an
+full of features, and a language that can handle all possible formats will be a general-purpose
+programming language.  Instead, we keep Nail itself small, and introduce an
 interface that allows complicated format structures to be handled by  plug-in \emph{transformations} 
-in a general purpose
+in a general-purpose
 language. Of course, we ship Nail with a handy library of common transformations to handle common
@@
 These \emph{transformations} consume and produce
-\emph{streams} -- sequences of bytes -- which can be further passed to other transformations and
+\emph{streams}---sequences of bytes---which can be further passed to other transformations and
 eventually parsed by a Nail rule. Transformations can also access values in dependent fields. 
 Streams can be subsets of other streams, for example the substream starting at an
-offset given in a dependent field to handle pointer fields, or computed at runtime
-, such as by decompressing another stream with zlib.
+offset given in a dependent field to handle pointer fields, or computed at runtime,
+such as by decompressing another stream with zlib.
 
@@
 is a magic number followed by an array of filenames and pointers to compressed files. A grammar for the real ZIP format, which
-has more layers of indirection, is presented in the full paper.
+has more layers of indirection, is presented in the full paper~\cite{bangert:nail}.
 
- The grammar first calls the zipdir transform on line 2, which finds the magic number and splits the file into
+The grammar first calls the \cc{zipdir} transform on line 2, which finds the magic number and splits the file into
 two streams, one containing the compressed files, the other the directory.  Streams are referred to
 with \cc{\$identifiers}, similar to dependent fields. A C prototype of the \cc{zipdir} transform is
- shown in\ref{fig:zip-transform}. 
+shown in Figure~\ref{fig:zip-transform}.
+
 \begin{figure}[tb]
@@
 \end{figure}
-When parsing input, this will call \cc{zipdir\_parse}, which takes \cc{\$current} -- an implicit
-identifier always referring to the stream currently being handled -- and returns \cc{\$files} and
-\cc{\$header}. When generating output, this will call \cc{zipdir\_generate}, which appends
+When parsing input, line 2 of Figure~\ref{fig:zip-extract} will call \cc{zipdir\_parse}, which takes \cc{\$current} (an implicit
+identifier always referring to the stream currently being handled) and returns \cc{\$files} and
+\cc{\$header}. When generating output, the same line will call \cc{zipdir\_generate}, which appends
 \cc{\$files} and \cc{\$header} to \cc{\$current}.
 
-Line 3 then applies the \cc{dir} rule to  the \cc{\$header} stream, passing it the \cc{\$files}
-stream. Within \cc{dir}, \cc{\$current} is now \cc{\$header} and input is parsed from and output
-generated to that stream. The \cc{dir} rule in turn describes the structure of the directory -- a
-magic number and a count field, followed by that many file descriptors. Each file descriptor, is then parsed with two
+Line 3 of Figure~\ref{fig:zip-extract} then applies the \cc{dir} rule to  the \cc{\$header} stream, passing it the \cc{\$files}
+stream as an argument. Within \cc{dir}, \cc{\$current} is now \cc{\$header} and input is parsed from and output
+generated to that stream. The \cc{dir} rule in turn describes the structure of the directory, which consists of a
+magic number and a count field, followed by that many file descriptors. Each file descriptor is then parsed with two
 transformations, the standard-library \cc{slice} which describes an offset and a size within another
@@
 In a more complicated example, such as an Office document, we could now specify grammars for each
-entry within an archive.
+file contained within an archive.
 

### 2014-12-03 paper/login/eval.tex
done with pass

@@
 
-\subsection{Real-World Formats}
+\subsection{Real-world formats}
 \label{s:eval-formats}
@@
 \begin{figure}[tb]
+\smaller[0.5]
 \centering
@@
 
-\paragraph{DNS} In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
+\paragraph{DNS} In \S\ref{s:design}, we introduced Nail's syntax with a grammar
 for DNS packets, shown in Figure~\ref{fig:dns-full}.
@@
 contains the magic number~\cite{wolf:berlinsides-zip}.
-This header  contains points to the \emph{ZIP directory}, contains an entry every file
+This header  contains pointers to the \emph{ZIP directory}, which contains an entry for every file
 in the archive.  Each entry stores file metadata in addition to the
@@
 most information from the directory entry header and is followed immediately by the compressed
-archive entry.
+file data.
 Duplicating information made sense when ZIP files were stored on floppy
 disks with slow seek times and high fault rates, but nowadays lead to parsers being confused, such
-as in the recent Android Master Key bug.
+as in the Android ZIP file vulnerability described in \S\ref{sec:intro}.
 
@@
 Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
-files transparently with transformations, and the programmer
+files transparently with transformations; the programmer
 is presented with a data structure containing the archive contents. Nail even allows parsing the
-contents of an archive file -- allowing formats based on ZIP, such as Microsoft Office documents, to
+contents of an archive file, allowing formats based on ZIP, such as Microsoft Office documents, to
 be handled with one grammar.
@@
 DNS server and an
-\cc{unzip} program---based on the above grammars, and will compare the effort involved and the
+\cc{unzip} program---based on the above grammars, and compare the effort involved and the
 resulting security to similar applications with hand-written parsers and with other parser generators.
-We will use lines of code as a proxy for programmer effort. To evaluate security, we will argue how
-our design avoids classes of vulnerabilities and fuzz-test one of our applications.
+We use lines of code as a proxy for programmer effort. To evaluate security, we argue how
+our design avoids classes of vulnerabilities and we also fuzz-test one of our applications.
 \begin{figure}[tb]
@@
 the zlib library, and one for finding the end-of-directory header).
- The \cc{unzip} utility contains a file \cc{extract.c}, which parses
+
+As a point of comparison, the existing Linux \cc{unzip} utility contains a file \cc{extract.c}, which parses
 ZIP metadata and calls various decompression routines in other files. This
@@
 In our full paper~\cite{bangert:nail}, we present a study of 15 ZIP parsing bugs.
-11 of these vulnerabilities involved memory corruption during input handling, which generated code
-is immune to by design. We also fuzz-tested our DNS server.
+11 of these vulnerabilities involved memory corruption during input handling,
+to which our generated code is immune by design.
+We also fuzz-tested our DNS server, finding no bugs.
 More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.

### 2014-12-03 paper/login/intro.tex
done with pass

@@
 \section{Introduction}
+\label{sec:intro}
 Binary file formats and network protocols are hard to parse safely: 

### 2014-12-04 paper/login/eval.tex

@@
 
-\paragraph{DNS} In \S\ref{s:design}, we introduced Nail's syntax with a grammar
-for DNS packets, shown in Figure~\ref{fig:dns-full}.
-The grammar corresponds almost directly to the diagrams in RFC 1035,
+\paragraph{DNS} Previously, we used a grammar for DNS packets shown in Figure~\ref{fig:dns-full} to
+show how to write Nail grammars.
+This example grammar corresponds almost directly to the diagrams in RFC 1035,
 which defines DNS\@. Nail's dependent fields handle DNS's count fields, and
@@
 
-Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
-files transparently with transformations; the programmer
-is presented with a data structure containing the archive contents. Nail even allows parsing the
+Nail captures these redundancies with dependent fields, eliminating the ambiguities. It also  decompresses
+archive contents transparently with transformations, which allows parsing the
 contents of an archive file, allowing formats based on ZIP, such as Microsoft Office documents, to

### 2015-05-29 paper/thesis/abstract.tex
Commit

% $Log: abstract.tex,v $
% Revision 1.1  93/05/14  14:56:25  starflt
% Initial revision
% 
% Revision 1.1  90/05/04  10:41:01  lwvanels
% Initial revision
% 
%
%% The text of your abstract and nothing else (other than comments) goes here.
%% It will be single-spaced and the rest of the text that is supposed to go on
%% the abstract page will be generated by the abstractpage environment.  This
%% file should be \input (not \include 'd) from cover.tex.
In this thesis, I designed and implemented a compiler which performs
optimizations that reduce the number of low-level floating point operations
necessary for a specific task; this involves the optimization of chains of
floating point operations as well as the implementation of a ``fixed'' point
data type that allows some floating point operations to simulated with integer
arithmetic.  The source language of the compiler is a subset of C, and the
destination language is assembly language for a micro-floating point CPU.  An
instruction-level simulator of the CPU was written to allow testing of the
code.  A series of test pieces of codes was compiled, both with and without
optimization, to determine how effective these optimizations were.

### 2015-05-29 paper/thesis/propcover.tex
Commit

%% This file is for producing a Doctoral Thesis proposal.  It should be fairly
%% self-explanatory.

\documentclass[a4paper]{article}

\usepackage[doublespacing]{setspace}
\usepackage[margin=1in]{geometry}
\usepackage{url}
\begin{document}

\bibliographystyle{plain}
\pagestyle{empty}
\markboth{{\sc thesis proposal}}{{\sc thesis proposal}}
\def\title{Nail, a practical tool for parsing and generating data formats.}
\def\author{Julian Bangert}
\def\addrone{550 Memorial Drive, 15E3}
\def\addrtwo{Cambridge, MA  02139}

\def\degree{Master of Science}
%% Added \deptname for PhD proposals in other fields.
%% Krishna Sethuraman (1990)
\def\deptname{Electrical Engineering \\ and Computer Science}
\def\laboratory{Computer Science and Artificial Intelligence Laboratory}

\def\submissiondate{\today}
\def\completiondate{September 2015}

\def\supervisor{Professor Nickolai Zeldovich}
\def\supertitleone{Associate Professor }

\def\abstract{ The proposed research is extending grammar-based parser
  generators to handle binary protocols and data formats.  Our
  prototype tool, Nail, introduces several new ideas to achieve that
  goal, including dependent fields and stream transforms to capture
  protocol features such as size and offset fields, checksums, and
  compressed data, which are impractical to express in existing
  protocol languages.  We will show several grammars for complicated
  real-world formats and demonstrate applications built with those. We
  argue that applications built with Nail are more resilient to memory
  corruption attacks as well as less-understood parser differential
  vulnerabilites.  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% You Should Not Need To Modify Anything Below Here %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Cover Page - Author signs %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{singlespace}
\begin{center}
{\Large \bf 
   Massachusetts Institute of Technology
\\ Department of \deptname \\}
\vspace{.25in}
{\Large \bf
   Proposal for Thesis Research in Partial Fulfillment
\\ of the Requirements for the Degree of
\\ \degree \\}
\end{center}

\vspace{.5in}

\def\sig{{\small \sc (Signature of Author)}}

\begin{tabular}{rlc}
   {\small \sc Title:}                       & \multicolumn{2}{l}{\title}
\\ {\small \sc Submitted by:}
                            & \author  & \\
                            & \addrone & \\
& \addrtwo & \\
& \\
& \\
\cline{2-2} & \\
			    & \makebox[2in][c]{\sig} & \\
\\ {\small \sc Date of Submission:}          & \multicolumn{2}{l}{\submissiondate}
\\ {\small \sc Expected Date of Completion:} & \multicolumn{2}{l}{\completiondate}
\\ {\small \sc Laboratory:}                  & \multicolumn{2}{l}{\laboratory}
\end{tabular}

\vspace{.75in}
{\bf \sc Brief Statement of the Problem:}

\abstract

\underline{\bf  Supervision Agreement}

\vspace{.25in}

\vspace{.25in}
The program outlined in this proposal is adequate for a Master's thesis. The supplies and facilities
required are available, and I am willing to supervise the research and evaluate the thesis report.
\vspace{.5in}
\begin{tabular}{crc}
  \hspace{2in} & {\sc Signed:} & \\ \cline{3-3}
               &               & {\small \sc \supervisor} \\
               &               &                             \\
               & {\sc Date:}   & \\ \cline{3-3}
\end{tabular}

\end{singlespace}
\newpage

\section{Motivation}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past.
Most of these attacks involve memory corruption; therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

However, as memory-safe languages and exploit mitigation  tricks are becoming more prevalent,
attackers are moving to a new class of attack called {\em parser differentials}. Many applications use
hand-written input processing code, which is often mixed with the rest of the application, e.g. by
passing a pointer to the raw input through the application. This (anti-)pattern makes it impossible
to figure out whether two implementations of the same format or protocol are identical, and input
handling code can't be easily reused between applications. As a result, different applications often
disagree in the corner cases of a protocol, which can have fatal security consequences. For example,
Android has two parsers for ZIP archives involved in securely installing applications. First, a Java
program checks the signatures of files contained within an app archive and then another tool
extracts them to the filesystem. Because the two ZIP parsers disagree in multiple places, attackers
can modify a valid file so that the verifier will see the original contents, but attacker-controlled
files will be extracted to the filesystem, bypassing Android's code signing. Similar issues showed
up on iOS~\cite{geohot-evasion} and SSL~\cite{DBLP:conf/fc/KaminskyPS10}.

Instead of attempting to parse inputs by hand and failing, 
a promising approach is to specify
a precise grammar for the input data format, and automatically generate parsing code from that with tools like {\tt yacc}. As long
as the parser generator is bug-free, the application will be safe from many
input processing vulnerabilities. Grammars can also be re-used between
applications, further reducing effort and eliminating inconsistencies.

This approach is typical in compiler design and in other applications handling text-based inputs, but not
common for binary inputs. The Hammer framework~\cite{hammer-parser} and data description languages such as
PADS~\cite{Fisher:2005:PDL:1064978.1065046} have been developing generated parsers for binary protocols.

However, if you wanted to use existing tools to parse PDF
or ZIP files, you would soon find that they cannot handle the complicated---and therefore most
error-prone---aspects of such formats, so you'd still have to
hand-write the riskiest bits of code. For example, existing parser generators cannot conveniently represent size or offset fields, and more complex features, such as data compression or checksums, cannot be expressed at all.

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't produce convenient data structures, and instead call
but call \emph{semantic actions} that the developer has to write to build up a data structure that
the rest of the program can use. Therefore, the developer must
describe the format up to three times: in the grammar, in the data structure, and in the semantic
actions. Second, most parser generators only address parsing inputs, so the developer would have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit, allow generating output, but require the developer to write another set of semantic actions. 
\section{Approach}

We address these challenges with Nail, a new parser generator for binary formats:

First, parser generators typically parse inputs into an abstract syntax
tree (AST) that corresponds to the grammar.  In order to produce a data
structure that the rest of the application code can easily process,
application developers must write explicit {\em semantic actions} that
update the application's internal representation of the data based on
each AST node.  
Writing these semantic actions requires the programmer to describe the
structure of the input three
times---once to describe the grammar, once to describe the internal data
structure, and once again in the semantic actions that translate the
grammar into the data structure---leading to potential bugs and
inconsistencies.
Furthermore, in a memory-unsafe language like C, these semantic
actions often involve error-prone manual memory allocation and
pointer manipulation.

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators focus on just parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is work-intensive and leads to
more code that could contain errors.
Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
back into an AST\@.

Third, many data formats contain redundancies, such as repeating
information in multiple structures.  Applications usually do not
explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft an input that
causes inconsistencies~\cite{evaders6}.  Furthermore, security
vulnerabilities can occur when an application assumes two repetitions
of the same data to be consistent, such as allocating a buffer based on
the value of one size field and copying into that buffer based on the
value of another~\cite{python-bug:20078}.

Finally, real-world data formats, such as PNG or PDF, are hard to
represent with existing parser generators.  Those parsers cannot directly
deal with length or checksum fields, so the programmer has to
either write potentially unsafe code to deal with such features, or build
contrived grammar constructs, such as introducing one grammar rule for
each possible value of a length field.  Offset fields, which specify
the position at which some data structure is located, usually require
the programmer to manipulate a parser's internal state to re-position
its input.  More complicated transformations, such as handling compressed
data, cannot be represented at all.

\section{Proposed Evaluation}
To show that Nail is practical to handle real world data formats, we
will demonstrate grammars for several real-world file formats and
protocols that have had security vulnerabilities in their
hand-implemented parsers. We will then show how these grammars can be
used to implement applications and compare their performance and
complexity to hand-written implementations. So far, we have grammars
for DNS packets and ZIP files, which we use to develop
proof-of-concept applications that have comparable performance to
existing systems used in practice.

We will also show a grammar-based IPv4 and IPv6  stack showing acceptable performance and reduced complexity.
\section{Previous work}

This proposal and the thesis are based on previously published papers about Nail, including \cite{bangert:nail-langsec}, \cite{bangert:nail-osdi14} and \cite{bangert:nail-login}.

Generating parsers and generators from an executable specification is the
core concept of interface generators, such as CORBA~\cite{omg:corba},
XDR~\cite{RFC:1832}, and Protocol Buffers~\cite{varda2008}.  However,
interface generators do not allow the programmer to specify the byte-level
data format; instead, they define their own data encoding that is
specific to a particular interface generator.  For instance, XDR-based
protocols are incompatible with Protocol Buffers.  Moreover, this means
that interface generators cannot be used to interact with existing protocols
that were not defined using that interface generator in the first place.
As a result, interface generators cannot be used to parse or generate
widely used formats such as DNS or ZIP, which is a goal for Nail.

Closely related work has been done in the field of data description languages, for example
PacketTypes~\cite{mccann2000packet} and DataScript~\cite{back2002datascript}; a
broader overview of data description languages can be found in Fisher et
al~\cite{Fisher:700DDL}.
PacketTypes implements a C-like structure model enhanced with
length fields and constraints, but works only as a parser, and
not as an output generator.
DataScript adds output generation and built-in support for offset fields.
A particularly sophisticated data description language, PADS~\cite{Fisher:2005:PDL:1064978.1065046},
which is targeted more towards offline analysis, even features built-in support for XML and
automatic grammar inference.
However, these systems cannot easily handle complicated encodings such as compressed data, which are
supported by Nail's stream transforms. While sophisticated languages like PADS allow for handling
particular variations of offset fields, compressed data, or even XML entities, each of these
features has to be implemented in the data description language and all associated tools. Nail's
transformations keep the core language small, while enabling the wide range of
features real-world protocols require.

Recently, the Hammer project~\cite{hammer-parser} introduced a security-focused parser framework for
binary protocols.
Hammer implements grammars as language-integrated parser combinators, an approach popularized by
Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
described by Burge~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
grammars~\cite{Danielsson:2010:TPC:1863543.1863585}
by composing them from one or multiple sub-parsers.\footnote{For more
background on the history of expressing grammars, see Bryan Ford's
masters thesis~\cite{ford2002packrat}, which also describes the
default parsing algorithm used by Hammer.}
Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
an AST\@.

\begin{singlespace}
\bibliography{n-str,main,locasto,langsec,n,n-conf}

\end{singlespace}

\end{document}

### 2015-08-08 network/parse.cc
license

@@ -1,3 +1,3 @@
 #include "proto/net.nail.hh"
-#include <nail/xform/size.hh>
+#include <nail/size.hh>
 #include "proto/net.nail.cc"

### 2015-08-10 paper/thesis/abstract.tex
thesis

@@
 %% file should be \input (not \include 'd) from cover.tex.
-In this thesis, I designed and implemented a compiler which performs
-optimizations that reduce the number of low-level floating point operations
-necessary for a specific task; this involves the optimization of chains of
-floating point operations as well as the implementation of a ``fixed'' point
-data type that allows some floating point operations to simulated with integer
-arithmetic.  The source language of the compiler is a subset of C, and the
-destination language is assembly language for a micro-floating point CPU.  An
-instruction-level simulator of the CPU was written to allow testing of the
-code.  A series of test pieces of codes was compiled, both with and without
-optimization, to determine how effective these optimizations were.
+
+\textit{Nail} is a tool that greatly reduces the
+programmer effort for safely parsing and generating data formats defined
+by a grammar.
+Nail introduces several key
+ideas to achieve its goal.  First, Nail uses a protocol grammar to
+define not just the data format, but also the internal object model
+of the data.  Second, Nail eliminates the notion of semantic actions,
+used by existing parser generators, which reduces the expressive power
+but allows Nail to both {\em parse} data formats and {\em generate}
+them from the internal object model, by establishing a {\em semantic
+bijection} between the data format and the object model.  Third,
+Nail introduces {\em dependent fields} and {\em stream transforms} to
+capture protocol features such as size and offset fields, checksums,
+and compressed data, which are impractical to express in existing
+protocol languages.  Using Nail, we implement an authoritative  DNS server in
+C in under 300 lines of code and grammar, and an {\tt unzip}
+program in C in 220 lines of code and grammar, demonstrating that Nail
+makes it easy to parse complex real-world data formats.  Performance
+experiments show that a Nail-based DNS server can outperform the widely
+used BIND DNS server on an authoritative workload,
+ demonstrating that systems built with Nail can achieve good
+performance.

### 2015-08-10 network/parse.cc
parse

@@ -31,9 +31,9 @@ template <> struct ip_checksum_parse<NailMemStream> {
     return 0;
   }
 };
-int ip_checksum_generate(NailArena *tmp, NailOutStream *current, uint16_t *ptr_checksum){
+int ip_checksum_generate(NailArena *tmp, NailOutStream *current, uint16_t *ptr_checksum, uint16_t extra_1=0, uint16_t extra_2 =0){
   
-  uint32_t checksum=0;
+  uint32_t checksum=extra_1+extra_2;
   const uint16_t *field =  (uint16_t *)current->data;
   for(size_t ihl = current->pos/2; ihl>0;ihl--){
       checksum += *field;
@@ -72,9 +72,10 @@ int ip_header_generate(NailArena *tmp,  const NailOutStream* in_options, const N
   *ihl /= 4;
   *ihl += 5;
 
-  ip_checksum_generate(tmp, current, out_checksum);
   if(n_fail(tail_generate(tmp, in_body, current))) return -1;
   *total_length = *ihl * 4 + in_body->pos;
+  
+  ip_checksum_generate(tmp, current, out_checksum, *ihl, *total_length);
   return 0;
 }
   

### 2015-08-10 paper/thesis/code/zip_out.tex
thesis

--- 2014-11-25 paper/osdi14/code/zip-out.tex
@@
 \PY{p}{\PYZcb{}}\PY{p}{;}\PY{c+cm}{/*..*/}
-\PY{c+c1}{//The programmer calls these generated functions.}
-\PY{k+kt}{int} \PY{n+nf}{gen\PYZus{}zip}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp\PYZus{}arena}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{out}\PY{p}{,}\PY{n}{zip} \PY{o}{*}\PY{n}{val}\PY{p}{)}\PY{p}{;}
-\PY{n}{zip}\PY{o}{*}\PY{n+nf}{parse\PYZus{}zip}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{arena}\PY{p}{,} 
-              \PY{k}{const} \PY{k+kt}{uint8\PYZus{}t} \PY{o}{*}\PY{n}{data}\PY{p}{,} \PY{k+kt}{size\PYZus{}t} \PY{n}{size}\PY{p}{)}\PY{p}{;}
-\PY{c+c1}{//The programmers implements these two transformation}
-\PY{c+c1}{//functions and two similar ones for output.}
-\PY{k}{extern} \PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}eod\PYZus{}parse}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,}
- \PY{n}{NailStream} \PY{o}{*}\PY{n}{filestream}\PY{p}{,} \PY{n}{NailStream} \PY{o}{*}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{current}\PY{p}{)}\PY{p}{;}
-\PY{k}{extern} \PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}compression\PYZus{}parse}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,}
- \PY{n}{NailStream} \PY{o}{*}\PY{n}{uncomp}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{compressed}\PY{p}{,}\PY{k+kt}{uint32\PYZus{}t}\PY{o}{*} \PY{n}{size}\PY{p}{)}\PY{p}{;}
+\PY{c+c1}{//The programmer calls these functions to interact with the Nail\PYZhy{}generated parser.}
+\PY{k+kt}{int} \PY{n+nf}{gen\PYZus{}zip}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp\PYZus{}arena}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{out}\PY{p}{,}\PY{n}{zip} \PY{o}{*} \PY{n}{val}\PY{p}{)}\PY{p}{;}
+\PY{n}{zip}\PY{o}{*}\PY{n+nf}{parse\PYZus{}zip}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{arena}\PY{p}{,} \PY{k}{const} \PY{k+kt}{uint8\PYZus{}t} \PY{o}{*}\PY{n}{data}\PY{p}{,} \PY{k+kt}{size\PYZus{}t} \PY{n}{size}\PY{p}{)}\PY{p}{;}
+\PY{c+c1}{//The programmer implements these two functions, and Nail calls them.}
+\PY{k}{extern} \PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}eod\PYZus{}parse}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{str\PYZus{}filestream}\PY{p}{,}
+   \PY{n}{NailStream} \PY{o}{*}\PY{n}{str\PYZus{}end\PYZus{}directory}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{current}\PY{p}{)}\PY{p}{;}
+\PY{k}{extern} \PY{k+kt}{int} \PY{n+nf}{zip\PYZus{}compression\PYZus{}parse}\PY{p}{(}\PY{n}{NailArena} \PY{o}{*}\PY{n}{tmp}\PY{p}{,}\PY{n}{NailStream} \PY{o}{*}\PY{n}{str\PYZus{}uncompressed}\PY{p}{,}
+   \PY{n}{NailStream} \PY{o}{*}\PY{n}{compressed}\PY{p}{,}\PY{k+kt}{uint32\PYZus{}t}\PY{o}{*} \PY{n}{size}\PY{p}{)}\PY{p}{;}
 \end{Verbatim}

### 2015-08-10 paper/thesis/eval.tex
thesis

--- 2014-09-10 paper/osdi14/eval.tex
@@
 In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
-for DNS packets, shown in Figure~\ref{fig:dns-xform}.
+for DNS packets, shown in Figure~\ref{fig:dns-full}.
 The grammar corresponds almost directly to the diagrams in RFC 1035,
@@
 Handling label compression in existing tools, such as Bison or Hammer,
-would at best be very awkward, because some ad-hoc trick would have to
+would be awkward at best, because some ad-hoc trick would have to
 be used to re-position the parser's input stream.  Keeping track of the
@@
 %(whose two C function signatures are shown in Figure~\ref{fig:zip-eod-xform}). 
+The corresponding C function
 \cc{zip\_end\_of\_directory_parse}
@@
 application that uses Nail, we implemented two applications---a
-DNS server, and an
+DNS server and an
 \cc{unzip} program---based on the above grammars, and compared code
@@
 \centering
-\begin{tabular}{@{~}lrr@{~}l@{~}}
+\smaller[0.5]
+\begin{tabular}{lrr@{~}l}
 \toprule
@@
 
-In \S\ref{s:motivation}, we found 15 input handling vulnerabilities
+In \S\ref{s:motivation}, we presented 15 input handling vulnerabilities
 related to ZIP files.
@@
 Because Nail's generated code checks offsets before reading and does not
-expose any untrusted pointers to the application, it is immune against
+expose any untrusted pointers to the application, it is immune to
 memory corruption attacks by design.
 
-Nail also protects against parsing inconsistency vulnerabilities like the four others we studied. First,
+Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
 Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
-that caused a vulnerability in the Python ZIP library. Second, the other three vulnerabilities exist
+that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
 because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
-parsers are not very reusable, as they build application specific data structures and are tightly
+parsers are not very reusable, as they build application-specific data structures and are tightly
 coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
@@
 and implements many features that are not present in our Nail-based DNS
-server and allow it to be used in more complicated configurations, 
+server and that allow it to be used in more complicated configurations, 
 we believe our results demonstrate that Nail's parsers are not

### 2015-08-10 paper/thesis/fig-stream-api.tex
thesis

\begin{figure*}
  \begin{tabular}{lp{2.75in}}
    \toprule
    \bf Method & \bf Description \\
    \midrule
    \texttt{bool check(uint n)} & Check if the current position is at least $n$ bits before the end of the stream. \\
    \texttt{uint read\_bits(uint n)} &  Read $n$ bits from the stream. Check($n$) must be true. Does not move the current position.\\
    \texttt{uint read\_bits\_little(uint n)} &  Like read\_bits, but interprets the value as little endian.\\
    \texttt{void advance(uint n)} & Advances the current position by $n$ bits.\\
    \texttt{void backup(uint n))} & Reverses the current position by $n$ bits.\\
    \texttt{pos\_t get\_pos(uint n))} & Returns the current position.\\
    \texttt{void reposition(pos\_t old))} & Reverts the current position to a value formerly returned from get\_pos().\\

    \bottomrule
    \end{tabular}
\label{fig:stream-api}
\caption{The operations Nail parsers invoke on their input streams.}
\end{figure*}

### 2015-08-10 paper/thesis/impl.tex
thesis

--- 2014-11-14 paper/login/impl.tex
@@
 
-The current prototype of the Nail parser generator supports the C programming
-language. The implementation parses Nail grammars with Nail itself, using a
+The current prototype of the Nail parser generator can emit parsers in either C or C++. The implementation parses Nail grammars with Nail itself, using a
 130-line Nail grammar feeding into a 2,000-line C++ program that
@@
 performed with a subset of the grammar implemented using conventional grammars. 
-An option for C++ STL data models is in development.  In
-this section, we will discuss some particular features of our parser
+In this section, we will discuss some particular features of our parser
 implementation.
@@
 model. Currently the parser uses a straightforward top-down algorithm, which can perform poorly on
-grammars that backtrack heavily. However, preparations have been made to add Packrat
-parsing~\cite{packrat-parsing:icfp02} that achieve linear time even in the worst case.
+grammars that backtrack heavily. However, Packrat
+parsing~\cite{packrat-parsing:icfp02} that achieves linear time even in the worst case can be added relatively easily.
+
+\paragraph{Parser input/output.}
+Nail-generated parsers interact with their input through an abstract stream data type, which must support the operations show in Figure~\ref{fig:stream-api}.
+In the C implementation, the programmer can provide an implementation of these operations, or fall back to using a default implementation backed by continuous buffers in memory.
+Both the input stream and all intermediate streams must use the same implementation.
+
+While the concrete implementation is transparent to the Nail parser itself, transformations can and often must modify the stream data type through other interfaces.
+The standard library of Nail transformations is at present only for memory-buffer backed streams and often directly manipulates the underlying buffers.
+This makes many operations, such as splitting a stream into multiple substreams, much faster than if the substreams had to be copied directly.
+
+\input{fig-stream-api}
+\paragraph{C++ implementation.}
+The C implementation forces all streams to have the same underlying data type.
+To address this shortcoming, we developed an optional C++ backend for Nail.
+Nail C++ parsers are template functions that are parametrized by the types of the streams they operate on.
+Transformations are template functors -- classes that can perform different computation  and produce a different type of stream depending on the types of their inputs.
+
+Useful transformations are often only possible to write when specialized to their input types, and the fully general will either be very slow or not even implemented.
+Consider for example the standard library offset transform, which creates a substream starting at a given offset.
+Given just the stream interface shown in Figure~\ref{fig:stream-api}, the only way to implement offset would be to either to make a copy of the underlying stream, using reposition and read, or to reposition the underlying stream before and after each read.
+By specializing the transformation for in-memory streams, we can construct another memory stream that points to the correct range in the same memory.
+Similar specializations could exist for file-descriptor backed streams using a duplicate descriptor or sparse-fragment streams.
+
+Transformations can also be expressed through chaining together other transformations -- which is equivalent to composing them directly in the Nail grammar, but results in shorter and easier to understand Nail grammars.
+
+In addition to allowing complicated file formats to be captured more naturally by mapping their layout onto different stream types, we will show in Section~\ref{sec:network-stack} that templated streams allow new classes of applications for Nail transformations, such as reassembling packets.
 

### 2015-08-10 paper/thesis/relwk.tex
thesis

--- 2014-09-10 paper/osdi14/relwk.tex
@@
 Closely related work has been done in the field of data description languages, for example
-PacketTypes~\cite{mccann2000packet} and DataScript~\cite{back2002datascript}. 
-A broader overview of data
-description languages can be found in~\cite{Fisher:700DDL}. However, PacketTypes works
-only as a parser, not as an output generator and implement a C-like structure model enhanced with
-length fields and constraints. DataScript adds output generation and built-in support for offset
-fields. A
-particularly sophisticated data description language, PADS~\cite{Fisher:2005:PDL:1064978.1065046},
+PacketTypes~\cite{mccann2000packet} and DataScript~\cite{back2002datascript}; a
+broader overview of data description languages can be found in Fisher et
+al~\cite{Fisher:700DDL}.
+PacketTypes implements a C-like structure model enhanced with
+length fields and constraints, but works only as a parser, and
+not as an output generator.
+DataScript adds output generation and built-in support for offset fields.
+A particularly sophisticated data description language, PADS~\cite{Fisher:2005:PDL:1064978.1065046},
 which is targeted more towards offline analysis, even features built-in support for XML and
 automatic grammar inference.
-
 However, these systems cannot easily handle complicated encodings such as compressed data, which are
 supported by Nail's stream transforms. While sophisticated languages like PADS allow for handling
-particular variations of offset fields, compressed data or even XML entities, each of these
-features has to be implemented in the data description language - and all associated tools. Nail's
-transformations keep the core language practicably small, while enabling the wide range of
+particular variations of offset fields, compressed data, or even XML entities, each of these
+features has to be implemented in the data description language and all associated tools. Nail's
+transformations keep the core language small, while enabling the wide range of
 features real-world protocols require.
@@
 
-Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
-generating output for OCaml.  Parsifal structures grammars as an OCaml
+Parsifal~\cite{ANSSI:parsifal} is a parser framework for OCaml that also supports
+generating output.  Parsifal structures grammars as an OCaml
 type that holds an internal model and functions for parsing input and

### 2015-08-10 paper/thesis/code/dns-api.c

@@ -1,0 +1,5 @@
+struct dnspacket *parse_dnspacket(NailArena *arena,
+  NailStream *in);
+int gen_dnspacket(NailArena *tmp_arena,
+  NailStream *out,
+  struct dnspacket *val);

### 2015-08-10 paper/thesis/code/dns-xform.c

@@ -1,0 +1,7 @@
+int dnscompress_parse(NailArena *tmp,
+  NailStream *out_decomp,
+  NailStream *in_current);
+
+int dnscompress_generate(NailArena *tmp,
+  NailStream *in_decomp,
+  NailStream *out_current);

### 2015-08-10 paper/thesis/code/offset-xform.c

@@ -1,0 +1,15 @@
+int offset_u32_parse(NailArena *tmp,
+  NailStream *out_str, NailStream *in_current,
+  const uint32_t *off)
+{
+  /* out_str = suffix of in_current
+               at offset *off */
+}
+
+int offset_u32_generate(NailArena *tmp,
+  NailStream *in_fragment,
+  NailStream *out_current, uint32_t *off)
+{
+  /* *off = position of out_current */
+  /* append in_fragment to out_current */
+}

### 2015-08-11 paper/thesis/eval.tex
DNS CPP benchmark

@@
-\section{Evaluation}
+\chapter{Evaluation}
 \label{s:eval}
@@
 
-\subsection{Data formats}
+\section{Data formats}
 \label{s:eval-formats}
@@
 generators would be unable to fully handle 5 out of the 7 data formats.
-In the rest of this subsection, we describe the DNS and Zip grammars in
+In the rest of this section, we describe the DNS and Zip grammars in
 more detail, focusing on how Nail's features enable us to support these
@@
 
-\subsection{Programmer effort}
+\section{Programmer effort}
 \label{s:eval-effort}
@@
 
-\subsection{Security}
+\section{Security}
 
@@
 
-\subsection{Performance}
+\section{Performance}
 
@@
 
-The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
-our Nail-based DNS server can achieve higher performance and lower latency
-than BIND\@.  Although BIND is a more sophisticated DNS server,
-and implements many features that are not present in our Nail-based DNS
-server and that allow it to be used in more complicated configurations, 
-we believe our results demonstrate that Nail's parsers are not
-a barrier to achieving good performance.
-
+The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate
+that our Nail-based DNS server can achieve higher performance and
+lower latency than BIND\@.  Although BIND is a more sophisticated DNS
+server, and implements many features that are not present in our
+Nail-based DNS server and that allow it to be used in more complicated
+configurations, we believe our results demonstrate that Nail's parsers
+are not a barrier to achieving good performance.  Previous papers on
+Nail presented results using a Nail parser generated as C, whereas
+these results were produced on the same configuration with the newer
+C++ parser generator, which improves performance by using a more
+efficient stream representation.
 \input{fig-perf-dns}

### 2015-08-11 paper/thesis/relwk.tex
DNS CPP benchmark

@@
-\section{Related work}
+\chapter{Related work}
 \label{s:relwk}
 
+This thesis is based on previously published papers about Nail, including \cite{bangert:nail-langsec}, \cite{bangert:nail-osdi14} and \cite{bangert:nail-login}.
+We presented an earlier design of Nail at a
+workshop~\cite{bangert:nail-langsec}.  At that stage, Nail had only
+limited support for dependent fields, and did not support stream
+transforms at all, which are crucial for supporting real-world formats
+like DNS and ZIP\@.  The workshop paper also did not provide a detailed
+design discussion or evaluation.
+Since presenting the conference paper~\cite{bangert:nail-osdi14}, we implemented the C++ backend and explored novel uses of Nail for network protocols. 
 \paragraph{Parsers.}
@@
 
-We presented an earlier design of Nail at a
-workshop~\cite{bangert:nail-langsec}.  At that stage, Nail had only
-limited support for dependent fields, and did not support stream
-transforms at all, which are crucial for supporting real-world formats
-like DNS and ZIP\@.  The workshop paper also did not provide a detailed
-design discussion or evaluation.
-
 \paragraph{Application use of parsers.}

### 2015-08-11 examples/dns_cpp/bench/extract.sh

@@ -1,0 +1,2 @@
+grep qps $1 | awk '{print $4};'  > $1.qps
+grep 'RTT average' $1 | awk '{print $3}' > $1.rttavg

### 2015-08-11 examples/dns_cpp/bench/paste.sh

@@ -1,0 +1,2 @@
+paste *.qps > all.qps
+paste *.rttavg > all.rttavg

### 2015-08-11 examples/dns_cpp/bench/queryperf/queryperf.c

@@ -1,0 +1,2237 @@
+/*
+ * Copyright (C) 2000, 2001  Nominum, Inc.
+ *
+ * Permission to use, copy, modify, and distribute this software for any
+ * purpose with or without fee is hereby granted, provided that the above
+ * copyright notice and this permission notice appear in all copies.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS" AND INTERNET SOFTWARE CONSORTIUM
+ * DISCLAIMS ALL WARRANTIES WITH REGARD TO THIS SOFTWARE INCLUDING ALL
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS. IN NO EVENT SHALL
+ * INTERNET SOFTWARE CONSORTIUM BE LIABLE FOR ANY SPECIAL, DIRECT,
+ * INDIRECT, OR CONSEQUENTIAL DAMAGES OR ANY DAMAGES WHATSOEVER RESULTING
+ * FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT,
+ * NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF OR IN CONNECTION
+ * WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.
+ */
+
+/***
+ ***	DNS Query Performance Testing Tool  (queryperf.c)
+ ***
+ ***	Version $Id: queryperf.c,v 1.12 2007/09/05 07:36:04 marka Exp $
+ ***
+ ***	Stephen Jacob <sj@nominum.com>
+ ***/
+
+#define BIND_8_COMPAT	/* Pull in <arpa/nameser_compat.h> */
+
+#include <sys/time.h>
+#include <sys/types.h>
+#include <sys/socket.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <limits.h>
+#include <time.h>
+#include <unistd.h>
+#include <netdb.h>
+#include <netinet/in.h>
+#include <arpa/nameser.h>
+#include <resolv.h>
+#include <math.h>
+#include <errno.h>
+
+#ifdef HAVE_CONFIG_H
+#include "config.h"
+#ifndef HAVE_GETADDRINFO
+#include "missing/addrinfo.h"
+#endif
+#endif
+
+/*
+ * Configuration defaults
+ */
+
[…]

### 2015-08-11 examples/dns_cpp/bench/run-benchmark.sh

@@ -1,0 +1,3 @@
+sudo ../server naildns.zone 1>&2 &
+sleep 5
+queryperf/queryperf -l 60 2>&1 < queryperf.input

### 2015-08-11 examples/dns_cpp/bench/run-naildns.sh

@@ -1,0 +1,5 @@
+./run-benchmark.sh > results.50000/results.naildns
+for i in $(seq 6); do
+    ./run-benchmark.sh >> results.50000/results.naildns
+done
+

### 2015-08-11 generator/mk_directparser.cc
proper support for checksums

@@ -164,7 +164,7 @@ public:
         ValExpr lval(std::string ("dep_") + name);
         out << type << " dep_" << name << ";\n";
         parser(field->dependency.parser->pr,lval,fail,newscope);
-        newscope.add_dependency_definition(name,type);
+        newscope.add_dependency_definition(name,type, 0);
         break;
       }
       case TRANSFORM:{

### 2015-08-11 generator/mk_gen.cc

@@ -104,7 +104,7 @@ class GenGenerator{
           }
           int width =  boost::lexical_cast<int>(mk_str(p.pr.integer.parser.unsign));
           out << type  << " dep_"<< name << ";";
-          newscope.add_dependency_definition(name,type);
+          newscope.add_dependency_definition(name,type, width);
           assert(post == "");
           out << "NailOutStreamPos rewind_"<< name << "=NailOutStream_getpos(str_current);";
           out << "NailOutStream_write(str_current,0,"<<width<<");";
@@ -149,15 +149,25 @@ class GenGenerator{
           FOREACH(stream, field->transform.left){
             newscope.add_stream_definition(mk_str(*stream));
           }
+          invocation << "{ NailOutStreamPos t_rewind_eot =NailOutStream_getpos(str_current);\n";
+
+          FOREACH(param, field->transform.right){
+            if(param->N_type != PDEPENDENCY) continue;
+            std::string name (mk_str(param->pdependency));
+            if(!newscope.is_local_dependency(name)) continue;
+            invocation << "NailOutStream_reposition(str_current, rewind_"<<name<<");";
+            invocation << "NailOutStream_write(str_current,dep_"<<name<<","<<newscope.dependency_width(name)<<");";
+          }
+          invocation << " NailOutStream_reposition(str_current, t_rewind_eot);\n }";
           transform_invocations.push_front(invocation.str());
         }
           break;
         }
       }
+      fixup <<  "NailOutStream_reposition(str_current, end_of_struct);";
       for(auto &transform: transform_invocations){
         out << transform;
       }
-      fixup <<  "NailOutStream_reposition(str_current, end_of_struct);";
       //TODO:  We need to properly interleave APPLY and TRANSFORM when generating! Transform at the
       //end is wrong in some cases (when the applied parser relies on a context from the transform)
       out << fixup.str();

### 2015-08-11 generator/mk_parser.cc

@@ -514,7 +514,7 @@ class CPrimitiveParser{
             out  << "NailStream *stream = &tmpstream;"; //TODO: parametrize this on action
             dependency_action.action(field->dependency.parser->pr,lval,end); 
             out << "}";
-            newscope.add_dependency_definition(name,type);
+            newscope.add_dependency_definition(name,type, 0);
             out << "n_tr_setpos(trace,trace_"<<name<<");\n";
             out << "n_tr_const(trace,str_current);\n";
           }

### 2015-08-11 generator/nailtool.h

@@ -44,6 +44,7 @@ NailException(const std::string &message) : std::logic_error(message){}
 struct Dependency{
   std::string name;
   std::string typedef_type;
+  uint64_t width; // XXX: Replace by a line that should be inserted to output this Dependency
 };
 class Scope{ 
   std::map<std::string,Dependency> dependencies;
@@ -57,10 +58,11 @@ public:
       throw NailException("Duplicate stream "+value);
     }
   }
-  void add_dependency_parameter(std::string value, std::string type){
+  void add_dependency_parameter(std::string value, std::string type,uint64_t width){
     Dependency dep;
     dep.name = std::string("dep_")+value;
     dep.typedef_type = type;
+    dep.width = width;
     if(!dependencies.emplace(value,dep).second){
       throw NailException("Duplicate dependency "+value);
     }
@@ -70,10 +72,11 @@ public:
       throw NailException("Duplicate dependency "+value);
     }
   }
-  void add_dependency_definition(std::string value,std::string type){
+  void add_dependency_definition(std::string value,std::string type,uint64_t width ){
     Dependency dep;
     dep.name = std::string("&dep_")+value;
     dep.typedef_type = type;
+    dep.width = width;
     if(!dependencies.emplace(value,dep).second){
       throw NailException("Duplicate dependency "+value);
     }
@@ -107,6 +110,19 @@ public:
         return parent->stream_ptr(name);
     }
     return i->second;                
+  }
+  bool is_local_dependency(std::string name){
+    return dependencies.count(name)>0;
+  }
+  uint64_t dependency_width(std::string name){
+    auto i = dependencies.find(name);
+    if(i==dependencies.end()){
+      if(!parent)
+              throw NailException(std::string("Undefined reference to ") +  name);
+      else
+        return parent->dependency_width(name);
+    }
+    return i->second.width;            
   }
 };
 

### 2015-08-11 generator/util.cc

@@ -154,7 +154,7 @@ std::string parameter_definition(const definition &def, Scope &scope){
             std::string name = mk_str(param->ddependency.name);
             assert(post == "");
             params << ","<<  type << "* dep_" << name << post;
-            scope.add_dependency_parameter(name,type);
+            scope.add_dependency_parameter(name,type, 0);
           }
             break;
           }

### 2015-08-11 network/parse.cc
working ip header checksum, now in clean

@@ -1,5 +1,6 @@
 #include "parse.hh"
 #include <nail/size.hh>
+#include <arpa/inet.h> //htons
 /* template<typename siz_t> struct NailCoroFragmentStream {
 private:
   typedef std::map<siz_t, NailMemStream> inner_t;
@@ -10,14 +11,10 @@ public:
     NailMemStream::pos_t pos;
   } 
   };*/
-template <typename str> struct ip_checksum_parse {
+template <typename str> struct internet_checksum_parse {
 };
-template <> struct ip_checksum_parse<NailMemStream> {
-  typedef NailMemStream out_1_t;
-  typedef NailMemStream out_2_t;
-  typedef NailMemStream out_3_t;
-  static int f(NailArena *tmp,NailMemStream *in_current,uint16_t *ptr_checksum){
-       
+template <> struct internet_checksum_parse<NailMemStream> {
+  static int f(NailArena *tmp,NailMemStream *in_current,uint16_t *ptr_checksum){       
     uint32_t checksum = 0;
     uint16_t *field = (uint16_t *)in_current->getBuf();
     for(size_t l = in_current->getSize()/2; l>0;l--){
@@ -31,9 +28,9 @@ template <> struct ip_checksum_parse<NailMemStream> {
     return 0;
   }
 };
-int ip_checksum_generate(NailArena *tmp, NailOutStream *current, uint16_t *ptr_checksum, uint16_t extra_1=0, uint16_t extra_2 =0){
-  
-  uint32_t checksum=extra_1+extra_2;
+
+int internet_checksum_generate(NailArena *tmp, NailOutStream *current, uint16_t *ptr_checksum){
+  uint32_t checksum=0;
   const uint16_t *field =  (uint16_t *)current->data;
   for(size_t ihl = current->pos/2; ihl>0;ihl--){
       checksum += *field;
@@ -42,28 +39,45 @@ int ip_checksum_generate(NailArena *tmp, NailOutStream *current, uint16_t *ptr_c
   while(checksum >> 16){
     checksum = (checksum & 0xFFFF)+ (checksum>>16);
   }
-  *ptr_checksum = ~((uint16_t)checksum);
+  *ptr_checksum = htons(~((uint16_t)checksum));
   return 0;
 }
+template <typename str> struct ip_header_checksum_parse {
+};
+template <> struct ip_header_checksum_parse<NailMemStream> {
+  static int f(NailArena *tmp,NailMemStream *in_current,uint16_t *ptr_checksum, uint8_t *ihl){
[…]

### 2015-08-11 network/tunclient.cc

@@ -7,6 +7,7 @@
 #include <stdlib.h>
 #include <stdio.h>
 #include <unistd.h>
+#include <arpa/inet.h>
 #include <linux/if.h>
 #include <linux/if_tun.h>
 #include <time.h>

### 2015-08-11 network/tunclient.cc
ICMP works!!

@@ -12,7 +12,7 @@
 #include <linux/if_tun.h>
 #include <time.h>
 
-#include <uthash.h>   
+#include <unordered_map> 
 #include "parse.hh"
 
 
@@ -67,7 +67,6 @@ typedef struct arpentry{
         uint32_t ip;
         macaddr mac; // All 0 when unused 
         time_t time;
-        UT_hash_handle hh;
 } arpentry;
 #define ARPSIZE 512
 struct phy_iface;
@@ -75,20 +74,18 @@ struct ip_iface{
   uint32_t ip;
   uint32_t broadcast;
   phy_iface *phy;
-};
+ip_iface(phy_iface *parent, uint32_t _ip, uint32_t _broadcast) : phy(parent), ip(_ip), broadcast(_broadcast){}
+    };
 typedef struct phy_iface{
-        int tun_fd;
-        macaddr mac;
+int tun_fd;
+macaddr mac;
 
-        arpentry *arp; 
-        struct ip_iface ip; // TODO: Add multi-IP support?
+std::unordered_map<uint32_t,arpentry> arp;
+struct ip_iface ip; // TODO: Add multi-IP support?
+  phy_iface(int fd,macaddr MAC, uint32_t ip_addr,uint32_t broadcast, uint32_t netmask,uint32_t gateway):  tun_fd(fd), ip(this,ip_addr,broadcast){
+     memcpy(mac, MAC, sizeof mac);
+  }
 } phy_iface;
-void phy_init(phy_iface *phy, macaddr *mac, uint32_t ip){
-        memcpy(phy->mac,mac,sizeof(phy->mac));
-        phy->arp = NULL;
-        phy->ip.ip = ip; 
-        phy->ip.phy = phy;
-}
 
 int send_frame(NailArena *arena,ethernet *packet,struct phy_iface *iface){
         const char *buf;
@@ -112,9 +109,11 @@ void prepare_ethernet(ethernet *eth, struct phy_iface *iface){
                 break;
         case IPFOUR:
                 //eth->dest =  0;//ARP here!
-                memcpy(eth->src,iface->mac,sizeof(eth->src)); 
-                break;
-        }
+          memset(eth->dest, 0xFF,sizeof eth->dest);// TODO: ARP!!!
+          
+             memcpy(eth->src,iface->mac,sizeof(eth->src)); 
+             break;
+             }
 }
[…]

### 2015-08-11 examples/network/tunclient.c
Get rid of old network example

@@ -1,234 +1,0 @@
-#include <sys/types.h>  
-#include <sys/socket.h>
-#include <sys/stat.h>
-#include <sys/ioctl.h>
-#include <fcntl.h>
-#include <string.h>
-#include <stdlib.h>
-#include <stdio.h>
-#include <unistd.h>
-#include <linux/if.h>
-#include <linux/if_tun.h>
-#include <time.h>
-
-#include <uthash.h>   
-
-#include "net.nail.h"
-//http://backreference.org/2010/03/26/tuntap-interface-tutorial/
-int tun_alloc(char *dev, int flags) {
-
-  struct ifreq ifr;
-  int fd, err;
-  char *clonedev = "/dev/net/tun";
-
-  /* Arguments taken by the function:
-   *
-   * char *dev: the name of an interface (or '\0'). MUST have enough
-   *   space to hold the interface name if '\0' is passed
-   * int flags: interface flags (eg, IFF_TUN etc.)
-   */
-
-   /* open the clone device */
-   if( (fd = open(clonedev, O_RDWR)) < 0 ) {
-     return fd;
-   }
-
-   /* preparation of the struct ifr, of type "struct ifreq" */
-   memset(&ifr, 0, sizeof(ifr));
-
-   ifr.ifr_flags = flags;   /* IFF_TUN or IFF_TAP, plus maybe IFF_NO_PI */
-
-   if (*dev) {
-     /* if a device name was specified, put it in the structure; otherwise,
-      * the kernel will try to allocate the "next" device of the
-      * specified type */
-     strncpy(ifr.ifr_name, dev, IFNAMSIZ);
-   }
-
-   /* try to create the device */
-   if( (err = ioctl(fd, TUNSETIFF, (void *) &ifr)) < 0 ) {
-     close(fd);
-     return err;
-   }
-
-  /* if the operation was successful, write back the name of the
-   * interface to the variable "dev", so the caller can know
-   * it. Note that the caller MUST reserve space in *dev (see calling
-   * code below) */
-  strcpy(dev, ifr.ifr_name);
-
-  /* this is the special file descriptor that the caller will use to talk
-   * with the virtual interface */
-  return fd;
-}
-typedef struct arpentry{
-        uint32_t ip;
-        macaddr mac; // All 0 when unused 
-        time_t time;
-        UT_hash_handle hh;
-} arpentry;
-#define ARPSIZE 512
-struct ip_iface{
-        uint32_t ip;
[…]

### 2015-08-13 paper/thesis/impl.tex
DNS CPP benchmark
foo

@@
-\section{Implementation}
+\chapter{Implementation}
 \label{s:impl}
@@
 Transformations can also be expressed through chaining together other transformations -- which is equivalent to composing them directly in the Nail grammar, but results in shorter and easier to understand Nail grammars.
-
-In addition to allowing complicated file formats to be captured more naturally by mapping their layout onto different stream types, we will show in Section~\ref{sec:network-stack} that templated streams allow new classes of applications for Nail transformations, such as reassembling packets.
+In addition to allowing complicated file formats to be captured more naturally by mapping their layout onto different stream types, for example, fragmented network packets can be stored as individual fragments and do not have to be assembled into a continuous buffer.
 

### 2015-08-13 network/parse.cc

@@ -1,7 +1,7 @@
 #include "parse.hh"
 #include <nail/size.hh>
 #include <arpa/inet.h> //htons
-/* template<typename siz_t> struct NailCoroFragmentStream {
+template<typename siz_t> struct NailCoroFragmentStream {
 private:
   typedef std::map<siz_t, NailMemStream> inner_t;
   std::map<siz_t, NailMemStream> streams;
@@ -10,7 +10,7 @@ public:
     inner_t::iterator iter;
     NailMemStream::pos_t pos;
   } 
-  };*/
+};
 template <typename str> struct internet_checksum_parse {
 };
 template <> struct internet_checksum_parse<NailMemStream> {

### 2015-08-13 network/tunclient.cc

@@ -195,7 +195,7 @@ void process_frame(NailArena *arena,ethernet *frame, struct phy_iface *i ){
 }
 unsigned char buffer[9000];
 uint32_t mkip(uint32_t a,uint32_t b,uint32_t c,uint32_t d){
-        return a<< 24 | b<<16 | c<<8 | d;
+        return a<< 24 | b<<16 | c<<8 | d; 
 }
 
 int main (int argc, char **argv){

### 2015-08-17 paper/thesis/concl.tex
DNS CPP benchmark
paper->thesis

--- 2014-09-11 paper/osdi14/concl.tex
@@
-\section{Conclusion}
+\chapter{Conclusion}
 \label{s:concl}
 
-This paper presented the design and implementation of \textit{Nail},
+This thesis presented the design and implementation of \textit{Nail},
 a tool for parsing and generating complex data formats based on a
@@
 fewer lines of code.
-Nail and all of the applications and grammars developed in this paper
+Nail and all of the applications and grammars developed in this thesis
 are released as open-source software, available at

### 2015-08-17 paper/thesis/intro.tex
thesis
paper->thesis

--- 2014-09-10 paper/osdi14/intro.tex
@@
-\section{Introduction}
+\chapter{Introduction}
 
@@
 
-This paper presents the design and implementation of Nail, a parser generator that greatly reduces
+This thesis presents the design and implementation of Nail, a parser generator that greatly reduces
 the programmer effort required to use grammars. Nail addresses the above challenges with
@@
 
-The rest of this paper is organized as follows.  \S\ref{s:relwk}
+The rest of this thesis is organized as follows.  \S\ref{s:relwk}
 puts Nail in the context of related work.

### 2015-08-17 paper/thesis/eval.tex
section->chapter

@@
 generators would be unable to fully handle 5 out of the 7 data formats.
-In the rest of this section, we describe the DNS and Zip grammars in
+In the rest of this chapter, we describe the DNS and Zip grammars in
 more detail, focusing on how Nail's features enable us to support these
@@
 
-In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
+In Chapter~\ref{s:design}, we introduced Nail's syntax with a grammar
 for DNS packets, shown in Figure~\ref{fig:dns-full}.

### 2015-08-17 paper/thesis/chap1.tex
Bibliography

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the 
%% \files=
%% prompt when you run the file main.tex through LaTeX.
\chapter{Introduction}

Micro-optimization is a technique to reduce the overall operation count of
floating point operations.  In a standard floating point unit, floating
point operations are fairly high level, such as ``multiply'' and ``add'';
in a micro floating point unit ($\mu$FPU), these have been broken down into
their constituent low-level floating point operations on the mantissas and
exponents of the floating point numbers.

Chapter two describes the architecture of the $\mu$FPU unit, and the
motivations for the design decisions made.

Chapter three describes the design of the compiler, as well as how the
optimizations discussed in chapter~\ref{ch1:opts} were implemented.

Chapter four describes the purpose of test code that was compiled, and which
statistics were gathered by running it through the simulator.  The purpose
is to measure what effect the micro-optimizations had, compared to
unoptimized code.  Possible future expansions to the project are also
discussed.

\section{Motivations for micro-optimization}

The idea of micro-optimization is motivated by the recent trends in computer
architecture towards low-level parallelism and small, pipelineable
instruction sets \cite{patterson:risc,rad83}.  By getting rid of more
complex instructions and concentrating on optimizing frequently used
instructions, substantial increases in performance were realized.

Another important motivation was the trend towards placing more of the
burden of performance on the compiler.  Many of the new architectures depend
on an intelligent, optimizing compiler in order to realize anywhere near
their peak performance
\cite{ellis:bulldog,pet87,coutant:precision-compilers}.  In these cases, the
compiler not only is responsible for faithfully generating native code to
match the source language, but also must be aware of instruction latencies,
delayed branches, pipeline stages, and a multitude of other factors in order
to generate fast code \cite{gib86}.

Taking these ideas one step further, it seems that the floating point
operations that are normally single, large instructions can be further broken
down into smaller, simpler, faster instructions, with more control in the
compiler and less in the hardware.  This is the idea behind a
micro-optimizing FPU; break the floating point instructions down into their
basic components and use a small, fast implementation, with a large part of
the burden of hardware allocation and optimization shifted towards
compile-time.

Along with the hardware speedups possible by using a $\mu$FPU, there are
also optimizations that the compiler can perform on the code that is
generated.  In a normal sequence of floating point operations, there are
many hidden redundancies that can be eliminated by allowing the compiler to
control the floating point operations down to their lowest level.  These
optimizations are described in detail in chapter
~\ref{ch1:opts}.

\section{Description of micro-optimization}\label{ch1:opts}

In order to perform a sequence of floating point operations, a normal FPU
performs many redundant internal shifts and normalizations in the process of
performing a sequence of operations.  However, if a compiler can
decompose the floating point operations it needs down to the lowest level,
it then can optimize away many of these redundant operations.

If there is some additional hardware support specifically for
micro-optimization, there are additional optimizations that can be
performed.  This hardware support entails extra ``guard bits'' on the
standard floating point formats, to allow several unnormalized operations to
be performed in a row without the loss information\footnote{A description of
the floating point format used is shown in figures~\ref{exponent-format}
and~\ref{mantissa-format}.}.  A discussion of the mathematics behind
unnormalized arithmetic is in appendix~\ref{unnorm-math}.

The optimizations that the compiler can perform fall into several categories:

\subsection{Post Multiply Normalization}

When more than two multiplications are performed in a row, the intermediate
normalization of the results between multiplications can be eliminated.
This is because with each multiplication, the mantissa can become
denormalized by at most one bit.  If there are guard bits on the mantissas
to prevent bits from ``falling off'' the end during multiplications, the
normalization can be postponed until after a sequence of several
multiplies\footnote{Using unnormalized numbers for math is not a new idea; a
good example of it is the Control Data CDC 6600, designed by Seymour Cray.
\cite{thornton:cdc6600} The CDC 6600 had all of its instructions performing
unnormalized arithmetic, with a separate {\tt NORMALIZE} instruction.}.

% This is an example of how you would use tgrind to include an example
% of source code; it is commented out in this template since the code
% example file does not exist.  To use it, you need to remove the '%' on the
% beginning of the line, and insert your own information in the call.
%
%\tagrind[htbp]{code/pmn.s.tex}{Post Multiply Normalization}{opt:pmn}

As you can see, the intermediate results can be multiplied together, with no
need for intermediate normalizations due to the guard bit.  It is only at
the end of the operation that the normalization must be performed, in order
to get it into a format suitable for storing in memory\footnote{Note that
for purposed of clarity, the pipeline delays were considered to be 0, and
the branches were not delayed.}.

\subsection{Block Exponent}

In a unoptimized sequence of additions, the sequence of operations is as
follows for each pair of numbers ($m_1$,$e_1$) and ($m_2$,$e_2$).
\begin{enumerate}
  \item Compare $e_1$ and $e_2$.
  \item Shift the mantissa associated with the smaller exponent $|e_1-e_2|$
        places to the right.
  \item Add $m_1$ and $m_2$.
  \item Find the first one in the resulting mantissa.
  \item Shift the resulting mantissa so that normalized
  \item Adjust the exponent accordingly.
\end{enumerate}

Out of 6 steps, only one is the actual addition, and the rest are involved
in aligning the mantissas prior to the add, and then normalizing the result
afterward.  In the block exponent optimization, the largest mantissa is
found to start with, and all the mantissa's shifted before any additions
take place.  Once the mantissas have been shifted, the additions can take
place one after another\footnote{This requires that for n consecutive
additions, there are $\log_{2}n$ high guard bits to prevent overflow.  In
the $\mu$FPU, there are 3 guard bits, making up to 8 consecutive additions
possible.}.  An example of the Block Exponent optimization on the expression
X = A + B + C is given in figure~\ref{opt:be}.

% This is an example of how you would use tgrind to include an example
% of source code; it is commented out in this template since the code
% example file does not exist.  To use it, you need to remove the '%' on the
% beginning of the line, and insert your own information in the call.
%
%\tgrind[htbp]{code/be.s.tex}{Block Exponent}{opt:be}

\section{Integer optimizations}

As well as the floating point optimizations described above, there are
also integer optimizations that can be used in the $\mu$FPU.  In concert
with the floating point optimizations, these can provide a significant
speedup.

\subsection{Conversion to fixed point}

Integer operations are much faster than floating point operations; if it is
possible to replace floating point operations with fixed point operations,
this would provide a significant increase in speed.

This conversion can either take place automatically or or based on a
specific request from the programmer.  To do this automatically, the
compiler must either be very smart, or play fast and loose with the accuracy
and precision of the programmer's variables.  To be ``smart'', the computer
must track the ranges of all the floating point variables through the
program, and then see if there are any potential candidates for conversion
to floating point.  This technique is discussed further in
section~\ref{range-tracking}, where it was implemented.

The other way to do this is to rely on specific hints from the programmer
that a certain value will only assume a specific range, and that only a
specific precision is desired.  This is somewhat more taxing on the
programmer, in that he has to know the ranges that his values will take at
declaration time (something normally abstracted away), but it does provide
the opportunity for fine-tuning already working code.

Potential applications of this would be simulation programs, where the
variable represents some physical quantity; the constraints of the physical
system may provide bounds on the range the variable can take.
\subsection{Small Constant Multiplications}

One other class of optimizations that can be done is to replace
multiplications by small integer constants into some combination of
additions and shifts.  Addition and shifting can be significantly faster
than multiplication.  This is done by using some combination of
\begin{eqnarray*}
a_i & = & a_j + a_k \\
a_i & = & 2a_j + a_k \\
a_i & = & 4a_j + a_k \\
a_i & = & 8a_j + a_k \\
a_i & = & a_j - a_k \\
a_i & = & a_j \ll m \mbox{shift}
\end{eqnarray*}
instead of the multiplication.  For example, to multiply $s$ by 10 and store
the result in $r$, you could use:
\begin{eqnarray*}
r & = & 4s + s\\
r & = & r + r
\end{eqnarray*}
Or by 59:
\begin{eqnarray*}
t & = & 2s + s \\
r & = & 2t + s \\
r & = & 8r + t
\end{eqnarray*}
Similar combinations can be found for almost all of the smaller
integers\footnote{This optimization is only an ``optimization'', of course,
when the amount of time spent on the shifts and adds is less than the time
that would be spent doing the multiplication.  Since the time costs of these
operations are known to the compiler in order for it to do scheduling, it is
easy for the compiler to determine when this optimization is worth using.}.
\cite{magenheimer:precision}

\section{Other optimizations}

\subsection{Low-level parallelism}

The current trend is towards duplicating hardware at the lowest level to
provide parallelism\footnote{This can been seen in the i860; floating point
additions and multiplications can proceed at the same time, and the RISC
core be moving data in and out of the floating point registers and providing
flow control at the same time the floating point units are active. \cite{byte:i860}}

Conceptually, it is easy to take advantage to low-level parallelism in the
instruction stream by simply adding more functional units to the $\mu$FPU,
widening the instruction word to control them, and then scheduling as many
operations to take place at one time as possible.

However, simply adding more functional units can only be done so many times;
there is only a limited amount of parallelism directly available in the
instruction stream, and without it, much of the extra resources will go to
waste.  One process used to make more instructions potentially schedulable
at any given time is ``trace scheduling''.  This technique originated in the
Bulldog compiler for the original VLIW machine, the ELI-512.
\cite{ellis:bulldog,colwell:vliw}  In trace scheduling, code can be
scheduled through many basic blocks at one time, following a single
potential ``trace'' of program execution.  In this way, instructions that
{\em might\/} be executed depending on a conditional branch further down in
the instruction stream are scheduled, allowing an increase in the potential
parallelism.  To account for the cases where the expected branch wasn't
taken, correction code is inserted after the branches to undo the effects of
any prematurely executed instructions.

\subsection{Pipeline optimizations}

In addition to having operations going on in parallel across functional
units, it is also typical to have several operations in various stages of
completion in each unit.  This pipelining allows the throughput of the
functional units to be increased, with no increase in latency.

There are several ways pipelined operations can be optimized.  On the
hardware side, support can be added to allow data to be recirculated back
into the beginning of the pipeline from the end, saving a trip through the
registers.  On the software side, the compiler can utilize several tricks to
try to fill up as many of the pipeline delay slots as possible, as
seendescribed by Gibbons. \cite{gib86}
